### 2023-01-15 paper/code/paper.tex
create proper gitignore + add paper

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Territory size, governance and economic performance: evidence from German dynasties}
\author{Elias Hadj Ammar}
\date{January 2023}

\begin{document}

\maketitle

Alternative title ideas: 
\begin{itemize}
    \item "Good riddance? What happens to states when dynasties go extinct" - catchy, but obscure, and also not what the paper is about.
    \item "What kind of ruler makes states prosper?" "The best kind of autocrat?"
    \item "How individual dynasties in the HRE promoted or hindered growth" - to the point, but detached from literature.
    \item Winners and losers
    \item Good riddance
    \item Thinly veiled dick joke
    \item Preview of results
\end{itemize}

\section{Introduction}

\end{document}

### 2023-01-15 paper/paper.tex
untrack latex auxiliary files

--- 2023-01-15 paper/code/paper.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage[a4paper,left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
 
-\title{Territory size, governance and economic performance: evidence from German dynasties}
+\usepackage{natbib} % we need this so we can use citation and bib properly
+\usepackage{enumitem} % allow you to customize your items (like margins etc.)
+\usepackage{xurl} % allow URL breaks at any alphanumerical character
+\usepackage{hyperref} % allows you to add hyperlink
+\usepackage{amsmath} % allows you to use most mathematical features
+\usepackage{float,fancyhdr}
+\usepackage{amssymb}
+\usepackage{setspace} % allows you to change the line spacing
+\usepackage{xcolor}
+\usepackage{graphicx} % we need this so we can add figures
+
+\title{Territory size, governance and economic performance: \\Evidence from German dynasties}
 \author{Elias Hadj Ammar}
@@
 \maketitle
+\thispagestyle{empty}
 
@@
 
+\newpage
+
+\setcounter{page}{1}
+
 \section{Introduction}
 
+This is like \cite{alesinaS1997}
+
+\newpage
+\bibliographystyle{chicago}
+\bibliography{references.bib}
+
 \end{document}

### 2023-01-15 build/code/exploration.Rmd
add R markdown for exploration

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
```

## Ruler changes, family extinctions and territory outcomes

Questions I can think of:
- how are family extinctions different from regular successions?
- how can I get a list of these events from the data? How many are there?
- what kind of outcome data can I find on a city or territory level?

```{r data import}
territories <- read.csv("build/input/territories.csv")
summary(territories)
```

Is the variable `sovereign_number` the same as `ruler_id` in my data? The first thing to do is really really understand the data I have at my disposal. If the published data set doesn't have what I need, I would have to talk to Matthias and Prof. Cantoni about using the unpublished data.

First of all, drop a shitton of useless variables.

```{r variable selection}
territories <- territories %>%
  select(city_id, terr_id, sovereign_number, time_point, observation, point_id, 
         type_reign, type_change, city_status, foreign_rule)

View(territories)
```

From just viewing this, I can already see that this is not the data I want, at least not fully. 
These are city-year level observations of `terr_id`, that is, of ruling dynasty. So this - the already-published `territories` data - is basically a key to link `rulers` and `cities` data. I know that the `rulers` data exist, because I worked on them, and I am sure there is city-level data on outcomes like economic performance (wages, population, urbanisation) for the HRE out there.

So I need to 
- ask Prof. Cantoni for permission to use the `rulers` data
- identify and filter out family extinction events
- compile my own `outcomes` data set from other sources
- match it to the cities in the published `territories` data
- run a staggered diff-in-diff on that.

### 2023-01-16 build/code/exploration/data_structure.Rmd
add exploration and GetRulers.R

---
title: "Target for the data"
output: html_document
date: "2023-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(haven)
```

## What would I run the regressions on?

The regression would look something like this:

```{r cars}
reg1 <- lm(pop ~ city_fe + year_fe + treat*post + controls)
```

Ideally, I want a city-year level dataset with an entry for every year, based on the ruler dataset, not the territories dataset. For any given year, I want to know immediately not only the city's ruling lineage, but also the ruler. And then, for these city-year observations, I can add my outcome variables.

## Relational database approach

How do I do that? It seems to me that this kind of works like a relational database.
Each city-year observation must have non-missing lineage. Each lineage-year observation must have a non-missing ruler. That's how I could get there.

### Lineage for all city-year observations

But first let me check whether this is already in `cities_families.dta`.

```{r cities data}
cities <- read_dta('build/input/cities_families_1300_1918.dta')
summary(cities)
view(cities)
```

Okay, perfect. That file already contains the lineages for all cities for all years. At first glance it seems to be exactly what I would have had to do with the territories data - expand it so there's an observation for each year, based on the exact years in the commentary fields. That would have been a huge pain.

### Ruler for all lineage-year observations

The file `family_rulers_imputed.dta` is a list of ruler-level observations. These rulers happen to be in chronological order, but they're not lineage-year observations. Yet. How do I get them there?

```{text pseudo-code}
lineages <- expand(terr_id x year)
lineages <- mutate(ruler_id = case_when(
  year %in% start_reign:end_reign ~ ruler_id
))
```

Something like that may get the job done, but it seems computationally intensive and unelegant. Mathematically, what am I trying to do? I'm not trying to map lineage-year observations to rulers. I'm

Maybe do a function, or a dictionary. You input a terr_id and a year. It filters the rulers data down to the terr_id. Then it drops all rulers with reigns ending before or starting after that year. it returns the ruler of that lineage of that year.

Then `map` year-lineage entries to rulers with that function.

### 2023-01-16 build/code/exploration/lineage_extinctions.Rmd
add exploration and GetRulers.R

---
title: "extinction events"
output: html_document
date: "2023-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(haven)
```

## Identifying family extinctions

```{r data import}
people <- read_dta("build/input/families_rulers_imputed.dta")
people$terr_id <- na_if(people$terr_id, '')

rulers <- people %>% 
  filter(!is.na(terr_id))

territories <- read.csv("build/input/territories.csv")
```

Okay. I have a list of all rulers. How do I get all extinction events? What *is* an extinction event, really?

I want all rulers whose successor is not from their own family.

---
Idea 1: start out from the `territories` data. This is a dataset at the city-year level. Make a list of every instance of a city switching `terr_id` between two time points. A city switch consists of a `city_id`, a `time_point`, the old `terr_id`, the new `terr_id`, and the `type_change`. To be clear, a city switch can be due to extinction, due to conquest, Then take this list of city-year observations over to the `rulers` data and

So. The `territories` data actually already codes extinction of lineage. I just need to filter for `type_change == 3`. Then I get all city-year observations in which a new family replaced an extinct one. Let's do that first.

```{r get post observations}
city_events <- territories %>% 
  filter(type_change == 3)

view(city_events)
```

Oh yeah, right. For each family extinction (a family-level event), this will give me *multiple* city-level events. First, because for each extinction, multiple cities change hands. Second, because the event shows up for all years afterwards until a new event happens, I also get all city-year observations *after* the city event.

### 2023-01-16 build/code/GetRulers.R

@@ -1,0 +1,26 @@
+library(tidyverse)
+library(haven)
+
+setwd("~/GitHub/BA")
+
+Main <- function(){
+  people <- ImportPeople()
+  rulers <- FilterRulers(people)
+  rulers %>% 
+    mutate(death_year = death_year+1, end_reign = end_reign+1) %>% 
+    write.csv("build/temp/rulers.csv")
+}
+
+ImportPeople <- function(){
+  people <- read_dta("build/input/families_rulers_imputed.dta")
+  return(people)
+}
+
+FilterRulers <- function(people){
+  people$terr_id <- na_if(people$terr_id, '')
+  rulers <- people %>% 
+    filter(!is.na(terr_id))
+  return(rulers)
+}
+
+Main()

### 2023-01-16 build/code/exploration/ExtinctionsFromTerritories.Rmd
first try to get extinctions from terr data

---
title: "Triangulating extinctions"
output: html_document
date: "2023-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(haven)
```

I have two separate sources for extinction events:
- the territories data (`observation` with `type_change`=3)
- the rulers data (final observation of a `terr_id`)
How much do they align?

```{r look at territories data}
territories <- read.csv("build/input/territories.csv")
takeovers <- territories %>% 
  select(city_id, terr_id, sovereign_number, time_point, observation, point_id, 
         type_reign, type_change, city_status, foreign_rule) %>% 
  filter(type_change==3) %>% 
  group_by(observation) %>% 
  summarise(first_year = min(time_point))
```

And with that, I might already have a list of all ruler extinctions in the territory data - the events are there, I just need to extract the previous territory (which will probably be a pain). Unless an observation can refer to multiple terr_ids. Then nevermind.

```{r check about observations and terrs}
territories <- read.csv("build/input/territories.csv")
takeovers <- territories %>% 
  select(city_id, terr_id, sovereign_number, time_point, observation, point_id, 
         type_reign, type_change, city_status, foreign_rule) %>% 
  filter(type_change==3) %>% 
  group_by(observation, time_point) %>% 
  summarise(count = n()) %>% summary()
print(takeovers)

```
Okay no, perfect. An observation only ever corresponds to one territory. Unless I did something wrong, which I definitely won't rule out. But if I did, I'm guessing / hoping that bugs will bring it to my attention.
This means I can use the observations as a key to first get the rest of the information back (year, dynasty usurping the city) and then the original dynasty as well.

I love that we get the exact cities from the territories data and the exact timing from the rulers data and the exact outcomes from the cities data.

```{r attempt at extracting previous territory}
territories %>% 
  select(city_id, terr_id, sovereign_number, time_point, observation, point_id, 
         type_reign, type_change, city_status, foreign_rule) %>%

```

### 2023-01-16 build/code/exploration/data_structure.Rmd
first try to get extinctions from terr data

@@
 ```{r cars}
-reg1 <- lm(pop ~ city_fe + year_fe + treat*post + controls)
+reg1 <- lm(pop ~ city_fe + year_fe + treat*post + treat*post*channel + controls)
 ```
+
+In the end, `treat`x`post` will be 1 for all city-years after that city's extinction year. That's my intuition. But I haven't read up about staggered diff-in-diff yet.
 
@@
 
-Something like that may get the job done, but it seems computationally intensive and unelegant. Mathematically, what am I trying to do? I'm not trying to map lineage-year observations to rulers. I'm
+I did something like that in `RulersToTerritoryYears.R`.
+Surprisingly, only about one-third of territory-year observation have a corresponding ruler who was alive at the time.
+But I don't think that should be an issue.
+I should look at some descriptive stats.
 
-Maybe do a function, or a dictionary.
-You input a terr_id and a year.
-It filters the rulers data down to the terr_id.
-Then it drops all rulers with reigns ending before or starting after that year.
-it returns the ruler of that lineage of that year.
+```{r descriptive stats of rulers_linked}
+rulers <- read.csv("build/temp/rulers_linked.csv")
+print(rulers %>% group_by(terr_id) %>% summarise(years = n()) %>% summary())
 
-Then `map` year-lineage entries to rulers with that function.
+```
+The median dynasty in my data lasted 106 years. That seems plausible.
+
+```{r more descriptive stats}
+print(rulers %>% group_by(year) %>% summarise(territories = n()) %>% summary())
+```
+If I did it right, this means that there were on average 75 dynasties with rulers coexisting in the same year. Also plausible.
+
+It would be interesting why the other 100,000 dynasty-year observations fell through.
+Is it that some years (e.g. early years) are often missing rulers?
+Is it that some dynasties (e.g. small ones) are often missing rulers?
+Could there be an explanation other than data quality?
+This is a potential source of bias; I should address it.
+Maybe limit the analysis to cities in territories larger than a certain threshold, and to recent centuries in which there are few missing links.
+
+One more note: `rulers_linked.csv` essentially contains the personal history of all these dynasties.
+It contains the extinction events and, crucially, the last ruler of that dynasty, whose characteristics I can link to the event from elsewhere.
+#

### 2023-01-16 build/code/ExtinctionsFromTerritories.R

@@ -1,0 +1,21 @@
+library(tidyverse)
+
+setwd("~/GitHub/BA")
+
+
+ImportTerritories <- function(){
+  territories <- read.csv("build/input/territories.csv")
+  territories <- territories %>% 
+     select(city_id, terr_id, sovereign_number, time_point, observation, point_id, 
+          type_reign, type_change, city_status, foreign_rule)
+  return(territories)
+}
+
+ExtinctionObservationsTiming <- function(territories){
+  observations <- territories %>% 
+    filter(type_change == 3) %>% 
+    group_by(observation) %>% 
+    summarise(first_year = min(time_point))
+  return(observations)
+}
+

### 2023-01-17 build/code/exploration/PopData.Rmd
add population data (not cited yet)

---
title: "Europop Data (De Vries 1984)"
output: html_document
date: "2023-01-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(haven)
```

## First look at the population data

```{r first look}
devries <- read.csv("build/input/devries1984.csv")
bairoch <- read.csv("build/input/bairoch1988.csv")
bairoch %>% drop_na(population) %>% view()
```

Um. Okay, that's really not super granular. 50 is the coarsest I would have gone, and it is apparently also the finest I can go. But it will do. I just need to limit my analysis to only those extinction events with at least 50 or 100 years of stability afterwards.

## Extracting the cities I want (all in `cities`)

## Adding a population variable to the city-year observations

### 2023-01-17 build/code/exploration/WageData.Rmd
add population data (not cited yet)

---
title: "cities with wage data"
output: html_document
date: "2023-01-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(haven)
```

## Cities with wage data

See below for a list of cities in the wage data (also in my data if `city_id` not NA):

```{r overview of Allen's data}
wage_list <- read.csv("build/input/CitiesWagesList.csv")
wage_list %>% drop_na(crafts) %>% view()
```

I'm thinking I should just manually add Vienna as a city to the build. It should be easy to verify that it never changes territory and always belongs to the main line of the Habsburgs. Adding it anyway may help with estimating year fixed effects (see Olken). But which `terr_id` are the main Habsburgs?

Anyways, the craftsmen data contains more cities - Augsburg, Leipzig, Munich, Vienna, Gdansk, Hamburg. The labourer data is missing Munich and Hamburg, so I guess the crafts wages will have to do. And since they're available for such a small amount of cities (some of which probably won't have any usable extinction events), I definitely should rely on population / urbanisation for the most part, and only use wages for specific things. For example, I could use it for a short-term event study in the style of Jones and Olken.

### 2023-01-20 paper/paper.tex
test citations and shorten labels
more paper experiments
add new ideas for titles

@@
 
-\title{Territory size, governance and economic performance: \\Evidence from German dynasties}
+\title{When Cities Change Hands: Extinctions of Dynasties and \\Breaks in Development}
 \author{Elias Hadj Ammar}
@@
 \begin{document}
-
 \maketitle
-\thispagestyle{empty}
-
+\footnote{I gratefully acknowledge helpful comments from Davide Cantoni, Matthias Weigand, and my fellow research assistants at the Chair of Economic History.}
+Keep in mind that the title doesn't need to say everything. It needs to be catchy and it needs to be not-misleading about the core of the paper.
 Alternative title ideas: 
@@
     \item "What kind of ruler makes states prosper?" "The best kind of autocrat?"
+    \item ``Territory size, leaders and economic development: Evidence from German dynasties'' {Leaders, territory size and growth: Evidence from extinctions of German dynasties}
+
     \item "How individual dynasties in the HRE promoted or hindered growth" - to the point, but detached from literature.
@@
     \item Good riddance
-    \item Thinly veiled dick joke
+    \item Thinly veiled dick joke about size
     \item Preview of results
+    \item Which country you belong to / ``When cities change hands''
+    \item {Extinctions of dynasties and changes in development}
+    \item {When Cities Change Hands: Extinctions of Dynasties and Changes in Development}
+    \item Family Affair / Family Matters / Family Business / something more fitting about families
+    \item Successions and success
+    \item {Family Matters: Extinctions of Dynasties and Changes in Development}
+    \item something about game of thrones / a silly pun
 \end{itemize}
@@
 
-This is like \cite{alesinaS1997}
+This is like \cite{as1997} and \cite{aw1998} and also \cite{as1997} and also \cite{olken2005}. hi there \cite{olken2005}
 

### 2023-01-31 paper/paper.tex
add theory notebook

@@
 
-\title{When Cities Change Hands: Extinctions of Dynasties and \\Breaks in Development}
+\title{When Cities Change Hands: Extinctions of Dynasties and \\Breaks in the Development of N German Cities}
 \author{Elias Hadj Ammar}
@@
 \maketitle
-\footnote{I gratefully acknowledge helpful comments from Davide Cantoni, Matthias Weigand, and my fellow research assistants at the Chair of Economic History.}
+\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}
 Keep in mind that the title doesn't need to say everything. It needs to be catchy and it needs to be not-misleading about the core of the paper.
@@
     \item {Extinctions of dynasties and changes in development}
-    \item {When Cities Change Hands: Extinctions of Dynasties and Changes in Development}
+    \item {When Cities Change Hands: Extinctions of Dynasties and Changes in Development} - very clean but also detached from literature (may be fine, though)
     \item Family Affair / Family Matters / Family Business / something more fitting about families
@@
     \item something about game of thrones / a silly pun
+    \item Adoptive Family / Foster Father or Evil Stepdad / In-laws
+    \item Foster family or evil in-laws [only if many takeovers are by inlaws]
+    \item ... of N German cities
 \end{itemize}

### 2023-01-31 analysis/code/theory.jl

@@ -1,0 +1,32 @@
+### A Pluto.jl notebook ###
+# v0.19.22
+
+using Markdown
+using InteractiveUtils
+
+# ╔═╡ 1b494746-a128-11ed-0c0e-6f5473fb4a8c
+md"""
+
+# Sketch of a theory
+
+What is the theory? There's a set of lineages and a set of cities belonging to each lineage. When a lineage goes extinct, its cities change possession. This means a qualitative change in leadership for these cities. It also means a change in territory size for all cities in the newly enlarged territory. I want to focus on the latter.
+
+Maybe I can do some sort of (aggregate but structural) model of cities and territory sizes. As in, _sketching_ a _very simple_ model of the channels through which territory size affects construction in cities. 
+
+Important: I don't need to come up with this by myself. If I can find something good and relevant in the literature, that would actually be better. Then it's just about exploring and visualising it in a notebook.
+"""
+
+# ╔═╡ 8305abd1-48f0-47c3-84fb-aa0ba930a403
+md"""
+### How to approach this
+- read the papers on territory size for inspiration
+- decide which channels you think are plausible
+- study example notebooks that derive models (like [this one](https://floswald.github.io/julia-bootcamp/09-introsir.html))
+- come up with the - important - *simplest possible way* of modeling small pieces of the problem
+- translate those concepts / aspects into sets of objects and functions
+- continue to tinker with them as you assemble the model
+"""
+
+# ╔═╡ Cell order:
+# ╟─1b494746-a128-11ed-0c0e-6f5473fb4a8c
+# ╟─8305abd1-48f0-47c3-84fb-aa0ba930a403

### 2023-02-16 paper/paper.tex
add abstract and initial writeups

@@
 
-This is like \cite{as1997} and \cite{aw1998} and also \cite{as1997} and also \cite{olken2005}. hi there \cite{olken2005}
+I estimate the effect of state size on development using a panel of construction events in cities of the Holy Roman Empire.
+I exploit variation from lineage extinctions, which caused cities to become part of different lineages' territorial holdings.
+I find that X.
+I discuss economic mechanisms that explain this effect.
 

### 2023-02-17 paper/paper.tex
add aspirational introduction

@@
 
-\title{When Cities Change Hands: Extinctions of Dynasties and \\Breaks in the Development of N German Cities}
+\title{State Size and City Growth: \\Evidence from Dynastic Takeovers}
 \author{Elias Hadj Ammar}
@@
     \item "How individual dynasties in the HRE promoted or hindered growth" - to the point, but detached from literature.
-    \item Winners and losers
     \item Good riddance
-    \item Thinly veiled dick joke about size
+    \item Dynastic Takeovers
+    \item Extinctions of Noble Families
     \item Preview of results
@@
     \item ... of N German cities
+    \item {When Cities Change Hands: Extinctions of Dynasties and \\Breaks in the Development of N German Cities}
 \end{itemize}
@@
 
-I estimate the effect of state size on development using a panel of construction events in cities of the Holy Roman Empire.
+Abstract: I estimate the effect of state size on development using a panel of construction events in cities of the Holy Roman Empire.
 I exploit variation from lineage extinctions, which caused cities to become part of different lineages' territorial holdings.
@@
 I discuss economic mechanisms that explain this effect.
+
+motivation for effects of size
+
+background on the hre / motivation for using city-level data and territory switches
+
+I use a generalised difference-in-differences framework with varying intensity of treatment.
+The treatment is the size difference between the old and the new rulers' territorial holdings.
+I will argue that the takeover constitutes a shock to territory size that is exogenous to development.
+I support this argument using an event study, showing that cities with different takeover experiences did not differ in growth (conditional on city and year fixed effects) before their respective takeovers.
+
+The outcome variable used throughout the analysis is yearly construction activity.
+I go to some lenghts to show that this is a good proxy for growth, comparing the present data from the Deutsches St\"adtebuch to the craftsmen real wage time series by Allen for available cities.
+The two measures generally align.
+
+Results indicate that territory size has a large, significant and positive effect on city growth.
+Cities that joined a twice-as-big territory experienced 10 \% higher construction activity in subsequent years.
+Splitting the sample into cities from big, medium-sized and small territories reveals that the effect is mostly driven by cities from small territories.
+However, cities in big territories generally changed hands less often, making it more difficult to find a meaningful and significant effect in that subsample.
+
+The literature says that benefits of larger states include free trade within a larger area.
+Trade within the HRE was not exactly free.
+I'm not currently aware of data on historical trade flows between German cities.
+However, I use the establishment of historical markets (Cantoni et al) and the presence of Jews (data from Voigtl\"ander and Voth) to investigate this channel.
+I find that trade increases (cities are (insignificantly) more likely to establish markets, and the presence of Jews increases) when cities join larger territories.
 

### 2023-02-19 paper/paper.tex
update intro and references

@@
 
-motivation for effects of size
+The political history of Europe, until fairly recently, was defined by ruling families passing their territorial holdings down in hereditary monarchies.
+Land and cities changed hands often throughout the history of Europe.
+Which territory a city belonged to had economic as well as political consequences, and it is not hard to imagine that the fortunes of whole stretches of land could be decided by such dynastic takeovers.
+For example, (good example of a takeover and its consequences here).
+
+In this paper I investigate changes in territory size as one possible channel through which dynastic takeovers could affect development.
+There exists a literature on the optimal size of state (e.g.
+\cite{as1997}, \cite{aw1998}, \cite{asw2000}, \cite{easterly2000}).
+This literature focuses on the tradeoffs between scale effects in government on one hand and conflict on the other, as well as trade and openness.
 

### 2023-02-23 paper/paper.tex
more lit, save progress on ID, structure paper

@@
 
-\title{State Size and City Growth: \\Evidence from Dynastic Takeovers}
+\title{Does Political Integration Make Cities Grow Faster?\\
+Evidence From Dynastic Takeovers\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}
+}
 \author{Elias Hadj Ammar}
@@
 \maketitle
-\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}
+
+\begin{abstract}
+I use historical data on construction activity and territorial history to estimate the effect of state size on city growth in the Holy Roman Empire.
+I exploit variation from lineage extinctions, which caused cities to become part of different lineages' territorial holdings.
+I find that an X\% increase in territory size is associated with a Y\% increase in construction activity.
+\end{abstract}
+
 Keep in mind that the title doesn't need to say everything. It needs to be catchy and it needs to be not-misleading about the core of the paper.
@@
     \item "How individual dynasties in the HRE promoted or hindered growth" - to the point, but detached from literature.
-    \item Good riddance
     \item Dynastic Takeovers
@@
     \item Successions and success
+    \item Power and prosperity
     \item {Family Matters: Extinctions of Dynasties and Changes in Development}
@@
     \item {When Cities Change Hands: Extinctions of Dynasties and \\Breaks in the Development of N German Cities}
+    \item {State Size and City Growth: Evidence from Dynastic Takeovers}
+    \item {Political Integration and Growth: \\Evidence from Dynastic Takeovers}
     \end{itemize}
@@
 
+\doublespacing
+
 \section{Introduction}
-
-Abstract: I estimate the effect of state size on development using a panel of construction events in cities of the Holy Roman Empire.
-I exploit variation from lineage extinctions, which caused cities to become part of different lineages' territorial holdings.
-I find that X.
-I discuss economic mechanisms that explain this effect.
 
@@
 
-I use a generalised difference-in-differences framework with varying intensity of treatment.
+I use a stacked event study framework with varying intensity of treatment.
 The treatment is the size difference between the old and the new rulers' territorial holdings.
 I will argue that the takeover constitutes a shock to territory size that is exogenous to development.
-I support this argument using an event study, showing that cities with different takeover experiences did not differ in growth (conditional on city and year fixed effects) before their respective takeovers.
+The event study shows that cities with different takeover experiences did not differ in growth (conditional on city and year fixed effects) before their respective takeovers.
 
@@
 
+\section{Cities, lineages, and territories}
+
+\section{Empirical approach}
+
+\subsection{Pooled OLS regression}
+
+\subsection{Obtaining a dataset of stacked natural experiments}
+
+\subsection{Stacked event study and extinction IV}
+
+\section{Results}
+
+\section{Validity and robustness}
+
+\section{Conclusion}
+
 \newpage
@@
 
+\newpage
+\section*{Tables and figures}
+
 \end{document}

### 2023-02-23 paper/paper.tex
make paper look nice

@@
 \begin{document}
+\onehalfspacing
 \maketitle
+\thispagestyle{empty}
 
@@
 
-Keep in mind that the title doesn't need to say everything. It needs to be catchy and it needs to be not-misleading about the core of the paper.
-Alternative title ideas: 
-\begin{itemize}
-    \item "Good riddance? What happens to states when dynasties go extinct" - catchy, but obscure, and also not what the paper is about.
-    \item "What kind of ruler makes states prosper?" "The best kind of autocrat?"
-    \item ``Territory size, leaders and economic development: Evidence from German dynasties'' {Leaders, territory size and growth: Evidence from extinctions of German dynasties}
-
-    \item "How individual dynasties in the HRE promoted or hindered growth" - to the point, but detached from literature.
-    \item Dynastic Takeovers
-    \item Extinctions of Noble Families
-    \item Preview of results
-    \item Which country you belong to / ``When cities change hands''
-    \item {Extinctions of dynasties and changes in development}
-    \item {When Cities Change Hands: Extinctions of Dynasties and Changes in Development} - very clean but also detached from literature (may be fine, though)
-    \item Family Affair / Family Matters / Family Business / something more fitting about families
-    \item Successions and success
-    \item Power and prosperity
-    \item {Family Matters: Extinctions of Dynasties and Changes in Development}
-    \item something about game of thrones / a silly pun
-    \item Adoptive Family / Foster Father or Evil Stepdad / In-laws
-    \item Foster family or evil in-laws [only if many takeovers are by inlaws]
-    \item ... of N German cities
-    \item {When Cities Change Hands: Extinctions of Dynasties and \\Breaks in the Development of N German Cities}
-    \item {State Size and City Growth: Evidence from Dynastic Takeovers}
-    \item {Political Integration and Growth: \\Evidence from Dynastic Takeovers}
-    \end{itemize}
-
 \newpage
@@
 \setcounter{page}{1}
-
 \doublespacing
@@
 \newpage
+\onehalfspacing
 \bibliographystyle{chicago}

### 2023-02-23 build/code/exploration/data_structure.Rmd
inventory of build code

@@
 
-The file `family_rulers_imputed.dta` is a list of ruler-level observations. These rulers happen to be in chronological order, but they're not lineage-year observations. Yet. How do I get them there?
+The file `family_rulers_imputed.dta` is a list of person-level observations. I turn this into a lineage-year panel of rulers in the same way I did for work: filter for non-missing terr_id. 
+This relies on what Matthias told me to do in summer being correct. I should verify that non-rulers in `family_rulers_imputed.dta` have missing terr_id.
+
+Discount the rest of this document. Commit it one more time, then remove it in a housekeeping commit. Also remove `RulersToTerritoryYears.R` in that commit.
+
+This is what I previously thought I had to do (but I think it's unnecessary):
 
@@
 
-I did something like that in `RulersToTerritoryYears.R`.
+This is implemented in `RulersToTerritoryYears.R`.
 Surprisingly, only about one-third of territory-year observation have a corresponding ruler who was alive at the time.

### 2023-02-23 build/code/ExtinctionsFromTerritories.R

@@ -1,3 +1,8 @@
+# Input: the *published* (coarse) territories data.
+# Output: the territory-timepoint observations following an extinction.
+# Note: this script can probably be discarded; I want to get extinctions from rulers
+# and takeovers from cities.
+
 library(tidyverse)
 
 setwd("~/GitHub/BA")

### 2023-02-23 build/code/GetRulers.R

@@ -1,3 +1,7 @@
+# Input: list of people.
+# Output: territory-year panel of rulers.
+
+
 library(tidyverse)
 library(haven)
 

### 2023-02-23 build/code/RulersToTerritoryYears.R

@@ -1,3 +1,9 @@
+# Input: ruler list.
+# Output: territory-year panel of ruler id.
+# Note: I no longer need this; I think there is a simpler approach. For the
+# rationale behind what happens in this file, look at `data_structure.rmd`. 
+# Commit one more time with this note, then delete in housekeeping commit.
+
 library(tidyverse)
 library(haven)
 

### 2023-02-23 build/code/exploration/data_structure.Rmd
housekeeping build folder

--- 2023-01-16 build/code/exploration/data_structure.Rmd
@@
 
-## What would I run the regressions on?
-
-The regression would look something like this:
-
-```{r cars}
-reg1 <- lm(pop ~ city_fe + year_fe + treat*post + controls)
-```
-
-Ideally, I want a city-year level dataset with an entry for every year, based on the ruler dataset, not the territories dataset.
-For any given year, I want to know immediately not only the city's ruling lineage, but also the ruler.
-And then, for these city-year observations, I can add my outcome variables.
+I want a city-year panel of
+- construction activity
+- ruling lineage (terr_id)
+- ruling lineage's extinction in that year
+- cities held by ruling lineage in that year
+- controls (capital status, conflict)
 
@@
 
-How do I do that? It seems to me that this kind of works like a relational database.
-Each city-year observation must have non-missing lineage. Each lineage-year observation must have a non-missing ruler. That's how I could get there.
+It seems to me that this kind of works like a relational database. Each city-year observation has non-missing ruling lineage. Each lineage-year observation has non-missing extinction dummy and size.
 
@@
 
-But first let me check whether this is already in `cities_families.dta`.
+This is already in `cities_families.dta`. The lineage is given for every city-year observation.
 
@@
 ```
+### Extinction for all lineage-year observations
 
-Okay, perfect.
-That file already contains the lineages for all cities for all years.
-At first glance it seems to be exactly what I would have had to do with the territories data - expand it so there's an observation for each year, based on the exact years in the commentary fields.
-That would have been a huge pain.
+The file `family_rulers_imputed.dta` is a list of person-level observations. I turn this into a lineage-year panel of rulers in the same way I did for work: filter for non-missing terr_id. 
+(This relies on what Matthias told me to do in summer being correct.
+I should verify my implicit assumptions about the dataset - that non-rulers in `family_rulers_imputed.dta` have missing terr_id, and that the ruler list is therefore a lineage-year panel of rulers.)
 
-### Ruler for all lineage-year observations
+Then I can obtain the extinction year of each lineage by filtering for the last ruler in each lineage (max start_reign).
+This ruler's death_year is the extinction year of the lineage (or where our data ends.
+In a future iteration of this, I could use the `territories.csv` data to cross-validate.)
 
-The file `family_rulers_imputed.dta` is a list of ruler-level observations. These rulers happen to be in chronological order, but they're not lineage-year observations. Yet. How do I get them there?
+### Size for all lineage-year observations
 
-```{text pseudo-code}
-lineages <- expand(terr_id x year)
-lineages <- mutate(ruler_id = case_when(
-  year %in% start_reign:end_reign ~ ruler_id
-))
-```
+Use number of cities. You can recycle code you wrote in the territories project.
 
-Something like that may get the job done, but it seems computationally intensive and unelegant. Mathematically, what am I trying to do? I'm not trying to map lineage-year observations to rulers. I'm
+## Merging / joining the relational data
 
-Maybe do a function, or a dictionary.
-You input a terr_id and a year.
-It filters the rulers data down to the terr_id.
-Then it drops all rulers with reigns ending before or starting after that year.
-it returns the ruler of that lineage of that year.
-
-Then `map` year-lineage entries to rulers with that function.
+Find out how this works once you have constructed everything.

### 2023-02-23 build/code/CombineData.R

@@ -1,0 +1,11 @@
+# Inputs:
+# - city-year panel of lineage and construction.
+# - lineage-year panel of size.
+# - lineage level list of extinction year.
+
+# Output: city-year panel of 
+# - lineage, 
+# - lineage extinction dummy,
+# - lineage size, 
+# - size change from last year,
+# - construction.

### 2023-02-23 build/code/Construction.R

@@ -1,0 +1,2 @@
+# Input: city-year panel of construction from Princes and Townspeople.
+# Output: city-year panel of custom aggregates of construction.

### 2023-02-23 build/code/ExtinctionsFromTerritories.R

@@ -1,26 +1,0 @@
-# Input: the *published* (coarse) territories data.
-# Output: the territory-timepoint observations following an extinction.
-# Note: this script can probably be discarded; I want to get extinctions from rulers
-# and takeovers from cities.
-
-library(tidyverse)
-
-setwd("~/GitHub/BA")
-
-
-ImportTerritories <- function(){
-  territories <- read.csv("build/input/territories.csv")
-  territories <- territories %>% 
-     select(city_id, terr_id, sovereign_number, time_point, observation, point_id, 
-          type_reign, type_change, city_status, foreign_rule)
-  return(territories)
-}
-
-ExtinctionObservationsTiming <- function(territories){
-  observations <- territories %>% 
-    filter(type_change == 3) %>% 
-    group_by(observation) %>% 
-    summarise(first_year = min(time_point))
-  return(observations)
-}
-

### 2023-02-23 build/code/LineageExtinctions.R

@@ -1,0 +1,2 @@
+# Input: lineage-year panel of rulers.
+# Output: lineage-level data set of last ruler (max start_reign) with death_year.

### 2023-02-23 build/code/LineageSize.R

@@ -1,0 +1,2 @@
+# Input: city-year panel of ruling lineage.
+# Output: lineage-year panel of no. of cities held.

### 2023-02-23 build/code/RulersToTerritoryYears.R

@@ -1,105 +1,0 @@
-# Input: ruler list.
-# Output: territory-year panel of ruler id.
-# Note: I no longer need this; I think there is a simpler approach. For the
-# rationale behind what happens in this file, look at `data_structure.rmd`. 
-# Commit one more time with this note, then delete in housekeeping commit.
-
-library(tidyverse)
-library(haven)
-
-setwd("~/GitHub/BA")
-
-Main <- function(){
-  rulers <- ImportRulers()
-  rulers_lite <<- SelectVariables(rulers) # global is necessary I tried everything
-  cities <- ImportCities()
-  lineage_years <- LineageYearObs(cities)
-  
-  rulers_linked <- MapToRulers(lineage_years) # takes 15-20 minutes
-  WriteSeparateCSV(rulers_linked)
-  
-  return(0)
-}
-
-TestOnSubset <- function(size){
-  rulers <- ImportRulers()
-  rulers_lite <<- SelectVariables(rulers) # same as above
-  cities <- ImportCities() 
-  lineage_years <- LineageYearObs(cities) # longest part if n<1000
-  
-  testdata <- slice_sample(lineage_years, n=size)
-  sample_linked <- MapToRulers(testdata)
-  WriteSeparateCSV(sample_linked)
-  #return(sample_linked)
-  return(0)
-}
-
-ImportRulers <- function(){
-  rulers <- read.csv("build/temp/rulers.csv")
-  return(rulers)
-}
-
-SelectVariables <- function(rulers){
-  rulers_lite <- rulers %>% 
-    select(id, terr_id, start_reign, end_reign)
-  return(rulers_lite)
-}
-
-ImportCities <- function(){
-  cities <- read_dta("build/input/cities_families_1300_1918.dta")
-  return(cities)
-}
-
-LineageYearObs <- function(cities){
-  # Remember, we want to know: Who was head of territory j in year t?
-  # The cities data contain many cities for each lineage-year. I don't need that here.
-  # I want unique lineage-year observations to which I can assign a ruler.
-  result <- cities %>% 
-    select(terr_id, year) %>% 
-    unique()
-  return(result)
-}
-
-CurrentRuler <- function(lineage, year){
-  # We want this to return
-  # - the unique ruler of that year and lineage if there is only one
-  # - the ruler who was in power at the end of the year if there are more
-  # end_reign is the year in which the ruler lost power.
-  # We use this in a map2() call in MapToRulers().
[…]

### 2023-02-23 analysis/code/SummaryStats.Rmd
build lineage size and extinctions

---
title: "SummaryStats"
output: html_document
date: "2023-02-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
```

## Import datasets

```{r import data}
ImportRulers <- function(){
  rulers <- read.csv("build/temp/rulers.csv")
  rulers <- rulers %>% 
    select(id, family, name, terr_id, birth_year, 
           death_year, start_reign, end_reign)
  return(rulers)
}

rulers <- ImportRulers()
```

## Summary statistics: lineages

```{r count rulers for each dynasty}
CountRulers <- function(rulers){
  last_rulers <- rulers %>% 
    group_by(terr_id) %>% 
    count()
}

ruler_counts <- CountRulers(rulers)
ruler_counts %>% View()
```

### 2023-02-23 build/code/LineageExtinctions.R

@@ -1,2 +1,34 @@
 # Input: lineage-year panel of rulers.
 # Output: lineage-level data set of last ruler (max start_reign) with death_year.
+
+library(tidyverse)
+
+setwd("~/GitHub/BA")
+
+Main <- function(){
+  rulers <- ImportRulers()
+  last_rulers <- GetLastRulersDeaths(rulers)
+  last_rulers %>% 
+    write.csv("build/temp/last_rulers.csv")
+}
+
+ImportRulers <- function(){
+  rulers <- read.csv("build/temp/rulers.csv")
+  rulers <- rulers %>% 
+    select(id, family, name, terr_id, birth_year, 
+           death_year, start_reign, end_reign)
+  return(rulers)
+}
+
+GetLastRulersDeaths <- function(rulers){
+  # Slightly different results when you filter by max start_reign (one-year reigns).
+  # Strictly speaking, what we care about is the death of the last family member,
+  # not the death of the last ruler. They should be the same though.
+  last_rulers <- rulers %>% 
+    group_by(terr_id) %>% 
+    mutate(last_death_year = max(death_year)) %>% 
+    filter(death_year == last_death_year)
+  return(last_rulers)
+}
+
+Main()

### 2023-02-23 build/code/LineageSize.R

@@ -1,2 +1,34 @@
 # Input: city-year panel of ruling lineage.
 # Output: lineage-year panel of no. of cities held.
+
+library(tidyverse)
+library(haven)
+
+setwd("~/GitHub/BA")
+
+Main <- function(){
+  cities <- ImportCities()
+  terr_sizes <- CountCities(cities)
+  terr_sizes %>% 
+    write.csv("build/temp/terr_sizes.csv")
+}
+
+
+ImportCities <- function(){
+  cities <- read_dta("build/input/cities_families_1300_1918.dta")
+  cities <- cities %>% 
+    select(city_id, year, terr_id, foreign_rule)
+  return(cities)
+}
+
+CountCities <- function(cities){
+  terr_sizes <- cities %>% 
+    group_by(terr_id, year) %>% 
+    summarise(count_cities = n())
+  return(terr_sizes)
+}
+
+Main()
+
+
+

### 2023-02-28 analysis/code/Exploration.Rmd
#1 first attempt at exploration

## Exploration

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(cowplot)
setwd('~/GitHub/BA')
```

```{r data import}
data <- read.csv("analysis/input/build.csv")
```

### Observations

```{r observations across time}
ggplot(data) + 
  geom_bar(aes(x=year))
```

Looks fairly smooth except for the big jump around 1800.

It's promising that there's a large number of observations (which should correspond to cities 1:1) for every single year.

```{r duplicate keys}
obs_vs_distinct <- data %>% count - data %>% distinct() %>% count()
obs_vs_city_year <- data %>% count() - data %>% distinct(city_id, year) %>% count()

sprintf("%s, %s, %s", obs_vs_distinct, obs_vs_city_year, (obs_vs_city_year/data %>% count))

```

There are no duplicate rows, but there are 3708 observations with duplicate city-year keys.
These observations make up 0.3% of the data set. I can probably drop them? Unless they contain all the extinction events for some reason.

### Lineages

```{r lineage level data}
terrs <- data %>% 
  distinct(terr_id, year, .keep_all=TRUE)
```

```{r number of lineages across time}
yearly_plot <- terrs %>% 
  ggplot(aes(x=year))+
  geom_bar()

decades_plot <- terrs %>% 
  mutate(decade = year - year %% 10) %>% # averaging would be better
  distinct(decade, terr_id) %>% # pretty sure this is necessary
  ggplot(aes(x=decade))+
  geom_bar()

plot_grid(yearly_plot, decades_plot, labels = c("Yearly", "Decades"))
```

This counts the number of lineages that show up in that year or decade.

That is pretty cool. The most striking thing is Napoleon, of course.
This is the Number of Nations. I also want the Size of Nations.

```{r territory size across time}
terrs_centuries <- terrs %>% 
  mutate(century = year - year %% 100) %>% 
  mutate(century = factor(century)) # necessary or else year = width

# create plot with outliers
size_time <- terrs_centuries %>% 
  ggplot() + 
  geom_boxplot(aes(x = century, y = count_cities))

# create plot with no outliers
size_time_no_outliers <- terrs_centuries %>% 
  ggplot() + 
  geom_boxplot(aes(x = century, y = count_cities), outlier.shape = NA) +
  coord_cartesian(ylim = quantile(terrs_centuries$count_cities, c(0, 0.96)))

# plot side by side
plot_grid(size_time, size_time_no_outliers, labels = c("Full plot", "Zoom"))
```

Note: grouping by century and calculating average size of each territory across that century would be better than flooring the year of each observation and dropping duplicates. The way it's currently implemented only leaves the earliest count_cities observation for a territory in a century.

### Extinctions

```{r extinction treatments across time}
data %>%
  filter(extinction == 1) %>% 
  ggplot(aes(x=year))+
  geom_bar()
```

(make a second plot side-by-side to the above with century bins)

The lineages for which an extinction year is recorded, how large are they compared to others?

```{r extinct vs non-extinct}
terrs_life <- terrs %>% 
  group_by(terr_id) %>% 
  summarise(
    extinction_ever = max(extinction), 
    max_size = max(count_cities), # compare maximum size across lifespan
    lifespan = max(year) - min(year)
    ) %>% 
  mutate(extinction_ever = factor(extinction_ever)) # necessary or else year = width

size_ext <- terrs_life %>% 
  ggplot() + 
  geom_boxplot(aes(x = extinction_ever, y=max_size))

size_ext_no_outliers <- terrs_life %>% 
  ggplot() + 
  geom_boxplot(aes(x = extinction_ever, y=max_size), outlier_shape = NA) +
  coord_cartesian(ylim = quantile(terrs_life$max_size, c(0, 0.9)))

plot_grid(size_ext, size_ext_no_outliers, labels = c("Full plot", "Zoom"))
```

How is the lifespan of lineages distributed?

```{r lifespan of lineages}
terrs_life %>% 
  ggplot()+
  geom_histogram(aes(x=lifespan))
```

Left: Is the lifespan of a lineage correlated to its (maximum lifetime) size? 
Right: Are lineages less likely to go extinct when they are larger? (extinction on contemp. size)

```{r}
lifespan_on_max_size <- terrs_life %>% 
  ggplot(aes(x=log(max_size), y=lifespan))+
  geom_point()
```

### Example cities

### 2023-03-01 analysis/code/DuplicateKeys.Rmd
#1 explore dupe keys issue

## Ensuring that city-year keys are unique

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(cowplot)
library(haven)
```

What I've established in Exploration.Rmd:

- In the current version of build.csv, there are 3708 rows that have duplicate city-year keys.
- It turns out that all these rows have `terr_id` "A1341" (Anhalt-Dessau, younger line).
- It turns out that exactly half of all observations with "A1341" are dupes.
- It turns out that the only things that vary for dupes are `X` (a key that was automatically generated in some join operation) and `count_diff` (the difference between `count_cities` and `lag(count_cities)`, which is calculated rowwise).
- It turns out that for each city-year, only the first dupe has the right `count_diff`.

It does seem like I could simply run distinct(city_id, year) on my current data build and get unique city-year keys without losing any information. But this could happen again, on a larger scale, for more territories, if I change the build code somehow. So I want to find out where the duplication happens - in the original cities data, or at any point in the build pipeline - and fix the bug properly.

```{r data import, message=FALSE}
cities_orig <- read_dta("build/input/cities_families_1300_1918.dta") %>% 
  select(city_id, year, terr_id)

cities <- read_csv("build/output/cities.csv", show_col_types = FALSE)
lineages <- read_csv("build/output/lineages.csv", show_col_types = FALSE)
build <- read_csv("build/output/build.csv", col_types = cols( # necessary because first 1000 are NA
  city="c", 
  real_wage="d", 
  welfare_ratio="d"
  ))
```

### Are there dupes in the original cities data?

```{r check cities_orig}
summary(cities_orig) # no missing terr_id for any city-year - good

cities_orig %>% filter(terr_id == "A1341") %>% count() # the dupes are not in the original file!

# double-check: yes, no duplicate keys in cities_orig.
cities_orig %>% count() - cities_orig %>% distinct(city_id, year) %>% count()
```

### Are there dupes in the output of CombineCityData.R?

No. As of now, the output of that file (cities.csv) has the same city-years as the original data. Also, it uses a left join to add wages to cities - I think this is inherently less problematic.

Just for good measure (and I can just re-run this when cities.csv changes):

```{r check cities.csv}
summary(cities) # yep, same as above

cities %>% filter(terr_id == "A1341") %>% count() # same as above

# no dupes in cities
cities %>% count() - cities %>% distinct(city_id, year) %>% count()
```

This means that the bug must be in CombineAll.R.

### Where in CombineAll.R does the duplication happen?

It must be in CombineTables because as I've seen, the AddSizeDiff already operates on data that contains duplicates. It's the inner join between the lineages and the cities data in CombineTables.

What if I just do a left join instead?

Expected result:

- a data frame that is the same length as cities (1.4 million obs. rather than 1.1 million)
- cities with `terr_id` that doesn't show up in the lineages data have NA values
- no duplicate city-year keys
- summary statistics identical to the previous build (after dropping `count_cities` = NA)

```{r try left join in CombineTables}
CombineTables <- function(cities, lineages){
  joined <- left_join(cities, lineages, by=c("terr_id", "year")) 
  return(joined)
}

new_build <- CombineTables(cities, lineages)
summary(new_build)
```

I did not get the expected result. Why?
I think it must be because "A1341" shows up twice in the lineage data.

### Are there dupes in the lineage data?

```{r check lineage data}
# Yes, there are dupes in lineages.csv:
lineages %>% count() - lineages %>% distinct(terr_id, year) %>% count()

# The 315 dupes in lineages.csv are 315 dupes of terr-years of A1341:
lineages %>% filter(terr_id == "A1341") %>% count()
lineages %>% filter(terr_id == "A1341") %>% distinct(terr_id, year) %>% count()

# The duplicate terr-year keys are full duplicates of rows:
lineages %>% select(-...1) %>% filter(terr_id == "A1341") %>% distinct() %>% count()
```

### Where do the dupes in the lineage data come from?

It must be CombineLineageData.R, and within that file it must be AddExtinctionDummies or AddLineageNames.

Old Hypothesis:
There is no reason to use an inner join in any of the functions there, and using a left join instead will solve the duplicates problem. EXCEPT if there are two last rulers (or two lineage names) for A1341.

New Hypothesis:
In the year Anhalt-Dessau went extinct, there were two rulers.

```{r check last_rulers.csv}
last_rulers <- read_csv("build/temp/last_rulers.csv")

last_rulers %>% filter(terr_id == "A1341") # HAH, I knew it.

# Very interesting - there are *two* lineages with two last rulers.
last_rulers %>% count() - last_rulers %>% distinct(terr_id) %>% count()

# Which ones:
anti_join(last_rulers, last_rulers %>% distinct(terr_id, .keep_all=TRUE), by="...1")

# The other one is of Wessex, terr_id "0W0079". Do we have cities in Wessex? No...
cities_orig %>% filter(terr_id == "0W0079")

# ... which is why this other duplicate extinction doesn't lead to dupes in the build.
```

Solution: modify LineageExtinctions.R so that last_rulers.csv has `terr_id` as a *unique* key. See if that fixes the problem.

### 2023-03-01 analysis/code/Exploration.Rmd
#1 explore dupe keys issue

@@
 library(cowplot)
-setwd('~/GitHub/BA')
 ```
+
+### Data import and summary
 
@@
 data <- read.csv("analysis/input/build.csv")
+
+summary(data)
 ```
 
+Some initial takeaways:
+
+- there are almost 1,100,000 observations in my dataset.
+- the min of final_year is 1101 while the min of year is 1300.
+This means that the ruler dataset is not complete, i.e.
+lack of a recorded ruler doesn't necessarily mean that the lineage is extinct.
+I should definitely cross-check my extinction list with the territories extinction list before picking out sub-experiments.
+- the median final ruler dies in 1920.
+- the share of city-year observations in which the ruling lineage goes extinct is 0.374%.
+
 ### Observations
+
+#### Number of observations per year
 
@@
 
-Looks fairly smooth except for the big jump around 1800.
+Looks fairly smooth except for the big jump around 1800, and the drop around 1350.
 
@@
 
+#### Is each observation a unique city-year?
+
 ```{r duplicate keys}
-obs_vs_distinct <- data %>% count - data %>% distinct() %>% count()
+obs_vs_distinct <- data %>% count() - data %>% distinct() %>% count()
 obs_vs_city_year <- data %>% count() - data %>% distinct(city_id, year) %>% count()
+share_dupes <- obs_vs_city_year/(data %>% count())
 
-sprintf("%s, %s, %s", obs_vs_distinct, obs_vs_city_year, (obs_vs_city_year/data %>% count))
+sprintf("Duplicate rows: %s", obs_vs_distinct)
+sprintf("Rows wih duplicate keys: %s", obs_vs_city_year)
+sprintf("Share of rows with duplicate keys: %s", share_dupes)
 
@@
 There are no duplicate rows, but there are 3708 observations with duplicate city-year keys.
-These observations make up 0.3% of the data set. I can probably drop them? Unless they contain all the extinction events for some reason.
+These observations make up 0.3% of the data set.
+I can probably drop them?
+Unless they are very different from the rest for some reason.
+*0.337% is suspiciously close to the share of city-years with extinctions.*
+
+```{r get obs with dupe keys}
+
+# if there are multiple observations for a city-year key, number them
+data_count_dupes <- data %>% 
+  group_by(city_id, year) %>% 
+  mutate(count_dupe = row_number()) %>% 
+  ungroup()
+
+# show that all observations with duplicate keys have terr_id A1341
+data_count_dupes %>% filter(count_dupe != 1) %>% ggplot() + geom_bar(aes(x=terr_id))
+
+# show that exactly half of all A1341 observations are dupes
+data_count_dupes %>% 
+  filter(terr_id == "A1341") %>% 
+  mutate(count_dupe = factor(count_dupe)) %>% 
+  ggplot(aes(x=count_dupe)) + 
+  geom_bar() +
+  geom_text(stat='count', aes(label=after_stat(count)), vjust=0)
+
+# suspicion: the only thing that varies for dupes is `X` (auto-generated while joining)
+# and `count_diff` (only calculated for the first instance of a city-year observation).
+data %>% filter(terr_id == "A1341") %>% select(-X, -count_diff) %>% distinct() %>% count()
+
+```
+
+Conclusion: for some reason, all observations with terr_id "A1341" got duplicated somewhere along the way. I *could* just drop all the dupes, because the `count_diff` is always with the first version:
+
+```{r}
+data_count_dupes %>% 
+  filter(terr_id == "A1341") %>% 
+  mutate(count_dupe = factor(count_dupe)) %>% 
+  ggplot(aes(x=count_dupe)) + 
+  geom_boxplot(aes(y=count_diff))
+```
+
+But because this isn't actually the final version of the data, and this might happen again, it's probably prudent to check where the duplication occurs. I do this in a separate Rmd.
 
@@
 
+#### Getting lineage-year observations
+
 ```{r lineage level data}
 terrs <- data %>% 
-  distinct(terr_id, year, .keep_all=TRUE)
+  distinct(terr_id, year, .keep_all=TRUE) %>% 
+  select(terr_id, year, territory, count_cities, final_year, extinction)
 ```
+
+Note: this only keeps the first city-year observation for each lineage-year.
+However, we only use this data set to look at variables that are identical for 
+all city-years in a lineage-year, so `count_cities` and `extinction`.
+
+#### Number of distinct lineages across time
 
@@
 
-decades_plot <- terrs %>% 
-  mutate(decade = year - year %% 10) %>% # averaging would be better
-  distinct(decade, terr_id) %>% # pretty sure this is necessary
-  ggplot(aes(x=decade))+
+# count distinct terr_ids that show up in a century
+centuries_plot <- terrs %>% 
+  mutate(century = year - year %% 100) %>% # round year down to century
+  distinct(century, terr_id) %>% # else counts territory-year observations
+  ggplot(aes(x=century))+
   geom_bar()
 
-plot_grid(yearly_plot, decades_plot, labels = c("Yearly", "Decades"))
+plot_grid(yearly_plot, centuries_plot, labels = c("Yearly", "Centuries"))
 ```
 
-This counts the number of lineages that show up in that year or decade.
+That is pretty cool. The most striking thing is Napoleon, of course.
+But it's also important to note that the number of territories continues to
+steeply decline throughout the entire 19th century. Keep that in mind when you
+summarise by century later.
 
-That is pretty cool. The most striking thing is Napoleon, of course.
 This is the Number of Nations. I also want the Size of Nations.
+
+#### Territory size across time
 
@@
   mutate(century = year - year %% 100) %>% 
-  mutate(century = factor(century)) # necessary or else year = width
+  mutate(century = factor(century)) %>% # necessary for boxplots or else year = width
+  group_by(century, terr_id) %>% # group and average because the total no. of cities per year
+  summarise(count_cities = mean(count_cities)) # jumps in 1800; I ASSUME this is an artifact
+  # comment out the last two lines to see the effect of that!
 
@@
 
-Note: grouping by century and calculating average size of each territory across that century would be better than flooring the year of each observation and dropping duplicates.
-The way it's currently implemented only leaves the earliest count_cities observation for a territory in a century.
+It would be good to zoom into the 17th to 20th century and use half-centuries, because this hides the considerable within-century variation in the 1800s.
 
@@
 
+#### Cities affected by lineage extinction across time
 ```{r extinction treatments across time}
-data %>%
+
+CityExtPlot <- function(full_data, bins){
+  plt <- full_data %>% 
     filter(extinction == 1) %>% 
     ggplot(aes(x=year))+
-  geom_bar()
+    geom_histogram(binwidth = bins)
+  return(plt)
+}
+
+plot_grid(CityExtPlot(data, 1), CityExtPlot(data, 10),
+          CityExtPlot(data, 30), CityExtPlot(data, 50))
+
 ```
 
-(make a second plot side-by-side to the above with century bins)
+#### Lineages that go extinct across time
+```{r lineage extinctions across time}
 
-The lineages for which an extinction year is recorded, how large are they compared to others?
+TerrExtPlot <- function(terrs, bins){
+  plt <- terrs %>% 
+    filter(extinction == 1) %>% 
+    ggplot(aes(x=year))+
+    geom_histogram(binwidth = bins)
+  return(plt)
+}
 
+plot_grid(TerrExtPlot(terrs, 1), TerrExtPlot(terrs, 10),
+          TerrExtPlot(terrs, 30), TerrExtPlot(terrs, 50))
+
+```
+
+#### Are larger lineages more or less likely to have a recorded extinction?
 ```{r extinct vs non-extinct}
@@
 
-How is the lifespan of lineages distributed?
+#### Distribution of lineage lifespan
 
@@
   ggplot(aes(x=log(max_size), y=lifespan))+
+  geom_point()+
+  geom_smooth(method="lm", se=FALSE)
+
+ext_on_cities <- terrs %>% 
+  ggplot(aes(x=count_cities, y=extinction))+
   geom_point()
+
+plot_grid(lifespan_on_max_size, ext_on_cities)
+
+#mod_lifespan_size <- lm(lifespan ~ log(max_size), data=terrs_life)
+#summary(mod_lifespan_size)
 ```

### 2023-03-01 build/code/LineageExtinctions.R
#1 fixes duplicate keys bug

@@ -8,7 +8,8 @@ setwd("~/GitHub/BA")
 Main <- function(){
   rulers <- ImportRulers()
   last_rulers <- GetLastRulersDeaths(rulers)
-  last_rulers %>% 
+  extinctions <- RemoveDupes(last_rulers)
+  extinctions %>% 
     write.csv("build/temp/last_rulers.csv")
 }
 
@@ -29,6 +30,20 @@ GetLastRulersDeaths <- function(rulers){
     mutate(last_death_year = max(death_year)) %>% 
     filter(death_year == last_death_year)
   return(last_rulers)
+}
+
+RemoveDupes <- function(last_rulers){
+  # Turns the input into a long dataset at the lineage level.
+  # Drops duplicate terr_ids and adds a dummy for affected lineages.
+  # Needed because some terr_ids have two "last rulers", i.e. the two last family
+  # members died in the same year. If we leave both in, this causes duplication
+  # of lineage-year and city-year observations downstream.
+  extinctions <- last_rulers %>% 
+    group_by(terr_id) %>% 
+    mutate(dupe_rulers = if_else(n() > 1, 1, 0)) %>%
+    ungroup() %>% 
+    distinct(terr_id, .keep_all=TRUE)
+  return(extinctions)
 }
 
 Main()

### 2023-03-01 analysis/code/Exploration.Rmd
#1 update exploration + housekeeping

@@
 
-There are no duplicate rows, but there are 3708 observations with duplicate city-year keys.
-These observations make up 0.3% of the data set.
-I can probably drop them?
-Unless they are very different from the rest for some reason.
-*0.337% is suspiciously close to the share of city-years with extinctions.*
-
-```{r get obs with dupe keys}
-
-# if there are multiple observations for a city-year key, number them
-data_count_dupes <- data %>% 
-  group_by(city_id, year) %>% 
-  mutate(count_dupe = row_number()) %>% 
-  ungroup()
-
-# show that all observations with duplicate keys have terr_id A1341
-data_count_dupes %>% filter(count_dupe != 1) %>% ggplot() + geom_bar(aes(x=terr_id))
-
-# show that exactly half of all A1341 observations are dupes
-data_count_dupes %>% 
-  filter(terr_id == "A1341") %>% 
-  mutate(count_dupe = factor(count_dupe)) %>% 
-  ggplot(aes(x=count_dupe)) + 
-  geom_bar() +
-  geom_text(stat='count', aes(label=after_stat(count)), vjust=0)
-
-# suspicion: the only thing that varies for dupes is `X` (auto-generated while joining)
-# and `count_diff` (only calculated for the first instance of a city-year observation).
-data %>% filter(terr_id == "A1341") %>% select(-X, -count_diff) %>% distinct() %>% count()
-
-```
-
-Conclusion: for some reason, all observations with terr_id "A1341" got duplicated somewhere along the way. I *could* just drop all the dupes, because the `count_diff` is always with the first version:
-
-```{r}
-data_count_dupes %>% 
-  filter(terr_id == "A1341") %>% 
-  mutate(count_dupe = factor(count_dupe)) %>% 
-  ggplot(aes(x=count_dupe)) + 
-  geom_boxplot(aes(y=count_diff))
-```
-
-But because this isn't actually the final version of the data, and this might happen again, it's probably prudent to check where the duplication occurs. I do this in a separate Rmd.
+Yes. each observation has a unique city-year key.
 
@@
   select(terr_id, year, territory, count_cities, final_year, extinction)
+
+# terrs <- read_csv("build/output/lineages.csv") # comment out for sanity check
 ```
@@
 However, we only use this data set to look at variables that are identical for 
-all city-years in a lineage-year, so `count_cities` and `extinction`.
+all city-years in a lineage-year, so `count_cities` and `extinction`. I did a sanity check: the following plots look exactly the same for the lineages.csv data.
 
@@
 
-Left: Is the lifespan of a lineage correlated to its (maximum lifetime) size? 
-Right: Are lineages less likely to go extinct when they are larger? (extinction on contemp. size)
+#### Correlation between size and survival
 
 ```{r}
+#mod_lifespan_size <- lm(lifespan ~ log(max_size), data=terrs_life)
+#summary(mod_lifespan_size)
+
 lifespan_on_max_size <- terrs_life %>% 
@@
 
-#mod_lifespan_size <- lm(lifespan ~ log(max_size), data=terrs_life)
-#summary(mod_lifespan_size)
 ```
 
-### Example cities
+Left: The lifespan of a lineage is weakly correlated to its maximum lifetime size.
+Right: This honestly doesn't say much. What it tries to say is that lineages may be less likely to go extinct when they are larger.
+
+### Cities
+
+#### Lifetime summary statistics for cities
+
+```{r}
+cities_life <- data %>% 
+  group_by(city_id) %>% 
+  summarise(
+    years_exists = max(year) - min(year),
+    rulers_count = n_distinct(terr_id),
+    extinctions_count = sum(extinction),
+    diffs_count = sum(count_diff != 0, na.rm=TRUE),
+    mean_integ = mean(count_cities),
+    lifetime_integ_diff = sum(count_diff, na.rm=TRUE),
+    mean_yearly_diff = mean(count_diff, na.rm=TRUE),
+    sd_yearly_diff = sd(count_diff, na.rm=TRUE)
+    )
+
+# check whether diffs_count works:
+# only1001 <- data[data$city_id == 1001,]
+# sum(only1001$count_diff != 0, na.rm = TRUE) # there are 57 nonzero countdiffs
+# cities_life %>% filter(city_id == 1001) %>% select(diffs_count) # yes it works
+
+summary(cities_life)
+```
+
+This is very insightful. Some takeaways:
+
+- I have almost 500 observations on average for each city.
+- **A normal city changes hands 3-4 times throughout its history.**
+- **Of these territory changes, 1-2 are due to extinction.**
+- **A normal city undergoes around 60 changes in integration.**
+- Less than a quarter of cities is integrated with less than 100 other cities on average throughout its history.
+- Most cities become drastically more integrated over time.
+- Yearly differences in integration are positive on average but vary a lot.

### 2023-03-01 analysis/code/theory.jl

@@ -1,32 +1,0 @@
-### A Pluto.jl notebook ###
-# v0.19.22
-
-using Markdown
-using InteractiveUtils
-
-# ╔═╡ 1b494746-a128-11ed-0c0e-6f5473fb4a8c
-md"""
-
-# Sketch of a theory
-
-What is the theory? There's a set of lineages and a set of cities belonging to each lineage. When a lineage goes extinct, its cities change possession. This means a qualitative change in leadership for these cities. It also means a change in territory size for all cities in the newly enlarged territory. I want to focus on the latter.
-
-Maybe I can do some sort of (aggregate but structural) model of cities and territory sizes. As in, _sketching_ a _very simple_ model of the channels through which territory size affects construction in cities. 
-
-Important: I don't need to come up with this by myself. If I can find something good and relevant in the literature, that would actually be better. Then it's just about exploring and visualising it in a notebook.
-"""
-
-# ╔═╡ 8305abd1-48f0-47c3-84fb-aa0ba930a403
-md"""
-### How to approach this
-- read the papers on territory size for inspiration
-- decide which channels you think are plausible
-- study example notebooks that derive models (like [this one](https://floswald.github.io/julia-bootcamp/09-introsir.html))
-- come up with the - important - *simplest possible way* of modeling small pieces of the problem
-- translate those concepts / aspects into sets of objects and functions
-- continue to tinker with them as you assemble the model
-"""
-
-# ╔═╡ Cell order:
-# ╟─1b494746-a128-11ed-0c0e-6f5473fb4a8c
-# ╟─8305abd1-48f0-47c3-84fb-aa0ba930a403

### 2023-03-02 analysis/code/Exploration.Rmd
clarify deaths, extinctions, treatments

@@
   distinct(terr_id, year, .keep_all=TRUE) %>% 
-  select(terr_id, year, territory, count_cities, final_year, extinction)
+  select(terr_id, year, territory, count_cities, final_full_year, extinction)
 
@@
   geom_boxplot(aes(x = century, y = count_cities), outlier.shape = NA) +
-  coord_cartesian(ylim = quantile(terrs_centuries$count_cities, c(0, 0.96)))
+  coord_cartesian(ylim = quantile(terrs_centuries$count_cities, c(0, 0.93)))
 
@@
   ggplot() + 
-  geom_boxplot(aes(x = extinction_ever, y=max_size), outlier_shape = NA) +
-  coord_cartesian(ylim = quantile(terrs_life$max_size, c(0, 0.9)))
+  geom_boxplot(aes(x = extinction_ever, y=max_size), outlier.shape = NA) +
+  coord_cartesian(ylim = quantile(terrs_life$max_size, c(0, 0.93)))
 

### 2023-03-02 build/code/CombineAll.R

@@ -18,6 +18,7 @@ Main <- function(){
   
   build <- CombineTables(cities, lineages)
   build <- AddSizeDiffs(build)
+  build <- PushExtinctions1Year(build)
   build <- TidyOutput(build)
   
   build %>% write.csv("build/output/build.csv")
@@ -38,10 +39,24 @@ AddSizeDiffs <- function(table){
   return(diffed)
 }
 
+PushExtinctions1Year <- function(table){
+  # reasoning: the extinction dummy shows up in the last year of the 
+  # dying lineage. That was necessary because the extinction list was 
+  # easier to join to the old lineage than to the lineage that takes its 
+  # place.
+  # But I want a treatment dummy to show up in the same year as the
+  # count_diff, i.e. in the first year under the new lineage.
+  pushed <- table %>% 
+    group_by(city_id) %>% 
+    mutate(treatment = lag(extinction)) %>% # there's no way to avoid leading NA
+    ungroup() # because I don't know the previous terr_id of those observations
+  return(pushed)
+}
+
 TidyOutput <- function(table){
   final <- table %>% 
-    select(city_id, year, city, terr_id, territory, final_year, extinction, 
-           count_cities, count_diff, real_wage, welfare_ratio)
+    select(city_id, year, city, terr_id, territory, final_full_year, extinction, 
+           treatment, count_cities, count_diff, real_wage, welfare_ratio)
   return(final)
 }
 

### 2023-03-02 build/code/CombineLineageData.R

@@ -16,9 +16,7 @@ Main <- function(){
   lineages <- read.csv("build/temp/terr_sizes.csv")
   terr_codes <- read.csv("build/input/territory_codes.csv")
   extinctions <- read.csv("build/temp/last_rulers.csv") %>% 
-    select(terr_id, death_year, id) %>% 
-    rename(extinction_year = death_year,
-           last_ruler_id = id)
+    select(terr_id, death_year)
   
   lineages <- AddLineageNames(lineages, terr_codes)
   lineages <- AddExtinctionDummies(lineages, extinctions)
@@ -27,27 +25,25 @@ Main <- function(){
   lineages %>% write.csv("build/output/lineages.csv")
 }
 
-AddExtinctionDummies <- function(lineages, extinctions){
-  lineages_with_ext_years <- inner_join(lineages, extinctions, by="terr_id")
-  # In this inner join I lose 100,000 lineage-year observations (presumably 
-  # lineages in the city data which do not show up in the ruler data.)
-  
-  lineages_with_dummies <- lineages_with_ext_years %>% 
-    mutate(final_year = extinction_year - 1) %>% 
-    # -1 because terr_ids are empty (no cities have them) in the actual extinction year
-    mutate(extinction = if_else(year == final_year, 1, 0))
-  
-  return(lineages_with_dummies)
-}
-
 AddLineageNames <- function(lineages, terr_codes){
   lineages_with_names <- inner_join(lineages, terr_codes, by="terr_id")
   return(lineages_with_names)
 }
 
+AddExtinctionDummies <- function(lineages, extinctions){
+  lineages_joined <- inner_join(lineages, extinctions, by="terr_id")
+  # In this inner join I lose 100,000 lineage-year observations
+  
+  lineages_with_dummies <- lineages_joined %>% 
+    mutate(final_full_year = death_year-1) %>% 
+    # -1 because terr_ids are empty (no cities have them) in the actual extinction year
+    mutate(extinction = if_else(year == final_full_year, 1, 0))
+  return(lineages_with_dummies)
+}
+
 TidyOutput <- function(lineages){
   lineages_final <- lineages %>% 
-    select(terr_id, year, terr_name, final_year, extinction, count_cities) %>% 
+    select(terr_id, year, terr_name, final_full_year, extinction, count_cities) %>% 
     rename(territory = terr_name)
   return(lineages_final)
 }

### 2023-03-02 build/code/LineageExtinctions.R

@@ -1,5 +1,6 @@
 # Input: lineage-year panel of rulers.
 # Output: lineage-level data set of last ruler (max start_reign) with death_year.
+# NOTE: death years are correct (not 1 year lower) because I added 1 in Rulers.R.
 
 library(tidyverse)
 

### 2023-03-02 analysis/code/SubExperiments.Rmd
#1 first attempt at sub-experiments

## Stacking sub-experiments

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(cowplot)
```

In this file I want to get a feeling for how I can get stackable sub-experiments from the data.

What I learned in Exploration.Rmd (based on 382 extinctions from the ruler data):

- I have almost 500 observations on average for each city.
- **A normal city changes hands 3-4 times throughout its history.**
- **Of these territory changes, 1-2 are due to extinction.**
- **A normal city undergoes around 60 changes in integration.**
- Most cities become drastically more integrated over time.

My process for getting sub-experiments should take an extinction list as input, apply transparent criteria, and return a list of sub-experiments.

### What kind of object is a sub-experiment?

A sub-experiment is a subset of the build, i.e. a city-year data set, so that

- the sub-experiment contains, for each city in it, all years between t-W and t+W, where t is an extinction year.
- no city in the sub-experiment undergoes an extinction (or size diff?) at any time other than t.

### Sketching the code

1. Identify sub-experiments (extinctions with W years of data before/after)
2. Define inclusion criteria to get a data set for each subexp
    - C1: within d all treated cities' only extinction within [t-2W, t+W] is t
    - C2: only units with no extinctions in [t-2W, t+W] are eligible controls
    - C3: only observations from years [t-W, t+W] are included
    Make dummies for each of those conditions
3. Stack the data

### Checking correlation between extinction treatment and count diff

```{r data import}
build <- read_csv("analysis/input/build.csv") %>% 
  select(city_id, year, terr_id, final_full_year, 
        extinction, treatment, extinction_of, count_diff)

```

```{r check extinct vs non-extinct}
# use x = extinction as sanity check (diffs will be 0)

diff_ext <- build %>% 
  mutate(treatment=factor(treatment)) %>% 
  ggplot() + 
  geom_boxplot(aes(x = treatment, y=count_diff))

diff_ext_no_outliers <- build %>% 
  mutate(treatment=factor(treatment)) %>% 
  ggplot() + 
  geom_boxplot(aes(x = treatment, y=count_diff), outlier.shape = NA) +
  coord_cartesian(ylim = quantile(build$count_diff, c(0.0026, 0.996), na.rm=TRUE))

diff_ext_big_diffs <- build %>% 
  mutate(treatment=factor(treatment)) %>% 
  filter(abs(count_diff) > 1) %>% 
  ggplot() + 
  geom_boxplot(aes(x = treatment, y=count_diff))

diff_ext_big_diffs_zoom <- build %>% 
  mutate(treatment=factor(treatment)) %>% 
  filter(abs(count_diff) > 1) %>% 
  ggplot() + 
  geom_boxplot(aes(x = treatment, y=count_diff), outlier.shape = NA) +
  coord_cartesian(ylim = quantile(build$count_diff, c(0.0026, 0.996), na.rm=TRUE))

plot_grid(diff_ext, diff_ext_no_outliers, diff_ext_big_diffs, diff_ext_big_diffs_zoom,
          labels = c("Full plot", "Zoom", "Diffs >1", "Diffs >1, zoom"))
```

Make this code neater. But you can definitely see that extinctions are correlated with bigger count diffs.

### Implementing stacked sub-experiments

```{r define window size}
W <- 20
```

```{r identify sub-experiments}

extinctions <- read_csv("build/temp/last_rulers.csv") %>% 
  select(terr_id, death_year, family, name) %>% 
  mutate(sub_exp = if_else(
    between(death_year, min(build$year) + W, max(build$year) - W),
    1, 0))

subexp <- extinctions %>% 
  rename(treat_year = death_year) %>% # last full year + 1 = death year = treat year
  filter(sub_exp == 1) %>% 
  select(terr_id, treat_year)

subexp <- subexp %>% 
  mutate(
    lower_inclusion_bound = treat_year - 2*W,
    upper_inclusion_bound = treat_year + W,
    exp_start = treat_year - W,
    exp_end = treat_year + W)

```

```{r get data for each subexp}

# let's pick one subexp to tinker on. let's pick the extinction of A1541.
d <- subexp[subexp$terr_id == "A1541",]

# first we want to get all cities that fall into the inclusion range of d:
inclusion_range <- d$lower_inclusion_bound:d$upper_inclusion_bound
d_cities <- build %>% 
  group_by(city_id) %>% 
  summarise(
    d_dummy = if_else(all(inclusion_range %in% min(year):max(year)), 1, 0)
    ) %>% 
  filter(d_dummy == 1)

# does this make sense? all cities with a 1 should have yearly observations
# within the inclusion range.

build %>% 
  semi_join(d_cities, by="city_id") %>% 
  filter(year %in% inclusion_range) %>% 
  ggplot()+
  geom_bar(aes(x=year))

# there is some slight variation but it looks good overall.
# how do I drop cities where not all years are in the data?

build %>% 
  semi_join(d_cities, by="city_id") %>% 
  filter(year %in% inclusion_range) %>% 
  group_by(city_id) %>% 
  filter(n() == 3*W+1) %>% 
  ggplot()+
  geom_bar(aes(x=year))

# this is how it works. But after this I still have to check the other criteria.
```

```{r get data for d}

d_base_data <- build %>% # this is all city-years in the inclusion range
  semi_join(d_cities, by="city_id") %>% 
  filter(year %in% inclusion_range) %>% 
  group_by(city_id) %>% 
  filter(n() == 3*W+1)

d_treat_cities <- d_base_data %>% # this is all cities satisfying:
  group_by(city_id) %>% 
  filter(sum(treatment) == 1) %>% # there must be exactly one treatment
  ungroup() %>% 
  filter(treatment == 1) %>% # d must be that one treatment
  filter(extinction_of == d$terr_id) %>% 
  mutate(d_treat = 1)

d_control_cities <- d_base_data %>% 
  group_by(city_id) %>% 
  filter(sum(treatment) == 0) %>% # there must be no treatment in the inclusion period
  summarise(d_treat = 0)

d_sample <- bind_rows(d_treat_cities, d_control_cities)

```

I am worried because count_diff is now 0 for all treatment cities.

```{r check whether it worked}
# check whether the cities actually underwent size diffs
build %>% filter(city_id == 11056 & between(year, 1660, 1670))

# check the sizes of the territories
lineages <- read_csv("analysis/input/lineages.csv")
lineages %>% 
  filter(terr_id %in% c("A1541", "A156") & between(year, 1663, 1667)) %>% 
  select(terr_id, year, final_full_year, extinction, count_cities)
```

This is very reassuring; the territories are actually the same size; A156 had nothing before and took everything A1541 had.

I guess now the next step is to abstract what I learned here into functions. I do this in a different file.

### 2023-03-02 build/code/CombineAll.R

@@ -48,15 +48,16 @@ PushExtinctions1Year <- function(table){
   # count_diff, i.e. in the first year under the new lineage.
   pushed <- table %>% 
     group_by(city_id) %>% 
-    mutate(treatment = lag(extinction)) %>% # there's no way to avoid leading NA
-    ungroup() # because I don't know the previous terr_id of those observations
+    mutate(treatment = lag(extinction)) %>% # there's no way to avoid leading NA because I don't know the previous terr_id of those observations
+    mutate(extinction_of = if_else(treatment==1, lag(terr_id), "")) %>% 
+    ungroup()
   return(pushed)
 }
 
 TidyOutput <- function(table){
   final <- table %>% 
     select(city_id, year, city, terr_id, territory, final_full_year, extinction, 
-           treatment, count_cities, count_diff, real_wage, welfare_ratio)
+           treatment, extinction_of, count_cities, count_diff, real_wage, welfare_ratio)
   return(final)
 }
 

### 2023-03-02 analysis/code/SubExperiments.Rmd
#1 working version of GenSubExps

@@
   filter(n() == 3*W+1) %>% 
-  ggplot()+
-  geom_bar(aes(x=year))
+  summarise() # %>% 
+  # ggplot()+
+  # geom_bar(aes(x=year))
 

### 2023-03-02 analysis/code/GetSubExps.R

@@ -1,0 +1,101 @@
+# Input: build data and lineage-level list of extinctions
+# Output: assignment of cities to groups in sub-experiments.
+# list of cities with a huge amount of dummies.
+
+library(tidyverse)
+setwd("~/GitHub/BA")
+
+
+Main <- function(){
+  data <- read_csv("analysis/input/build.csv")
+  extinctions <- read_csv("build/temp/last_rulers.csv")
+  
+  W <- 10
+  
+  subexps <- IdentifySubExps(extinctions, data, W)
+  assignment <- AssignCitiesAll(subexps, data)
+  
+  assignment <- DropEmptySubExps(assignment, threshold=4)
+  
+  assignment %>% write.csv("analysis/temp/assignment.csv")
+  #return(assignment)
+}
+
+
+IdentifySubExps <- function(extinctions, build, W){
+  # returns extinctions with (2*)W years of data before/after,
+  # and their sub-experiment and inclusion windows.
+  
+  subexps <- extinctions %>% 
+    filter(between(death_year, min(build$year) + W, max(build$year) - W)) %>% 
+    rename(treat_year = death_year) %>% 
+    select(terr_id, treat_year) %>% 
+    mutate(
+      lower_inclusion_bound = treat_year - 2*W, # this is a choice
+      upper_inclusion_bound = treat_year + W,
+      exp_start = treat_year - W,
+      exp_end = treat_year + W)
+  return(subexps)
+}
+
+AssignCitiesSubExp <- function(d, build){
+  # returns the city-year observations for sub-experiment d.
+  incl_range <- d$lower_inclusion_bound:d$upper_inclusion_bound
+  
+  # find eligible cities (full available data) and ungroup back to city-year
+  d_base_data <- build %>% 
+    filter(year %in% incl_range) %>% 
+    group_by(city_id) %>% 
+    filter(n() == length(incl_range)) %>% 
+    ungroup()
+  
+  # get treatment group
+  d_treat_cities <- d_base_data %>% # this is all cities satisfying:
+    group_by(city_id) %>% 
+    filter(sum(treatment) == 1) %>% # 1) there must be exactly one treatment AND
+    ungroup() %>% 
+    filter(treatment == 1) %>% 
+    filter(extinction_of == d$terr_id) %>% # 2) that one treatment is d
+    mutate(d_treat = TRUE)
+  
+  # get control group
[…]

### 2023-03-02 paper/paper.tex
#2 add sketch to paper.tex
#2 sketch literature review

@@
 
+\section{Literature review}
+
+You should ultimately boil this down to one page.
+Identify strengths and weaknesses of related studies as they relate to my project.
+Contrast sharply.
+Highlight your contribution in the context of the rest.
+Give your best answer to the ``So What?'' question \textit{now}.
+
+Begin with comments about the body of research as a whole. 
+First of all, which field is your project located in? Or at the intersection of which fields?
+
+Have there been many or few studies?
+Have they focused on similar questions, or different?
+Are they empirical or theoretical?
+Which kind of setting do they use?
+What kind of data do they use?
+Do they use fancy structural estimation of specific parameters in trade models?
+What about the others, that are more like yours?
+What are recent papers on?
+When was this topic popular?
+
+Once you've set the stage, locate your project in it.
+As of now, I think this project makes two main contributions: it shows the effects of integration in a historical context, and it's novel in using detailed regionally disaggregated panel data to attempt causal inference.
+
+Then sharpen this localisation by comparing your project to salient papers from the different strands (thematic groups): first papers from a).
+then from b).
+then c).
+Explain merits and shortcomings of these papers in relation to the question.
+Explain how your study hopes to avoid flaws in their analyses, and, in stark terms, what it ultimately contributes.
+This concludes your lit review.
+
+Strands of relevant literature (gradually copy these upward):
+
+\subsection{State size and capacity}
+
+Theory: \cite{as1997}
+
+Historical evidence: \cite{cervellati2022}, \cite{koyama2017}, \cite{schoenholzer2018}, \cite{sng2014}
+
+\subsection{Economic and political integration}
+
+Theory: \cite{sapir2011}, \cite{asw2000}
+
+Historical evidence: \cite{chilosi2013}, \cite{kindleberger1975}, \cite{wolf2009}, \cite{chen2004}
+
+Effects: \cite{campos2019}
+
+\subsection{HNEs in the HRE}
+
+Papers using similar approaches: \cite{cantoni2011}
+
+\subsection{Hereditary rule as a HNE}
+
+Theory: \cite{besley2017}
+
+Evidence: \cite{olken2005}, \cite{dube2020}, \cite{ottinger2022}
+
 \section{Cities, lineages, and territories}

### 2023-03-09 analysis/code/Exploration.Rmd
count buildings + explore city integ

@@
 - Yearly differences in integration are positive on average but vary a lot.
+
+#### City size across time
+
+```{r}
+
+plot_integ <- function(data, width, startyear=0, endyear=2000){
+  cities <- data %>% 
+    filter(year >= startyear & year <= endyear) %>% 
+    mutate(period = year - year %% width) %>% 
+    mutate(period = factor(period)) %>%
+    group_by(period, city_id) %>% 
+    summarise(mean_integ = mean(count_cities))
+
+  plot <- cities %>% 
+    ggplot() +
+    geom_boxplot(aes(x = period, y = mean_integ))
+
+  return(plot)
+}
+
+#plot_integ(data, 10, startyear=1800)
+
+integ_50y <- plot_integ(data, 50)
+integ_1600_25y <- plot_integ(data, 25, startyear=1600)
+integ_1800_10y <- plot_integ(data, 10, startyear=1750)
+
+plot_grid(integ_50y, integ_1600_25y, integ_1800_10y,
+          align="v", ncol=1, labels = c("Full", "1600+", "1800+"))
+```
+
+The political integration of cities stagnates until 1750, starts to quickly and steadily increase from the late 18th to mid-19th century, and then jumps abruptly with Prussia getting the Norddeutscher Bund in 1866.
+
+Because we average within decades, this is kind of misleading.
+It appears like the jump happens in two stages: one in the 1860s and one in the 1870s.
+As we see below, that's not what's really in the data: there is only one jump for Prussia, in 1866.
+Because it's in the middle of the 1860s, the 1860 boxplot is halfway between 1850 and 1870, making it seem like there's another jump in the 1870s.
+This isn't actually the case, though (at least not for Prussia).
+
+#### Dynamics of the largest territories
+```{r}
+plot_top_terrs <- function(data, how_many, window=0:2000){
+  threshold = 2
+
+  top_terrs <- data %>% 
+    group_by(terr_id) %>% 
+    summarise(
+      max_size = max(count_cities),
+      years_in_window = sum(year %in% window)) %>% 
+    filter(years_in_window > threshold) %>% 
+    arrange(desc(max_size)) %>% 
+    slice(1:how_many)
+
+  top_data <- data %>% 
+    filter(year %in% window) %>% 
+    filter(terr_id %in% top_terrs$terr_id) %>% 
+    distinct(year, terr_id, .keep_all = TRUE)
+
+  plot <- top_data %>% 
+    ggplot(aes(x=year, y=count_cities))+
+    geom_line(aes(colour=terr_id))
+
+  return(plot)
+}
+
+full_plot <- plot_top_terrs(data, 9)
+zoom_plot <- plot_top_terrs(data, 6, window = 1800:1900)
+
+plot_grid(full_plot, zoom_plot, align="v", ncol=1, labels=c("Full", "1800s"))
+
+```
+
+The top panel shows that there's a problem with H010.
+Apparently there are some missing values, i.e.
+the territory disappears from the dataset for around 300 years and then reappears.
+Before I output any exploration graphs that will go in the paper, I should take care of this in Cleaning.R.
+
+Also, this is very messy. If I put something like this in the paper, I should probably use facets of a single ggplot, so the colour-territory-correspondence is the same and there is only one legend.

### 2023-03-09 build/code/CombineCityData.R

@@ -13,10 +13,10 @@ Main <- function(){
   cities <- read_dta("build/input/cities_families_1300_1918.dta") %>% 
     select(city_id, year, terr_id)
   wages <- read.csv("build/temp/wages.csv")
-  # construction <- read.csv("build/temp/construction_new.csv")
+  construction <- read.csv("build/temp/construction_new.csv")
   
   
-  cities <- AddConstruction(cities, 0)
+  cities <- AddConstruction(cities, construction)
   cities <- AddWages(cities, wages)
   cities <- TidyOutput(cities)
   
@@ -24,7 +24,11 @@ Main <- function(){
 }
 
 AddConstruction <- function(cities, construction){
-  return(cities)
+  # For now, just adds construction at the given time_points.
+  # Will change this later.
+  joined <- left_join(cities, construction, by=c("city_id", "year"))
+  
+  return(joined)
 }
 
 AddWages <- function(cities, wages){
@@ -34,7 +38,7 @@ AddWages <- function(cities, wages){
 
 TidyOutput <- function(cities){
   final <- cities %>% 
-    select(city_id, year, city, terr_id, real_wage, welfare_ratio)
+    select(city_id, year, city, terr_id, real_wage, welfare_ratio, buildings)
   return(final)
 }
 

### 2023-03-09 build/code/CombineLineageData.R

@@ -15,7 +15,7 @@ setwd("~/GitHub/BA")
 Main <- function(){
   lineages <- read.csv("build/temp/terr_sizes.csv")
   terr_codes <- read.csv("build/input/territory_codes.csv")
-  extinctions <- read.csv("build/temp/last_rulers.csv") %>% 
+  extinctions <- read.csv("build/output/extinctions.csv") %>% 
     select(terr_id, death_year)
   
   lineages <- AddLineageNames(lineages, terr_codes)

### 2023-03-09 build/code/Construction.R

@@ -7,24 +7,48 @@ library(tidyverse)
 setwd("~/GitHub/BA")
 
 Main <- function(){
-  construction <- ImportConstruction()
-  construction %>% write.csv("build/temp/construction_new.csv")
-  return(0)
+  cons_raw <- read.csv("build/input/construction.csv")
+  cons <- CountBuildings(cons_raw)
+  cons_gaps <- AddGaps(cons)
+  cons_clean <- TidyOutput(cons_gaps)
+  cons_clean %>% write.csv("build/temp/construction_new.csv")
 }
 
-ImportConstruction <- function(){
-  construction <- read.csv("build/input/construction.csv")
-  return(construction)
+
+CountBuildings <- function(cons_raw){
+  # experiment with filter settings!
+  cons <- cons_raw %>% 
+    filter(building %in% c(6:9, 13:16) & range == 2) %>% 
+    group_by(city_id, time_point) %>% 
+    summarise(buildings = n())
+  
+  return(cons)
 }
 
+AddGaps <- function(cons){
+  all_keys <- cons %>% 
+    group_by(city_id) %>% 
+    expand(full_seq(time_point, 25)) %>% # change to 1 later
+    rename(time_point = `full_seq(time_point, 25)`)
+  
+   with_gaps <- cons %>% 
+     right_join(all_keys, by=c("city_id", "time_point")) %>% 
+     replace_na(list(buildings = 0)) %>% 
+     arrange(city_id, time_point)
+   
+   return(with_gaps)
+}
+
+TidyOutput <- function(cons){
+  # change this when you get the full data
+  clean <- cons %>% 
+    rename(year = time_point) %>% 
+    select(city_id, year, buildings)
+
+  return(clean)
+}
 
 # Turns out that the publicly available construction data is coarsened.
 # I have to ask Prof. Cantoni for the raw data before I can proceed 
-
-# Or make fake data.
-
-FakeConstructionData <- function(dat){
-  return(0)
-}
 
 Main()

### 2023-03-09 analysis/code/Exploration.Rmd
add minimal outcome exploration

@@
 ```{r}
+
 cities_life <- data %>% 
@@
     mean_yearly_diff = mean(count_diff, na.rm=TRUE),
-    sd_yearly_diff = sd(count_diff, na.rm=TRUE)
+    sd_yearly_diff = sd(count_diff, na.rm=TRUE),
+    lifetime_construction = sum(construction, na.rm=TRUE),
+    across( # count non-missing observations of each outcome
+      c(real_wage, welfare_ratio, construction), 
+      ~ sum(!is.na(.x)),
+      .names = "years_{.col}"
       )
+    ) %>% 
+  mutate(years_construction = years_construction * 25)
 

### 2023-03-09 analysis/code/ExploreOutcomes.Rmd
add minimal outcome exploration

## Exploration: Outcomes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(cowplot)
```

### Data import and summary

```{r data import}
data <- read.csv("analysis/input/build.csv")
```

### Naive plot

```{r}

ListCitiesNotNA <- function(data, outcome, window=1300:1918){
  ids <- data %>% 
    filter(year %in% window) %>% 
    drop_na({{outcome}}) %>% 
    distinct(city_id) %>% 
    .$city_id
  return(ids)
}

const_cities <- ListCitiesNotNA(data, construction)
allen_cities <- ListCitiesNotNA(data, real_wage)
both <- intersect(const_cities, allen_cities)

PlotOutcomeLines <- function(outcome, dt=data, city_list=NULL, window=1300:1918){
  if (length(city_list) == 0){
    city_list <- ListCitiesNotNA(dt, {{outcome}}, window)
  }

  plot <- dt %>% 
    filter(city_id %in% city_list & year %in% window) %>% 
    mutate(city_id = factor(city_id)) %>% 
    drop_na({{outcome}}) %>% 
    ggplot(aes(x=year, colour=city_id))+
    geom_line(aes(y={{outcome}}))+
    scale_x_continuous(limits=c(min(window), max(window)))

  return(plot)
}

wage_plot <- PlotOutcomeLines(real_wage)
wratio_plot <- PlotOutcomeLines(welfare_ratio)
const_plot <- PlotOutcomeLines(construction, city_list = both)

plot_grid(wage_plot, wratio_plot, const_plot, align="v", axis="b", ncols=1)
```

The line plots work really well for the wage data, but not so much for the 25-yearly construction data. It may look better for the yearly construction data, though.

What's striking and concerning is that they really don't match at all. Maybe I need to control for something? How would that work, though?

Now I want to see construction for some more cities, but not for multiple thousand.

```{r}

TopConstCities <- function(data, howmany=8, by="obs", window=1300:1918){
  cities_life <- data %>% 
    filter(year %in% window) %>% 
    group_by(city_id) %>% 
    summarise(
      lifetime_construction = sum(construction, na.rm=TRUE),
      observations = sum(!is.na(construction))
      )

  if (by == "obs"){
    top_ids <- cities_life %>% 
      arrange(desc(observations), desc(lifetime_construction))
  } else {
    top_ids <- cities_life %>% 
      arrange(desc(lifetime_construction), desc(observations))
  }

  return(top_ids$city_id[1:howmany])
}

city_list <- TopConstCities(data, howmany=5)

PlotOutcomeLines(construction, city_list = city_list)
```

This looks messy as fuck, even for only five cities. Again, I hope that this will look better when the changes happen yearly, not simultaneously every 25 years. Actually, since most cities in most years will have only a single event recorded, I will probably plot it as a histogram and try out various bin sizes. Write a separate Plot function for that.

### 2023-03-09 build/code/CombineAll.R

@@ -57,7 +57,7 @@ PushExtinctions1Year <- function(table){
 TidyOutput <- function(table){
   final <- table %>% 
     select(city_id, year, city, terr_id, territory, final_full_year, extinction, 
-           treatment, extinction_of, count_cities, count_diff, real_wage, welfare_ratio)
+           treatment, extinction_of, count_cities, count_diff, construction, real_wage, welfare_ratio)
   return(final)
 }
 

### 2023-03-09 build/code/CombineCityData.R

@@ -13,20 +13,20 @@ Main <- function(){
   cities <- read_dta("build/input/cities_families_1300_1918.dta") %>% 
     select(city_id, year, terr_id)
   wages <- read.csv("build/temp/wages.csv")
-  construction <- read.csv("build/temp/construction_new.csv")
+  cons <- read.csv("build/temp/construction_new.csv")
   
   
-  cities <- AddConstruction(cities, construction)
+  cities <- AddConstruction(cities, cons)
   cities <- AddWages(cities, wages)
   cities <- TidyOutput(cities)
   
   cities %>% write_csv("build/output/cities.csv")
 }
 
-AddConstruction <- function(cities, construction){
+AddConstruction <- function(cities, cons){
   # For now, just adds construction at the given time_points.
   # Will change this later.
-  joined <- left_join(cities, construction, by=c("city_id", "year"))
+  joined <- left_join(cities, cons, by=c("city_id", "year"))
   
   return(joined)
 }
@@ -38,7 +38,7 @@ AddWages <- function(cities, wages){
 
 TidyOutput <- function(cities){
   final <- cities %>% 
-    select(city_id, year, city, terr_id, real_wage, welfare_ratio, buildings)
+    select(city_id, year, city, terr_id, real_wage, welfare_ratio, construction)
   return(final)
 }
 

### 2023-03-09 build/code/Construction.R

@@ -18,9 +18,9 @@ Main <- function(){
 CountBuildings <- function(cons_raw){
   # experiment with filter settings!
   cons <- cons_raw %>% 
-    filter(building %in% c(6:9, 13:16) & range == 2) %>% 
+    filter(building %in% c(5:9, 13:16) & uncertainty==0) %>% 
     group_by(city_id, time_point) %>% 
-    summarise(buildings = n())
+    summarise(construction = n())
   
   return(cons)
 }
@@ -33,7 +33,7 @@ AddGaps <- function(cons){
   
    with_gaps <- cons %>% 
      right_join(all_keys, by=c("city_id", "time_point")) %>% 
-     replace_na(list(buildings = 0)) %>% 
+     replace_na(list(construction = 0)) %>% 
      arrange(city_id, time_point)
    
    return(with_gaps)
@@ -43,7 +43,7 @@ TidyOutput <- function(cons){
   # change this when you get the full data
   clean <- cons %>% 
     rename(year = time_point) %>% 
-    select(city_id, year, buildings)
+    select(city_id, year, construction)
 
   return(clean)
 }

### 2023-03-26 analysis/code/Exploration.Rmd
#3 ggsave most exploration plots

@@
 ggplot(data) + 
-  geom_bar(aes(x=year))
+  geom_bar(aes(x=year)) +
+  labs(
+    title="Number of observations across time"
+  )
+
+ggsave("obs_yearly.png", device="png", path="analysis/output/exploration")
 ```
@@
   ggplot(aes(x=year))+
-  geom_bar()
+  geom_bar()+
+  labs(
+    title="Distinct lineages across time"
+  )
 
@@
   ggplot(aes(x=century))+
-  geom_bar()
+  geom_bar()+
+  labs(
+    title="Distinct lineages in each century"
+  )
 
-plot_grid(yearly_plot, centuries_plot, labels = c("Yearly", "Centuries"))
+plot_grid(yearly_plot, centuries_plot)
+ggsave("terrs_yearly_cents.png", device="png", path="analysis/output/exploration")
 ```
@@
   ggplot() + 
-  geom_boxplot(aes(x = century, y = count_cities))
+  geom_boxplot(aes(x = century, y = count_cities))+
+  labs(
+    title="Territory size across time",
+    y = "Mean city count"
+  )
 
@@
   geom_boxplot(aes(x = century, y = count_cities), outlier.shape = NA) +
-  coord_cartesian(ylim = quantile(terrs_centuries$count_cities, c(0, 0.93)))
+  coord_cartesian(ylim = quantile(terrs_centuries$count_cities, c(0, 0.93)))+
+    labs(
+    title="Zoom",
+    y = "Mean city count"
+  )
 
 # plot side by side
-plot_grid(size_time, size_time_no_outliers, labels = c("Full plot", "Zoom"))
+plot_grid(size_time, size_time_no_outliers)
+ggsave("terrsize_cents_boxplot.png", device="png", path="analysis/output/exploration")
+
 ```
@@
 
-CityExtPlot <- function(full_data, bins){
+CityExtPlot <- function(full_data, bins, title){
   plt <- full_data %>% 
@@
     ggplot(aes(x=year))+
-    geom_histogram(binwidth = bins)
+    geom_histogram(binwidth = bins)+
+    labs(
+      title=title
+    )
   return(plt)
@@
 
-plot_grid(CityExtPlot(data, 1), CityExtPlot(data, 10),
-          CityExtPlot(data, 30), CityExtPlot(data, 50))
+plot_grid(CityExtPlot(data, 1, "Cities affected by extinction, yearly"), CityExtPlot(data, 10, "10-year bins"),
+          CityExtPlot(data, 30, "30-year bins"), CityExtPlot(data, 50, "50-year bins"))
 
+ggsave("cities_ext.png", device="png", path="analysis/output/exploration")
 ```
@@
 
-TerrExtPlot <- function(terrs, bins){
+TerrExtPlot <- function(terrs, bins, title){
   plt <- terrs %>% 
@@
 
-plot_grid(TerrExtPlot(terrs, 1), TerrExtPlot(terrs, 10),
-          TerrExtPlot(terrs, 30), TerrExtPlot(terrs, 50))
+plot_grid(TerrExtPlot(terrs, 1, "Lineage extinctions, yearly"), TerrExtPlot(terrs, 10, "10-year bins"),
+          TerrExtPlot(terrs, 30, "30-year bins"), TerrExtPlot(terrs, 50, "50-year bins"))
+ggsave("terrs_ext.png", device="png", path="analysis/output/exploration")
 
@@
   ggplot() + 
-  geom_boxplot(aes(x = extinction_ever, y=max_size))
+  geom_boxplot(aes(x = extinction_ever, y=max_size))+
+  coord_flip()+
+  labs(
+    title="Maximum size of lineages with recorded extinction"
+  )
 
@@
   geom_boxplot(aes(x = extinction_ever, y=max_size), outlier.shape = NA) +
-  coord_cartesian(ylim = quantile(terrs_life$max_size, c(0, 0.93)))
+  coord_flip(ylim = quantile(terrs_life$max_size, c(0, 0.915)))+
+  labs(
+    title="Zoom"
+  )
 
-plot_grid(size_ext, size_ext_no_outliers, labels = c("Full plot", "Zoom"))
+plot_grid(size_ext, size_ext_no_outliers, ncol = 1)
+ggsave("ext_max_size.png", device="png", path="analysis/output/exploration")
+
 ```
@@
   ggplot()+
-  geom_histogram(aes(x=lifespan))
+  geom_histogram(aes(x=lifespan))+
+  labs(
+    title="Distribution of lineage lifespan",
+    subtitle="(Last year minus first year that lineage shows up in the city data)"
+  )
+
+ggsave("lifespan_histo.png", device="png", path="analysis/output/exploration")
 ```
@@
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 #mod_lifespan_size <- lm(lifespan ~ log(max_size), data=terrs_life)
@@
 
+(The below refers to the output of a deactivated cell.)
 Left: The lifespan of a lineage is weakly correlated to its maximum lifetime size.
@@
 
-plot_integ <- function(data, width, startyear=0, endyear=2000){
+plot_integ <- function(data, width, startyear=1300, endyear=1918){
   cities <- data %>% 
@@
     ggplot() +
-    geom_boxplot(aes(x = period, y = mean_integ))
+    geom_boxplot(aes(x = period, y = mean_integ))+
+    labs(
+      title=sprintf("Integration of cities (period mean), %i-%i", startyear, endyear)
+    )
 
@@
 plot_grid(integ_50y, integ_1600_25y, integ_1800_10y,
-          align="v", ncol=1, labels = c("Full", "1600+", "1800+"))
+          align="v", ncol=1)
+ggsave("city_integ_boxplots.png", device="png", path="analysis/output/exploration")
 ```
@@
 
+# figure out how to unify the colour legends before saving this!
+
 ```

### 2023-03-29 slides/main.tex
#3 add slides and ggsaves

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass{beamer}

\mode<presentation> 
{

\usetheme{Madrid}

}
\usepackage{graphicx} % we need this so we can add figures
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
\usepackage{natbib} % we need this so we can use citation and bib properly
\usepackage{hyperref} % allows you to add hyperlink
\usepackage{amsmath} % allows you to use most mathematical features
\usepackage{setspace} % allows you to change the line spacing
\usepackage{longtable}
\usepackage{xr}
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
\usepackage{longtable}
\usepackage{threeparttable}
\usepackage{ragged2e}

\title{Lineage extinctions in the HRE} % The short title appears at the bottom of every slide, the full title is only on the title page

\subtitle{Data exploration and ideas}

\author{Elias Hadj Ammar} % Your name

\institute[LMU,Munich] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
LMU, Munich % Your institution for the title page

}
\date{March 29, 2023} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

\section{Original concept}
\begin{frame}
\frametitle{Original concept}
    \begin{itemize}
        \item Question: how do geopolitical takeovers affect economic growth?
        \item Background: say the rulers of some medium-sized duchy die out and the territory is taken over by the Habsburgs. Does it prosper, having become part of a large and powerful empire? Or does it suffer from worse fiscal governance and conflict? What does that depend on?
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Original concept}
\begin{itemize}
    \item Research design: to estimate the causal effect of takeovers, use lineage extinctions as exogenous shocks to the borders of German states.
    \item Treat each extinction like its own natural experiment and run a staggered difference-in-differences regression.
    \item Use cities as the unit of observation in a yearly panel, and territory size changes due to extinctions as the treatment.
\end{itemize}
\end{frame}

\section{The dataset}
\begin{frame}
\frametitle{Data used}
\begin{enumerate}
    \item biographical data on rulers in the HRE
    \item city-level panel of ruling lineage from 1300-1918
    \item city-level panel of construction activity (coarsened to 25 year intervals) from \cite{construction2020}
    \item city-level panel of territorial history from \cite{territories2020}
    \item real wages of craftsmen for some cities from \cite{allen2001}
\end{enumerate}

\end{frame}

\begin{frame}
\frametitle{What I did with it}
Basic data build:
\begin{enumerate}
    \item combined data sets into one big city-level panel
    \item obtained a list of lineage extinction events from ruler and territories data
    \item added territory size / "integration" variable: total number of cities in the territory a city belongs to
    \item added treatment dummies to city-years in which the ruling lineage goes extinct
\end{enumerate}
\end{frame}

\section{Summary statistics}

\begin{frame}
\frametitle{No. of observations (cities)}
\begin{figure}
    \centering
    \includegraphics[scale=0.6]{exploration/obs_yearly.png}
    \label{fig:obs_yearly}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{No. of lineages}
\begin{figure}
    \centering
    \includegraphics[scale=0.55]{exploration/terrs_yearly_cents.png}
    \label{fig:terrs_yearly_cents}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Distribution of territory size}
\begin{figure}
    \centering
    \includegraphics[scale=0.55]{exploration/terrsize_cents_boxplot.png}
    \label{fig:terrsize_boxplot}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Political integration of cities}
\begin{figure}
    \centering
    \includegraphics[scale=0.55]{exploration/city_integ_boxplots.png}
    \label{fig:city_integ}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Lineage extinction events}
\begin{figure}
    \centering
    \includegraphics[scale=0.55]{exploration/terrs_ext.png}
    \label{fig:terrs_ext}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Staggered adoption diff-in-diff}
Estimate the following regression equation:
\begin{equation}\label{eq:did}
\begin{split}
    construction_{itd} = \theta_{id} + \delta_{td} + \beta (Treat_{id} \times Post_{td}) + X_{itd} \Gamma + \epsilon_{itd}
\end{split}
\end{equation}
\end{frame}

\begin{frame}
\frametitle{Sub-experiments}
\begin{figure}
    \centering
    \includegraphics[scale=0.5]{exploration/subexps.png}
    \caption{Assignment of cities to sub-experiments}
    \label{fig:subexps}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Outcome variables}
\begin{figure}
    \centering
    \includegraphics[scale=0.52]{exploration/const_wages.png}
    \caption{Outcomes for Gdansk (1013), Leipzig (9070), and Munich (23092)}
    \label{fig:const_wages}
\end{figure}
\end{frame}

\section{Ideas}

\begin{frame}
\frametitle{Ideas on how to proceed}
\begin{itemize}
    \item double down on the original concept and write that paper, using construction as the outcome
    \item write an econometrics paper on staggered adoption designs, and use what I've got so far as a demonstration / application
    \item use different outcome data?
\end{itemize}
\end{frame}

\begin{frame}
\Huge{\centerline{The End}}
\end{frame}

\section{References}

\begin{frame}
\frametitle{References}
\justifying

\bibliography{references.bib}
\bibliographystyle{chicago}

\end{frame}

\section{Appendix}

\begin{frame}
\frametitle{Appendix: More summary stats}
\begin{figure}
    \centering
    \includegraphics[scale=0.3]{exploration/build_summary.png}
    \label{fig:build_sum}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Cities affected by extinction events}
\begin{figure}
    \centering
    \includegraphics[scale=0.55]{exploration/cities_ext.png}
    \label{fig:cities_ext}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{More summary stats}
\begin{figure}
    \centering
    \includegraphics[scale=0.55]{exploration/ext_max_size.png}
    \label{fig:ext_max_size}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Stacked event study + IV spec}
\begin{equation}\label{eq:eventstudy}
\begin{split}
    construction_{itd} = \theta_{id} +  \sum_{\tau=-10}^{10}\beta_\tau \times I(TSE_{td}=\tau) + \\
    \sum_{\tau=-10}^{10} \delta_\tau \times (\widehat{treat}_{id} \times I(TSE_{td}=\tau)) + X_{it} \Gamma + \epsilon_{it}
\end{split}
\end{equation}
\bigskip
\begin{equation}\label{eq:firststage}
    \widehat{treat}_{id} = \alpha_d + \phi \times I(extinction_{id}) + X_{id} \Gamma + u_{id}
\end{equation}
\end{frame}

\begin{frame}
\frametitle{Example construction series}
\begin{figure}
    \centering
    \includegraphics[scale=0.52]{exploration/const_four.png}
    \caption{Construction time series for the 16 cities with the most observations}
    \label{fig:const_four}
\end{figure}    
\end{frame}

\end{document}

### 2023-03-29 analysis/code/exploration/CheckExtinctionLists.Rmd
#4 organise analysis folder

## Sanity check extinctions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(cowplot)
```

What I know so far:

- I record 382 extinctions in the ruler dataset.
- I can tell from the summary statistics of the build that the ruler dataset is not complete (i.e. the extinction dataset contains too many extinctions) because the min of final_year is 1101 while the min of year is 1300. Lack of a recorded ruler doesn't necessarily mean that the lineage is extinct. I should definitely cross-check my extinction list with the territories extinction list before picking out sub-experiments.
- There is another source for extinctions: `territories.csv`, which is published in Princes and Townspeople. It's not yearly, but I can still use it as a comparison.

```{r data import}
territories <- read_csv("build/input/territories.csv") %>% 
  select(city_id, time_point, terr_id, observation, type_reign, 
         type_change, commentary_primary, commentary_secondary)
```
### Understanding where the variation in this dataset comes from
```{r}
# Hunch: type_change varies at the observation level
mins <- territories %>% 
  group_by(observation) %>% 
  summarise(type_change = min(type_change))

maxs <- territories %>% 
  group_by(observation) %>% 
  summarise(type_change = max(type_change))

colSums(mins != maxs) # yes they're the same

```

The `observation` column has the actual observations, which are then expanded across the time_points. So what I actually want is, for each observation with type_change==3, the row with the earliest year. That yields a data set of extinction takeover observations.

### Getting type 3 takeovers
```{r get extinction takeover observations}
takeovers <- territories %>% 
  group_by(observation) %>% 
  mutate(
    min_year = min(time_point),
    type_change = min(type_change)
  ) %>% 
  filter(type_change == 3) %>% 
  ungroup() %>% 
  filter(time_point == min_year)

takeovers
```

That's 2,028 extinction observations! I feel like that's too many? That's an order of magnitude more than I have in the rulers data. Ah, but this is cities. There are 2,028 city-time-points with extinctions. So it makes more sense.

Note that `terr_id` is the lineage that takes over. I would like to know the lineage that went extinct. I need, for each observation, the terr_id of the previous time point.

### Getting the lineages that go extinct
```{r get extinct lineage}
takeover_dummies <- takeovers %>% 
  mutate(takeover = 1) %>% 
  select(observation, time_point, takeover)

terrs_with_takeovers <- left_join(territories, takeover_dummies, by=c("observation", "time_point"))

extinctions <- terrs_with_takeovers %>% 
  group_by(city_id) %>% 
  mutate(prev_terr_id = lag(terr_id, order_by = time_point)) %>% 
  ungroup() %>% 
  filter(takeover == 1) %>% 
  distinct(prev_terr_id, .keep_all=TRUE) # try including time_point here

extinctions
```

I have 392 distinct `prev_terr_id`-`time_point observations` (of which 6 previous terrs are NA), but only 306 distinct `prev_terr_id` observations. Does this mean that some lineages go extinct multiple times? Essentially yes. It means there are cities owned by Habsburg that change hands due to extinction in 1650, and there are also cities owned by Habsburg that change hands due to extinction in 1700. That doesn't make sense, does it?

But no matter - these are the extinctions that are in the territories data. How many of these are in *my* extinctions list?

### Comparing the two lists
```{r import my extinction list}
ext_rulers <- read_csv("build/temp/last_rulers.csv") %>% 
  select(terr_id, death_year, end_reign, family, name)

# Make datasets compatible
ext_rulers <- ext_rulers %>% 
  rename(prev_terr_id = terr_id)
```
```{r join and compare}
innerjoined <- inner_join(ext_rulers, extinctions, by="prev_terr_id") %>% 
  select(prev_terr_id, death_year, end_reign, time_point,
         commentary_primary, commentary_secondary)

leftjoined <- left_join(ext_rulers, extinctions, by="prev_terr_id") %>%
  select(prev_terr_id, death_year, end_reign, time_point,
         commentary_primary, commentary_secondary)

innerjoined
```

What I take away from this:
- of 382 extinctions in the ruler list, only 125 show up in the territories data.
- among those 125 extinctions there are many cases where the territories data directly confirms my data because the exact `death_year` is mentioned in a `commentary`.
- there are some cases (e.g. "0W0070") where the data sets contradict each other: there are rulers recorded until the 20th century, but there exists a city which was ruled by that lineage until it went extinct in 1639 (according to the comment)

Consider limiting the set of potential treatments to those extinctions from the ruler data which have a `commentary` from the terr data that contains the `death_year` from the ruler data.

### 2023-03-29 analysis/code/exploration/ExploreOutcomes.Rmd
#4 organise analysis folder

## Exploration: Outcomes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(cowplot)
library(viridisLite)
```

### Data import and summary

```{r data import}
data <- read.csv("analysis/input/build.csv")
```

### Naive plot

```{r}

ListCitiesNotNA <- function(data, outcome, window=1300:1918){
  ids <- data %>% 
    filter(year %in% window) %>% 
    drop_na({{outcome}}) %>% 
    distinct(city_id) %>% 
    .$city_id
  return(ids)
}

const_cities <- ListCitiesNotNA(data, construction)
allen_cities <- ListCitiesNotNA(data, real_wage)
both <- intersect(const_cities, allen_cities)
colour_vec <- c("gold", "cornflowerblue", "coral")
names(colour_vec) <- allen_cities

PlotOutcomeLines <- function(outcome, colour_vec, dt=data, city_list=NULL, window=1300:1918){
  if (length(city_list) == 0){
    city_list <- ListCitiesNotNA(dt, {{outcome}}, window)
  }

  plot <- dt %>% 
    filter(city_id %in% city_list & year %in% window) %>% 
    mutate(city_id = factor(city_id)) %>% 
    drop_na({{outcome}}) %>% 
    ggplot(aes(x=year, y={{outcome}}))+
    geom_line(aes(colour=city_id))+
    scale_color_manual("City", values=colour_vec)+
    scale_x_continuous(limits=c(min(window), max(window)))

  return(plot)
}

wage_plot <- PlotOutcomeLines(real_wage, colour_vec)
wratio_plot <- PlotOutcomeLines(welfare_ratio, colour_vec)
const_plot <- PlotOutcomeLines(construction, colour_vec, city_list = both)

plot_grid(wage_plot, wratio_plot, const_plot, align="v", axis="b", ncols=1)
ggsave("const_wages.png", device = "png", path="analysis/output/exploration")
```

The line plots work really well for the wage data, but not so much for the 25-yearly construction data. It may look better for the yearly construction data, though.

What's striking and concerning is that they really don't match at all. Maybe I need to control for something? How would that work, though?

Now I want to see construction for some more cities, but not for multiple thousand.

```{r}

TopConstCities <- function(data, which=1:8, by="obs", window=1300:1918){
  cities_life <- data %>% 
    filter(year %in% window) %>% 
    group_by(city_id) %>% 
    summarise(
      lifetime_construction = sum(construction, na.rm=TRUE),
      observations = sum(!is.na(construction))
      )

  if (by == "obs"){
    top_ids <- cities_life %>% 
      arrange(desc(observations), desc(lifetime_construction))
  } else {
    top_ids <- cities_life %>% 
      arrange(desc(lifetime_construction), desc(observations))
  }

  return(top_ids$city_id[which])
}

Plot4 <- function(which4){
  upper <- 1 + 4*which4
  lower <- 4 + 4*which4
  city_list <- TopConstCities(data, lower:upper)
  colour_vec <- viridis(4)
  names(colour_vec) <- city_list
  PlotOutcomeLines(construction, colour_vec, city_list = city_list)
}

plot1 <- Plot4(0)
plot2 <- Plot4(1)
plot3 <- Plot4(2)
plot4 <- Plot4(3)

plot_grid(plot1, plot2, plot3, plot4)
ggsave("const_four.png", device = "png", path="analysis/output/exploration")

```

This looks messy as fuck, even for only five cities. Again, I hope that this will look better when the changes happen yearly, not simultaneously every 25 years. Actually, since most cities in most years will have only a single event recorded, I will probably plot it as a histogram and try out various bin sizes. Write a separate Plot function for that.

### 2023-04-10 analysis/code/exploration/ExploreOutcomes.Rmd
#4 rework outcome plotting

@@
 
-### Naive plot
+### List cities to plot
 
 ```{r}
-
 ListCitiesNotNA <- function(data, outcome, window=1300:1918){
@@
 
-const_cities <- ListCitiesNotNA(data, construction)
-allen_cities <- ListCitiesNotNA(data, real_wage)
-both <- intersect(const_cities, allen_cities)
-colour_vec <- c("gold", "cornflowerblue", "coral")
-names(colour_vec) <- allen_cities
-
-PlotOutcomeLines <- function(outcome, colour_vec, dt=data, city_list=NULL, window=1300:1918){
-  if (length(city_list) == 0){
-    city_list <- ListCitiesNotNA(dt, {{outcome}}, window)
-  }
-
-  plot <- dt %>% 
-    filter(city_id %in% city_list & year %in% window) %>% 
-    mutate(city_id = factor(city_id)) %>% 
-    drop_na({{outcome}}) %>% 
-    ggplot(aes(x=year, y={{outcome}}))+
-    geom_line(aes(colour=city_id))+
-    scale_color_manual("City", values=colour_vec)+
-    scale_x_continuous(limits=c(min(window), max(window)))
-
-  return(plot)
-}
-
-wage_plot <- PlotOutcomeLines(real_wage, colour_vec)
-wratio_plot <- PlotOutcomeLines(welfare_ratio, colour_vec)
-const_plot <- PlotOutcomeLines(construction, colour_vec, city_list = both)
-
-plot_grid(wage_plot, wratio_plot, const_plot, align="v", axis="b", ncols=1)
-ggsave("const_wages.png", device = "png", path="analysis/output/exploration")
-```
-
-The line plots work really well for the wage data, but not so much for the 25-yearly construction data. It may look better for the yearly construction data, though.
-
-What's striking and concerning is that they really don't match at all. Maybe I need to control for something? How would that work, though?
-
-Now I want to see construction for some more cities, but not for multiple thousand.
-
-```{r}
-
-TopConstCities <- function(data, which=1:8, by="obs", window=1300:1918){
+ListTopCities <- function(data, which=1:8, by="obs", window=1300:1918){
   cities_life <- data %>% 
@@
 }
+```
 
-Plot4 <- function(which4){
-  upper <- 1 + 4*which4
-  lower <- 4 + 4*which4
-  city_list <- TopConstCities(data, lower:upper)
-  colour_vec <- viridis(4)
-  names(colour_vec) <- city_list
-  PlotOutcomeLines(construction, colour_vec, city_list = city_list)
-}
+```{r}
+# city list for comparison
+const_cities <- ListCitiesNotNA(data, construction)
+allen_cities <- ListCitiesNotNA(data, real_wage)
+both <- intersect(const_cities, allen_cities)
 
-plot1 <- Plot4(0)
-plot2 <- Plot4(1)
-plot3 <- Plot4(2)
-plot4 <- Plot4(3)
-
-plot_grid(plot1, plot2, plot3, plot4)
-ggsave("const_four.png", device = "png", path="analysis/output/exploration")
+# city lists for exploring construction
+set.seed(151101)
+top9 <- ListTopCities(data, which=1:9)
+top16 <- ListTopCities(data, which=1:16)
+random9 <- sample(const_cities, 9)
+random16 <- sample(const_cities, 16)
 
@@
 
-This looks messy as fuck, even for only five cities.
-Again, I hope that this will look better when the changes happen yearly, not simultaneously every 25 years.
-Actually, since most cities in most years will have only a single event recorded, I will probably plot it as a histogram and try out various bin sizes.
-Write a separate Plot function for that.
+### Plotting functions
+
+```{r}
+PlotOutcomeLines <- function(data, citylist=NULL, outcome, colour_vec, window=1300:1918){
+  if (length(citylist) == 0){
+    city_list <- ListCitiesNotNA(data, {{outcome}}, window)
+  }
+  plot <- data %>% 
+    filter(city_id %in% citylist & year %in% window) %>% 
+    mutate(city_id = factor(city_id)) %>% 
+    drop_na({{outcome}}) %>% 
+    ggplot(aes(x=year, y={{outcome}}))+
+    geom_line(aes(colour=city_id))+
+    scale_color_manual("City", values=colour_vec)+
+    scale_x_continuous(limits=c(min(window), max(window)))
+
+  return(plot)
+}
+
+PlotHistograms <- function(data, citylist, binw=10){
+  data |>
+    filter(city_id %in% citylist) |> 
+    mutate(city_id = as_factor(city_id)) |> 
+    filter(construction > 0) |> 
+    ggplot(aes(x=year))+
+    geom_histogram(aes(y=after_stat(density)), binwidth = binw)+
+    geom_density(aes(y=after_stat(density), colour="red"), show.legend = FALSE)+
+    facet_wrap(~city_id)
+}
+
+PlotDensity <- function(data, citylist, colour_vec, window=1300:1918){
+  data |>
+    filter(city_id %in% citylist & year %in% window) |> 
+    mutate(city_id = as_factor(city_id)) |> 
+    filter(construction > 0) |> 
+    ggplot(aes(x=year))+
+    geom_density(aes(y=after_stat(density), colour=city_id))+
+    scale_color_manual("City", values=colour_vec)+
+    scale_x_continuous(limits=c(min(window), max(window)))
+}
+
+```
+
+### Comparison to real wages
+
+```{r plots for Allen cities}
+colour_vec <- c("gold", "cornflowerblue", "coral")
+names(colour_vec) <- allen_cities
+
+wage_plot <- PlotOutcomeLines(data, citylist=both, real_wage, colour_vec, window=1400:1800)
+wratio_plot <- PlotOutcomeLines(data, citylist=both, welfare_ratio, colour_vec, window=1400:1800)
+const_plot <- PlotDensity(data, citylist = both, colour_vec, window=1400:1800)
+
+plot_grid(wage_plot, wratio_plot, const_plot, align="v", axis="b", ncols=1)
+ggsave("const_wages.png", device = "png", path="analysis/output/exploration")
+```
+
+What's striking and concerning is that they really don't match at all. Maybe I need to control for something? How would that work, though?
+
+Now I want to see construction for some more cities, but not for multiple thousand.
+
+```{r}
+
+PlotHistograms(data, random16) +
+  ggtitle("Construction activity: random cities")
+PlotHistograms(data, top16) +
+  ggtitle("Construction activity: cities with the most observations")
+
+```
+
+I want to see how the number of recorded events is related to
+- a city's population in 1875
+- a city's age

### 2023-04-10 build/code/Construction.R

@@ -1,6 +1,5 @@
 # Input: city-year panel of construction from Princes and Townspeople.
 # Output: city-year panel of custom aggregates of construction.
-# Currently not functional.
 
 library(tidyverse)
 

### 2023-04-20 analysis/code/exploration/ExploreOutcomes.Rmd
#4 construction outcomes

@@
 data <- read.csv("analysis/input/build.csv")
+
+cities_life <- data %>% 
+  group_by(city_id) %>% 
+  summarise(
+    lifetime_construction = sum(construction, na.rm=TRUE),
+    observations = sum(!is.na(construction)), # not super meaningful
+    pop1875 = min(pop1875), # should be same as max(pop1875)
+    datayears = max(year) - min(year)
+    )
+
 ```
@@
 ListTopCities <- function(data, which=1:8, by="obs", window=1300:1918){
-  cities_life <- data %>% 
-    filter(year %in% window) %>% 
-    group_by(city_id) %>% 
+  cities_life <- data |> 
+    filter(year %in% window) |> 
+    group_by(city_id) |> 
     summarise(
       lifetime_construction = sum(construction, na.rm=TRUE),
-      observations = sum(!is.na(construction))
+      observations = sum(!is.na(construction)) # not super meaningful
       )
@@
   if (by == "obs"){
-    top_ids <- cities_life %>% 
+    top_ids <- cities_life |> 
       arrange(desc(observations), desc(lifetime_construction))
   } else {
-    top_ids <- cities_life %>% 
+    top_ids <- cities_life |> 
       arrange(desc(lifetime_construction), desc(observations))
@@
 - a city's age
+
+```{r}
+cities_life |> 
+  ggplot(aes(x=log(pop1875), y=lifetime_construction))+
+  geom_point()+
+  geom_smooth(method = "lm")
+
+```
+
+That's definitely a positive association right there.
+The slope is about 8, which means that all else being equal, a twice-as-large city will have eight more construction events recorded in the Städtebuch (if we extrapolate down from the 89 most populous cities in 1875).
+That sounds reasonable.
+Maybe the quality of the data isn't so bad after all?
+Maybe there just wasn't that much construction of notable public buildings in all these tiny cities?
+
+If construction happens rarely, and possibly in bursts / waves (insert mental model of construction here), then it could well be that these bursts happen to be closer in time, on average, to territorial takeovers.
+
+Two possible channels: takeover per se (gaining favour), or size.
+
+Another idea: add ordinal minimum "<20000" for pop1875 so it's never missing

### 2023-04-20 build/code/Construction.R

@@ -6,9 +6,8 @@ suppressPackageStartupMessages(library(tidyverse))
 setwd("~/GitHub/BA")
 
 Main <- function(){
-  cons_raw <- read_csv("build/input/construction.csv", show_col_types = F) |> 
-    select(city_id, city_name, year_clear, building, buildgen, newbuild, uncertainty) |> 
-    rename(year = year_clear)
+  cons_raw <- read_csv("build/input/construction_all.csv", show_col_types = F) |> 
+    rename(year = time_point)
   cons <- CountBuildings(cons_raw)
   cons_gaps <- AddGaps(cons)
   cons_clean <- TidyOutput(cons_gaps)
@@ -49,7 +48,8 @@ TidyOutput <- function(cons){
   return(clean)
 }
 
-# Turns out that the publicly available construction data is coarsened.
-# I have to ask Prof. Cantoni for the raw data before I can proceed 
+# There are some impurities in the new data - specifically, some entries have comments
+# as city ids and single-digit years. But they should fall away with the join in
+# CombineCityData.R, I think.
 
 test <- Main()

### 2023-05-09 paper/paper.tex
count buildings + explore city integ
#8 update parts of paper.tex

@@
 
-\title{Does Political Integration Make Cities Grow Faster?\\
-Evidence From Dynastic Takeovers\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}
+\title{Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}
 }
 \author{Elias Hadj Ammar}
-\date{January 2023}
+\date{May 2023}
 
@@
 \begin{abstract}
-I use historical data on construction activity and territorial history to estimate the effect of state size on city growth in the Holy Roman Empire.
-I exploit variation from lineage extinctions, which caused cities to become part of different lineages' territorial holdings.
-I find that an X\% increase in territory size is associated with a Y\% increase in construction activity.
+I use historical data on construction activity and territorial history to estimate the effect of switching between territories on urban construction in the Holy Roman Empire.
+I exploit variation from dynastic extinctions, which caused cities to become part of different lineages' territorial holdings.
+I find that an X\% increase in the number of cities possessed by the ruling family is associated with a Y\% increase in construction activity.
 \end{abstract}
@@
 \doublespacing
+
+% ########################################################
 
@@
 
+% ########################################################
+
 \section{Literature review}
@@
 
+% --------------------------------------------------------
+
 \subsection{State size and capacity}
@@
 Historical evidence: \cite{cervellati2022}, \cite{koyama2017}, \cite{schoenholzer2018}, \cite{sng2014}
+
+% --------------------------------------------------------
 
@@
 
+\cite{chilosi2013} use  to find . for instance Give a figure: within central europe, what was the dispersion of prices in the first, second and third period each?
+
 Effects: \cite{campos2019}
+
+% --------------------------------------------------------
 
@@
 Papers using similar approaches: \cite{cantoni2011}
+
+% --------------------------------------------------------
 
@@
 
+% ########################################################
+
 \section{Cities, lineages, and territories}
+
+% --------------------------------------------------------
+
+\subsection{What do I measure?}
+
+% ########################################################
 
@@
 
+% --------------------------------------------------------
+
 \subsection{Pooled OLS regression}
+
+% --------------------------------------------------------
 
@@
 
+% --------------------------------------------------------
+
 \subsection{Stacked event study and extinction IV}
+
+% ########################################################
 
@@
 
+% ########################################################
+
 \section{Validity and robustness}
 
+% ########################################################
+
 \section{Conclusion}
+
+% ########################################################
 

### 2023-05-09 paper/paper.tex
#9 add thoughts on baseline sample selection

@@
 
+\subsection{Description of the sample}
+
+I have an unbalanced panel of cities.
+
+The basic version of my build has two problems that stop me from running a staggered DiD on it straight-out-the-box:
+\begin{itemize}
+    \item there are missing values to be dealt with [where?]
+    \item it is unclear how to deal with cities that get treated multiple times    
+\end{itemize}
+
+To deal with these issues, I do the following:
+\begin{itemize}
+    \item don't bother with a DiD and do an event study from the start
+    \item version 1: exclude all cities with multiple treatments and, if that doesn't shrink the sample by a ridiculous amount, run the regression on that
+    \item version 2: exclude cities with multiple treatments only if the event windows overlap
+    \item worry about missing values when they actually become a problem.
+\end{itemize}
+
+Since the non-stacked version is biased, I shouldn't be super worried if I see pre-trends here.
+Run it for good measure, but then argue (based on the literature) that while this looks weird, we should not take it seriously.
+
+After I've implemented my restricted sample, I could run the Exploration.Rmd notebook on that build, too, and see which graphics and stats to include here.
+However, if there are two different builds to discuss (the "build" build and the sample for the baseline), I should be mindful of redundancy and think about the optimal arrangement.
+
 \subsection{Pooled OLS regression}

### 2023-05-11 utils/BaselineSample.R
#9 update baseline regression
note: the baseline estimates change because the sample size increased because I fixed construction_all.csv, which added 3,230 construction events

@@ -1,20 +1,13 @@
 BaselineSample <- function(build){
-  
-  # Drop observations with essential variables missing
-  clean <- build |> 
-    DropNACount("year") |>
-    DropNACount("city_id") |> 
-    DropNACount("construction") |> 
-    DropNACount("treatment")
-  
+
   # Drop cities that were treated more than once from the sample
-  max1treat <- clean |>  
+  max1treat <- build |>  
     group_by(city_id) |> 
     mutate(treatments_total = sum(treatment)) |> 
     filter(treatments_total < 2)
   
-  n_dropped <- nrow(clean)-nrow(max1treat)
-  cities_dropped <- n_distinct(clean$city_id)-n_distinct(max1treat$city_id)
+  n_dropped <- nrow(build)-nrow(max1treat)
+  cities_dropped <- n_distinct(build$city_id)-n_distinct(max1treat$city_id)
   
   print(sprintf("Dropped %d observations with multiple treatments (%d cities)", 
                 n_dropped, cities_dropped))

### 2023-05-11 analysis/code/baseline/baseline_eventstudy.R
#9 fix path to sample util in baseline event study

@@ -8,7 +8,7 @@ library(fixest)
 
 setwd("~/GitHub/BA")
 
-source("analysis/code/lib/SampleSelection.R")
+source("utils/BaselineSample.R")
 
 Main <- function(){
   

### 2023-05-11 analysis/code/exploration/Exploration.Rmd

--- 2023-03-26 analysis/code/Exploration.Rmd
@@
 ```{r data import}
-data <- read.csv("analysis/input/build.csv")
+data <- read.csv("analysis/input/build_full.csv")
 
@@
 
-ggsave("obs_yearly.png", device="png", path="analysis/output/exploration")
+#ggsave("obs_yearly.png", device="png", path="analysis/output/exploration")
 ```
@@
 plot_grid(yearly_plot, centuries_plot)
-ggsave("terrs_yearly_cents.png", device="png", path="analysis/output/exploration")
+#ggsave("terrs_yearly_cents.png", device="png", path="analysis/output/exploration")
 ```
@@
 plot_grid(size_time, size_time_no_outliers)
-ggsave("terrsize_cents_boxplot.png", device="png", path="analysis/output/exploration")
+#ggsave("terrsize_cents_boxplot.png", device="png", path="analysis/output/exploration")
 
@@
 
-ggsave("cities_ext.png", device="png", path="analysis/output/exploration")
+#ggsave("cities_ext.png", device="png", path="analysis/output/exploration")
 ```
@@
           TerrExtPlot(terrs, 30, "30-year bins"), TerrExtPlot(terrs, 50, "50-year bins"))
-ggsave("terrs_ext.png", device="png", path="analysis/output/exploration")
+#ggsave("terrs_ext.png", device="png", path="analysis/output/exploration")
 
@@
 plot_grid(size_ext, size_ext_no_outliers, ncol = 1)
-ggsave("ext_max_size.png", device="png", path="analysis/output/exploration")
+#ggsave("ext_max_size.png", device="png", path="analysis/output/exploration")
 
@@
 
-ggsave("lifespan_histo.png", device="png", path="analysis/output/exploration")
+#ggsave("lifespan_histo.png", device="png", path="analysis/output/exploration")
 ```
@@
           align="v", ncol=1)
-ggsave("city_integ_boxplots.png", device="png", path="analysis/output/exploration")
+#ggsave("city_integ_boxplots.png", device="png", path="analysis/output/exploration")
 ```

### 2023-05-11 analysis/code/baseline/BaselineDiD_countdiff.R
#10 implement DiD with count diff treatment

@@ -1,0 +1,43 @@
+suppressPackageStartupMessages(library(tidyverse))
+suppressPackageStartupMessages(library(fixest))
+
+setwd("~/GitHub/BA")
+
+source("utils/BaselineSample.R")
+
+Main <- function(){
+  build <- read_csv("analysis/input/build.csv", show_col_types = FALSE)
+  clean <- BaselineSample(build)
+  with_timing <- AddTreatDummies(clean)
+  
+  
+  mod <- fixest::feols(construction ~ D | city_id + year, with_timing)
+  
+  setFixest_dict(c(city_id = "City", year = "Year",
+                   construction = "Construction events", D = "Size change x POST"))
+  
+  #summary(mod)
+  
+  tex_output <- etable(mod, tex=TRUE)
+  write(tex_output, file="analysis/output/tables/baseline_did_size.tex")
+}
+
+
+AddTreatDummies <- function(clean){
+  timing <- clean |> 
+    filter(treatment == 1) |> 
+    mutate(treat_year = year) |> 
+    select(city_id, treat_year, count_diff) |> 
+    rename(treat_intensity = count_diff)
+  
+  clean_with_timing <- clean |> 
+    left_join(timing, by="city_id") |> 
+    mutate(D = case_when(
+      is.na(treat_year) ~ 0,
+      year < treat_year ~ 0,
+      year >= treat_year ~ treat_intensity
+    ))
+}
+
+
+Main()

### 2023-05-11 utils/BaselineSample.R

@@ -14,7 +14,7 @@ BaselineSample <- function(build){
   
   # Remove superfluous columns
   clean <- max1treat |> 
-    select(year, city_id, construction, treatment) |> 
+    select(year, city_id, construction, treatment, count_cities, count_diff) |> 
     as.data.frame()
   
   return(clean)

### 2023-05-11 paper/paper.tex
#8 trial automatic table sourcing
#8 delete old draft

--- 2023-05-09 paper/paper.tex
@@
 \usepackage{graphicx} % we need this so we can add figures
+\usepackage{booktabs}
 
@@
 
-% ########################################################
+% ######################################################
 
@@
 
-The political history of Europe, until fairly recently, was defined by ruling families passing their territorial holdings down in hereditary monarchies.
-Land and cities changed hands often throughout the history of Europe.
-Which territory a city belonged to had economic as well as political consequences, and it is not hard to imagine that the fortunes of whole stretches of land could be decided by such dynastic takeovers.
-For example, (good example of a takeover and its consequences here).
-
-In this paper I investigate changes in territory size as one possible channel through which dynastic takeovers could affect development.
-There exists a literature on the optimal size of state (e.g.
-\cite{as1997}, \cite{aw1998}, \cite{asw2000}, \cite{easterly2000}).
-This literature focuses on the tradeoffs between scale effects in government on one hand and conflict on the other, as well as trade and openness.
-
-background on the hre / motivation for using city-level data and territory switches
-
-I use a stacked event study framework with varying intensity of treatment.
-The treatment is the size difference between the old and the new rulers' territorial holdings.
-I will argue that the takeover constitutes a shock to territory size that is exogenous to development.
-The event study shows that cities with different takeover experiences did not differ in growth (conditional on city and year fixed effects) before their respective takeovers.
-
-The outcome variable used throughout the analysis is yearly construction activity.
-I go to some lenghts to show that this is a good proxy for growth, comparing the present data from the Deutsches St\"adtebuch to the craftsmen real wage time series by Allen for available cities.
-The two measures generally align.
-
-Results indicate that territory size has a large, significant and positive effect on city growth.
-Cities that joined a twice-as-big territory experienced 10 \% higher construction activity in subsequent years.
-Splitting the sample into cities from big, medium-sized and small territories reveals that the effect is mostly driven by cities from small territories.
-However, cities in big territories generally changed hands less often, making it more difficult to find a meaningful and significant effect in that subsample.
-
-The literature says that benefits of larger states include free trade within a larger area.
-Trade within the HRE was not exactly free.
-I'm not currently aware of data on historical trade flows between German cities.
-However, I use the establishment of historical markets (Cantoni et al) and the presence of Jews (data from Voigtl\"ander and Voth) to investigate this channel.
-I find that trade increases (cities are (insignificantly) more likely to establish markets, and the presence of Jews increases) when cities join larger territories.
-
-% ########################################################
+% ######################################################
 
@@
 
-You should ultimately boil this down to one page.
-Identify strengths and weaknesses of related studies as they relate to my project.
-Contrast sharply.
-Highlight your contribution in the context of the rest.
-Give your best answer to the ``So What?'' question \textit{now}.
-
-Begin with comments about the body of research as a whole. 
-First of all, which field is your project located in? Or at the intersection of which fields?
-
-Have there been many or few studies?
-Have they focused on similar questions, or different?
-Are they empirical or theoretical?
-Which kind of setting do they use?
-What kind of data do they use?
-Do they use fancy structural estimation of specific parameters in trade models?
-What about the others, that are more like yours?
-What are recent papers on?
-When was this topic popular?
-
-Once you've set the stage, locate your project in it.
-As of now, I think this project makes two main contributions: it shows the effects of integration in a historical context, and it's novel in using detailed regionally disaggregated panel data to attempt causal inference.
-
-Then sharpen this localisation by comparing your project to salient papers from the different strands (thematic groups): first papers from a).
-then from b).
-then c).
-Explain merits and shortcomings of these papers in relation to the question.
-Explain how your study hopes to avoid flaws in their analyses, and, in stark terms, what it ultimately contributes.
-This concludes your lit review.
-
-Strands of relevant literature (gradually copy these upward):
-
-% --------------------------------------------------------
-
-\subsection{State size and capacity}
-
-Theory: \cite{as1997}
-
-Historical evidence: \cite{cervellati2022}, \cite{koyama2017}, \cite{schoenholzer2018}, \cite{sng2014}
-
-% --------------------------------------------------------
-
-\subsection{Economic and political integration}
-
-Theory: \cite{sapir2011}, \cite{asw2000}
-
-Historical evidence: \cite{chilosi2013}, \cite{kindleberger1975}, \cite{wolf2009}, \cite{chen2004}
-
-\cite{chilosi2013} use  to find . for instance Give a figure: within central europe, what was the dispersion of prices in the first, second and third period each?
-
-Effects: \cite{campos2019}
-
-% --------------------------------------------------------
-
-\subsection{HNEs in the HRE}
-
-Papers using similar approaches: \cite{cantoni2011}
-
-% --------------------------------------------------------
-
-\subsection{Hereditary rule as a HNE}
-
-Theory: \cite{besley2017}
-
-Evidence: \cite{olken2005}, \cite{dube2020}, \cite{ottinger2022}
-
-% ########################################################
+% ######################################################
 
@@
 
-% --------------------------------------------------------
-
-\subsection{What do I measure?}
-
-% ########################################################
+% #####################################################
 
@@
 
-% --------------------------------------------------------
-
-\subsection{Pooled OLS regression}
-
-% --------------------------------------------------------
-
-\subsection{Obtaining a dataset of stacked natural experiments}
-
-% --------------------------------------------------------
-
-\subsection{Stacked event study and extinction IV}
-
-% ########################################################
+% #####################################################
 
@@
 
-% ########################################################
+% #####################################################
 

### 2023-05-11 analysis/code/stacked_did/SubExperiments.Rmd
#16 go over existing subexps code

--- 2023-03-02 analysis/code/SubExperiments.Rmd
@@
 build <- read_csv("analysis/input/build.csv") %>% 
-  select(city_id, year, terr_id, final_full_year, 
-        extinction, treatment, extinction_of, count_diff)
+  select(city_id, year, terr_id, treatment, extinction_of, count_diff)
 

### 2023-05-11 analysis/code/stacked_did/GetSubExps.R

@@ -31,7 +31,7 @@ IdentifySubExps <- function(extinctions, build, W){
   
   subexps <- extinctions %>% 
     filter(between(death_year, min(build$year) + W, max(build$year) - W)) %>% 
-    rename(treat_year = death_year) %>% 
+    rename(treat_year = death_year) %>% # not clear whether this needs to be +1!
     select(terr_id, treat_year) %>% 
     mutate(
       lower_inclusion_bound = treat_year - 2*W, # this is a choice
@@ -45,7 +45,7 @@ IdentifySubExps <- function(extinctions, build, W){
 AssignToAllSubExps <- function(build, subexps){
   # this environment gets exported to all the workers
   
-  subexps.list <- subexps |> 
+  subexps.list <- subexps |> # get list of rows of subexps
     select(-treat_year, -exp_start, -exp_end) |> 
     {\(.) split(., seq(nrow(.)))}()
   
@@ -59,15 +59,15 @@ AssignToAllSubExps <- function(build, subexps){
 
 AssignCitiesSubExp <- function(d, build){
   # returns the assignment of cities to treat, control, or exclude for subexp d.
-
-  # get city-year observations from build for sub-experiment d
-  incl_range <- d$lower_inclusion_bound:d$upper_inclusion_bound
   
-  # find eligible cities (full available data) and ungroup back to city-year
+  # get city-year observations from build for sub-experiment d
+  incl_window <- d$lower_inclusion_bound:d$upper_inclusion_bound
+  
+  # find eligible cities (data for full window) and ungroup back to city-year
   d_base_data <- build %>% 
-    filter(year %in% incl_range) %>% 
+    filter(year %in% incl_window) %>% 
     group_by(city_id) %>% 
-    filter(n() == length(incl_range)) %>% 
+    filter(n() == length(incl_window)) %>% 
     ungroup()
   
   # get treatment group

### 2023-05-12 analysis/code/stacked_did/GetSubExps.R
#16 turn assignment to stacked data + run DiD

@@ -20,7 +20,7 @@ Main <- function(){
   
   assignment <- DropEmptySubExps(assignment, threshold=10)
   
-  assignment %>% write.csv("analysis/temp/assignment.csv")
+  assignment |> write_csv("analysis/temp/assignment.csv")
   #return(assignment)
 }
 
@@ -29,10 +29,10 @@ IdentifySubExps <- function(extinctions, build, W){
   # returns extinctions with (2*)W years of data before/after,
   # and their sub-experiment and inclusion windows.
   
-  subexps <- extinctions %>% 
-    filter(between(death_year, min(build$year) + W, max(build$year) - W)) %>% 
-    rename(treat_year = death_year) %>% # not clear whether this needs to be +1!
-    select(terr_id, treat_year) %>% 
+  subexps <- extinctions |> 
+    filter(between(death_year, min(build$year) + W, max(build$year) - W)) |> 
+    rename(treat_year = death_year) |>  # not clear whether this needs to be +1!
+    select(terr_id, treat_year) |>  
     mutate(
       lower_inclusion_bound = treat_year - 2*W, # this is a choice
       upper_inclusion_bound = treat_year + W,
@@ -64,29 +64,30 @@ AssignCitiesSubExp <- function(d, build){
   incl_window <- d$lower_inclusion_bound:d$upper_inclusion_bound
   
   # find eligible cities (data for full window) and ungroup back to city-year
-  d_base_data <- build %>% 
-    filter(year %in% incl_window) %>% 
-    group_by(city_id) %>% 
-    filter(n() == length(incl_window)) %>% 
+  d_base_data <- build |>  
+    filter(year %in% incl_window) |>  
+    group_by(city_id) |> 
+    filter(n() == length(incl_window)) |> 
     ungroup()
   
   # get treatment group
-  d_treat_cities <- d_base_data %>% # this is all cities satisfying:
-    group_by(city_id) %>% 
-    filter(sum(treatment) == 1) %>% # 1) there must be exactly one treatment AND
-    ungroup() %>% 
-    filter(treatment == 1) %>% 
-    filter(extinction_of == d$terr_id) %>% # 2) that one treatment is d
+  d_treat_cities <- d_base_data |>  # this is all cities satisfying:
+    group_by(city_id) |>  
+    filter(sum(treatment) == 1) |>  # 1) there must be exactly one treatment AND
+    ungroup() |> 
+    filter(treatment == 1) |> 
[…]

### 2023-05-12 analysis/code/stacked_did/StackedDiD.R

@@ -1,0 +1,39 @@
+# The point of this script is to add the necessary dummies to the build for regression.
+# If we start out with a simple stacked DiD, then that's one TREAT dummy for each d,
+# and one POST dummy for each d.
+
+# How do I get that?
+
+# I think joining to the build is messy, and it may not even work, because the IU slides
+# say that in a stacked design one city may show up multiple times.
+
+# Alternative: get the data set for each sub-experiment, and then bind_rows() them.
+
+
+library(tidyverse) |> suppressPackageStartupMessages()
+library(fixest) |> suppressPackageStartupMessages()
+
+source("utils/GetStackedData.R")
+
+setwd("~/GitHub/BA")
+
+Main <- function(){
+  build <- read_csv("analysis/input/build.csv", show_col_types = F)
+  extinctions <- read_csv("analysis/input/extinctions.csv", show_col_types = F)
+  assignment <- read_csv("analysis/temp/assignment.csv", show_col_types = F)
+
+  W <- 10
+  
+  stacked_data <- GetStackedData(assignment, build, extinctions, W)
+
+  mod <- fixest::feols(construction ~ D | city_id^subexp + year^subexp, stacked_data)
+  
+  setFixest_dict(c(city_id = "City", year = "Year", subexp = "Extinction",
+                   construction = "Construction events", D = "TREAT x POST"))
+  
+  tex_output <- etable(mod, tex = TRUE)
+  write(tex_output, file="analysis/output/tables/stacked_did.tex")
+}
+
+
+Main()

### 2023-05-12 utils/GetStackedData.R

@@ -1,0 +1,49 @@
+GetStackedData <- function(assignment, build, extinctions, W){
+  
+  # create list of subexp names to map over
+  subexps_list <- assignment |> select(-city_id) |> colnames()
+  
+  # get the dataset for each subexp and rowbind them all together
+  stacked_data <- subexps_list |> 
+    purrr::map(GetDataForOneSubExp, build, extinctions, W) |> 
+    bind_rows()
+  
+  return(stacked_data)
+}
+
+
+GetDataForOneSubExp <- function(colname, build, extinctions, W){
+  
+  # obtain the terr_id corresponding to this subexp
+  terr <- str_split_1(colname, "_")[2]
+  
+  # obtain the treat_year corresponding to this subexp
+  treat_year <- extinctions |> 
+    filter(terr_id == terr) |> 
+    pull(death_year)
+  
+  # drop observations outside the experiment window
+  data_in_window <- build |> 
+    filter(between(year, treat_year-W, treat_year+W))
+  
+  # add treatment dummies and drop cities that are not in this subexp
+  treat_col <- assignment |> 
+    select(city_id, !!sym(colname)) |> 
+    rename(TREAT = !!sym(colname))
+  
+  data_with_treat <- data_in_window |> 
+    left_join(treat_col, by="city_id") |> 
+    drop_na(TREAT)
+  
+  # add post dummies and interaction
+  data_with_post <- data_with_treat |> 
+    mutate(POST = if_else(year >= treat_year, 1, 0)) |> 
+    mutate(D = if_else(TREAT * POST == 1, 1, 0))
+  
+  # add subexp info (for the stacked table)
+  output <- data_with_post |> 
+    mutate(subexp = terr, treat_year = treat_year)
+  
+  return(output)
+}
+

### 2023-05-12 paper/paper.tex
#15 structure into economics and econometrics

@@
 
-% ######################################################
+What am I even writing this for?
+Who am I writing this for?
+What is my role here?
 
-\section{Literature review}
+I think it's an interesting question, per se.
+When a ruling dynasty goes extinct, what happens to their land?
+That is the most interesting way I can phrase this.
+
+What are the most interesting aspects of that, for the purpose of this thesis? Which are the issues I want to discuss?
+
+The details of the empirical strategy which I use to find the answer.
+
+What it says and what it doesn't say.
+
+Explanations \textit{why} there might be an effect, and why it may have the sign it has. Possible channels for this effect to go through.
+
+The economics that are at play. Economic mechanisms, decision makers, markets, costs, benefits, institutions, incentives, strategic behaviour.
+
+I think these are really good already.
+Note that "historical context" is not on the list.
+That's something to spend one to two paragraphs on.
+
+No, the way to make this interesting is to go really deep into a) the economics and b) the econometrics.
+
+Make me care.
+Make the reader care. That's your challenge.
 
@@
 
-\section{Cities, lineages, and territories}
+\section{The Economics}
 
-% #####################################################
+% ######################################################
 
-\section{Empirical approach}
-
-% #####################################################
-
-\section{Results}
-
-% #####################################################
-
-\section{Validity and robustness}
+\section{The Econometrics}
 

### 2023-05-13 paper/paper.tex
#15 structure rquestion-economics-metrics
#15 add thoughts on research question

--- 2023-05-11 paper/paper.tex
@@
 
-\section{Introduction}
+Make me care.
+Make the reader care. That's your challenge.
 
@@
 
-\section{Literature review}
+\section{Research Question}
+
+When a ruling dynasty goes extinct, what happens to their lands?
+
+\subsection{Disambiguation}
+Let me be more precise. Concepts to be defined:
+\begin{itemize}
+    \item[ruling dynasty] In the context of the HRE. Until the 19th century geopolitics was not done in terms of nation-states, but in terms of the territorial holdings of noble families.
+    \item[their lands] The cities in that dynasty's possession at the time of its extinction.
+    \item[what happens] It's complicated and precisely this is what I need to clarify first.
+\end{itemize}
+
+What \textit{could} it mean?
+\begin{itemize}
+    \item economic prosperity.
+    \item public goods provision.
+    \item quality of government.
+\end{itemize}
+
+The whole thing is that construction by itself is just barely not interesting enough. i need just one intermediate step to motivate me using construction.
+
+i could be intellectually honest and just state the process i went through.
+i initially thought about economic prosperity, then it didn't line up with the wages, then i focused on public goods and ruler commitment instead.
+
+and i can actually do all of those by splitting my construction variable up!
+one outcome variable with all buildings (as a benchmark), another with public goods buildings, and the other with representative buildings.
+
+I love that idea.
+
+So my chain of reasoning will go the following way:
+\begin{itemize}
+    \item get to stating the research question in the shortest possible way
+    \item explain ruling dynasty
+    \item explain their lands
+    \item explain the aspects i'm interested in: economic prosperity, public goods provision, ruler commitment / quality of government
+\end{itemize}
+
+Where in this do i discuss the issue with the wages?
+The intro is no place for it, but i can't bring economic prosperity up as a serious goal of the analysis in the intro if i'm going to discredit it two sections later.
+i need to be cautious about it from the start, maybe mention it as a detour / side note but refer to the wages figure in the appendix.
+
+John H. Cochrane says "Figure out the \textit{one} central and novel contribution of your paper. Write this down in one paragraph. As with all your writing, this must be concrete."
+
+My central contribution is the finding that dynastic extinctions have (no/an/X) effect on urban economic development.
+
+\subsection{Issues to address}
+
+What are the most interesting aspects of that, for the purpose of this thesis? Which are the issues I want to discuss? What are secondary research questions?
+
+The effect of size changes is a secondary research question.
 
@@
 
-\section{Cities, lineages, and territories}
+\section{The Economics / Hypotheses}
 
-% #####################################################
+% ######################################################
 
-\section{Empirical approach}
-
-% #####################################################
-
-\section{Results}
-
-% #####################################################
-
-\section{Validity and robustness}
+\section{The Econometrics / Evidence}
 

### 2023-05-18 paper/paper.tex
#15 add introduction with business metaphor

@@
 \begin{abstract}
+When a ruling dynasty goes extinct, what happens to their lands?
 I use historical data on construction activity and territorial history to estimate the effect of switching between territories on urban construction in the Holy Roman Empire.
@@
 
-\section{Research Question}
+\section{Introduction}
 
-When a ruling dynasty goes extinct, what happens to their lands?
+% one paragraph to get the reader from where he stands, to understanding and caring about the research question.
+The geopolitics of central Europe, throughout its history and until very recently, have been defined by noble families struggling over control of a vast array of fragmented territories - unlike, for example, China.
+Rulers have played an important role in the fates of their nations.
+Examples come to mind.
+But did such geopolitical drama have real effects on the lives of people?
+
+Let us consider Example city A.
+From 1XXX on, A belonged to the dukedom of Saxe-Weimar-Eisenach, until its rulers, the House of Ascania, went extinct in 1XYZ.
+The territory of Saxe-Weimar-Eisenach was then subsumed into the much larger Electorate of Saxony, ruled by the Albertine branch of the House of Wettin.
+By the same token, A changed hands from the Ascanians to the Albertines.
+
+What happens on the political stage is clear; the economic ("real") implications are not.
+In the language of business acquisitions: do such extinction-and-succession events manifest similar to a strategic acquisition, with some degree of reorientation?
+Do the participants benefit from synergies and reduced transaction costs?
+Or are these events more akin to mere ownership transfers?
 
@@
 \begin{itemize}
-    \item[ruling dynasty] In the context of the HRE. Until the 19th century geopolitics was not done in terms of nation-states, but in terms of the territorial holdings of noble families.
+    \item[ruling dynasty] In the context of the HRE.
+Until the 19th century geopolitics was not done in terms of nation-states, but in terms of the territorial holdings of noble families.
+Leadership changes can be split into within-family and across-family - we care about the latter.
     \item[their lands] The cities in that dynasty's possession at the time of its extinction.
@@
 
+The effects of different types of rule.
+
+Leaning on Dube and Harish (2022): family-individual characteristics vs family-leadership characteristics (Habsburgs may be docile by nature but make aggressive decisions in the interest of their realm out of geopolitical necessity)
+
 % ######################################################
@@
 \section{The Econometrics / Evidence}
+
+Do mention that, because construction is a rare event, the time series for individual cities aren't as meaningful as averages.
 

### 2023-05-31 paper/paper.tex
#20 add skeleton of minimal lit review

@@
 
-\title{Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}
+\title{Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.
+I thank Matthias Weigand in particular for providing me with his own data and for pointing me towards the paper by \cite{schoenholzer2018}.}
 }
@@
 
+\section{Literature review}
+
+A seminal paper is that of \cite{as1997}. It suggests a theoretical model for determining the optimal size and number of states. The authors model the tradeoff between BLANK and BLANK.
+
+More recently, a lot of empirical research has been done using city-level historical data and investigating the effects of Europe's political landscape on the economy.
+In a way, this started with the Neo-Institutionalists - e.g.
+\cite{delong1993}, who investigate the effect of absolutist vs non-absolutist rule and find that self-governance is advantageous.
+
+\cite{diamond1997} discusses some ways in which the geography of a place, via its geopolitical implications, can affect the long-run development of civilisation: he contrasts remote and unified China with geographically and politically fragmented Europe, suggesting that BLANK.
+
+\cite{cervellati2022} provide a very rich survey of the literature on the state system in pre-industrial Europe. They also provide empirical evidence for BLANK.
+
+The next big thing in the economic history literature appears to be fiscal capacity. I'm not yet sure how that is relevant to my project, but I feel like I ought to mention it.
+
+Finally, \cite{schoenholzer2018} are the closest paper to this one.
+They ask a very similar question: How does switching between states affect long-run urban growth?
+To answer this, they do BLANK.
+They find that BLANK.
+The present project differs from theirs in a few ways: firstly, conflict is explicitly part of their analysis; it is part of what causes the tradeoff between short-run costs and long-run benefits.
+In my survey, conflict is merely a confounder whose effect I wish to exclude.
+
+2-3 pubs each on
+public goods provision
+transaction costs
+strategic motives
+
+% ######################################################
+
 \section{The Economics / Hypotheses}

### 2023-06-07 paper/paper.tex
#22 clear everything + update abstract

--- 2023-05-13 paper/paper.tex
@@
 
-\title{Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}
+\title{Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{
+I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.
+I thank Matthias Weigand in particular for providing me with his own data and for pointing me towards the paper by \cite{schoenholzer2022}.
+}
 }
@@
 \begin{abstract}
-I use historical data on construction activity and territorial history to estimate the effect of switching between territories on urban construction in the Holy Roman Empire.
-I exploit variation from dynastic extinctions, which caused cities to become part of different lineages' territorial holdings.
-I find that an X\% increase in the number of cities possessed by the ruling family is associated with a Y\% increase in construction activity.
+How does switching between states affect the long-run development of cities, and why?
+I study this question in the setting of the Holy Roman Empire, using historical data on construction activity and the territorial history of cities.
+I specifically test the hypothesis that competition between states, in which higher-quality states dominated, caused cities to benefit from permanent ownership changes.
+I find evidence in support of this.
 \end{abstract}
@@
 
-Make me care.
-Make the reader care. That's your challenge.
+\section{Introduction}
+
+% one paragraph to get the reader from where he stands, to understanding and caring about the research question.
 
@@
 
-\section{Research Question}
-
-When a ruling dynasty goes extinct, what happens to their lands?
-
-\subsection{Disambiguation}
-Let me be more precise. Concepts to be defined:
-\begin{itemize}
-    \item[ruling dynasty] In the context of the HRE. Until the 19th century geopolitics was not done in terms of nation-states, but in terms of the territorial holdings of noble families.
-    \item[their lands] The cities in that dynasty's possession at the time of its extinction.
-    \item[what happens] It's complicated and precisely this is what I need to clarify first.
-\end{itemize}
-
-What \textit{could} it mean?
-\begin{itemize}
-    \item economic prosperity.
-    \item public goods provision.
-    \item quality of government.
-\end{itemize}
-
-The whole thing is that construction by itself is just barely not interesting enough. i need just one intermediate step to motivate me using construction.
-
-i could be intellectually honest and just state the process i went through.
-i initially thought about economic prosperity, then it didn't line up with the wages, then i focused on public goods and ruler commitment instead.
-
-and i can actually do all of those by splitting my construction variable up!
-one outcome variable with all buildings (as a benchmark), another with public goods buildings, and the other with representative buildings.
-
-I love that idea.
-
-So my chain of reasoning will go the following way:
-\begin{itemize}
-    \item get to stating the research question in the shortest possible way
-    \item explain ruling dynasty
-    \item explain their lands
-    \item explain the aspects i'm interested in: economic prosperity, public goods provision, ruler commitment / quality of government
-\end{itemize}
-
-Where in this do i discuss the issue with the wages?
-The intro is no place for it, but i can't bring economic prosperity up as a serious goal of the analysis in the intro if i'm going to discredit it two sections later.
-i need to be cautious about it from the start, maybe mention it as a detour / side note but refer to the wages figure in the appendix.
-
-John H. Cochrane says "Figure out the \textit{one} central and novel contribution of your paper. Write this down in one paragraph. As with all your writing, this must be concrete."
-
-My central contribution is the finding that dynastic extinctions have (no/an/X) effect on urban economic development.
-
-\subsection{Issues to address}
-
-What are the most interesting aspects of that, for the purpose of this thesis? Which are the issues I want to discuss? What are secondary research questions?
-
-The effect of size changes is a secondary research question.
-
 % ######################################################
-
-\section{The Economics / Hypotheses}
-
-% ######################################################
-
-\section{The Econometrics / Evidence}
-
-% ######################################################
-
-\section{Conclusion}
 

### 2023-06-07 paper/paper.tex
#22 add handbook articles and snippets

@@
 
-% one paragraph to get the reader from where he stands, to understanding and caring about the research question.
+The state system of early modern Europe was fraught with conflict.
+
+This paper studies a different interaction between conflict and growth.
+I study how conflict affected the institutional quality of states.
+In particular, I investigate whether wars and conflict served as a selection mechanism - whether the constant competition caused better states to grow and worse states to shrink and eventually disappear.
+
+These ideas are not new -
+
+Institutional factors that benefit a state's economy also benefit a state's ability to win conflicts with others.
+
+While frequent wars brought devastation over the people and destruction to property, \cite{malthus1798} and others after him have argued, frequent wars may have kept population in check, leading to higher land-labour ratios and therefore a higher level of income per capita.
+\cite{voigtlnder2013} argue that the relative wealth of Europe after 1350 can be explained by a feedback loop centered on this Malthusian mechanism: after the huge loss of life caused by the Black Death, which served as a catalyst, taxable incomes rose, which made rulers richer, which allowed them to afford more wars, which in turn kept population in check.
 

### 2023-06-07 paper/paper.tex
#22 add levine modica snippets

@@
 
-Institutional factors that benefit a state's economy also benefit a state's ability to win conflicts with others.
+"In conflict luck matters, but success begets success.
+Conquering a city or a province strengthens the winner and weakens the loser.
+Hence as war and invasion unfold we expect that eventually one side gets lucky enough, strong enough, and faces opponents weak enough, that it wins outright." (\citealp{levine2021}, p.
+436).
+
+The model developed by \cite{levine2013} finds that "evolution selects societies with high state power - the ability to prevail in conflict over rivals." (\citealp{levine2021}, p. 436).
+
+Malthus vs conflict: Levine and Modica hate Malthus. They argue that social evolution is driven by conflict.
+
+In particular: evolution selects for aggression.
+"we show that those societies that simply seek to defend their territory are doomed to domination by those who aggressively seek expansion.
+Hence successful societies such as Spain and England have survived precisely because they are aggressive." (p.
+438) That is why Spain still cares about regaining Gibraltar and keeping Catalonia, and why the UK wants to keep Scotland and the Falklands.
+
+Their answer to Why Europe and not China or India: "Ghengis Khan, the cannon, and the English.
+Ghengis Khan accomplished two things: he brought cannons to Europe, and he depopulated Mongolia.
+In China the removal of the outside threat of the Mongols led - as the theory asserts - to an extractive hegemony.
+In Europe the advent of the cannon together with the preying of the English on the French and the Spanish resulted - as the theory asserts - in an inclusive balance of power.
+Here as is often asserted the English played a key role in the industrial revolution - albeit a rather different one than is commonly assumed." (p.
+439).
+
+How they define state power (pp.
+439-440): "the potential ability of a social organisation to resist and influence other societies.
+This depends on institutional characteristics of the society, including stability of the government and the ability to collect taxes and conscript soldiers, which depend in turn on the law-abidingness of citizens, the efficiency of the courts, and the overall economic strength of the society.
+
+They *assume* in their model (p.443) that state power is directly proportional to free resources (resources in excess of subsistence).
+"Then (unless the outsiders are very strong) the long-run stochastically stable outcome will maximise free resources.
+[in contrast to Malthus, where the long-run stable outcome is subsistence income.]" This is pretty much the result I also try to show.
+They do allow that not all free resources are spent on things that improve the chance of success in conflict (state power).
+
+"The residual - profits if you will - may be turned to many uses and only some of these uses - swords not jewelry - enhance state power.
+Hence a profit maximising monarchy that through social norms is bound to use its profits for fighting and conflict is the type we expect to survive; a profit maximising monarchy that spends its profits on large and beautiful palaces has less of a future." (p.
+444)
+
+Levine and Modica in a nutshell: societies with higher state power prevail.
+
+My own point, sharply: conquering a city implies state power, while inheriting it does not.
+
+NEXT SECTION: 15.6 PROSPERITY
+
+stylised fact to be explained in 15.6 (p.455): competition between inclusive institutions in Europe, extractive hegemony in China
+
+competing societies in India - why did they generate little innovation?
+
+they introduce another dimension alongside hegemony vs balance of power: societies can now be inclusive vs extractive
+
+puzzle: "since extractive institutions generally levy higher taxes and have larger armies, if evolution is driven by conflict how come these "strong" extractive institutions do not predominate over "weaker" inclusive institutions?
+
+model: inclusive institutions (commercial elites have upper hand) or extractive institutions (military elites have upper hand) within each society.
+Transfer from commercial to military elites in each period - smaller in inclusive societies.
+
+results:
+- inclusive hegemonies and heterogenous balances of power cannot be stochastically stable
+- if inclusive institutions do not differ much from extractive ones, only extractive institutions are stable in general.
+(The commerical elites won't make much effort to defend inclusive institutions that barely make a difference).
+whether it's a hegemony or balance of power then depends on the strength of the outsiders.
+- if the inclusive institutions are worth it for the commercial elites, but outsiders are weak, only extractive institutions are stable.
+if outsiders are strong, there is a balance of power, but whether it's inclusive or extractive depends on military technology: effective fortifications (defending is not costly) favour extractive institutions, while good siege technology (defending comes at a high cost) favours inclusive institutions.
+(pp.
+459-460)
+
+Result: 
+- arrival of cannon and strong outsiders made stable inclusive balance of power in Europe
+- weak outsiders favoured extractive hegemony in China
+- strong outsiders but primitive siege technology [cheap defensibility of cities] in India led to extractive balance of power
+
+-
+
+Institutional factors that benefit a state's economy also benefit a state's ability to win conflicts with others. Namely fiscal capacity
 

### 2023-06-07 paper/paper.tex
#22 edit down Levine in intro

--- 2023-06-07 paper/paper.tex
@@
 
-The state system of early modern Europe was fraught with conflict.
+\cite{levine2013, levine2021} argue that social evolution is driven by conflict.
+Their model finds that "evolution selects societies with high state power - the ability to prevail in conflict over rivals." (\citealp{levine2021}, p.
+436).Their answer to Why Europe and not China or India: "Ghengis Khan, the cannon, and the English.
+Ghengis Khan accomplished two things: he brought cannons to Europe, and he depopulated Mongolia.
+In China the removal of the outside threat of the Mongols led - as the theory asserts - to an extractive hegemony.
+In Europe the advent of the cannon together with the preying of the English on the French and the Spanish resulted - as the theory asserts - in an inclusive balance of power." (p.
+439).
+strong outsiders but primitive siege technology [cheap defensibility of cities] in India led to extractive balance of power.
 
-This paper studies a different interaction between conflict and growth.
-I study how conflict affected the institutional quality of states.
-In particular, I investigate whether wars and conflict served as a selection mechanism - whether the constant competition caused better states to grow and worse states to shrink and eventually disappear.
+How they define state power (pp.
+439-440): "the potential ability of a social organisation to resist and influence other societies.
+This depends on institutional characteristics of the society, including stability of the government and the ability to collect taxes and conscript soldiers, which depend in turn on the law-abidingness of citizens, the efficiency of the courts, and the overall economic strength of the society." They *assume* in their model (p.443) that state power is directly proportional to free resources (resources in excess of subsistence).
 
-These ideas are not new -
+I do not study their explanation for the European equilibrium that we observe in history; rather, I take a step back by taking the state system and the process of conflict-driven evolution as given.
+What interests me are the consequences of this process for economic growth.
+Is it, as \cite{schoenholzer2022} argue explicitly, beneficial creative destruction - does state power equal state quality?
+Or are powerful states detrimental after all - because they engage in more extraction?
+\cite{tilly1985}
 
-Institutional factors that benefit a state's economy also benefit a state's ability to win conflicts with others.
+Why we would expect the hypothesis to be true: state power, defined as the ability to win conflicts with rivals, consists of factors like fiscal capacity, efficient courts, law-abiding citizens (\citealp{levine2021}, pp.439-449) - institutional factors that are also widely held to be beneficial for growth (see the vast literature on institutions).
+Add some institutions lit here.
+
+My central identifying assumption: conquering a city implies more state power than inheriting it.
+
+\subsection*{Other literature}
 

### 2023-06-07 paper/paper.tex
#22 write banger first paragraph

@@
 
+The history of Europe was defined by conflict.
+No other continent at any time experienced more war than Europe in the early modern era (\citealp{voigtlnder2013}, p.
+174).
+Wars brought death and devastation to the societies affected by them, but their effects also manifested on the political map --- they play out in a diverse ecosystem of polities that come and go and change size over time as the borders get redrawn and territory changes hands.
+One can indeed imagine this as an evolutionary process: states compete for resources and powerful states eventually absorb weaker ones.
+As a result, traits that increase state power --- the ability to win conflicts against rivals --- get selected for, while other traits are doomed to disappear \citep{levine2013, levine2021}.
+
+In this paper I study the role of this process in the economic development of Europe. In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development.
+
 \cite{levine2013, levine2021} argue that social evolution is driven by conflict.
@@
 
+Key motivation papers: \cite{levine2013, levine2021}, \cite{diamond1997}, \cite{schoenholzer2022}, \cite{cervellati2022}, \cite{voigtlnder2013}, \cite{tilly1985}
+
+The idea that today's states are just the most successful protection rackets of the past already shows up in \cite{tilly1985}.
+
 While frequent wars brought devastation over the people and destruction to property, \cite{malthus1798} and others after him have argued, frequent wars may have kept population in check, leading to higher land-labour ratios and therefore a higher level of income per capita.
@@
 % ######################################################
+
+\section{Historical Background}
 

### 2023-06-08 paper/paper.tex
#22 add project description to intro

@@
 
-In this paper I study the role of this process in the economic development of Europe. In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development.
+In this paper I study the role of this process in the economic development of Europe.
+In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development.
+If the effect is positive, then selection of "better" states through conflict-driven evolution may partially explain the ascendancy of Europe.
 
-\cite{levine2013, levine2021} argue that social evolution is driven by conflict.
+I do this by looking at the development of cities after they switch states.
+In a paper from which I draw heavily, \cite{schoenholzer2022} find that switching states affects population negatively in the short run but positively in the long run.
+I take their analysis a step further by identifying special cases of switches: in \textit{conquests} a city is forcefully absorbed into a new state, while in \textit{successions} a city is absorbed into a new state after the rulers of the old state have gone extinct.
+Using these concepts I make a key assumption: conquests imply state power differentials but successions do not. If state A takes a city from state B,
+
+absorbing a city by conquest is more likely if the new state is more powerful than the old one, but absorbing a city by succession underlies no such limitations - it only requires
+
+the new state after a conquest is selected for state power, but the new state after a succession is not. Roughly speaking,
+
+If a city is absorbed via conquest,
+
+Past treatments of this topic have suffered from a lack of available historical data.
+To empirically tackle it one needs at least two things: the detailed territorial history of each city, and some city-level proxy for development, both reaching far enough into the past and ideally recorded in short intervals.
+
+\cite{diamond1997} and \cite{landes2006} present narrative arguments for the economic benefits of state competition but stop short of formal analysis. 
+More recently \cite{schoenholzer2022}, from whom I borrow extensively in this paper, use a dataset with a bigger geographical scope but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is only available in intervals of 50 or 100 years.
+They are still able to estimate short-run (same-period) and long-run (next two centuries) effects of switching states.
+The dataset I use has several advantages. First, it traces the territorial history of cities in the Holy Roman Empire precisely up to a single year. Second, it contains
+
+Rather than m
+
+\subsection*{Other literature}
+
+\cite{diamond1997} asserts that the geographical remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history.
+\cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled innovation in medieval China would not have survived long in a more fragmented state system.
+
+\cite{levine2013, levine2021} provide the theoretical foundation for conflict-driven selection, as well as compelling evidence for it. 
+They argue that social evolution is driven by conflict.
 Their model finds that "evolution selects societies with high state power - the ability to prevail in conflict over rivals." (\citealp{levine2021}, p.
@@
 My central identifying assumption: conquering a city implies more state power than inheriting it.
-
-\subsection*{Other literature}
 

### 2023-06-08 paper/paper.tex
#22 rearrange schönholzer

--- 2023-06-07 paper/paper.tex
@@
 
-In this paper I study the role of this process in the economic development of Europe. In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development.
+In this paper I study the role of this process in the economic development of Europe.
+In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development.
+If the effect is positive, then selection of "better" states through conflict-driven evolution may partially explain the ascendancy of Europe.
 
-\cite{levine2013, levine2021} argue that social evolution is driven by conflict.
+Past treatments of this topic have suffered from a lack of available historical data.
+\cite{diamond1997} and \cite{landes2006} each present narrative arguments for the economic benefits of state competition but stop short of formal empirical analysis.
+Such a study requires at the very least three things: 1) a detailed territorial history of European cities, 2) some city-level measure of development reaching far enough into the past, 3) some measure of state power, institutional quality, or whatever else one wishes to explain effects with.
+These data limitations are starting to be overcome.
+Recently \cite{schoenholzer2022} have investigated how switching between states affects the development of cities.
+They find that switching states affects population negatively in the short run but positively in the long run.
+Similar to the evolution metaphor, they interpret this as "creative destruction": the short run cost of destruction that is compensated by the long-run benefit of becoming part of a "better" state.
+However, they do not directly estimate the effect of anything other than switching to different states.
+
+I do this by looking at the development of cities after they switch states.
+I take their analysis a step further by identifying special cases of switches: in \textit{conquests} a city is forcefully absorbed into a new state, while in \textit{successions} a city is absorbed into a new state after the rulers of the old state have gone extinct.
+Comparing the effects of these two types of switches allows me to tease out the effect of state power.
+
+I make a key assumption: conquests imply state power differentials but successions do not. If state A takes a city from state B, 
+absorbing a city by conquest is more likely if the new state is more powerful than the old one, but absorbing a city by succession underlies no such limitations - it only requires 
+the new state after a conquest is selected for state power, but the new state after a succession is not. Roughly speaking, 
+If a city is absorbed via conquest,
+
+The dataset I use has several advantages. First, it traces the territorial history of cities in the Holy Roman Empire precisely up to a single year. Second, it contains
+
+Rather than m
+
+\subsection*{Other literature}
+
+\cite{diamond1997} asserts that the geographical remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history.
+\cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled innovation in medieval China would not have survived long in a more fragmented state system.
+
+\cite{levine2013, levine2021} provide the theoretical foundation for conflict-driven selection, as well as compelling evidence for it. 
+They argue that social evolution is driven by conflict.
 Their model finds that "evolution selects societies with high state power - the ability to prevail in conflict over rivals." (\citealp{levine2021}, p.
@@
 
-\subsection*{Other literature}
+\cite{schoenholzer2022} use a dataset with a bigger geographical scope but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is only available in intervals of 50 or 100 years, and a series of political maps.
 
@@
 \section{Historical Background}
+
+One paragraph about the history of Europe's state system; this is the place for Cervellati
+
+One story about a city's territory switch
+
+Two pithy figures, and a no-nonsense table
+
+NO MORE THAN TWO PAGES TOTAL
 

### 2023-06-08 paper/paper.tex
#22 explain key assumption

@@
 They find that switching states affects population negatively in the short run but positively in the long run.
-Similar to the evolution metaphor, they interpret this as "creative destruction": the short run cost of destruction that is compensated by the long-run benefit of becoming part of a "better" state.
-However, they do not directly estimate the effect of anything other than switching to different states.
+Similar to the evolution metaphor, they interpret this as "creative destruction": the short run cost of destruction is compensated by the long-run benefit of becoming part of a "better" state.
+However, their
 
-I do this by looking at the development of cities after they switch states.
-I take their analysis a step further by identifying special cases of switches: in \textit{conquests} a city is forcefully absorbed into a new state, while in \textit{successions} a city is absorbed into a new state after the rulers of the old state have gone extinct.
+The analysis in this paper extends that of \cite{schoenholzer2022} by identifying special cases of switching events: in \textit{conquests} a city is forcefully absorbed into a new state, while in \textit{successions} a city is absorbed into a new state after the rulers of the old state have gone extinct.
 Comparing the effects of these two types of switches allows me to tease out the effect of state power.
 
-I make a key assumption: conquests imply state power differentials but successions do not. If state A takes a city from state B, 
-absorbing a city by conquest is more likely if the new state is more powerful than the old one, but absorbing a city by succession underlies no such limitations - it only requires 
-the new state after a conquest is selected for state power, but the new state after a succession is not. Roughly speaking, 
-If a city is absorbed via conquest,
+I make a key assumption: conquests imply state power differentials but successions do not.
+In other words, if we consider the universe of all switching events, the average increase in state power from the old to the new owner would be higher for the conquests than for the successions.
+The reasoning is that conflict --- as in a conquest --- selects the new state for power, but conflict-free successions, depending mainly on dynastic marriages, are more "random".
+This assumption allows me to use conquests as a stand-in for an increase in state power.
 
-The dataset I use has several advantages. First, it traces the territorial history of cities in the Holy Roman Empire precisely up to a single year. Second, it contains
-
-Rather than m
+The rest of the paper proceeds as follows.
 
@@
 
-\section{Historical Background}
+\section{Background}
 

### 2023-06-09 paper/paper.tex
#22 add a lot more literature to intro

@@
 
-\title{Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{
+\title{Conflict Driven Evolution and Urban Development: \\ Evidence from the Holy Roman Empire\footnote{
 I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.
@@
 
-The history of Europe was defined by conflict.
+The European state system was forged in conflict.
 No other continent at any time experienced more war than Europe in the early modern era (\citealp{voigtlnder2013}, p.
 174).
-Wars brought death and devastation to the societies affected by them, but their effects also manifested on the political map --- they play out in a diverse ecosystem of polities that come and go and change size over time as the borders get redrawn and territory changes hands.
+Wars brought death and devastation to the societies affected by them, but their effects also manifested on the political map --- they play out in a diverse ecosystem of polities that come and go and change size over time.
+Borders get redrawn and territory changes hands.
 One can indeed imagine this as an evolutionary process: states compete for resources and powerful states eventually absorb weaker ones.
-As a result, traits that increase state power --- the ability to win conflicts against rivals --- get selected for, while other traits are doomed to disappear \citep{levine2013, levine2021}.
+As a result, traits that increase state power --- the ability to win conflicts against rivals --- get selected for, while unfavourable traits are doomed to disappear \citep{levine2013, levine2021}.
 
@@
 In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development.
-If the effect is positive, then selection of "better" states through conflict-driven evolution may partially explain the ascendancy of Europe.
+If the effect is positive, then the conflict driven evolution of states may partially explain the eventual ascendancy of Europe: by selecting for military success, Europe's violent history may have accidentally produced fiscal capacity, efficient legal systems, and prospering middle classes.
 
-Past treatments of this topic have suffered from a lack of available historical data.
-\cite{diamond1997} and \cite{landes2006} each present narrative arguments for the economic benefits of state competition but stop short of formal empirical analysis.
-Such a study requires at the very least three things: 1) a detailed territorial history of European cities, 2) some city-level measure of development reaching far enough into the past, 3) some measure of state power, institutional quality, or whatever else one wishes to explain effects with.
-These data limitations are starting to be overcome.
-Recently \cite{schoenholzer2022} have investigated how switching between states affects the development of cities.
+The idea that war and conflict can raise the standard of living dates back to \cite{malthus1798}.
+He reasoned that "positive checks" on the population (war, disease, famine) lead to higher land-labour ratios and therefore higher levels of income per capita.
+\cite{voigtlnder2013} build upon this Malthusian mechanism and add an explanation for the incessant wars: taxes on those higher incomes filled the treasuries of the rulers, which increased their demand for war --- a luxury good ---, which in turn kept population low.
+
+My story is not about population but institutions (although it is compatible with Malthusianism).
+Explaining long-run growth with institutional factors has a long tradition in economics: North, Acemoglu, property rights, civil law, middle class.
+The key institutional change brought about by conflict was fiscal capacity \citep{tilly1985, dincecco2012}, or the state's ability to raise taxes.
+A newer strand of research focuses on fiscal capacity:  \cite{cantoni2023}
+
+Theory of wars and states: \cite{tilly1985} argues that war was the main reason for governments to invest in fiscal capacity. The formal theory of conflict-driven evolution is from \cite{levine2013}.
+
+\cite{diamond1997} and \cite{landes1969, landes2006} have argued for the economic benefits of political fragmentation and state competition. \cite{diamond1997} asserts that geography 
+remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history.
+\cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled the spread of innovation in medieval China would not have survived long in a more fragmented state system.
+
+The closest relative to this paper is \cite{schoenholzer2022} who investigate how switching between states affects the development of cities.
 They find that switching states affects population negatively in the short run but positively in the long run.
-Similar to the evolution metaphor, they interpret this as "creative destruction": the short run cost of destruction is compensated by the long-run benefit of becoming part of a "better" state.
-However, their
+Similar to the evolution metaphor, they interpret this as "creative destruction": the short run cost of destruction is compensated by the long-run benefit of becoming part of a higher-quality state.
+However, they do not show explicitly how these improvements in state quality are driven by selection through conflict.
 
-The analysis in this paper extends that of \cite{schoenholzer2022} by identifying special cases of switching events: in \textit{conquests} a city is forcefully absorbed into a new state, while in \textit{successions} a city is absorbed into a new state after the rulers of the old state have gone extinct.
+\hspace{0.5cm}
+
+The analysis in this paper builds upon that of \cite{schoenholzer2022} and extends it by identifying special cases of switching events: in \textit{conquests} a city is forcefully absorbed into a new state, while in \textit{successions} a city is absorbed into a new state after the rulers of the old state have gone extinct.
 Comparing the effects of these two types of switches allows me to tease out the effect of state power.
+I find a positive and significant effect of state power on the long-run development of cities.
+This result suggests that conflict-driven evolution indeed played a role in the rise of Europe relative to other regions of the world.
 
 I make a key assumption: conquests imply state power differentials but successions do not.
-In other words, if we consider the universe of all switching events, the average increase in state power from the old to the new owner would be higher for the conquests than for the successions.
+Specifically, if we consider the universe of all switching events, the average increase in state power from the old to the new owner would be higher for the conquests than for the successions.
 The reasoning is that conflict --- as in a conquest --- selects the new state for power, but conflict-free successions, depending mainly on dynastic marriages, are more "random".
 This assumption allows me to use conquests as a stand-in for an increase in state power.
+Several other assumptions are needed to interpret my results causally, which I will discuss in detail below.
 
@@
 \subsection*{Other literature}
-
-\cite{diamond1997} asserts that the geographical remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history.
-\cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled innovation in medieval China would not have survived long in a more fragmented state system.
 
@@
 
+\section{Data}
+
+Empirical tests have suffered from a lack of available historical data.
+Such a study requires at the very least three things: 1) a detailed territorial history of European cities, 2) some city-level measure of development reaching far enough into the past, and 3) some measure of state power, institutional quality, or whatever else one wishes to explain effects with.
+These data limitations are starting to be overcome.
+
 % ######################################################

### 2023-06-09 paper/paper.tex
#22 add levine explanation

@@
 One can indeed imagine this as an evolutionary process: states compete for resources and powerful states eventually absorb weaker ones.
-As a result, traits that increase state power --- the ability to win conflicts against rivals --- get selected for, while unfavourable traits are doomed to disappear \citep{levine2013, levine2021}.
+As a result, traits that increase state power --- the ability to win conflicts against rivals --- get selected for, while unfavourable traits are doomed to disappear \citep{levine2021}.
 
@@
 If the effect is positive, then the conflict driven evolution of states may partially explain the eventual ascendancy of Europe: by selecting for military success, Europe's violent history may have accidentally produced fiscal capacity, efficient legal systems, and prospering middle classes.
+
+I build my paper on the theory of conflict driven evolution, outlined above, developed by \cite{levine2013, levine2021, levine2022}.
+Their model explains why hegemonies arose in East Asia while balances of power could prevail in Europe: unlike China, the major powers of continental Europe were frequently preyed upon by strong outsiders --- the Vikings, the Swedes, and the English --- which prevented any one state from growing too strong (\citealp{levine2021}, p.
+439).
+However, it focuses more on these aggregate patterns than on the traits that cause states to succeed.
+\footnote{The basic idea of Darwinian selection has previously been borrowed by economists, though - for example \cite{galor2002} do the literal version, where humans are subject to selection and accumulate favourable traits over the course of history.
+More infamously, \cite{clark2007} employs it in an attempt to explain why the Industrial Revolution took place in Europe, arguing that the rich out-bred the poor and passed favourable traits on to their offspring.
+in Europe but not elsewhere.}
 
@@
 My story is not about population but institutions (although it is compatible with Malthusianism).
-Explaining long-run growth with institutional factors has a long tradition in economics: North, Acemoglu, property rights, civil law, middle class.
-The key institutional change brought about by conflict was fiscal capacity \citep{tilly1985, dincecco2012}, or the state's ability to raise taxes.
-A newer strand of research focuses on fiscal capacity:  \cite{cantoni2023}
-
-Theory of wars and states: \cite{tilly1985} argues that war was the main reason for governments to invest in fiscal capacity. The formal theory of conflict-driven evolution is from \cite{levine2013}.
+A longstanding tradition in economics emphasises the role of institutional factors in long-run growth (e.g.
+\cite{north1970}, \cite{delong1993}, \cite{ajr2001}).
+From this vast literature one theme is especially relevant here: the institution of fiscal capacity, or the state's ability to collect taxes.
+\cite{tilly1985} argues that the main reason for states to collect taxes is to be able to wage war.
+In other words, fiscal capacity and state power are tightly linked --- if conflict selects for state power, then it also selects for fiscal capacity.
+\cite{cantoni2023} show that drove state consolidation and played a role in growth.
+\cite{dincecco2012}.
 
@@
 \cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled the spread of innovation in medieval China would not have survived long in a more fragmented state system.
+Fragmentation also plays a central role in \cite{cervellati2022}'s theory.
 
@@
 However, they do not show explicitly how these improvements in state quality are driven by selection through conflict.
-
-\hspace{0.5cm}
 

### 2023-06-09 paper/paper.tex
#22 finish introduction

@@
 I build my paper on the theory of conflict driven evolution, outlined above, developed by \cite{levine2013, levine2021, levine2022}.
-Their model explains why hegemonies arose in East Asia while balances of power could prevail in Europe: unlike China, the major powers of continental Europe were frequently preyed upon by strong outsiders --- the Vikings, the Swedes, and the English --- which prevented any one state from growing too strong (\citealp{levine2021}, p.
+Its main contribution is to explain why hegemonies arose in East Asia while balances of power could prevail in Europe: unlike China, the major powers of continental Europe were frequently preyed upon by strong outsiders --- the Vikings, the Swedes, and the English --- which prevented any one state from growing too strong (\citealp{levine2021}, p.
 439).
-However, it focuses more on these aggregate patterns than on the traits that cause states to succeed.
-\footnote{The basic idea of Darwinian selection has previously been borrowed by economists, though - for example \cite{galor2002} do the literal version, where humans are subject to selection and accumulate favourable traits over the course of history.
-More infamously, \cite{clark2007} employs it in an attempt to explain why the Industrial Revolution took place in Europe, arguing that the rich out-bred the poor and passed favourable traits on to their offspring.
-in Europe but not elsewhere.}
+It is worth highlighting how the authors model war between states: you win with state power, and state power is simply free resources, i.e.
+resources above subsistence.
+If the most successful states are the ones with the biggest surplus, selection should have made Europe increasingly prosperous.\footnote{The basic idea of Darwinian selection has previously been borrowed by economists - for example \cite{galor2002} do the literal version, where humans are subject to selection and accumulate favourable traits over the course of history.
+More infamously, \cite{clark2007} employs it in an attempt to explain why the Industrial Revolution took place in Europe, arguing that the rich out-bred the poor and passed favourable traits on to their offspring (in Europe but not elsewhere for some reason.)}
 
@@
 \cite{tilly1985} argues that the main reason for states to collect taxes is to be able to wage war.
-In other words, fiscal capacity and state power are tightly linked --- if conflict selects for state power, then it also selects for fiscal capacity.
-\cite{cantoni2023} show that drove state consolidation and played a role in growth.
-\cite{dincecco2012}.
+In other words, fiscal capacity and state power are tightly linked --- so if conflict selects for state power, then it also selects for fiscal capacity.
+\cite{cantoni2023} show that conflict drove state consolidation.
+\cite{dincecco2012} show that fiscal capacity had a positive effect on growth.
 
@@
 However, they do not show explicitly how these improvements in state quality are driven by selection through conflict.
+They use a dataset with a bigger geographical scope (all of Europe) but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is only available in intervals of 50 or 100 years, and a series of political maps.
 
@@
 The rest of the paper proceeds as follows.
-
-\subsection*{Other literature}
-
-\cite{levine2013, levine2021} provide the theoretical foundation for conflict-driven selection, as well as compelling evidence for it. 
-They argue that social evolution is driven by conflict.
-Their model finds that "evolution selects societies with high state power - the ability to prevail in conflict over rivals." (\citealp{levine2021}, p.
-436).Their answer to Why Europe and not China or India: "Ghengis Khan, the cannon, and the English.
-Ghengis Khan accomplished two things: he brought cannons to Europe, and he depopulated Mongolia.
-In China the removal of the outside threat of the Mongols led - as the theory asserts - to an extractive hegemony.
-In Europe the advent of the cannon together with the preying of the English on the French and the Spanish resulted - as the theory asserts - in an inclusive balance of power." (p.
-439).
-strong outsiders but primitive siege technology [cheap defensibility of cities] in India led to extractive balance of power.
-
-How they define state power (pp.
-439-440): "the potential ability of a social organisation to resist and influence other societies.
-This depends on institutional characteristics of the society, including stability of the government and the ability to collect taxes and conscript soldiers, which depend in turn on the law-abidingness of citizens, the efficiency of the courts, and the overall economic strength of the society." They *assume* in their model (p.443) that state power is directly proportional to free resources (resources in excess of subsistence).
-
-I do not study their explanation for the European equilibrium that we observe in history; rather, I take a step back by taking the state system and the process of conflict-driven evolution as given.
-What interests me are the consequences of this process for economic growth.
-Is it, as \cite{schoenholzer2022} argue explicitly, beneficial creative destruction - does state power equal state quality?
-Or are powerful states detrimental after all - because they engage in more extraction?
-\cite{tilly1985}
-
-Why we would expect the hypothesis to be true: state power, defined as the ability to win conflicts with rivals, consists of factors like fiscal capacity, efficient courts, law-abiding citizens (\citealp{levine2021}, pp.439-449) - institutional factors that are also widely held to be beneficial for growth (see the vast literature on institutions).
-Add some institutions lit here.
-
-My central identifying assumption: conquering a city implies more state power than inheriting it.
-
-\cite{schoenholzer2022} use a dataset with a bigger geographical scope but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is only available in intervals of 50 or 100 years, and a series of political maps.
-
-Key motivation papers: \cite{levine2013, levine2021}, \cite{diamond1997}, \cite{schoenholzer2022}, \cite{cervellati2022}, \cite{voigtlnder2013}, \cite{tilly1985}
-
-The idea that today's states are just the most successful protection rackets of the past already shows up in \cite{tilly1985}.
-
-While frequent wars brought devastation over the people and destruction to property, \cite{malthus1798} and others after him have argued, frequent wars may have kept population in check, leading to higher land-labour ratios and therefore a higher level of income per capita.
-\cite{voigtlnder2013} argue that the relative wealth of Europe after 1350 can be explained by a feedback loop centered on this Malthusian mechanism: after the huge loss of life caused by the Black Death, which served as a catalyst, taxable incomes rose, which made rulers richer, which allowed them to afford more wars, which in turn kept population in check.
+Section 2 presents some background information.
+Section 3 describes the data used in the analysis.
+Section 4 details the empirical approach.
+Section 5 presents results.
+Section 6 discusses implications and robustness of the results.
+Section 7 concludes.
 

### 2023-06-10 paper/paper.tex

@@
 In other words, fiscal capacity and state power are tightly linked --- so if conflict selects for state power, then it also selects for fiscal capacity.
-\cite{cantoni2023} show that conflict drove state consolidation.
+\cite{gennaioli2015} and \cite{cantoni2023} show that conflict drove state consolidation.
 \cite{dincecco2012} show that fiscal capacity had a positive effect on growth.
@@
 
-Empirical tests have suffered from a lack of available historical data.
-Such a study requires at the very least three things: 1) a detailed territorial history of European cities, 2) some city-level measure of development reaching far enough into the past, and 3) some measure of state power, institutional quality, or whatever else one wishes to explain effects with.
-These data limitations are starting to be overcome.
+I use data from the \textit{Princes and Townspeople} database which is largely based on the \textit{Deutsches Städtebuch} \citep{keyser1939} and \cite{kobler2007}'s \textit{Historisches Lexikon der Deutschen Länder}.
+The specific datasets I use are those on city locations \citep{pt1}, territorial histories \citep{pt2}, construction activity \citep{pt5} and conflict incidents \citep{pt6}.\footnote{I thank Davide Cantoni and Matthias Weigand for giving me access to unpublished yearly versions of the construction and territorial histories data.} I aggregate data from these sources to obtain a city-by-year panel of switching events, construction activity, and conflict.
+
+Switching: describe format; total number of recorded switches; territory id.
+
+Construction: describe format / split and rationale. This can be two paragraphs.
+
+Conflict: describe format (just a dummy).
+
+I also build coarsened versions of this dataset in which I aggregate years into longer periods.
+Apart from the yearly version there is also a 10-yearly version, a 50-yearly version and a 100-yearly version.
+Not only does this make event studies nicer to look at, it also allows me to use more of the conflict and construction data: the 50-year and 100-year versions contain events that aren't recorded with a precise year, but only "beginning of the 16th century" and the like.
 

### 2023-06-10 paper/paper.tex
#23 clarify aggregation description

@@
 
-I also build coarsened versions of this dataset in which I aggregate years into longer periods.
-Apart from the yearly version there is also a 10-yearly version, a 50-yearly version and a 100-yearly version.
-Not only does this make event studies nicer to look at, it also allows me to use more of the conflict and construction data: the 50-year and 100-year versions contain events that aren't recorded with a precise year, but only "beginning of the 16th century" and the like.
+I also build coarsened versions of this dataset in which I aggregate years into longer periods (10, 50, and 100 years).
+Not only does this aggregation make event studies nicer to look at, it also allows me to use more of the conflict and construction data as periods get longer: the 50-year and 100-year versions contain events that aren't recorded with a precise year, but only "beginning of the 16th century" and the like.
+While the 1-year and 10-year data builds contain BLANK construction events with sufficiently precise timing, the 50-year version contains BLANK and the 100-year version contains BLANK.
 

### 2023-06-10 source/analysis/exploration/ExploreOutcomes.Rmd
#23 HUGE RESTRUCTURE

--- 2023-04-20 analysis/code/exploration/ExploreOutcomes.Rmd
@@
 plot_grid(wage_plot, wratio_plot, const_plot, align="v", axis="b", ncols=1)
-ggsave("const_wages.png", device = "png", path="analysis/output/exploration")
+# ggsave("const_wages.png", device = "png", path="analysis/output/exploration")
 ```

### 2023-06-10 utils/AggregateYears.R

@@ -1,30 +1,0 @@
-AggregateYears <- function(build, spacing){
-  # The steps are:
-  # - map years to periods.
-  # - group yearly observations by period and map yearly values to period values.
-  #   - for construction, use the group mean (better than sum bc of truncated periods!).
-  #   - for treatment, use the group maximum.
-  # After that, you can compute time_to_treat just like before, with years -> periods.
-  aggregated <- build |> 
-    mutate(period = year - year %% spacing) |>
-    group_by(city_id, period) |> 
-    summarise(
-      construction = mean(construction, na.rm = T),
-      treatment = max(treatment)
-    )
-  
-  return(aggregated)
-}
-
-AggregateStacked <- function(stacked_data, spacing){
-  aggregated <- stacked_data |> 
-    mutate(
-      period = year - year %% spacing, 
-      treat_period = treat_year - treat_year %% spacing) |> 
-    group_by(city_id, period, subexp) |> 
-    mutate(construction = mean(construction, na.rm = T)) |> 
-    select(city_id, period, construction, TREAT, subexp, treat_period) |> 
-    distinct()
-  
-  return(aggregated)
-}

### 2023-06-10 utils/BaselineSample.R

@@ -1,21 +1,0 @@
-BaselineSample <- function(build){
-
-  # Drop cities that were treated more than once from the sample
-  max1treat <- build |>  
-    group_by(city_id) |> 
-    mutate(treatments_total = sum(treatment)) |> 
-    filter(treatments_total < 2)
-  
-  n_dropped <- nrow(build)-nrow(max1treat)
-  cities_dropped <- n_distinct(build$city_id)-n_distinct(max1treat$city_id)
-  
-  print(sprintf("Dropped %d observations with multiple treatments (%d cities)", 
-                n_dropped, cities_dropped))
-  
-  # Remove superfluous columns
-  clean <- max1treat |> 
-    select(year, city_id, construction, treatment, count_cities, count_diff) |> 
-    as.data.frame()
-  
-  return(clean)
-}

### 2023-06-12 paper/paper.tex
#25 add results to manuscript

@@
 
+\section{Empirical strategy}
+
+First I replicate the long-term finding of \cite{schoenholzer2022} using the following equation:
+
+\begin{equation}
+    construction_{it} = \alpha_i + \delta_t + \mathbf{1}(S_{it} > 0) + 
+    \sum_{\tau = -100}^{200} \mathbf{1}(t = e^{NewState}_i + \tau)\pi_t + \varepsilon_{it}
+\end{equation}
+
+I can barely believe it myself, but I replicate their finding:
+
+\begin{figure}
+    \centering
+    \includegraphics[scale = 0.4]{paper/output/regressions/SW22_replication_50y.png}
+    \caption{It looks just like it's supposed to!}
+    \label{fig:SW_replication}
+\end{figure}
+
+Next I tried to find the effect of state power - the effect of the difference in selection for state power between conquests and successions:
+
+\begin{multline}
+    construction_{it} = \alpha_i + \delta_t + \mathbf{1}(S_{it} > 0) +
+    \mathbf{1}(t \ge e^{NewState}_i)\beta +
+    \mathbf{1}(t \ge e^{NewState}_i) \times C_i \theta + \varepsilon_{it}
+\end{multline}
+
+Unfortunately it's not as convincing:
+
+\input{paper/output/regressions/baseline_50y}
+
 \newpage

### 2023-06-12 source/analysis/baseline/Baseline.R
#25 add conflict control to baseline

@@ -22,16 +22,24 @@ Main <- function(){
                      binarise_switches = T)
   
   # Estimate my baseline interaction equation
-  mod <- fixest::feols(
+  mod_basic <- fixest::feols(
     c_all ~ i(treat_type, D, ref=2) + D + switches | city_id + period, 
     data = dat)
   
+  mod_conflict <- fixest::feols(
+    c_all ~ i(treat_type, D, ref=2) + D + switches + conflict | city_id + period, 
+    data = dat)
+  
+  
+  
   # Produce regression table and export to LaTeX
   setFixest_dict(c(city_id = "City", period = "Period",
-                   c_all = "All construction (binary)", D = "TREAT x POST",
-                   treat_type = "Type", switches = "Switch"))
+                   c_all = "Construction", D = "1(t > e^NewState)",
+                   treat_type = "Switch Type", switches = "1(S_{it} > 0"),
+                   conflict = "Conflict")
   
-  tex_output <- etable(mod, tex=TRUE)
+  etable(mod_basic, mod_conflict)
+  tex_output <- etable(mod_basic, mod_conflict, tex=TRUE)
   
   filename <- sprintf("paper/output/regressions/baseline_%iy.tex", t)
   write(tex_output, file=filename)

### 2023-06-12 source/analysis/baseline/Baseline.R
#25 make etableDefaults utility

@@ -7,6 +7,7 @@ setwd("~/GitHub/BA")
 
 source("source/utils/HandleCommandArgs.R")
 source("source/utils/DataPrepSuite.R")
+source("source/utils/etableDefaults.R")
 
 Main <- function(){
   
@@ -30,16 +31,10 @@ Main <- function(){
     c_all ~ i(treat_type, D, ref=2) + D + switches + conflict | city_id + period, 
     data = dat)
   
-  
-  
   # Produce regression table and export to LaTeX
-  setFixest_dict(c(city_id = "City", period = "Period",
-                   c_all = "Construction", D = "1(t > e^NewState)",
-                   treat_type = "Switch Type", switches = "1(S_{it} > 0"),
-                   conflict = "Conflict")
-  
-  etable(mod_basic, mod_conflict)
-  tex_output <- etable(mod_basic, mod_conflict, tex=TRUE)
+  etableDefaults()
+  tex_output <- etable(mod_basic, mod_conflict, tex=TRUE,
+                       title="Test Title")
   
   filename <- sprintf("paper/output/regressions/baseline_%iy.tex", t)
   write(tex_output, file=filename)

### 2023-06-12 source/utils/etableDefaults.R

@@ -1,0 +1,15 @@
+etableDefaults <- function(){
+  
+  # uniform defaults (across all tables)
+  setFixest_etable(
+    se.below = FALSE,
+    digits = 2,
+    digits.stats = 4,
+  )
+  
+  # set default dict
+  setFixest_dict(c(city_id = "City", period = "Period",
+                   c_all = "Construction", D = "Switching to another state",
+                   treat_type = "Selection", switches = "Switching indicator"),
+                 conflict = "Conflict")
+  }

### 2023-06-17 source/analysis/baseline/ReplicateSW22.R
#25 add SW22 table

@@ -7,6 +7,7 @@ setwd("~/GitHub/BA")
 
 source("source/utils/HandleCommandArgs.R")
 source("source/utils/DataPrepSuite.R")
+source("source/utils/etableDefaults.R")
 
 Main <- function(){
   
@@ -29,10 +30,19 @@ Main <- function(){
     data = dat)
 
   
+  # Produce regression table and export to LaTeX
+  etableDefaults()
+  tex_output <- etable(mod, tex=TRUE,
+                       title="Dynamic effects of switching")
+  
+  filename <- sprintf("paper/output/regressions/SWW22_replication_%iy.tex", t)
+  write(tex_output, file=filename)
+  
+  
   # Produce replication of Fig. 5 and save as PNG
   filename <- sprintf("paper/output/regressions/SW22_replication_%iy.png", t)
   png(filename=filename, width = 800, height = 800, pointsize = 20)
-  iplot(mod)   
+  iplot(mod) 
   dev.off()
   
   return(0)

### 2023-06-17 paper/paper.tex
Updates from Overleaf

@@
 Comparing the effects of these two types of switches allows me to tease out the effect of state power.
-I find a positive and significant effect of state power on the long-run development of cities.
-This result suggests that conflict-driven evolution indeed played a role in the rise of Europe relative to other regions of the world.
-
-I make a key assumption: conquests imply state power differentials but successions do not.
+Identification is based on the assumption that conquests imply state power differentials but successions do not.
 Specifically, if we consider the universe of all switching events, the average increase in state power from the old to the new owner would be higher for the conquests than for the successions.
@@
 This assumption allows me to use conquests as a stand-in for an increase in state power.
-Several other assumptions are needed to interpret my results causally, which I will discuss in detail below.
+Several other assumptions are needed to interpret my results causally.
+
+While I am able to replicate \cite{schoenholzer2022}'s result that the long-run effect of switching to another state is positive, I fail to identify a statistically significant effect of state power on urban development.
+This does not necessarily mean that conflict-driven evolution played no role in the rise of Europe --- several issues limit the statistical power of my test, which might prevent me from finding an effect even if there is one.
+The biggest one is that the data I use is arguably inadequate for answering the research question.
 
@@
 
+\subsection{Long-run effect of switching}
+
 First I replicate the long-term finding of \cite{schoenholzer2022} using the following equation:
@@
 
-I can barely believe it myself, but I replicate their finding:
-
-\begin{figure}
-    \centering
-    \includegraphics[scale = 0.4]{paper/output/regressions/SW22_replication_50y.png}
-    \caption{It looks just like it's supposed to!}
-    \label{fig:SW_replication}
-\end{figure}
+\subsection{Effect of state power}
 
 Next I tried to find the effect of state power - the effect of the difference in selection for state power between conquests and successions:
+
+I use a triple interaction. I extend the previous regression
 
@@
     \mathbf{1}(t \ge e^{NewState}_i)\beta +
-    \mathbf{1}(t \ge e^{NewState}_i) \times C_i \theta + \varepsilon_{it}
+    \mathbf{1}(t \ge e^{NewState}_i) \times C_i \times \theta + \varepsilon_{it}
 \end{multline}
 
-Unfortunately it's not as convincing:
+Blah blah.
 
-\input{paper/output/regressions/baseline_50y}
+\section{Results}
+
+I can barely believe it myself, but I replicate their finding. Table 
+Figure \ref{fig:SW_replication} shows the dynamic effect of switching. Cities were on average
+
+The it's not as convincing:
 
@@
 
+\subsection{Tables}
+
+\input{paper/output/regressions/baseline_50y}
+
+\subsection{Figures}
+
+\begin{figure}
+    \centering
+    \includegraphics[scale = 0.4]{paper/output/regressions/SW22_replication_50y.png}
+    \caption{It looks just like it's supposed to!}
+    \label{fig:SW_replication}
+\end{figure}
+
 \end{document}

### 2023-06-17 source/analysis/baseline/Baseline.R
#25 add tex labels to reg tables

@@ -34,7 +34,8 @@ Main <- function(){
   # Produce regression table and export to LaTeX
   etableDefaults()
   tex_output <- etable(mod_basic, mod_conflict, tex=TRUE,
-                       title="Test Title")
+                       title="Test Title",
+                       label = sprintf("tab:baseline_%iy", t))
   
   filename <- sprintf("paper/output/regressions/baseline_%iy.tex", t)
   write(tex_output, file=filename)

### 2023-06-17 source/analysis/baseline/ReplicateSW22.R

@@ -32,10 +32,12 @@ Main <- function(){
   
   # Produce regression table and export to LaTeX
   etableDefaults()
-  tex_output <- etable(mod, tex=TRUE,
-                       title="Dynamic effects of switching")
+  tex_output <- etable(mod, 
+                       tex=TRUE,
+                       title="Dynamic effects of switching",
+                       label = sprintf("tab:SW22_replication_%iy", t))
   
-  filename <- sprintf("paper/output/regressions/SWW22_replication_%iy.tex", t)
+  filename <- sprintf("paper/output/regressions/SW22_replication_%iy", t)
   write(tex_output, file=filename)
   
   

### 2023-06-17 source/analysis/baseline/ReplicateSW22.R
#25 fix file path

@@ -37,7 +37,7 @@ Main <- function(){
                        title="Dynamic effects of switching",
                        label = sprintf("tab:SW22_replication_%iy", t))
   
-  filename <- sprintf("paper/output/regressions/SW22_replication_%iy", t)
+  filename <- sprintf("paper/output/regressions/SW22_replication_%iy.tex", t)
   write(tex_output, file=filename)
   
   

### 2023-06-17 source/analysis/exploration/CitiesMap.R
#27 messy map code

@@ -1,0 +1,52 @@
+library(tidyverse)
+library(sf)
+library(mapview)
+library(rnaturalearth)
+library(tmap)
+
+setwd("~/GitHub/BA")
+
+source("source/utils/HandleCommandArgs.R")
+
+
+
+# Read data
+t <- HandleCommandArgs(default_length = 50)
+read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
+build <- read_csv(read_path, show_col_types = F)
+map_points <- st_read("drive/raw/attributes/city_locations")
+worldmap <- ne_countries(scale = 'small', 
+                         type = 'map_units',
+                         continent = 'europe',
+                         returnclass = 'sf')
+worldmap.fix <- eu_map |>  
+  st_crop(c(xmin= 3, ymin = 45, xmax = 25.5, ymax = 56))
+
+
+
+# worldmap.fix <- st_make_valid(worldmap)
+st_is_valid(worldmap, reason=T)
+
+
+worldmap <- worldmap |> 
+  st_transform(crs=st_crs(map_points)) #|> 
+  #st_set_crs(st_crs(map_eu))
+map_joined <- st_join(map_eu, map_points)
+
+build_with_locations <- build |> 
+  group_by(city_id) |> 
+  summarise(lifetime_c_all = sum(c_all)) |> 
+  filter(lifetime_c_all > 20) |> 
+  left_join(shapefile, by="city_id") |> 
+  st_as_sf()
+
+set_st_crs <- st_crs(shapefile)
+shapefile
+eu_map
+
+mapView(map_points)
+
+
+
+tm_shape(worldmap)+
+  tm_polygons()

### 2023-06-17 paper/paper.tex
Updates from Overleaf
#25 proper labels for figs and tabs

@@
 
-My story is not about population but institutions (although it is compatible with Malthusianism).
+By contrast, my story is not about population dynamics but institutions (although it is compatible with Malthusianism).
 A longstanding tradition in economics emphasises the role of institutional factors in long-run growth (e.g.
@@
 
-One paragraph about the history of Europe's state system; this is the place for Cervellati
+Early modern Europe consisted of many large and small states.
 
@@
 
-The it's not as convincing:
+The it's not as convincing: \ref{tab:SW22_replication_50y}
 
@@
 
-\subsection{Tables}
+\subsection*{Tables}
+
+\input{paper/output/regressions/SW22_replication_50y}
 
@@
 
-\subsection{Figures}
+\clearpage
+\subsection*{Figures}
 

### 2023-06-17 paper/paper.tex
#25 explain regression 1

@@
 
+\cite{schoenholzer2022} find a short-run cost of switching (presumably due to conflict) and a long-run benefit.
+They get the short-term result by looking at cities that only switched states briefly before switching back, and the long-run result by comparing that to cities that permanently stayed with their new territory.
+I operate in the same framework and I borrow their notation: the variable $S_{it}$ is the number of times a city switched states in period $t$.
+By contrast, the variable $e^{NewState}_i$ denotes the first period in which city $i$ belongs to a different state than the previous.
+Intuitively, in periods with $S_{it} > 0$ a city suffers short-run effects of contemporary switching, while in periods following $e^{NewState}_i$ it will be under new rule and therefore experience long-run effects of switching.
+
 First I replicate the long-term finding of \cite{schoenholzer2022} using the following equation:
@@
     construction_{it} = \alpha_i + \delta_t + \mathbf{1}(S_{it} > 0) + 
-    \sum_{\tau = -100}^{200} \mathbf{1}(t = e^{NewState}_i + \tau)\pi_t + \varepsilon_{it}
+    \sum_{\tau = -100}^{200} \mathbf{1}(t = e^{NewState}_i + \tau)\pi_t + \varepsilon_{it},
 \end{equation}
+
+where the outcome $construction_{it}$ is an indicator that equals 1 if any construction activity was recorded in city $i$ and period $t$,  the variable $S_{it}$ is the number of times a city switched states in period $t$, the variable $e^{NewState}_i$ denotes the first period in which city $i$ belongs to a different state than the previous, and $\alpha_i$ and $\delta_t$ denote city and period fixed effects.
+
+Some comments on this.
+First, it may seem like a good idea to include state fixed effects.
+But I cannot because the hypothesised selection effect works through an improvement in state quality.
+State fixed effects would absorb the variation in time-invariant state quality.
 

### 2023-06-18 paper/paper.tex
#25 first half of empirics section

@@
 \begin{equation}
-    construction_{it} = \alpha_i + \delta_t + \mathbf{1}(S_{it} > 0) + 
-    \sum_{\tau = -100}^{200} \mathbf{1}(t = e^{NewState}_i + \tau)\pi_t + \varepsilon_{it},
+\label{eq:sw22}
+    construction_{it} = \alpha_i + \delta_t + \mathbf{1}(S_{it} > 0)\beta + 
+    \sum_{\tau = -100}^{200} \mathbf{1}(t = e^{NewState}_i + \tau)\pi_\tau + \varepsilon_{it},
 \end{equation}
@@
 where the outcome $construction_{it}$ is an indicator that equals 1 if any construction activity was recorded in city $i$ and period $t$, the variable $S_{it}$ is the number of times a city switched states in period $t$, the variable $e^{NewState}_i$ denotes the first period in which city $i$ belongs to a different state than the previous, and $\alpha_i$ and $\delta_t$ denote city and period fixed effects.
+The coefficients of interest are the dynamic effects $\pi_\tau$ of switching to a new state permanently.
 
-Some comments on this.
-First, it may seem like a good idea to include state fixed effects.
-But I cannot because the hypothesised selection effect works through an improvement in state quality.
-State fixed effects would absorb the variation in time-invariant state quality.
+Since the outcome is binary, equation \ref{eq:sw22} represents a linear probability model: we model $construction_{it}$ as a random variable that follows a Bernoulli distribution with a probability parameter $p$ that is conditional on the right-hand side variables.
+This way we can interpret the values of coefficients as differences in the probability of recorded construction activity in a given period.
+
+The coefficient $\pi_\tau$ for given $\tau$ is the difference in the probability of construction between two cities $i$ and $j$ with identical time-invariant characteristics, in the same period, who undergo the same amount of switching turmoil in this period.
+The only difference is that one city underwent a lasting switch to another state $\tau$ periods ago, while the other did not.
+
+% Since comparisons between cities at different \tau are problematic, we also use the Sun and Abraham estimator (as do SW22).
+
+Remember, we expect this difference to stem from the fact that switching states generally means switching to a higher-quality state (due to selection).
+As such, even though it would be possible in principle to include state fixed effects, it is not a good idea here: state fixed effects would absorb the variation in state quality whose effects we want to capture in $\pi_\tau$.
 
@@
 
-Next I tried to find the effect of state power - the effect of the difference in selection for state power between conquests and successions:
+Next I tried to find the effect of state power - the effect of the difference in selection for state power between conquests and successions.
 
-I use a triple interaction. I extend the previous regression
+I switch to a simple difference-in-differences specification for clarity. I extend the previous regression by interacting with the type of the ownership change:
 
-\begin{multline}
-    construction_{it} = \alpha_i + \delta_t + \mathbf{1}(S_{it} > 0) +
-    \mathbf{1}(t \ge e^{NewState}_i)\beta +
-    \mathbf{1}(t \ge e^{NewState}_i) \times C_i \times \theta + \varepsilon_{it}
-\end{multline}
+\begin{equation}
+\label{eq:baseline}
+    construction_{it} = \alpha_i + \delta_t + \mathbf{1}(S_{it} > 0)\beta +
+    \mathbf{1}(t \ge e^{NewState}_i)\pi +
+    \mathbf{1}(t \ge e^{NewState}_i) T_i \theta + \varepsilon_{it}
+\end{equation}
 
-Blah blah.
+a second-order interaction necessitates that the first-order interaction pattern and associated coefficient (whether zero or non-zero) is itself different depending on the level of a third predictor.
+(from stackexchange)
+
+Blah blah. Equation \ref{eq:baseline}
+
+% ######################################################
 
@@
 The it's not as convincing: \ref{tab:SW22_replication_50y}
+
+% ######################################################
+
+\section{Discussion}
+
+% ######################################################
+
+\section{Conclusion}
 

### 2023-06-21 source/analysis/baseline/Baseline.R
#25 recode type_change

@@ -24,18 +24,19 @@ Main <- function(){
   
   # Estimate my baseline interaction equation
   mod_basic <- fixest::feols(
-    c_all ~ i(treat_type, D, ref=2) + D + switches | city_id + period, 
+    c_all ~ i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + D + switches | city_id + period, 
     data = dat)
   
   mod_conflict <- fixest::feols(
-    c_all ~ i(treat_type, D, ref=2) + D + switches + conflict | city_id + period, 
+    c_all ~ i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + D + switches + conflict | city_id + period, 
     data = dat)
   
   # Produce regression table and export to LaTeX
   etableDefaults()
   tex_output <- etable(mod_basic, mod_conflict, tex=TRUE,
                        title="Test Title",
-                       label = sprintf("tab:baseline_%iy", t))
+                       label = sprintf("tab:baseline_%iy", t),
+                       postprocess.tex = PostProcessBaseline)
   
   filename <- sprintf("paper/output/regressions/baseline_%iy.tex", t)
   write(tex_output, file=filename)
@@ -53,7 +54,7 @@ PrepareData <- function(build, max_switches,
   # Identify across-period switches
   with_e_dummies <- AddEAnother(selected)
   
-  # Add TREAT x POST dummies
+  # Add TREAT x POST dummies D
   with_D <- AddTreatXPost(with_e_dummies)
   
   # Binarise construction outcomes, if specified
@@ -67,11 +68,21 @@ PrepareData <- function(build, max_switches,
   
   # Rearrange columns (for readability)
   clean <- with_types |> select(
-    city_id, treat_time, period, terr_id, switches, e_another, D, treat_type,
+    city_id, treat_time, period, terr_id, switches, e_another, D, 
+    rule_conquest, rule_succession, rule_other,
     conflict, c_all, c_state, c_private, c_public
   )
   
   return(clean)
+}
+
+
+PostProcessBaseline <- function(tex_output){
+  # Post-processing the TeX string to make it neater
+  tex_post <- tex_output |> 
+    str_replace("Conquest \\$=\\$ 1", "Conquest") |> 
+    str_replace("Other \\$=\\$ 1", "Other")
+  return(tex_post)
 }
 
 

### 2023-06-21 source/analysis/baseline/ReplicateSW22.R

@@ -72,7 +72,7 @@ PrepareData <- function(build, max_switches, years_pre, years_post,
   # Rearrange columns (for readability)
   clean <- with_leads_lags |> select(
       city_id, treat_time, period, terr_id, switches, e_another, treat, time_to_treat, 
-      type_change, conflict, c_all, c_state, c_private, c_public
+      conflict, c_all, c_state, c_private, c_public
       )
 
   return(clean)

### 2023-06-21 source/build/Aggregate.R

@@ -41,7 +41,9 @@ AggregateSwitches <- function(switches, t){
     summarise(
       terr_id = first(terr_id), # set to owner at the start of the period
       switches = sum(switch),
-      type_change = first(type_change), # we want to know about *this* owner
+      rule_conquest = first(rule_conquest), # we want to know about *this* owner
+      rule_succession = first(rule_succession),
+      rule_other = first(rule_other),
       duration = n() # some periods may be cut off in the data
     )
   

### 2023-06-21 source/build/Switches.R

@@ -22,7 +22,7 @@ Main <- function(){
   # This means: 
   # - `terr_id` is that of the initial state in each year
   # - `switch` codes whether a switch took place during this year
-  # - `type_change` is how the state in `terr_id` came into power (as before)
+  # - `rule_{type}` codes how the state in `terr_id` came into power (as before)
 
   cities_with_switch_dummies <- cities |> 
     arrange(city_id, year) |> 
@@ -39,11 +39,10 @@ Main <- function(){
   OTHER_CATS <- 0:13 |> base::setdiff(CONQUEST_CATS) |> base::setdiff(SUCCESSION_CATS)
   
   cities_with_switch_types <- cities_with_switch_dummies |> 
-    mutate(type_change = case_when(
-      type_change %in% CONQUEST_CATS ~ 3,
-      type_change %in% SUCCESSION_CATS ~ 2,
-      type_change %in% OTHER_CATS ~ 1
-    )
+    mutate(
+      rule_conquest = if_else(type_change %in% CONQUEST_CATS, 1, 0),
+      rule_succession = if_else(type_change %in% SUCCESSION_CATS, 1, 0),
+      rule_other = if_else(type_change %in% OTHER_CATS, 1, 0)
       )
   
   write_csv(cities_with_switch_types, "drive/derived/cities_switches.csv")

### 2023-06-21 source/utils/DataPrepSuite.R

@@ -74,12 +74,11 @@ ProcessTypeChange <- function(with_dummies){
 
   treat_types <- with_dummies |>
     filter(period == treat_time) |>
-    select(city_id, treat_time, type_change) |> 
-    rename(treat_type = type_change)
+    select(city_id, treat_time, rule_conquest, rule_succession, rule_other)
 
   with_types <- with_dummies |>
-    left_join(treat_types, by=c("city_id", "treat_time")) |> 
-    select(-type_change)
+    select(-rule_conquest, -rule_succession, -rule_other) |> 
+    left_join(treat_types, by=c("city_id", "treat_time"))
   
   # I may need to replace NAs in `treat_type` with 0, or anything really.
   # It doesn't matter because if `treat_type` is NA then `treat` = 0 and
@@ -90,7 +89,7 @@ ProcessTypeChange <- function(with_dummies){
   # that NA is a separate thing you can use as a reference category.
   
   with_types <- with_types |> 
-    replace_na(list(treat_type = 1))
+    replace_na(list(rule_conquest = 0, rule_succession = 0, rule_other = 0))
   
   return(with_types)
 }

### 2023-06-21 source/utils/etableDefaults.R

@@ -9,7 +9,8 @@ etableDefaults <- function(){
   
   # set default dict
   setFixest_dict(c(city_id = "City", period = "Period",
-                   c_all = "Construction", D = "Switching to another state",
-                   treat_type = "Selection", switches = "Switching indicator"),
-                 conflict = "Conflict")
-  }
+                   c_all = "Construction", D = "Switch to another state",
+                   rule_conquest = "Conquest", rule_succession = "Succession",
+                   rule_other = "Other", switches = "Switching indicator",
+                   conflict = "Conflict"))
+}

### 2023-06-21 source/analysis/baseline/Baseline.R
#25 and different outcomes and improve baseline table

@@ -23,20 +23,44 @@ Main <- function(){
                      binarise_switches = T)
   
   # Estimate my baseline interaction equation
-  mod_basic <- fixest::feols(
-    c_all ~ i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + D + switches | city_id + period, 
+  mod_all_basic <- fixest::feols(
+    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
     data = dat)
   
-  mod_conflict <- fixest::feols(
-    c_all ~ i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + D + switches + conflict | city_id + period, 
+  mod_state_basic <- fixest::feols(
+    c_state ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
     data = dat)
+  
+  mod_priv_basic <- fixest::feols(
+    c_private ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat)
+  
+  mod_pub_basic <- fixest::feols(
+    c_public ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat)
+  
+  # mod_all_conflict <- fixest::feols(
+  #   c_all ~ i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + D + switches + conflict | city_id + period, 
+  #   data = dat)
+  
   
   # Produce regression table and export to LaTeX
   etableDefaults()
-  tex_output <- etable(mod_basic, mod_conflict, tex=TRUE,
-                       title="Test Title",
+  
+  note <- paste("Table presents results of estimation equation
+  \\eqref{eq:baseline}.", PeriodInsert(1), "Observations are at the city-period 
+  level. The dependent variables are indicators that take the value 1 if 
+  construction activity of the respective type was recorded. Standard errors are 
+  clustered at the city level. The switch type \"Succession\" is omitted as the 
+  reference category.", sep = " ") |> str_replace_all("\n ", "")
+  
+  tex_output <- etable(mod_all_basic, mod_state_basic,
+                       mod_priv_basic, mod_pub_basic,
+                       tex=TRUE,
+                       title="Differences in trends after switching, by switch type",
                        label = sprintf("tab:baseline_%iy", t),
[…]

### 2023-06-21 source/utils/etableDefaults.R

@@ -2,15 +2,25 @@ etableDefaults <- function(){
   
   # uniform defaults (across all tables)
   setFixest_etable(
-    se.below = FALSE,
+    se.below = TRUE,
     digits = 2,
     digits.stats = 4,
   )
   
   # set default dict
   setFixest_dict(c(city_id = "City", period = "Period",
-                   c_all = "Construction", D = "Switch to another state",
+                   c_all = "All construction", 
+                   c_state = "State construction",
+                   c_private = "Private construction",
+                   c_public = "Public goods",
+                   D = "Switch to another state",
                    rule_conquest = "Conquest", rule_succession = "Succession",
                    rule_other = "Other", switches = "Switching indicator",
                    conflict = "Conflict"))
+}
+
+PeriodInsert <- function(t){
+  period_insert <- sprintf("Yearly data was aggregated into periods of %i years.", t)
+  if (t == 1) { period_insert <- "Time periods have a length of 1 year." }
+  return(period_insert)
 }

### 2023-06-21 source/analysis/baseline/Baseline.R
#25 fix and improve baseline table

@@ -47,12 +47,13 @@ Main <- function(){
   # Produce regression table and export to LaTeX
   etableDefaults()
   
-  note <- paste("Table presents results of estimation equation
-  \\eqref{eq:baseline}.", PeriodInsert(1), "Observations are at the city-period 
+  note <- paste("Note: Table presents results of estimation equation
+  \\eqref{eq:baseline}. The switch type \"Succession\" is omitted as the 
+  reference category.", PeriodInsert(t), "Observations are at the city-period 
   level. The dependent variables are indicators that take the value 1 if 
   construction activity of the respective type was recorded. Standard errors are 
-  clustered at the city level. The switch type \"Succession\" is omitted as the 
-  reference category.", sep = " ") |> str_replace_all("\n ", "")
+  clustered at the city level.", SignifInsert(), sep = " ") |> 
+    str_replace_all("\n ", "")
   
   tex_output <- etable(mod_all_basic, mod_state_basic,
                        mod_priv_basic, mod_pub_basic,
@@ -103,9 +104,13 @@ PrepareData <- function(build, max_switches,
 
 PostProcessBaseline <- function(tex_output){
   # Post-processing the TeX string to make it neater
+  # If the number of columns changes, the last two lines may show up again
+  # because they specifically contain \multicolumn{5}
   tex_post <- tex_output |> 
     str_replace("Conquest \\$=\\$ 1", "Conquest") |> 
-    str_replace("Other \\$=\\$ 1", "Other")
+    str_replace("Other \\$=\\$ 1", "Other") |> 
+    str_replace(fixed("\\multicolumn{5}{l}{\\emph{Clustered (City) standard-errors in parentheses}}\\\\"), "") |> 
+    str_replace(fixed("\\multicolumn{5}{l}{\\emph{Signif. Codes: ***: 0.01, **: 0.05, *: 0.1}}\\\\"), "")
   return(tex_post)
 }
 

### 2023-06-21 source/utils/etableDefaults.R

@@ -4,7 +4,7 @@ etableDefaults <- function(){
   setFixest_etable(
     se.below = TRUE,
     digits = 2,
-    digits.stats = 4,
+    digits.stats = 4
   )
   
   # set default dict
@@ -20,7 +20,15 @@ etableDefaults <- function(){
 }
 
 PeriodInsert <- function(t){
+  # Utility to insert in notes below regression tables
   period_insert <- sprintf("Yearly data was aggregated into periods of %i years.", t)
   if (t == 1) { period_insert <- "Time periods have a length of 1 year." }
   return(period_insert)
+}
+
+SignifInsert <- function(){
+  # Utility to insert in notes below regression tables
+  insert <- "*, **, and *** denote significance on the 10 percent, 5 percent, and 1 percent 
+  level, respectively."
+  return(insert)
 }

### 2023-06-21 source/analysis/baseline/Baseline.R
#25 try abbreviating outcome names

@@ -47,6 +47,16 @@ Main <- function(){
   # Produce regression table and export to LaTeX
   etableDefaults()
   
+  setFixest_dict(c(city_id = "City", period = "Period",
+                   c_all = "All construction", 
+                   c_state = "State",
+                   c_private = "Private",
+                   c_public = "Public goods",
+                   D = "Switch to another state",
+                   rule_conquest = "Conquest", rule_succession = "Succession",
+                   rule_other = "Other", switches = "Switching indicator",
+                   conflict = "Conflict"))
+  
   note <- paste("Note: Table presents results of estimation equation
   \\eqref{eq:baseline}. The switch type \"Succession\" is omitted as the 
   reference category.", PeriodInsert(t), "Observations are at the city-period 

### 2023-06-21 source/utils/etableDefaults.R

@@ -10,8 +10,8 @@ etableDefaults <- function(){
   # set default dict
   setFixest_dict(c(city_id = "City", period = "Period",
                    c_all = "All construction", 
-                   c_state = "State construction",
-                   c_private = "Private construction",
+                   c_state = "State",
+                   c_private = "Private",
                    c_public = "Public goods",
                    D = "Switch to another state",
                    rule_conquest = "Conquest", rule_succession = "Succession",

### 2023-06-21 paper/paper.tex
#25 get equation 2 right
#25 push overleaf changes

@@
 
-Since the outcome is binary, equation \ref{eq:sw22} represents a linear probability model: we model $construction_{it}$ as a random variable that follows a Bernoulli distribution with a probability parameter $p$ that is conditional on the right-hand side variables.
+Since the outcome is binary, equation \eqref{eq:sw22} represents a linear probability model: we model $construction_{it}$ as a random variable that follows a Bernoulli distribution with a probability parameter $p$ that is conditional on the right-hand side variables.
 This way we can interpret the values of coefficients as differences in the probability of recorded construction activity in a given period.
@@
 I switch to a simple difference-in-differences specification for clarity. I extend the previous regression by interacting with the type of the ownership change:
-
 \begin{equation}
 \label{eq:baseline}
-    construction_{it} = \alpha_i + \delta_t + \mathbf{1}(S_{it} > 0)\beta +
-    \mathbf{1}(t \ge e^{NewState}_i)\pi +
-    \mathbf{1}(t \ge e^{NewState}_i) T_i \theta + \varepsilon_{it}
+\begin{split}
+    construction_{it} &= \alpha_i + \delta_t + \mathbf{1}(S_{it} > 0)\beta +
+    \mathbf{1}(t \ge e^{NewState}_i)\pi \\
+    &+ \mathbf{1}(t \ge e^{NewState}_i) \times
+    \mathbf{1}(Switch Type = conquest) \times \theta_{conquest} \\
+    &+ \mathbf{1}(t \ge e^{NewState}_i) \times
+    \mathbf{1}(Switch Type = other) \times \theta_{other} + \varepsilon_{it}
+\end{split}
 \end{equation}
+
+with the only addition being $T_i$, which denotes the type of switch that brought the new state into power.
+I distinguish three different categories of switches: conquests, successions, and other.
+In the above regression, successions are the omitted category.
+The coefficient of interest is the
 
@@
 
-\subsection*{Tables}
-
-\input{paper/output/regressions/SW22_replication_50y}
-
-\input{paper/output/regressions/baseline_50y}
-
-\clearpage
 \subsection*{Figures}
 
-\begin{figure}
+\begin{figure}[ht]
     \centering
@@
 
+\clearpage
+\subsection*{Tables}
+
+% \input{paper/output/regressions/SW22_replication_50y}
+
+\input{paper/output/regressions/baseline_50y}
+
 \end{document}

### 2023-06-21 source/analysis/baseline/Baseline.R
#25 fix baseline note

@@ -47,20 +47,10 @@ Main <- function(){
   # Produce regression table and export to LaTeX
   etableDefaults()
   
-  setFixest_dict(c(city_id = "City", period = "Period",
-                   c_all = "All construction", 
-                   c_state = "State",
-                   c_private = "Private",
-                   c_public = "Public goods",
-                   D = "Switch to another state",
-                   rule_conquest = "Conquest", rule_succession = "Succession",
-                   rule_other = "Other", switches = "Switching indicator",
-                   conflict = "Conflict"))
-  
   note <- paste("Note: Table presents results of estimation equation
   \\eqref{eq:baseline}. The switch type \"Succession\" is omitted as the 
   reference category.", PeriodInsert(t), "Observations are at the city-period 
-  level. The dependent variables are indicators that take the value 1 if 
+  level. Dependent variables are indicators that take the value 1 if 
   construction activity of the respective type was recorded. Standard errors are 
   clustered at the city level.", SignifInsert(), sep = " ") |> 
     str_replace_all("\n ", "")

### 2023-06-21 paper/paper.tex

@@
 They find that switching states affects population negatively in the short run but positively in the long run.
+Further, they show that this long-run benefit is driven by improved state quality after switching.
 Similar to the evolution metaphor, they interpret this as "creative destruction": the short run cost of destruction is compensated by the long-run benefit of becoming part of a higher-quality state.
 However, they do not show explicitly how these improvements in state quality are driven by selection through conflict.
-They use a dataset with a bigger geographical scope (all of Europe) but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is only available in intervals of 50 or 100 years, and a series of political maps.
+They use a dataset with a bigger geographical scope (all of Europe) but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is only available in intervals of 50 or 100 years.
 
@@
 
-\cite{schoenholzer2022} find a short-run cost of switching (presumably due to conflict) and a long-run benefit.
-They get the short-term result by looking at cities that only switched states briefly before switching back, and the long-run result by comparing that to cities that permanently stayed with their new territory.
-I operate in the same framework and I borrow their notation: the variable $S_{it}$ is the number of times a city switched states in period $t$.
-By contrast, the variable $e^{NewState}_i$ denotes the first period in which city $i$ belongs to a different state than the previous.
-Intuitively, in periods with $S_{it} > 0$ a city suffers short-run effects of contemporary switching, while in periods following $e^{NewState}_i$ it will be under new rule and therefore experience long-run effects of switching.
+First I attempt to replicate one of the key results of \cite{schoenholzer2022}: that switching states had a positive effect on the growth of cities in the long run.
+This is an important piece of evidence in support of my hypothesis: states that have successfully taken territory from another are "higher quality", i.e.
+better for growth.
+\cite{schoenholzer2022} arrive at this result using an event study with staggered adoption.\footnote
+{
+    Equation (3), \citealp{schoenholzer2022}, p. 12.
+}
+The outcome is city population, recorded in intervals of 50 to 100 years.
+The treatment is switching to another state from one observation to the next.
+Importantly, a city may change hands multiple times between observations but belong to the same state in the end --- such an event does \textit{not} count as treatment.
+As the authors show throughout their paper, the short-run effect of switching (either permanently or intermittently) is distinct from the long-run effect of permanently switching to a different state.
 
-First I replicate the long-term finding of \cite{schoenholzer2022} using the following equation:
+I will borrow extensively from this framework. The first regression I run is almost identical to the one described above:
 
@@
 
-where the outcome $construction_{it}$ is an indicator that equals 1 if any construction activity was recorded in city $i$ and period $t$, the variable $S_{it}$ is the number of times a city switched states in period $t$, the variable $e^{NewState}_i$ denotes the first period in which city $i$ belongs to a different state than the previous, and $\alpha_i$ and $\delta_t$ denote city and period fixed effects.
+where the outcome $construction_{it}$ is an indicator that equals 1 if any construction activity was recorded in city $i$ and period $t$, the variable $S_{it}$ is the number of times a city switched states in period $t$, the variable $e^{NewState}_i$ denotes the first period in which city $i$ belongs to a different state than the previous, and $\alpha_i$ and $\delta_t$ are city and period fixed effects.
 The coefficients of interest are the dynamic effects $\pi_\tau$ of switching to a new state permanently.
 
-Since the outcome is binary, equation \eqref{eq:sw22} represents a linear probability model: we model $construction_{it}$ as a random variable that follows a Bernoulli distribution with a probability parameter $p$ that is conditional on the right-hand side variables.
+This is a standard TWFE regression with staggered adoption and dynamic effects.
+Treatment events are \textit{lasting} switches to another state, i.e.
+a switching event only counts as treatment if the city belongs to a different state at the beginning of the next period.\footnote
+{
+    If a city undergoes multiple lasting switches, I create duplicate entries for each different $e^{NewState}_i$.
+I exclude cities that switch states (permanently or not) more than twice over the sample period.
+I report results obtained with different exclusion thresholds in Appendix BLANK.
+}
+This is because I want the coefficients $\pi_\tau$ to capture the effects of the new state's governance, not just of the political and economic turmoil that may be associated with a transition of power.
+I separately control for other switches by including the term $\mathbf{1}(S_{it} > 0)$ on the right-hand side so that the coefficient $\beta$ absorbs the effect of switching states in the same period.
+
+Since the outcome is binary in this case, we can interpret Equation \eqref{eq:sw22} as a linear probability model: we model $construction_{it}$ as a random variable that follows a Bernoulli distribution with a probability parameter $p$ that is conditional on the right-hand side variables.
 This way we can interpret the values of coefficients as differences in the probability of recorded construction activity in a given period.
-
 The coefficient $\pi_\tau$ for given $\tau$ is the difference in the probability of construction between two cities $i$ and $j$ with identical time-invariant characteristics, in the same period, who undergo the same amount of switching turmoil in this period.
@@
 
-Remember, we expect this difference to stem from the fact that switching states generally means switching to a higher-quality state (due to selection).
-As such, even though it would be possible in principle to include state fixed effects, it is not a good idea here: state fixed effects would absorb the variation in state quality whose effects we want to capture in $\pi_\tau$.
+Based on the theory of conflict-driven evolution we would expect any such difference to stem from the fact that switching states generally means switching to a higher-quality state (due to selection).
+As such, even though it would be possible in principle to include state fixed effects, it is not a good idea here: state fixed effects would absorb the effect of differences in state quality which we want to capture in $\pi_\tau$.
 
@@
 
-The it's not as convincing: \ref{tab:SW22_replication_50y}
-
 % ######################################################
@@
 \section{Discussion}
+
+differences within HRE may not have been large enough
+
+successions may also select for state power
+
+different switch sizes for conquests vs extinctions; maybe larger switch sizes caused neglect (-) or integration (+)
 

### 2023-06-21 source/analysis/sunab/SunabBasic.R
#25 add sketch of sunab event study

@@ -1,0 +1,70 @@
+#!/usr/bin/env Rscript --vanilla
+
+library(tidyverse) |> suppressPackageStartupMessages()
+library(fixest) |> suppressPackageStartupMessages()
+
+setwd("~/GitHub/BA")
+
+source("source/utils/HandleCommandArgs.R")
+source("source/utils/DataPrepSuite.R")
+source("source/utils/etableDefaults.R")
+
+Main <- function(){
+  
+  # Read data
+  t <- HandleCommandArgs(default_length = 50)
+  read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
+  build <- read_csv(read_path, show_col_types = F)
+  
+  # Select sample and process data (exact steps documented below)
+  dat <- PrepareData(build,
+                     max_switches = 2,
+                     binarise_construction = T,
+                     binarise_switches = T)
+  
+  mod <- fixest::feols(
+    c_all ~ sunab(treat_time, period) + switches | city_id + period, 
+    data = dat
+  )
+  
+  summary(mod, agg="att")
+
+  iplot(mod)  
+}
+
+
+PrepareData <- function(build, max_switches,
+                        binarise_construction, binarise_switches){
+  # Copied from baseline/Baseline.R
+  
+  
+  # Filter out cities with too many lifetime switches
+  selected <- FilterBySwitches(build, max_switches)
+  
+  # Identify across-period switches
+  with_e_dummies <- AddEAnother(selected)
+  
+  # Add TREAT x POST dummies D
+  with_D <- AddTreatXPost(with_e_dummies)
+  
+  # Binarise construction outcomes, if specified
+  if (binarise_construction) { with_D <- with_D |> BinariseOutcomes() }
+  
+  # Binarise no. of switches (control), if specified
+  if (binarise_switches) { with_D <- with_D |> BinariseSwitches() }
+  
+  # Process type change into treatment type
+  with_types <- ProcessTypeChange(with_D)
+  
+  # Rearrange columns (for readability)
+  clean <- with_types |> select(
+    city_id, treat_time, period, terr_id, switches, e_another, D, 
+    rule_conquest, rule_succession, rule_other,
+    conflict, c_all, c_state, c_private, c_public
+  )
+  
+  return(clean)
+}
+
+
+Main()

### 2023-06-21 paper/paper.tex

@@
 
-\subsection{Long-run effect of switching}
+\subsection{Dynamic effects of switching to a new state}
 
@@
 
-where the outcome $construction_{it}$ is an indicator that equals 1 if any construction activity was recorded in city $i$ and period $t$, the variable $S_{it}$ is the number of times a city switched states in period $t$, the variable $e^{NewState}_i$ denotes the first period in which city $i$ belongs to a different state than the previous, and $\alpha_i$ and $\delta_t$ are city and period fixed effects.
+where the outcome $construction_{it}$ is an indicator that equals 1 if any construction activity was observed in city $i$ and period $t$, the variable $S_{it}$ is the number of times a city switched states in period $t$, the variable $e^{NewState}_i$ denotes the first period in which city $i$ belongs to a different state than the previous, and $\alpha_i$ and $\delta_t$ are city and period fixed effects.
 The coefficients of interest are the dynamic effects $\pi_\tau$ of switching to a new state permanently.
@@
 Since the outcome is binary in this case, we can interpret Equation \eqref{eq:sw22} as a linear probability model: we model $construction_{it}$ as a random variable that follows a Bernoulli distribution with a probability parameter $p$ that is conditional on the right-hand side variables.
-This way we can interpret the values of coefficients as differences in the probability of recorded construction activity in a given period.
+This way we can interpret the values of coefficients as differences in the probability of construction in a given period.
 The coefficient $\pi_\tau$ for given $\tau$ is the difference in the probability of construction between two cities $i$ and $j$ with identical time-invariant characteristics, in the same period, who undergo the same amount of switching turmoil in this period.
@@
 
-\subsection{Effect of state power}
+\subsection{Differential effects of switching}
 
 Next I tried to find the effect of state power - the effect of the difference in selection for state power between conquests and successions.
+
+Effect of state power differentials on the effect of switching
 
@@
 
+limit post-window to 200 years
+
 % ######################################################

### 2023-06-21 paper/output/regressions/window_100y.tex
#25 implement baseline with 200y window

\begin{table}[htbp]
   \caption{\label{tab:window_100y} Differences in trends after switching, by switch type}
   \centering
   \begin{tabular}{lcccc}
      \tabularnewline \midrule \midrule
      Dependent Variables:                       & All construction & State       & Private & Public goods\\  
      Model:                                     & (1)              & (2)         & (3)     & (4)\\  
      \midrule
      \emph{Variables}\\
      Switch to another state                    & 0.07$^{*}$       & 0.004       & 0.02    & -0.05$^{*}$\\   
                                                 & (0.04)           & (0.03)      & (0.03)  & (0.03)\\   
      Switch to another state $\times$ Conquest  & -0.03            & -0.08$^{*}$ & -0.03   & -0.005\\   
                                                 & (0.04)           & (0.05)      & (0.04)  & (0.03)\\   
      Switch to another state $\times$ Other     & -0.03            & -0.01       & -0.04   & 0.03\\   
                                                 & (0.04)           & (0.04)      & (0.03)  & (0.03)\\   
      Switching indicator                        & 0.01             & 0.05$^{*}$  & 0.006   & -0.02\\   
                                                 & (0.03)           & (0.03)      & (0.02)  & (0.03)\\   
      \midrule
      \emph{Fixed-effects}\\
      City                                       & Yes              & Yes         & Yes     & Yes\\  
      Period                                     & Yes              & Yes         & Yes     & Yes\\  
      \midrule
      \emph{Fit statistics}\\
      Observations                               & 2,955            & 2,955       & 2,955   & 2,955\\  
      R$^2$                                      & 0.4287           & 0.4530      & 0.4802  & 0.4493\\  
      Within R$^2$                               & 0.0020           & 0.0046      & 0.0007  & 0.0017\\  
      \midrule \midrule

   \end{tabular}

   \par \raggedright 
   Note: Table presents results of estimation equation \eqref{eq:baseline}. The switch type "Succession" is omitted as the  reference category. Yearly data was aggregated into periods of 100 years. Observations are at the city-period  level. Dependent variables are indicators that take the value 1 if  construction activity of the respective type was recorded. Standard errors are  clustered at the city level. *, **, and *** denote significance on the 10 percent, 5 percent, and 1 percent  level, respectively.
\end{table}

### 2023-06-21 paper/output/regressions/window_10y.tex
#25 implement baseline with 200y window

--- 2023-06-21 paper/output/regressions/window_100y.tex
@@
 \begin{table}[htbp]
-   \caption{\label{tab:window_100y} Differences in trends after switching, by switch type}
+   \caption{\label{tab:window_10y} Differences in trends after switching, by switch type}
    \centering
@@
       \emph{Variables}\\
-      Switch to another state                    & 0.07$^{*}$       & 0.004       & 0.02    & -0.05$^{*}$\\   
-                                                 & (0.04)           & (0.03)      & (0.03)  & (0.03)\\   
-      Switch to another state $\times$ Conquest  & -0.03            & -0.08$^{*}$ & -0.03   & -0.005\\   
-                                                 & (0.04)           & (0.05)      & (0.04)  & (0.03)\\   
-      Switch to another state $\times$ Other     & -0.03            & -0.01       & -0.04   & 0.03\\   
-                                                 & (0.04)           & (0.04)      & (0.03)  & (0.03)\\   
-      Switching indicator                        & 0.01             & 0.05$^{*}$  & 0.006   & -0.02\\   
-                                                 & (0.03)           & (0.03)      & (0.02)  & (0.03)\\   
+      Switch to another state                    & 0.01$^{*}$       & 0.003   & -0.004                & -0.008$^{**}$\\   
+                                                 & (0.009)          & (0.004) & (0.003)               & (0.003)\\   
+      Switch to another state $\times$ Conquest  & -0.008           & -0.01   & 0.003                 & 0.004\\   
+                                                 & (0.01)           & (0.007) & (0.005)               & (0.005)\\   
+      Switch to another state $\times$ Other     & -0.01            & -0.005  & 0.002                 & 0.01$^{**}$\\   
+                                                 & (0.01)           & (0.005) & (0.004)               & (0.005)\\   
+      Switching indicator                        & -0.03$^{***}$    & -0.004  & -0.005                & -0.010$^{**}$\\   
+                                                 & (0.01)           & (0.005) & (0.004)               & (0.004)\\   
       \midrule
@@
       \emph{Fit statistics}\\
-      Observations                               & 2,955            & 2,955       & 2,955   & 2,955\\  
-      R$^2$                                      & 0.4287           & 0.4530      & 0.4802  & 0.4493\\  
-      Within R$^2$                               & 0.0020           & 0.0046      & 0.0007  & 0.0017\\  
+      Observations                               & 35,168           & 35,168  & 35,168                & 35,168\\  
+      R$^2$                                      & 0.1770           & 0.0835  & 0.1278                & 0.0890\\  
+      Within R$^2$                               & 0.0004           & 0.0001  & $8.13\times 10^{-5}$  & 0.0004\\  
       \midrule \midrule
@@
 The switch type "Succession" is omitted as the  reference category.
-Yearly data was aggregated into periods of 100 years.
+Yearly data was aggregated into periods of 10 years.
 Observations are at the city-period  level.

### 2023-06-21 paper/output/regressions/window_50y.tex
#25 implement baseline with 200y window

--- 2023-06-21 paper/output/regressions/window_100y.tex
@@
 \begin{table}[htbp]
-   \caption{\label{tab:window_100y} Differences in trends after switching, by switch type}
+   \caption{\label{tab:window_50y} Differences in trends after switching, by switch type}
    \centering
@@
       \emph{Variables}\\
-      Switch to another state                    & 0.07$^{*}$       & 0.004       & 0.02    & -0.05$^{*}$\\   
-                                                 & (0.04)           & (0.03)      & (0.03)  & (0.03)\\   
-      Switch to another state $\times$ Conquest  & -0.03            & -0.08$^{*}$ & -0.03   & -0.005\\   
-                                                 & (0.04)           & (0.05)      & (0.04)  & (0.03)\\   
-      Switch to another state $\times$ Other     & -0.03            & -0.01       & -0.04   & 0.03\\   
-                                                 & (0.04)           & (0.04)      & (0.03)  & (0.03)\\   
-      Switching indicator                        & 0.01             & 0.05$^{*}$  & 0.006   & -0.02\\   
-                                                 & (0.03)           & (0.03)      & (0.02)  & (0.03)\\   
+      Switch to another state                    & 0.02             & 0.02          & 0.002   & -0.04$^{***}$\\   
+                                                 & (0.02)           & (0.02)        & (0.01)  & (0.01)\\   
+      Switch to another state $\times$ Conquest  & 0.04             & -0.07$^{***}$ & -0.005  & 0.02\\   
+                                                 & (0.04)           & (0.03)        & (0.02)  & (0.02)\\   
+      Switch to another state $\times$ Other     & 0.008            & -0.02         & 0.006   & 0.04$^{**}$\\   
+                                                 & (0.03)           & (0.02)        & (0.01)  & (0.02)\\   
+      Switching indicator                        & -0.01            & 0.007         & -0.005  & 0.0006\\   
+                                                 & (0.02)           & (0.02)        & (0.01)  & (0.01)\\   
       \midrule
@@
       \emph{Fit statistics}\\
-      Observations                               & 2,955            & 2,955       & 2,955   & 2,955\\  
-      R$^2$                                      & 0.4287           & 0.4530      & 0.4802  & 0.4493\\  
-      Within R$^2$                               & 0.0020           & 0.0046      & 0.0007  & 0.0017\\  
+      Observations                               & 6,635            & 6,635         & 6,635   & 6,635\\  
+      R$^2$                                      & 0.3553           & 0.2798        & 0.3257  & 0.2917\\  
+      Within R$^2$                               & 0.0009           & 0.0012        & 0.0002  & 0.0011\\  
       \midrule \midrule
@@
 The switch type "Succession" is omitted as the  reference category.
-Yearly data was aggregated into periods of 100 years.
+Yearly data was aggregated into periods of 50 years.
 Observations are at the city-period  level.

### 2023-06-21 source/analysis/baseline/Window.R

@@ -1,0 +1,130 @@
+#!/usr/bin/env Rscript --vanilla
+
+library(tidyverse) |> suppressPackageStartupMessages()
+library(fixest) |> suppressPackageStartupMessages()
+
+setwd("~/GitHub/BA")
+
+source("source/utils/HandleCommandArgs.R")
+source("source/utils/DataPrepSuite.R")
+source("source/utils/etableDefaults.R")
+
+Main <- function(){
+  
+  # Read data
+  t <- HandleCommandArgs(default_length = 50)
+  read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
+  build <- read_csv(read_path, show_col_types = F)
+  
+  # Select sample and process data (exact steps documented below)
+  dat <- PrepareData(build,
+                     max_switches = 2,
+                     years_post = 200,
+                     binarise_construction = T,
+                     binarise_switches = T)
+  
+  # Estimate my baseline interaction equation
+  mod_all_basic <- fixest::feols(
+    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat)
+  
+  mod_state_basic <- fixest::feols(
+    c_state ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat)
+  
+  mod_priv_basic <- fixest::feols(
+    c_private ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat)
+  
+  mod_pub_basic <- fixest::feols(
+    c_public ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat)
+  
+  # mod_all_conflict <- fixest::feols(
+  #   c_all ~ i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + D + switches + conflict | city_id + period, 
+  #   data = dat)
+  
+  
+  # Produce regression table and export to LaTeX
+  etableDefaults()
+  
+  note <- paste("Note: Table presents results of estimation equation
+  \\eqref{eq:baseline}. The switch type \"Succession\" is omitted as the 
+  reference category.", PeriodInsert(t), "Observations are at the city-period 
+  level. Dependent variables are indicators that take the value 1 if 
+  construction activity of the respective type was recorded. Standard errors are 
[…]

### 2023-06-21 paper/paper.tex
update document style
#25 structure initial baseline notes

@@
-\documentclass{article}
+\documentclass[11pt, a4paper]{article}
 \usepackage[utf8]{inputenc}
-\usepackage[a4paper,left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
+\usepackage[left=2.5cm,right=3cm,top=2.5cm,bottom=2.5cm]{geometry}  
+% set 11pt and right=6cm for submission
 
@@
 Next I tried to find the effect of state power - the effect of the difference in selection for state power between conquests and successions.
+The framework of \cite{schoenholzer2022} only tells us BLANK, but it doesn't tell us BLANK.
+The reason may be that, having sourced their data on territorial histories from historical maps, they lack information about individual switching events.
+Because the Städtebuch data compiled by \cite{pt2} records exactly how each city was acquired by the territory holding it, I am able to look at differential trends after different types of ownership changes.
 
-Effect of state power differentials on the effect of switching
-
-I switch to a simple difference-in-differences specification for clarity. I extend the previous regression by interacting with the type of the ownership change:
+For this purpose I extend Equation \eqref{eq:sw22} by interacting the treatment indicator with the type of switch:
 \begin{equation}
@@
     &+ \mathbf{1}(t \ge e^{NewState}_i) \times
-    \mathbf{1}(Switch Type = conquest) \times \theta_{conquest} \\
+    \mathbf{1}(Switchtype = conquest) \times \theta_{conquest} \\
     &+ \mathbf{1}(t \ge e^{NewState}_i) \times
-    \mathbf{1}(Switch Type = other) \times \theta_{other} + \varepsilon_{it}
+    \mathbf{1}(Switchtype = other) \times \theta_{other} + \varepsilon_{it},
 \end{split}
@@
 
-with the only addition being $T_i$, which denotes the type of switch that brought the new state into power.
+For clarity I have converted the event study into a basic difference-in-differences specification; the lead and lag dummies have simply been replaced with $\mathbf{1}(t \ge e^{NewState}_i)$ and a single coefficient $\pi$, which is an average treatment effect.
+The key difference from Equation \eqref{eq:sw22} is the addition of interaction terms.
+
 I distinguish three different categories of switches: conquests, successions, and other.
 In the above regression, successions are the omitted category.
-The coefficient of interest is the
+In all cases this refers to how the state currently owning the city came into power.
+For example, $conquest_{it} = 1$ means that the state $\mathbf{J}(it)$ that holds city $i$ at time $t$ acquired city $i$ by conquest.
 
-a second-order interaction necessitates that the first-order interaction pattern and associated coefficient (whether zero or non-zero) is itself different depending on the level of a third predictor.
-(from stackexchange)
-
-Blah blah. Equation \ref{eq:baseline}
+The coefficient of interest in this regression is no longer $\pi$ but $\theta_{conquest}$. It is the effect of state power differentials on the effect of switching.
 

### 2023-06-21 paper/paper.tex
sketch discussion section

@@
 
-For this purpose I extend Equation \eqref{eq:sw22} by interacting the treatment indicator with the type of switch:
+For this purpose I extend Equation \eqref{eq:sw22} by interacting the treatment indicator with the type of switching event:
 \begin{equation}
@@
 
+How robust this null result is to other specifications: I limit post-window to 200 years; I try Sun and Abraham, I use population data.
+
 differences within HRE may not have been large enough
@@
 different switch sizes for conquests vs extinctions; maybe larger switch sizes caused neglect (-) or integration (+)
-
-limit post-window to 200 years
 

### 2023-06-21 paper/paper.tex
#25 add identification discussion

--- 2023-06-21 paper/paper.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage[left=2.5cm,right=3cm,top=2.5cm,bottom=2.5cm]{geometry}  
+\usepackage[left=2cm,right=6cm,top=2.5cm,bottom=2.5cm]{geometry}  
 % set 11pt and right=6cm for submission
@@
 
-\section{Background}
+\section{Historical background}
 
@@
 
-\section{Empirical strategy}
+\section{Regression equations}
 
@@
 This way we can interpret the values of coefficients as differences in the probability of construction in a given period.
-The coefficient $\pi_\tau$ for given $\tau$ is the difference in the probability of construction between two cities $i$ and $j$ with identical time-invariant characteristics, in the same period, who undergo the same amount of switching turmoil in this period.
+The coefficient $\pi_\tau$ for given $\tau$ is the difference in the probability of construction between two cities $i$ and $j$ with identical time-invariant characteristics, in the same period, who undergo the same kind of switching turmoil in this period.
 The only difference is that one city underwent a lasting switch to another state $\tau$ periods ago, while the other did not.
@@
 
+The basic assumption that needs to be satisfied for a causal interpretation of TWFE estimates is parallel trends.
+This means that, absent treatment, the outcome in cities that received treatment would have followed the same time trend as the outcome in never-treated cities.
+While it is impossible to empirically prove this assumption (because we cannot observe the counterfactual trend), it is common to check its validity by estimating pre-treatment effects: if the treatment and control group already showed different trends before being treated, there may be an unobserved variable correlated both with receiving treatment and with the outcome, biasing the estimate.
+For this reason I also include leads reaching 100 years into the past in my regression (as do \cite{schoenholzer2022}.)
+
 \subsection{Differential effects of switching}
 
-Next I tried to find the effect of state power - the effect of the difference in selection for state power between conquests and successions.
-The framework of \cite{schoenholzer2022} only tells us BLANK, but it doesn't tell us BLANK.
-The reason may be that, having sourced their data on territorial histories from historical maps, they lack information about individual switching events.
-Because the Städtebuch data compiled by \cite{pt2} records exactly how each city was acquired by the territory holding it, I am able to look at differential trends after different types of ownership changes.
+Next I attempt to identify the effect of state power --- the effect of the difference in selection for state power between conquests and successions.
+In the best case, if the identifying assumption is satisfied, the estimates from the previous regression tell us what the effect of switching states is - however, they cannot tell us why.
 
-For this purpose I extend Equation \eqref{eq:sw22} by interacting the treatment indicator with the type of switch:
+Because the \textit{Städtebuch} data compiled by \cite{pt2} records exactly how each city was acquired by the territory holding it, I am able to look at differential trends after different types of ownership changes.\footnote
+{
+Presumably \cite{schoenholzer2022}, having sourced their data on territorial histories from historical maps, lack detailed information about individual switching events.
+They instead use clever panel decomposition methods (pp.
+14-20) and data on state-level characteristics (pp.
+21-23) to explain what drives their estimates.
+}
+For this purpose I extend Equation \eqref{eq:sw22} by interacting the treatment indicator with the type of switching event:
 \begin{equation}
@@
 
-For clarity I have converted the event study into a basic difference-in-differences specification; the lead and lag dummies have simply been replaced with $\mathbf{1}(t \ge e^{NewState}_i)$ and a single coefficient $\pi$, which is an average treatment effect.
+For clarity I have turned the event study into a basic difference-in-differences specification; this way there is only a single coefficient of interest.
+The lead and lag dummies have simply been replaced with the post-treatment indicator $\mathbf{1}(t \ge e^{NewState}_i)$.
 The key difference from Equation \eqref{eq:sw22} is the addition of interaction terms.
@@
 In all cases this refers to how the state currently owning the city came into power.
-For example, $conquest_{it} = 1$ means that the state $\mathbf{J}(it)$ that holds city $i$ at time $t$ acquired city $i$ by conquest.
+For example, $conquest_{it} = 1$ means that the state $\mathbf{J}(i, t)$ that holds city $i$ at time $t$ acquired city $i$ by conquest.
 
-The coefficient of interest in this regression is no longer $\pi$ but $\theta_{conquest}$. It is the effect of state power differentials on the effect of switching.
+The coefficient of interest in this regression is no longer $\pi$ but $\theta_{conquest}$.
 
@@
 
+How robust this null result is to other specifications: I limit post-window to 200 years; I try Sun and Abraham, I use population data; I use different granularity.
+
 differences within HRE may not have been large enough

### 2023-06-21 paper/paper.tex
#25 finally explain baseline spec

@@
 
-For clarity I have turned the event study into a basic difference-in-differences specification; this way there is only a single coefficient of interest.
-The lead and lag dummies have simply been replaced with the post-treatment indicator $\mathbf{1}(t \ge e^{NewState}_i)$.
-The key difference from Equation \eqref{eq:sw22} is the addition of interaction terms.
+For clarity I move from the event study to a basic difference-in-differences specification; this way there is only a single coefficient of interest.
+I simply replace the lead and lag dummies with the post-treatment indicator $\mathbf{1}(t \ge e^{NewState}_i)$.
+The more important novelty are the interaction terms: I interact the type of switch with the post-treatment indicator to obtain estimates of how the trends in construction activity differ after different types of switches.
+I distinguish three categories: conquests, successions, and other.
+In all cases this refers to how the state currently owning the city came into power --- for example $conquest_{it} = 1$ means that the state that holds city $i$ at time $t$ acquired city $i$ by conquest.
+It will be useful to denote the state ruling city $i$ at time $t$ by $\mathbf{J}(i, t)$ in the following.
 
-I distinguish three different categories of switches: conquests, successions, and other.
-In the above regression, successions are the omitted category.
-In all cases this refers to how the state currently owning the city came into power.
-For example, $conquest_{it} = 1$ means that the state $\mathbf{J}(i, t)$ that holds city $i$ at time $t$ acquired city $i$ by conquest.
+A key economic assumption is needed, in addition to the econometric ones, for these interaction terms to tell us anything about state power.
+It is this: the state power of a city's owner $\mathbf{J}(i, t)$ is correlated with the manner in which the city was acquired by that owner.
+In particular, I assume that conquests are more strongly positively correlated with state power than successions.
+Support for this assumption comes from \cite{levine2021}'s model of conflict-driven evolution.
+BLANK.
+I will discuss its validity at length later on.
 
 The coefficient of interest in this regression is no longer $\pi$ but $\theta_{conquest}$.
+Note that there is no interaction term between the post-treatment indicator and $\mathbf{1}(Switchtype = succession)$ --- successions are the omitted category.
+$\theta_{conquest}$, then, captures the difference between the trend of a city acquired by conquest and the trend of a city acquired by succession.
 

### 2023-06-22 paper/paper.tex
#25 finish empirics section!

@@
     &+ \mathbf{1}(t \ge e^{NewState}_i) \times
-    \mathbf{1}(Switchtype = conquest) \times \theta_{conquest} \\
+    \mathbf{1}(Conquest_{it}) \times \theta_{conquest} \\
     &+ \mathbf{1}(t \ge e^{NewState}_i) \times
-    \mathbf{1}(Switchtype = other) \times \theta_{other} + \varepsilon_{it},
+    \mathbf{1}(Other_{it}) \times \theta_{other} + \varepsilon_{it},
 \end{split}
@@
 The more important novelty are the interaction terms: I interact the type of switch with the post-treatment indicator to obtain estimates of how the trends in construction activity differ after different types of switches.
-I distinguish three categories: conquests, successions, and other.
-In all cases this refers to how the state currently owning the city came into power --- for example $conquest_{it} = 1$ means that the state that holds city $i$ at time $t$ acquired city $i$ by conquest.
-It will be useful to denote the state ruling city $i$ at time $t$ by $\mathbf{J}(i, t)$ in the following.
+I distinguish three categories: $Conquest$, $Succession$, and $Other$.
+In all cases this refers to how the state currently owning the city came into power --- for example $Conquest_{it} = 1$ means that the state that holds city $i$ at time $t$ acquired city $i$ by conquest.
+It will be useful in the following to adopt one more bit of notation from \cite{schoenholzer2022} and let $\mathbf{J}(i, t)$ denote the state ruling city $i$ at time $t$.
 
-A key economic assumption is needed, in addition to the econometric ones, for these interaction terms to tell us anything about state power.
+A key economic assumption is needed, even before the econometric ones, for these interaction terms to tell us anything about state power.
 It is this: the state power of a city's owner $\mathbf{J}(i, t)$ is correlated with the manner in which the city was acquired by that owner.
@@
 
-The coefficient of interest in this regression is no longer $\pi$ but $\theta_{conquest}$.
-Note that there is no interaction term between the post-treatment indicator and $\mathbf{1}(Switchtype = succession)$ --- successions are the omitted category.
-$\theta_{conquest}$, then, captures the difference between the trend of a city acquired by conquest and the trend of a city acquired by succession.
+If this assumption holds true, the effect of state power should manifest itself in the difference between the trend of cities acquired by conquest and the trend of a cities acquired by succession.
+Note that successions are the omitted category --- there is no interaction term between the post-treatment indicator and $\mathbf{1}(Succession_{it})$ in Equation \eqref{eq:baseline} --- so that the coefficient $\theta_{conquest}$ captures exactly what we want: the difference in trends after conquests vis-a-vis successions.
+
+The econometric assumption necessary to allow causal interpretation of $\theta_{conquest}$ is again a version of the parallel trends assumption.
+In particular, it must hold that, had a conquered city i counterfactually been acquired by succession rather than conquest, it would have trended the same as the other cities acquired by succession.
+BLANK.
 
@@
 
-I can barely believe it myself, but I replicate their finding. Table 
+\subsection{More construction events after switching states}
 Figure \ref{fig:SW_replication} shows the dynamic effect of switching. Cities were on average
@@
 different switch sizes for conquests vs extinctions; maybe larger switch sizes caused neglect (-) or integration (+)
+
+new construction vs repairs (!!!)
 

### 2023-06-22 paper/paper.tex
#25 push overleaf changes
#28 begin results section

@@
 
-The econometric assumption necessary to allow causal interpretation of $\theta_{conquest}$ is again a version of the parallel trends assumption.
+The econometric assumption necessary to allow causal interpretation of $\theta_{conquest}$ is a different version of the parallel trends assumption.
 In particular, it must hold that, had a conquered city i counterfactually been acquired by succession rather than conquest, it would have trended the same as the other cities acquired by succession.
@@
 \subsection{More construction events after switching states}
-Figure \ref{fig:SW_replication} shows the dynamic effect of switching. Cities were on average
+Table \ref{tab:SW22_replication_50y} shows estimates for Equation \eqref{eq:sw22}. Figure \ref{fig:SW_replication} plots the coefficients to visualise how the effect evolves dynamically.
 
@@
 
-% \input{paper/output/regressions/SW22_replication_50y}
+\input{paper/output/regressions/SW22_replication_50y}
 

### 2023-06-22 paper/output/regressions/SW22_replication_100y.tex
#28 proper SW22 results table

--- 2023-06-21 paper/output/regressions/window_100y.tex
@@
 \begin{table}[htbp]
-   \caption{\label{tab:window_100y} Differences in trends after switching, by switch type}
+   \caption{\label{tab:SW22_replication_100y} Dynamic effects of switching}
    \centering
@@
       \emph{Variables}\\
-      Switch to another state                    & 0.07$^{*}$       & 0.004       & 0.02    & -0.05$^{*}$\\   
-                                                 & (0.04)           & (0.03)      & (0.03)  & (0.03)\\   
-      Switch to another state $\times$ Conquest  & -0.03            & -0.08$^{*}$ & -0.03   & -0.005\\   
-                                                 & (0.04)           & (0.05)      & (0.04)  & (0.03)\\   
-      Switch to another state $\times$ Other     & -0.03            & -0.01       & -0.04   & 0.03\\   
-                                                 & (0.04)           & (0.04)      & (0.03)  & (0.03)\\   
-      Switching indicator                        & 0.01             & 0.05$^{*}$  & 0.006   & -0.02\\   
+      Treat $\times$ Period $=$ 0    & 0.02             & -0.03      & 0.0006        & -0.04$^{**}$\\   
+                                     & (0.02)           & (0.03)     & (0.02)        & (0.02)\\   
+      Treat $\times$ Period $=$ 100  & 0.09$^{***}$     & 0.02       & 0.01          & -0.008\\   
+                                     & (0.03)           & (0.03)     & (0.03)        & (0.03)\\   
+      Treat $\times$ Period $=$ 200  & 0.08$^{***}$     & -0.02      & -0.10$^{***}$ & -0.06$^{*}$\\   
+                                     & (0.03)           & (0.04)     & (0.03)        & (0.04)\\   
+      Switching indicator            & 0.02             & 0.06$^{*}$ & 0.009         & -0.02\\   
                                      & (0.03)           & (0.03)     & (0.02)        & (0.03)\\   
@@
       Observations                   & 2,955            & 2,955      & 2,955         & 2,955\\  
-      R$^2$                                      & 0.4287           & 0.4530      & 0.4802  & 0.4493\\  
-      Within R$^2$                               & 0.0020           & 0.0046      & 0.0007  & 0.0017\\  
+      R$^2$                          & 0.4303           & 0.4530     & 0.4832        & 0.4499\\  
+      Within R$^2$                   & 0.0048           & 0.0047     & 0.0064        & 0.0027\\  
       \midrule \midrule
@@
    \par \raggedright 
-   Note: Table presents results of estimation equation \eqref{eq:baseline}.
-The switch type "Succession" is omitted as the  reference category.
+   Note: Table presents results of estimation equation \eqref{eq:sw22}.
 Yearly data was aggregated into periods of 100 years.

### 2023-06-22 source/analysis/baseline/ReplicateSW22.R

@@ -25,17 +25,39 @@ Main <- function(){
                      binarise_switches = T)
   
   # Estimate (3) from Schoenholzer and Weese (2022), p. 12
-  mod <- fixest::feols(
+  mod_all <- fixest::feols(
     c_all ~ i(time_to_treat, treat, ref = -t) + switches | city_id + period, 
     data = dat)
-
+  
+  mod_state <- fixest::feols(
+    c_state ~ i(time_to_treat, treat, ref = -t) + switches | city_id + period, 
+    data = dat)
+  
+  mod_priv <- fixest::feols(
+    c_private ~ i(time_to_treat, treat, ref = -t) + switches | city_id + period, 
+    data = dat)
+  
+  mod_pub <- fixest::feols(
+    c_public ~ i(time_to_treat, treat, ref = -t) + switches | city_id + period, 
+    data = dat)
   
   # Produce regression table and export to LaTeX
   etableDefaults()
-  tex_output <- etable(mod, 
+  
+  note <- paste("Note: Table presents results of estimation equation
+  \\eqref{eq:sw22}.", PeriodInsert(t), "Observations are at the city-period 
+  level. Dependent variables are indicators that take the value 1 if 
+  construction activity of the respective type was recorded. Standard errors are 
+  clustered at the city level.", SignifInsert(), sep = " ") |> 
+    str_replace_all("\n ", "")
+  
+  tex_output <- etable(mod_all, mod_state, 
+                       mod_priv, mod_pub,
                        tex=TRUE,
                        title="Dynamic effects of switching",
-                       label = sprintf("tab:SW22_replication_%iy", t))
+                       label = sprintf("tab:SW22_replication_%iy", t),
+                       postprocess.tex = PostProcessSW22,
+                       notes = note)
   
   filename <- sprintf("paper/output/regressions/SW22_replication_%iy.tex", t)
   write(tex_output, file=filename)
@@ -44,7 +66,7 @@ Main <- function(){
   # Produce replication of Fig. 5 and save as PNG
   filename <- sprintf("paper/output/regressions/SW22_replication_%iy.png", t)
   png(filename=filename, width = 800, height = 800, pointsize = 20)
-  iplot(mod) 
+  iplot(mod_all)
   dev.off()
   
   return(0)
@@ -76,6 +98,17 @@ PrepareData <- function(build, max_switches, years_pre, years_post,
       )
 
[…]

### 2023-06-22 source/analysis/exploration/SummaryTable.R

@@ -1,0 +1,17 @@
+#!/usr/bin/env Rscript --vanilla
+
+library(tidyverse) |> suppressPackageStartupMessages()
+library(vtable) |> suppressPackageStartupMessages()
+library(gtsummary)
+
+setwd("~/GitHub/BA")
+
+source("source/utils/HandleCommandArgs.R")
+
+# Read data
+t <- HandleCommandArgs(default_length = 50)
+read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
+build <- read_csv(read_path, show_col_types = F)
+
+vtable::sumtable(build)
+gtsummary::tbl_summary(build)

### 2023-06-22 source/utils/DataPrepSuite.R

@@ -28,7 +28,7 @@ AddLeadsLags <- function(with_e_dummies, years_pre, years_post){
   
   # Join treatment periods to data and add TREAT
   with_timing <- with_e_dummies |> 
-    left_join(timing, by="city_id", multiple = "all") |> # duplicate cities with multiple switches
+    left_join(timing, by="city_id", relationship = "many-to-many") |> # duplicate cities with multiple switches
     arrange(city_id, treat_time, period) |> 
     mutate(key = sprintf("id%i_t%i_s%i", city_id, period, treat_time)) |> # unique keys just in case
     mutate(treat = ifelse(is.na(treat_time), 0, 1)) # add TREAT dummy
@@ -53,7 +53,7 @@ AddTreatXPost <- function(with_e_dummies){
   
   # Join treatment periods to data and add D (TREAT x POST)
   with_D <- with_e_dummies |> 
-    left_join(timing, by="city_id", multiple = "all") |> # duplicate cities with multiple switches
+    left_join(timing, by="city_id", relationship = "many-to-many") |> # duplicate cities with multiple switches
     arrange(city_id, treat_time, period) |> 
     mutate(key = sprintf("id%i_t%i_s%i", city_id, period, treat_time)) |> # unique keys just in case
     mutate(D = case_when(

### 2023-06-22 source/utils/etableDefaults.R

@@ -16,7 +16,9 @@ etableDefaults <- function(){
                    D = "Switch to another state",
                    rule_conquest = "Conquest", rule_succession = "Succession",
                    rule_other = "Other", switches = "Switching indicator",
-                   conflict = "Conflict"))
+                   conflict = "Conflict",
+                   time_to_treat = "Period",
+                   treat = "Treat"))
 }
 
 PeriodInsert <- function(t){

### 2023-06-22 paper/output/regressions/SW22_replication_100y.tex
#28 round results to two decimals

@@
       \emph{Variables}\\
-      Treat $\times$ Period $=$ 0    & 0.02             & -0.03      & 0.0006        & -0.04$^{**}$\\   
-                                     & (0.02)           & (0.03)     & (0.02)        & (0.02)\\   
-      Treat $\times$ Period $=$ 100  & 0.09$^{***}$     & 0.02       & 0.01          & -0.008\\   
+      Treat $\times$ Period $=$ 0    & 0.02             & -0.03      & 0.00          & -0.04$^{**}$\\   
+                                     & (0.02)           & (0.02)     & (0.02)        & (0.02)\\   
+      Treat $\times$ Period $=$ 100  & 0.09$^{***}$     & 0.02       & 0.01          & -0.01\\   
                                      & (0.03)           & (0.03)     & (0.03)        & (0.03)\\   
@@
                                      & (0.03)           & (0.04)     & (0.03)        & (0.04)\\   
-      Switching indicator            & 0.02             & 0.06$^{*}$ & 0.009         & -0.02\\   
-                                     & (0.03)           & (0.03)     & (0.02)        & (0.03)\\   
+      Switching indicator            & 0.02             & 0.06$^{*}$ & 0.01          & -0.02\\   
+                                     & (0.03)           & (0.03)     & (0.02)        & (0.02)\\   
       \midrule

### 2023-06-22 paper/output/regressions/window_100y.tex
#28 round results to two decimals

@@
       \emph{Variables}\\
-      Switch to another state                    & 0.07$^{*}$       & 0.004       & 0.02    & -0.05$^{*}$\\   
+      Switch to another state                    & 0.07$^{*}$       & 0.00        & 0.02    & -0.05$^{*}$\\   
                                                  & (0.04)           & (0.03)      & (0.03)  & (0.03)\\   
-      Switch to another state $\times$ Conquest  & -0.03            & -0.08$^{*}$ & -0.03   & -0.005\\   
+      Switch to another state $\times$ Conquest  & -0.03            & -0.08$^{*}$ & -0.03   & -0.01\\   
                                                  & (0.04)           & (0.05)      & (0.04)  & (0.03)\\   
@@
                                                  & (0.04)           & (0.04)      & (0.03)  & (0.03)\\   
-      Switching indicator                        & 0.01             & 0.05$^{*}$  & 0.006   & -0.02\\   
-                                                 & (0.03)           & (0.03)      & (0.02)  & (0.03)\\   
+      Switching indicator                        & 0.02             & 0.05$^{*}$  & 0.01    & -0.02\\   
+                                                 & (0.03)           & (0.03)      & (0.02)  & (0.02)\\   
       \midrule

### 2023-06-22 paper/output/regressions/window_10y.tex
#28 round results to two decimals

@@
       \emph{Variables}\\
-      Switch to another state                    & 0.01$^{*}$       & 0.003   & -0.004                & -0.008$^{**}$\\   
-                                                 & (0.009)          & (0.004) & (0.003)               & (0.003)\\   
-      Switch to another state $\times$ Conquest  & -0.008           & -0.01   & 0.003                 & 0.004\\   
-                                                 & (0.01)           & (0.007) & (0.005)               & (0.005)\\   
-      Switch to another state $\times$ Other     & -0.01            & -0.005  & 0.002                 & 0.01$^{**}$\\   
-                                                 & (0.01)           & (0.005) & (0.004)               & (0.005)\\   
-      Switching indicator                        & -0.03$^{***}$    & -0.004  & -0.005                & -0.010$^{**}$\\   
-                                                 & (0.01)           & (0.005) & (0.004)               & (0.004)\\   
+      Switch to another state                    & 0.02$^{*}$       & 0.00   & 0.00                  & -0.01$^{**}$\\   
+                                                 & (0.01)           & (0.00) & (0.00)                & (0.00)\\   
+      Switch to another state $\times$ Conquest  & -0.01            & -0.01  & 0.00                  & 0.00\\   
+                                                 & (0.02)           & (0.01) & (0.01)                & (0.01)\\   
+      Switch to another state $\times$ Other     & -0.02            & -0.01  & 0.00                  & 0.01$^{**}$\\   
+                                                 & (0.01)           & (0.00) & (0.00)                & (0.00)\\   
+      Switching indicator                        & -0.03$^{***}$    & 0.00   & 0.00                  & -0.01$^{**}$\\   
+                                                 & (0.01)           & (0.01) & (0.00)                & (0.00)\\   
       \midrule

### 2023-06-22 paper/output/regressions/window_50y.tex
#28 round results to two decimals

@@
       \emph{Variables}\\
-      Switch to another state                    & 0.02             & 0.02          & 0.002   & -0.04$^{***}$\\   
+      Switch to another state                    & 0.02             & 0.02          & 0.00    & -0.04$^{***}$\\   
                                                  & (0.02)           & (0.02)        & (0.01)  & (0.01)\\   
-      Switch to another state $\times$ Conquest  & 0.04             & -0.07$^{***}$ & -0.005  & 0.02\\   
+      Switch to another state $\times$ Conquest  & 0.04             & -0.07$^{***}$ & 0.00    & 0.02\\   
                                                  & (0.04)           & (0.03)        & (0.02)  & (0.02)\\   
-      Switch to another state $\times$ Other     & 0.008            & -0.02         & 0.006   & 0.04$^{**}$\\   
+      Switch to another state $\times$ Other     & 0.01             & -0.02         & 0.01    & 0.04$^{**}$\\   
                                                  & (0.03)           & (0.02)        & (0.01)  & (0.02)\\   
-      Switching indicator                        & -0.01            & 0.007         & -0.005  & 0.0006\\   
+      Switching indicator                        & -0.01            & 0.01          & -0.01   & 0.00\\   
                                                  & (0.02)           & (0.02)        & (0.01)  & (0.01)\\   

### 2023-06-22 source/utils/etableDefaults.R

@@ -3,7 +3,7 @@ etableDefaults <- function(){
   # uniform defaults (across all tables)
   setFixest_etable(
     se.below = TRUE,
-    digits = 2,
+    digits = "r2",
     digits.stats = 4
   )
   

### 2023-06-22 source/analysis/baseline/Baseline.R
#28 change title of baseline table

@@ -58,7 +58,7 @@ Main <- function(){
   tex_output <- etable(mod_all_basic, mod_state_basic,
                        mod_priv_basic, mod_pub_basic,
                        tex=TRUE,
-                       title="Differences in trends after switching, by switch type",
+                       title="Differential effects of switching, by switch type",
                        label = sprintf("tab:baseline_%iy", t),
                        postprocess.tex = PostProcessBaseline,
                        notes = note)

### 2023-06-22 paper/paper.tex

@@
 \subsection{More construction events after switching states}
-Table \ref{tab:SW22_replication_50y} shows estimates for Equation \eqref{eq:sw22}. Figure \ref{fig:SW_replication} plots the coefficients to visualise how the effect evolves dynamically.
+
+Table \ref{tab:SW22_replication_50y} shows estimates for Equation \eqref{eq:sw22}. Figure \ref{fig:SW_replication} plots the coefficients in column (1) to visualise how the effect evolves dynamically.
+
+The
 

### 2023-06-22 paper/paper.tex
#28 add first results paragraph

@@
 
-\section{Data}
+\section{Data} \label{sec:data}
 
@@
 While it is impossible to empirically prove this assumption (because we cannot observe the counterfactual trend), it is common to check its validity by estimating pre-treatment effects: if the treatment and control group already showed different trends before being treated, there may be an unobserved variable correlated both with receiving treatment and with the outcome, biasing the estimate.
-For this reason I also include leads reaching 100 years into the past in my regression (as do \cite{schoenholzer2022}.)
+For this reason I include a lead reaching 100 years into the past in my regression (as do \cite{schoenholzer2022}.)
 
@@
 
-Table \ref{tab:SW22_replication_50y} shows estimates for Equation \eqref{eq:sw22}. Figure \ref{fig:SW_replication} plots the coefficients in column (1) to visualise how the effect evolves dynamically.
+Table \ref{tab:SW22_replication_50y} contains estimates from Equation \eqref{eq:sw22}.
+Each column reports the effect of switching states on a different subset of observed construction activity: Column (1) for all types of buildings, Column (2) for administrative, military and representative buildings, Column (3) for private construction, and Column (4) for buildings that provide public goods.
+The exact definitions of these categories are as described in Section \ref{sec:data}.
 
-The
+We can see in Column (1) that switching states has no immediate effect (relative to the period that precedes the switch) on the probability of recorded construction activity; over the following 200 years, however, the effect grows positive and significant.
+
+Figure \ref{fig:SW_replication} plots the coefficients in Column (1) to visualise how the effect evolves dynamically.
 

### 2023-06-22 paper/paper.tex
#28 begin writing baseline results

@@
 
-\subsection{Differential effects of switching}
+\subsection{Heterogeneous effects of switching}
 
-Next I attempt to identify the effect of state power --- the effect of the difference in selection for state power between conquests and successions.
-In the best case, if the identifying assumption is satisfied, the estimates from the previous regression tell us what the effect of switching states is - however, they cannot tell us why.
+In the best case, if the identifying assumption is satisfied, the estimates from the previous regression tell us what the effect of switching states is --- however, they cannot tell us why.
+In the following I attempt to identify the effect of state power, or more accurately, the effect of the difference in selection for state power between conquests and successions.
 
-Because the \textit{Städtebuch} data compiled by \cite{pt2} records exactly how each city was acquired by the territory holding it, I am able to look at differential trends after different types of ownership changes.\footnote
+Because the \textit{Städtebuch} data compiled by \cite{pt2} records exactly how each city was acquired by the territory holding it, I am able to estimate differential trends after different types of ownership changes.\footnote
 {
@@
 
-\subsection{More construction events after switching states}
+\subsection{Construction activity increases after switching states}
 
-Table \ref{tab:SW22_replication_50y} contains estimates from Equation \eqref{eq:sw22}.
+Table \ref{tab:SW22_replication_50y} contains estimates for the coefficients $\pi_\tau$ from Equation \eqref{eq:sw22}.
 Each column reports the effect of switching states on a different subset of observed construction activity: Column (1) for all types of buildings, Column (2) for administrative, military and representative buildings, Column (3) for private construction, and Column (4) for buildings that provide public goods.
 The exact definitions of these categories are as described in Section \ref{sec:data}.
+Note also that a $Period$ is an interval of 50 years that begins with the given year.
+$Period = -50$ is omitted as the reference category for treated cities.
+Reassuringly, as the coefficients on $Treat \times Period = -100$ show, there are no significant pre-trends in any of the outcome variables.
 
-We can see in Column (1) that switching states has no immediate effect (relative to the period that precedes the switch) on the probability of recorded construction activity; over the following 200 years, however, the effect grows positive and significant.
+We can see in Column (1), beginning  that switching states has no immediate effect on the probability of recorded construction activity.
+Over the next two centuries, however, the effect grows positive and significant: 100 years after a city switches states, the probability of construction activity is 5 percentage points higher than in cities that did not switch.
+After 200 years this difference grows to 9 percentage points.
+These estimates are significant at the 5 percent and 1 percent level, respectively.
+Figure \ref{fig:SW_replication} plots the coefficients in Column (1) to visualise these dynamics.
 
-Figure \ref{fig:SW_replication} plots the coefficients in Column (1) to visualise how the effect evolves dynamically.
+By contrast, almost all the coefficients in Columns (2), (3) and (4) are insignificant, suggesting that no particular subset of construction is driving the results.
+
+\subsection{No evidence for the state power hypothesis}
+
+Table \ref{tab:baseline_50y} shows estimates from Equation \eqref{eq:baseline}.
+This is a difference-in-differences specification with interactions to capture heterogeneous treatment effects.
+As before, different columns use different subsets of construction as the outcome variable.
+
+I report estimates of the coefficient of interest $\theta_{conquest}$ in the second row.
 

### 2023-06-22 paper/paper.tex
#28 fix typo

@@
 
-If this assumption holds true, the effect of state power should manifest itself in the difference between the trend of cities acquired by conquest and the trend of a cities acquired by succession.
+If this assumption holds true, the effect of state power should manifest itself in the difference between the trend of cities acquired by conquest and the trend of cities acquired by succession.
 Note that successions are the omitted category --- there is no interaction term between the post-treatment indicator and $\mathbf{1}(Succession_{it})$ in Equation \eqref{eq:baseline} --- so that the coefficient $\theta_{conquest}$ captures exactly what we want: the difference in trends after conquests vis-a-vis successions.

### 2023-06-22 paper/paper.tex
#28 clarification

@@
 
-By contrast, almost all the coefficients in Columns (2), (3) and (4) are insignificant, suggesting that no particular subset of construction is driving the results.
+By contrast, almost all the coefficients in Columns (2), (3) and (4) are insignificant, suggesting that neither of these specific subsets of construction is driving the results.
 

### 2023-06-22 source/analysis/baseline/Controls.R
#29 write script for controls table

@@ -1,0 +1,118 @@
+#!/usr/bin/env Rscript --vanilla
+
+library(tidyverse) |> suppressPackageStartupMessages()
+library(fixest) |> suppressPackageStartupMessages()
+
+setwd("~/GitHub/BA")
+
+source("source/utils/HandleCommandArgs.R")
+source("source/utils/DataPrepSuite.R")
+source("source/utils/etableDefaults.R")
+
+Main <- function(){
+  
+  # Read data
+  t <- HandleCommandArgs(default_length = 50)
+  read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
+  build <- read_csv(read_path, show_col_types = F)
+  
+  # Select sample and process data (exact steps documented below)
+  dat <- PrepareData(build,
+                     max_switches = 2,
+                     binarise_construction = T,
+                     binarise_switches = T)
+  
+  # Estimate Equation (2) with different sets of controls
+  mod_baseline <- fixest::feols(
+    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat)
+  
+  mod_noswitches <- fixest::feols( # neither
+    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) | city_id + period, 
+    data = dat)
+  
+  mod_conflict <- fixest::feols( # both
+    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches + conflict | city_id + period, 
+    data = dat)
+  
+  mod_conf_nosw <- fixest::feols( # only conflict
+    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + conflict | city_id + period, 
+    data = dat)
+  
+  
+  # Produce regression table and export to LaTeX
+  etableDefaults()
+  
+  note <- paste("Note: Table presents results of estimation equation
+  \\eqref{eq:baseline} but using different control variables. The switch type 
+  \"Succession\" is omitted as the reference category.", PeriodInsert(t), 
+  "Observations are at the city-period level. Dependent variables are indicators 
+  that take the value 1 if construction activity of the respective type was 
+  recorded. Standard errors are clustered at the city level.", SignifInsert(), 
+  sep = " ") |> 
+    str_replace_all("\n ", "")
+  
+  tex_output <- etable(mod_baseline, mod_noswitches,
[…]

### 2023-06-22 paper/output/regressions/controls_50y.tex
#29 output controls table

\begin{table}[htbp]
   \caption{\label{tab:controls_50y} Heterogeneous effects - different controls}
   \centering
   \begin{tabular}{lcccc}
      \tabularnewline \midrule \midrule
      Dependent Variable: & \multicolumn{4}{c}{All construction}\\
      Model:                                     & (1)    & (2)    & (3)           & (4)\\  
      \midrule
      \emph{Variables}\\
      Switch to another state                    & 0.02   & 0.02   & -0.01         & 0.03\\   
                                                 & (0.03) & (0.02) & (0.05)        & (0.05)\\   
      Switch to another state $\times$ Conquest  & 0.04   & 0.04   & -0.15$^{***}$ & -0.14$^{**}$\\   
                                                 & (0.04) & (0.04) & (0.05)        & (0.06)\\   
      Switch to another state $\times$ Other     & 0.01   & 0.01   & 0.05          & 0.05\\   
                                                 & (0.03) & (0.03) & (0.05)        & (0.05)\\   
      Switching indicator                        & -0.01  &        & -0.09         &   \\   
                                                 & (0.02) &        & (0.06)        &   \\   
      Conflict                                   &        &        & 0.01          & 0.01\\   
                                                 &        &        & (0.03)        & (0.03)\\   
      \midrule
      \emph{Fixed-effects}\\
      City                                       & Yes    & Yes    & Yes           & Yes\\  
      Period                                     & Yes    & Yes    & Yes           & Yes\\  
      \midrule
      \emph{Fit statistics}\\
      Observations                               & 6,635  & 6,635  & 1,955         & 1,955\\  
      R$^2$                                      & 0.3553 & 0.3553 & 0.4245        & 0.4230\\  
      Within R$^2$                               & 0.0009 & 0.0008 & 0.0071        & 0.0045\\  
      \midrule \midrule

   \end{tabular}

   \par \raggedright 
   Note: Table presents results of estimation equation \eqref{eq:baseline} but using different control variables. The switch type  "Succession" is omitted as the reference category. Yearly data was aggregated into periods of 50 years. Observations are at the city-period level. Dependent variables are indicators  that take the value 1 if construction activity of the respective type was  recorded. Standard errors are clustered at the city level. *, **, and *** denote significance on the 10 percent, 5 percent, and 1 percent  level, respectively.
\end{table}

### 2023-06-22 source/analysis/baseline/Controls.R

@@ -61,8 +61,6 @@ Main <- function(){
                        notes = note
                        )
   
-  print(tex_output)
-  
   filename <- sprintf("paper/output/regressions/controls_%iy.tex", t)
   write(tex_output, file=filename)
   

### 2023-06-23 paper/paper.tex
Updates from Overleaf

@@
 Comparing the effects of these two types of switches allows me to tease out the effect of state power.
-Identification is based on the assumption that conquests imply state power differentials but successions do not.
+Identification is based on the economic assumption that conquests imply state power differentials but successions do not.
 Specifically, if we consider the universe of all switching events, the average increase in state power from the old to the new owner would be higher for the conquests than for the successions.
@@
 This does not necessarily mean that conflict-driven evolution played no role in the rise of Europe --- several issues limit the statistical power of my test, which might prevent me from finding an effect even if there is one.
-The biggest one is that the data I use is arguably inadequate for answering the research question.
+One of them is that the data I use may not be fully adequate for answering the research question.
 
@@
 Based on the theory of conflict-driven evolution we would expect any such difference to stem from the fact that switching states generally means switching to a higher-quality state (due to selection).
-As such, even though it would be possible in principle to include state fixed effects, it is not a good idea here: state fixed effects would absorb the effect of differences in state quality which we want to capture in $\pi_\tau$.
+As such, even though it would be possible in principle to include state fixed effects, it is not a good idea here: state fixed effects would absorb the effect of differences in state quality which we want to capture in $\pi_\tau$ \citep[p.
+13]{schoenholzer2022}.
 
@@
 While it is impossible to empirically prove this assumption (because we cannot observe the counterfactual trend), it is common to check its validity by estimating pre-treatment effects: if the treatment and control group already showed different trends before being treated, there may be an unobserved variable correlated both with receiving treatment and with the outcome, biasing the estimate.
-For this reason I include a lead reaching 100 years into the past in my regression (as do \cite{schoenholzer2022}.)
+For this reason the regression also includes a lead reaching 100 years into the past.
 
@@
 
-Because the \textit{Städtebuch} data compiled by \cite{pt2} records exactly how each city was acquired by the territory holding it, I am able to estimate differential trends after different types of ownership changes.\footnote
+Because the \textit{Städtebuch} data compiled by \cite{pt2} records exactly how each city was acquired by the territory holding it, I am able to estimate heterogeneous effects for different types of ownership changes.\footnote
 {
@@
 I simply replace the lead and lag dummies with the post-treatment indicator $\mathbf{1}(t \ge e^{NewState}_i)$.
-The more important novelty are the interaction terms: I interact the type of switch with the post-treatment indicator to obtain estimates of how the trends in construction activity differ after different types of switches.
+The more important novelty are the interaction terms: I interact the post-treatment indicator with the type of switch to obtain estimates of how the trends in construction activity differ after different types of switches.
 I distinguish three categories: $Conquest$, $Succession$, and $Other$.
 In all cases this refers to how the state currently owning the city came into power --- for example $Conquest_{it} = 1$ means that the state that holds city $i$ at time $t$ acquired city $i$ by conquest.
-It will be useful in the following to adopt one more bit of notation from \cite{schoenholzer2022} and let $\mathbf{J}(i, t)$ denote the state ruling city $i$ at time $t$.
+It will be useful to adopt one final bit of notation from \cite{schoenholzer2022} and let $\mathbf{J}(i, t)$ denote the state ruling city $i$ at time $t$.
 
@@
 
-We can see in Column (1), beginning  that switching states has no immediate effect on the probability of recorded construction activity.
+We can see in Column (1) that switching states has no immediate effect on the probability of recorded construction activity.
 Over the next two centuries, however, the effect grows positive and significant: 100 years after a city switches states, the probability of construction activity is 5 percentage points higher than in cities that did not switch.

### 2023-06-23 source/analysis/baseline/Controls.R
#28 fix note in control table

@@ -46,10 +46,9 @@ Main <- function(){
   note <- paste("Note: Table presents results of estimation equation
   \\eqref{eq:baseline} but using different control variables. The switch type 
   \"Succession\" is omitted as the reference category.", PeriodInsert(t), 
-  "Observations are at the city-period level. Dependent variables are indicators 
-  that take the value 1 if construction activity of the respective type was 
-  recorded. Standard errors are clustered at the city level.", SignifInsert(), 
-  sep = " ") |> 
+  "Observations are at the city-period level. The dependent variable is an 
+  indicator that takes the value 1 if any construction activity was recorded. 
+  Standard errors are clustered at the city level.", SignifInsert(), sep = " ") |> 
     str_replace_all("\n ", "")
   
   tex_output <- etable(mod_baseline, mod_noswitches,

### 2023-06-23 source/analysis/baseline/Baseline.R
#28 fix baseline note, add controls table to scons

@@ -58,7 +58,7 @@ Main <- function(){
   tex_output <- etable(mod_all_basic, mod_state_basic,
                        mod_priv_basic, mod_pub_basic,
                        tex=TRUE,
-                       title="Differential effects of switching, by switch type",
+                       title="Heterogeneous effects of switching, by switch type",
                        label = sprintf("tab:baseline_%iy", t),
                        postprocess.tex = PostProcessBaseline,
                        notes = note)

### 2023-06-23 paper/output/regressions/controls_50y.tex
#28 fix note in control table
#28 change titles of tables

@@
 \begin{table}[htbp]
-   \caption{\label{tab:controls_50y} Heterogeneous effects - different controls}
+   \caption{\label{tab:controls_50y} Heterogeneous effects of switching: Robustness}
    \centering
@@
 Observations are at the city-period level.
-Dependent variables are indicators  that take the value 1 if construction activity of the respective type was  recorded.
+The dependent variable is an  indicator that takes the value 1 if any construction activity was recorded.
 Standard errors are clustered at the city level.

### 2023-06-23 source/analysis/baseline/Baseline.R

@@ -58,7 +58,7 @@ Main <- function(){
   tex_output <- etable(mod_all_basic, mod_state_basic,
                        mod_priv_basic, mod_pub_basic,
                        tex=TRUE,
-                       title="Heterogeneous effects of switching, by switch type",
+                       title="Heterogeneous effects of switching: Baseline",
                        label = sprintf("tab:baseline_%iy", t),
                        postprocess.tex = PostProcessBaseline,
                        notes = note)

### 2023-06-23 source/analysis/baseline/Controls.R

@@ -54,7 +54,7 @@ Main <- function(){
   tex_output <- etable(mod_baseline, mod_noswitches,
                        mod_conflict, mod_conf_nosw,
                        tex=TRUE,
-                       title="Heterogeneous effects - different controls",
+                       title="Heterogeneous effects of switching: Robustness",
                        label = sprintf("tab:controls_%iy", t),
                        postprocess.tex = PostProcessBaseline,
                        notes = note

### 2023-06-24 source/analysis/robustness/ConflictNA.R
#29 add conflict map (not in scons)

@@ -1,0 +1,143 @@
+#!/usr/bin/env Rscript --vanilla
+
+library(tidyverse) |> suppressPackageStartupMessages()
+library(sf)
+library(tmap)
+library(viridis)
+
+setwd("~/GitHub/BA")
+
+source("source/utils/HandleCommandArgs.R")
+source("source/utils/DataPrepSuite.R")
+source("source/utils/PrepareBaselineData.R")
+
+
+Main <- function(){
+  # Read data
+  t <- HandleCommandArgs(default_length = 50)
+  read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
+  build <- read_csv(read_path, show_col_types = F)
+  
+  # Prepare data (baseline)
+  dat <- PrepareBaselineData(build,
+                      max_switches = 2,
+                      binarise_construction = T,
+                      binarise_switches = T)
+  
+  
+  # Link obsevations to regions and obtain NA share for each region
+  locs <- read_csv("drive/raw/attributes/city_locations.csv", show_col_types = F) |> 
+    select(city_id, region_id, name, nat)
+  
+  na_shares <- GetNAShares(dat, locs)
+  
+  
+  # Get a spatial outline (convex hull) of each region
+  locs_spatial <- st_read("drive/raw/attributes/city_locations")
+  
+  regions <- unique(locs_spatial$region_id)
+  
+  polygons <- regions |> 
+    purrr::map(\(id)(RegionIDtoPolygon(id, locs_spatial))) |> 
+    bind_rows()
+  
+  # Bind the NA shares to the spatial data
+  na_shares_sf <- bind_cols(na_shares, polygons) |> st_as_sf()
+  
+  ### Create heatmap of NA shares
+  # Increase the bounding box of the spatial data to leave space for title
+  bbox_new <- BiggerBoundingBox(na_shares_sf, scale_top = 0.1)
+  
+  map1 <- tm_shape(na_shares_sf, 
+                   bbox = bbox_new)+
+    tm_polygons(col="share", 
+                title="Share missing",
+                palette = "Reds")+
+    tm_layout(title = "Sample composition after controlling for conflict",
+              title.size = 1,
+              title.position = c("center", "TOP"),
+              title.fontface = 2,
+              frame = FALSE)
+  
[…]

### 2023-06-24 source/utils/PrepareBaselineData.R

@@ -1,0 +1,30 @@
+PrepareBaselineData <- function(build, max_switches,
+                        binarise_construction, binarise_switches){
+  
+  # Filter out cities with too many lifetime switches
+  selected <- FilterBySwitches(build, max_switches)
+  
+  # Identify across-period switches
+  with_e_dummies <- AddEAnother(selected)
+  
+  # Add TREAT x POST dummies D
+  with_D <- AddTreatXPost(with_e_dummies)
+  
+  # Binarise construction outcomes, if specified
+  if (binarise_construction) { with_D <- with_D |> BinariseOutcomes() }
+  
+  # Binarise no. of switches (control), if specified
+  if (binarise_switches) { with_D <- with_D |> BinariseSwitches() }
+  
+  # Process type change into treatment type
+  with_types <- ProcessTypeChange(with_D)
+  
+  # Rearrange columns (for readability)
+  clean <- with_types |> select(
+    city_id, treat_time, period, terr_id, switches, e_another, D, 
+    rule_conquest, rule_succession, rule_other,
+    conflict, c_all, c_state, c_private, c_public
+  )
+  
+  return(clean)
+}

### 2023-06-24 paper/paper.tex
#28 include controls table
Updates from Overleaf

@@
 \begin{split}
-    construction_{it} &= \alpha_i + \delta_t + \mathbf{1}(S_{it} > 0)\beta +
-    \mathbf{1}(t \ge e^{NewState}_i)\pi \\
+    construction_{it} &= \alpha_i + \delta_t + \mathbf{X}_{it} \beta + 
+    \mathbf{1}(t \ge e^{NewState}_i) \pi \\
     &+ \mathbf{1}(t \ge e^{NewState}_i) \times
-    \mathbf{1}(Conquest_{it}) \times \theta_{conquest} \\
+    \mathbf{1}(Conquest_{it}) \cdot \theta_{conquest} \\
     &+ \mathbf{1}(t \ge e^{NewState}_i) \times
-    \mathbf{1}(Other_{it}) \times \theta_{other} + \varepsilon_{it},
+    \mathbf{1}(Other_{it}) \cdot \theta_{other} + \varepsilon_{it},
 \end{split}
@@
 I simply replace the lead and lag dummies with the post-treatment indicator $\mathbf{1}(t \ge e^{NewState}_i)$.
+Also, this equation features a vector of city-year level control variables $\mathbf{X}_{it}$.
+In the baseline version of this specification I control only for the dummy $\mathbf{1}(S_{it} > 0)$ that indicates switches in the same period, exactly like in Equation \eqref{eq:sw22}.
+In a robustness check I will additionally control for conflict incidents.
+
 The more important novelty are the interaction terms: I interact the post-treatment indicator with the type of switch to obtain estimates of how the trends in construction activity differ after different types of switches.
@@
 
+table 3: why neg? conflict NA may be systematocally correlated
+
+are some regions more common in the conflict NAs?
+
+The first explanation is that conflict is an important omitted variable, in the sense that it is correlated both with the treatment and with construction activity.
+
+If omitting  bias in $\theta_{conflict}$ from omitting conflict has the same sign as 
+$$\beta_{conflict} \cdot \frac{\mathbf{Cov}({Conflict},\: {Switching\: to\: another\: state \times Conquest})}{\mathbf{Var}(Conflict)}$$
+
+The whole term is negative.
+We can see in Columns (3) and (4) of Table \ref{tab:controls_50y} that $\beta_{conflict}$ has a positive sign.
+It follows that the correlation between must be negative (which makes sense - an owner with higher state power is likely better able to protect a city from attacks.
+
+But if $\beta_{conflict}$ is around zero, then why does it introduce bias at all?
+
+It doesn't make sense. It has to be something about the sample. Johannes must have been spot on.
+
 % ######################################################
@@
 
+\input{paper/output/regressions/controls_50y}
+
 \end{document}

### 2023-06-24 source/analysis/robustness/ConflictNA.R
#29 ConflictNA to scons

@@ -1,16 +1,14 @@
 #!/usr/bin/env Rscript --vanilla
 
 library(tidyverse) |> suppressPackageStartupMessages()
-library(sf)
-library(tmap)
-library(viridis)
+library(sf) |> suppressPackageStartupMessages()
+library(tmap) |> suppressPackageStartupMessages()
 
 setwd("~/GitHub/BA")
 
 source("source/utils/HandleCommandArgs.R")
 source("source/utils/DataPrepSuite.R")
 source("source/utils/PrepareBaselineData.R")
-
 
 Main <- function(){
   # Read data
@@ -55,9 +53,8 @@ Main <- function(){
                 palette = "Reds")+
     tm_layout(title = "Sample composition after controlling for conflict",
               title.size = 1,
-              title.position = c("center", "TOP"),
-              title.fontface = 2,
-              frame = FALSE)
+              title.position = c("center", "top"),
+              title.fontface = 2)
   
   # Add some city names as reference points
   citynames <- c("Muenchen", "Berlin", "Koenigsberg Pr.", "Koeln", 
@@ -76,9 +73,9 @@ Main <- function(){
   
   # Save map as PNG
   filename <- sprintf("paper/output/descriptive/map_conflict_NA_%iy.png", t)
-  png(filename=filename, width = 800, height = 800, pointsize = 20)
-  map2
-  dev.off()
+  tmap_save(map2, filename)
+  
+  return(0)
 }
 
 

### 2023-06-24 source/analysis/robustness/ConflictNA.R
#27 add cities map

@@ -9,6 +9,7 @@ setwd("~/GitHub/BA")
 source("source/utils/HandleCommandArgs.R")
 source("source/utils/DataPrepSuite.R")
 source("source/utils/PrepareBaselineData.R")
+source("source/utils/MapUtils.R")
 
 Main <- function(){
   # Read data
@@ -98,46 +99,6 @@ GetNAShares <- function(dat, locs){
     filter(conflict_na == TRUE)
   
   return(sum_dat)
-}
-
-
-RegionIDtoPolygon <- function(id, locs){
-  # Returns the convex hull of the set of cities in each region
-  
-  region_df <- locs |> filter(region_id == id)
-  nc <- st_centroid(region_df)
-  ch <- st_convex_hull(st_union(nc))
-  ch_as_sf <- st_sf(ch)
-  return(ch_as_sf)
-}
-
-
-CleanNames <- function(locs){
-  # Fixes spelling of city names.
-  clean <- locs |> 
-    mutate(name = str_replace_all(name, "ue", "ü")) |> 
-    mutate(name = str_replace_all(name, "oe", "ö")) |> 
-    mutate(name = str_replace(name, " Pr.", ""))
-  return(clean)
-}
-
-
-BiggerBoundingBox <- function(map_data, scale_top){
-  # From stackoverflow
-  bbox_new <- st_bbox(map_data) # current bounding box
-  
-  xrange <- bbox_new$xmax - bbox_new$xmin # range of x values
-  yrange <- bbox_new$ymax - bbox_new$ymin # range of y values
-  
-  # bbox_new[1] <- bbox_new[1] - (0.25 * xrange) # xmin - left
-  # bbox_new[3] <- bbox_new[3] + (0.25 * xrange) # xmax - right
-  # bbox_new[2] <- bbox_new[2] - (0.25 * yrange) # ymin - bottom
-  bbox_new[4] <- bbox_new[4] + (scale_top * yrange) # ymax - top
-  
-  bbox_new <- bbox_new |>   # take the bounding box ...
-    st_as_sfc() # ... and make it a sf polygon
-  
-  return(bbox_new)
 }
 
 

### 2023-06-24 source/analysis/summarystats/BasicStats.R

@@ -1,52 +1,142 @@
 #!/usr/bin/env Rscript --vanilla
 
 library(tidyverse) |> suppressPackageStartupMessages()
+library(sf)
+library(tmap)
 
 setwd("~/GitHub/BA")
 
 source("source/utils/HandleCommandArgs.R")
-source("source/utils/DataPrepSuite.R")
-source("source/utils/PrepareBaselineData.R")
+source("source/utils/MapUtils.R")
 
-# Read data
-t <- HandleCommandArgs(default_length = 1)
-read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
-build <- read_csv(read_path, show_col_types = F)
+Main <- function(){
+  # Read data
+  t <- HandleCommandArgs(default_length = 1)
+  read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
+  build <- read_csv(read_path, show_col_types = F)
+  
+  raw_build <- read_csv("drive/derived/cities_switches.csv",
+                        show_col_types = F) |> 
+    rename(
+      period = year,
+      switches = switch) |> 
+    mutate(conflict = 0) # just for convenience
+  
+  
+  GetBuildLevelInfo <- function(build){
+    build |> 
+      summarise(
+        min_year = min(period),
+        max_year = max(period),
+        avg_year = mean(period),
+        med_year = median(period),
+        n_cities = n_distinct(city_id),
+        n_terrs = n_distinct(terr_id)
+      )
+  }
+  
+  GetCityLevelInfo <- function(build){
+    build |> 
+      group_by(city_id) |> 
+      summarise(
+        duration = max(period) - min(period),
+        switches = sum(switches),
+        conflict = sum(conflict),
+        across(starts_with("c_"), sum)) |> 
+      summarise(
+        across(-c(city_id), \(col)(mean(col, na.rm=T)))
+      )
+  }
+  
+  
+  dsets <- list("build" = build, "raw" = raw_build)
+  dsets |> map(GetBuildLevelInfo)
+  dsets |> map(GetCityLevelInfo)
+}
 
-raw_build <- read_csv("drive/derived/cities_switches.csv",
-                      show_col_types = F) |> 
-  rename(
-    period = year,
-    switches = switch) |> 
-  mutate(conflict = 0) # just for convenience
 
+## Region level
 
-GetBuildLevelInfo <- function(build){
-  build |> 
[…]

### 2023-06-24 source/analysis/summarystats/MapCities.R

@@ -1,0 +1,100 @@
+#!/usr/bin/env Rscript --vanilla
+
+library(tidyverse) |> suppressPackageStartupMessages()
+library(sf)
+library(tmap)
+
+setwd("~/GitHub/BA")
+
+source("source/utils/HandleCommandArgs.R")
+source("source/utils/MapUtils.R")
+
+Main <- function(){
+  # Read data
+  t <- HandleCommandArgs(default_length = 50)
+  read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
+  build <- read_csv(read_path, show_col_types = F)
+  locs <- read_csv("drive/raw/attributes/city_locations.csv", show_col_types = F) |> 
+    select(city_id, region_id, name, nat)
+  locs_spatial <- st_read("drive/raw/attributes/city_locations")
+  
+  # Compute no. of cities in each region
+  region_stats <- GetRegionLevelInfo(build, locs)
+  
+  # Make a spatial dataset of region polygons
+  regions <- unique(locs_spatial$region_id)
+  
+  polygons <- regions |> 
+    purrr::map(\(id)(RegionIDtoPolygon(id, locs_spatial))) |> 
+    bind_rows()
+  
+  # Bind the city counts to the spatial region data
+  stats_sf <- bind_cols(region_stats, polygons) |> st_as_sf()
+  
+  ### Create heatmap of city counts
+  # Increase the bounding box of the spatial data to leave space for title
+  bbox_new <- BiggerBoundingBox(stats_sf, scale_top = 0.1)
+  
+  map1 <- tm_shape(stats_sf, 
+                   bbox = bbox_new)+
+    tm_polygons(col="n_cities", 
+                title="No. of cities",
+                palette = "Greens",
+                legend.reverse = TRUE) +
+    tm_layout(title = "Spatial distribution of cities",
+              title.size = 1.3,
+              title.position = c("center", "top"),
+              title.fontface = 2,
+              legend.title.size = 1.3,
+              legend.text.size = 1)
+  
+  # Add an unlabelled dot for each city in the build
+  allpoints_sf <- locs_spatial |> 
+    filter(city_id %in% unique(build$city_id))
+  
+  map2 <- map1 + 
+    tm_shape(allpoints_sf)+
+    tm_dots(alpha = 0.3)
+
+  
+  # Add some city names as reference points
[…]

### 2023-06-24 source/utils/MapUtils.R

@@ -1,0 +1,38 @@
+RegionIDtoPolygon <- function(id, locs){
+  # Returns the convex hull of the set of cities in each region
+  
+  region_df <- locs |> filter(region_id == id)
+  nc <- st_centroid(region_df)
+  ch <- st_convex_hull(st_union(nc))
+  ch_as_sf <- st_sf(ch)
+  return(ch_as_sf)
+}
+
+
+CleanNames <- function(locs){
+  # Fixes spelling of city names.
+  clean <- locs |> 
+    mutate(name = str_replace_all(name, "ue", "ü")) |> 
+    mutate(name = str_replace_all(name, "oe", "ö")) |> 
+    mutate(name = str_replace(name, " Pr.", ""))
+  return(clean)
+}
+
+
+BiggerBoundingBox <- function(map_data, scale_top){
+  # From stackoverflow
+  bbox_new <- st_bbox(map_data) # current bounding box
+  
+  xrange <- bbox_new$xmax - bbox_new$xmin # range of x values
+  yrange <- bbox_new$ymax - bbox_new$ymin # range of y values
+  
+  # bbox_new[1] <- bbox_new[1] - (0.25 * xrange) # xmin - left
+  # bbox_new[3] <- bbox_new[3] + (0.25 * xrange) # xmax - right
+  # bbox_new[2] <- bbox_new[2] - (0.25 * yrange) # ymin - bottom
+  bbox_new[4] <- bbox_new[4] + (scale_top * yrange) # ymax - top
+  
+  bbox_new <- bbox_new |>   # take the bounding box ...
+    st_as_sfc() # ... and make it a sf polygon
+  
+  return(bbox_new)
+}

### 2023-06-24 paper/paper.tex
Updates from Overleaf

@@
 The rest of the paper proceeds as follows.
-Section 2 presents some background information.
-Section 3 describes the data used in the analysis.
-Section 4 details the empirical approach.
-Section 5 presents results.
-Section 6 discusses implications and robustness of the results.
-Section 7 concludes.
-
-% ######################################################
-
-\section{Historical background}
-
-Early modern Europe consisted of many large and small states.
-
-One story about a city's territory switch
-
-Two pithy figures, and a no-nonsense table
-
-NO MORE THAN TWO PAGES TOTAL
+Section \ref{sec:data} describes the data used in the analysis.
+Section \ref{sec:empirics} details the empirical approach.
+Section \ref{sec:results} presents results.
+Section \ref{sec:discussion} discusses implications and robustness of the results.
+Section \ref{sec:conclusion} concludes.
 
@@
 I use data from the \textit{Princes and Townspeople} database which is largely based on the \textit{Deutsches Städtebuch} \citep{keyser1939} and \cite{kobler2007}'s \textit{Historisches Lexikon der Deutschen Länder}.
-The specific datasets I use are those on city locations \citep{pt1}, territorial histories \citep{pt2}, construction activity \citep{pt5} and conflict incidents \citep{pt6}.\footnote{I thank Davide Cantoni and Matthias Weigand for giving me access to unpublished yearly versions of the construction and territorial histories data.} I aggregate data from these sources to obtain a city-by-year panel of switching events, construction activity, and conflict.
+The specific datasets I use are those on city locations \citep{pt1}, territorial histories \citep{pt2}, construction activity \citep{pt5} and conflict incidents \citep{pt6}.\footnote{I thank Davide Cantoni and Matthias Weigand for giving me access to unpublished yearly versions of the construction and territorial histories data.}
 
-Switching: describe format; total number of recorded switches; territory id.
+I aggregate data from these sources to obtain a panel of
+
+- unit of observation: city-year / city-period
+- precision of time points
+- aggregation of years
+- unbalanced panel
+- definition of territory holding
+- definition of switch in period
+- justification for using construction
+- different groupings of construction
+- scope of the data / regions included
+- time period recorded
+- missing values / full sequence of event years
+- dropNA construction -> sample size
 
@@
 
-\section{Regression equations}
+\section{Regression equations} \label{sec:empirics}
 
@@
 
-\section{Results}
+\section{Results} \label{sec:results}
 
@@
 
-\section{Discussion}
+\section{Discussion} \label{sec:discussion}
 
@@
 
-\section{Conclusion}
+\section{Conclusion} \label{sec:conclusion}
 
@@
 
+\begin{figure}
+    \centering
+    \includegraphics{}
+    \caption{Caption}
+    \label{fig:enter-label}
+\end{figure}
+
 \begin{figure}[ht]
@@
     \label{fig:SW_replication}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[scale = 0.15]{paper/output/descriptive/map_conflict_NA_50y.png}
+    \caption{Share of observations remaining from each \textit{Städtebuch} region when controlling for conflict incidents in Equation \eqref{eq:baseline}.
+Geo-referenced borders of the regions were not available; this makeshift visualisation shows convex hulls of the city locations in each region (which results in some overlap.)}
+    \label{fig:conflict_map}
 \end{figure}

### 2023-06-24 source/analysis/robustness/ConflictNA.R
#29 add dots to NA map

@@ -61,6 +61,15 @@ Main <- function(){
               legend.title.size = 1.3,
               legend.text.size = 1)
   
+  
+  # Add an unlabelled dot for each city in the build
+  allpoints_sf <- locs_spatial |> 
+    filter(city_id %in% unique(dat$city_id))
+  
+  map2 <- map1 + 
+    tm_shape(allpoints_sf)+
+    tm_dots(alpha = 0.3)
+  
   # Add some city names as reference points
   citynames <- c("Muenchen", "Berlin", "Koenigsberg Pr.", "Koeln", 
                  "Hannover", "Breslau", "Wuerzburg", "Dresden",
@@ -70,7 +79,7 @@ Main <- function(){
     filter(name %in% citynames) |> 
     CleanNames()
   
-  map2 <- map1 +
+  map3 <- map2 +
     tm_shape(points_sf)+
     tm_dots()+
     tm_text("name", size=1, just="bottom", ymod=0.35
@@ -78,7 +87,7 @@ Main <- function(){
   
   # Save map as PNG
   filename <- sprintf("paper/output/descriptive/map_conflict_NA_%iy.png", t)
-  tmap_save(map2, filename)
+  tmap_save(map3, filename)
   
   return(0)
 }

### 2023-06-24 paper/paper.tex

@@
 - different groupings of construction
-- scope of the data / regions included
+- scope of the data / regions included: 1937 germany
 - time period recorded
@@
 
-\begin{figure}
+\begin{figure}[ht]
     \centering
-    \includegraphics{}
-    \caption{Caption}
-    \label{fig:enter-label}
+    \includegraphics[scale=0.15]{paper/output/descriptive/map_cities.png}
+    \caption{Distribution of cities in the assembled dataset.
+Polygons on the map do \textbf{not} represent historical borders; they merely represent regions by which cities are clustered in the \textit{Städtebuch}.
+The polygons are computed as convex hulls of the city locations in each region because geo-referenced borders of regions were not available --- this results in some overlap.}
+    \label{fig:map_cities}
 \end{figure}

### 2023-06-24 paper/paper.tex
#27 all figures and table are in

@@
     \includegraphics[scale = 0.4]{paper/output/regressions/SW22_replication_50y.png}
-    \caption{It looks just like it's supposed to!}
+    \caption{Dynamic effects of switching to another state.
+Coefficient estimates from Table \ref{eq:sw22}, Column (1).
+The outcome variable is the probability that construction activity of any type is observed in the period.
+The period beginning 50 years before the switch is omitted.}
     \label{fig:SW_replication}
@@
     \caption{Share of observations remaining from each \textit{Städtebuch} region when controlling for conflict incidents in Equation \eqref{eq:baseline}.
-Geo-referenced borders of the regions were not available; this makeshift visualisation shows convex hulls of the city locations in each region (which results in some overlap.)}
+Note that this map shows fewer cities than Figure \ref{fig:map_cities} because Figure \ref{fig:map_cities} was produced from the full dataset, while this map only plots cities that satisfy the baseline sample restrictions.}
     \label{fig:conflict_map}

### 2023-06-24 source/analysis/robustness/MapConflictNA.R
#27 fix critical bug in maps

@@ -29,13 +29,14 @@ Main <- function(){
     select(city_id, region_id, name, nat)
   
   na_shares <- GetNAShares(dat, locs) |> 
-    mutate(remaining = 1-share)
+    mutate(remaining = 1-share) |> 
+    arrange(region_id)
   
   
   # Get a spatial outline of each region
   locs_spatial <- st_read("drive/raw/attributes/city_borders")
   
-  regions <- unique(locs_spatial$region_id)
+  regions <- unique(locs_spatial$region_id) |> sort()
   
   polygons <- regions |> 
     purrr::map(\(id)(RegionIDtoUnion(id, locs_spatial))) |> 
@@ -60,7 +61,6 @@ Main <- function(){
               title.fontface = 2,
               legend.title.size = 1.3,
               legend.text.size = 1)
-  
   
   # Add an unlabelled dot for each city in the build
   allpoints_sf <- locs_spatial |> 

### 2023-06-24 source/analysis/summarystats/MapCitiesRaw.R

@@ -15,10 +15,11 @@ Main <- function(){
   locs_spatial <- st_read("drive/raw/attributes/city_borders")
   
   # Compute no. of cities in each region
-  region_stats <- GetRegionLevelInfo(build, locs)
+  region_stats <- GetRegionLevelInfo(build, locs) |> 
+    arrange(region_id)
   
   # Make a spatial dataset of region polygons
-  regions <- unique(locs_spatial$region_id)
+  regions <- unique(locs_spatial$region_id) |> sort()
   
   polygons <- regions |> 
     purrr::map(\(id)(RegionIDtoUnion(id, locs_spatial))) |> 

### 2023-06-24 source/analysis/summarystats/MapCitiesSample.R

@@ -24,10 +24,11 @@ Main <- function(){
   locs_spatial <- st_read("drive/raw/attributes/city_borders")
   
   # Compute no. of cities in each region
-  region_stats <- GetRegionLevelInfo(dat, locs)
+  region_stats <- GetRegionLevelInfo(dat, locs) |> 
+    arrange(region_id)
   
   # Make a spatial dataset of region polygons
-  regions <- unique(locs_spatial$region_id)
+  regions <- unique(locs_spatial$region_id) |> sort()
   
   polygons <- regions |> 
     purrr::map(\(id)(RegionIDtoUnion(id, locs_spatial))) |> 

### 2023-06-24 paper/paper.tex
#27 fix caption
Updates from Overleaf

@@
 The rest of the paper proceeds as follows.
+Section \ref{sec:background} gives essential historical background information.
 Section \ref{sec:data} describes the data used in the analysis.
@@
 
+\section{State system of the HRE} \label{sec:background}
+
+The Holy Roman Empire was a powerful monarchy that existed in Central Europe from the Middle Ages until its dissolution in 1806.
+Though ruled by the Holy Roman Emperor at the Imperial level, it consisted of many small and large territories with their own regional rulers.
+Before Germany first became a republic in 1918, it is not helpful to think of "states" in the modern sense: while territories may have existed as political entities on paper, they were ruled by aristocrats who handled power in their familial structures \citep[pp.
+2-3]{pt2}.
+These aristocratic families defined the territorial history of the HRE: territories were split by inheritances, joined by marriages and extinctions, and driven to war by struggles for dominance among the noble lineages ruling them.
+
+When I say "state" in the context of this paper I mean its early modern German equivalent --- the territory of a lineage of rulers, or the lineage itself.
+
 \section{Data} \label{sec:data}
 
-I use data from the \textit{Princes and Townspeople} database which is largely based on the \textit{Deutsches Städtebuch} \citep{keyser1939} and \cite{kobler2007}'s \textit{Historisches Lexikon der Deutschen Länder}.
+I use data from the \textit{Princes and Townspeople} database which is based on the \textit{Deutsches Städtebuch} \citep{keyser1939} and \cite{kobler2007}'s \textit{Historisches Lexikon der Deutschen Länder}.
+\cite{keyser1939} claim to  list "every settlement that at one point in the history of the Holy Roman Empire was awarded the status of a city" \citep[p.
+2]{pt1}.
+The raw \textit{Princes and Townspeople} database contains those 2,390 cities.
+The reporting boundaries are those of the German Empire in 1937 plus the Free City of Danzig.
 The specific datasets I use are those on city locations \citep{pt1}, territorial histories \citep{pt2}, construction activity \citep{pt5} and conflict incidents \citep{pt6}.\footnote{I thank Davide Cantoni and Matthias Weigand for giving me access to unpublished yearly versions of the construction and territorial histories data.}
 
-I aggregate data from these sources to obtain a panel of
+The most important There are detailed records of the territorial histories of cities (among many other things): we know to which lineage's territory a city belonged to at every time point in which it shows up in the data.
+Records begin in 1300 at the earliest and end at the latest with the abolition of the German monarchy and the foundation of the Weimar Republic in 1918.
 
@@
     \caption{Distribution of cities in the assembled dataset.
+Each dot represents a city.
+Some major cities are identified with labels.
 Polygons on the map do \textbf{not} represent historical borders; they merely represent regions by which cities are clustered in the \textit{Städtebuch}.

### 2023-06-24 paper/paper.tex
#27 improve captions

@@
     \centering
-    \includegraphics[scale=0.15]{paper/output/descriptive/map_cities.png}
-    \caption{Distribution of cities in the assembled dataset.
+    \includegraphics[scale=0.15]{paper/output/descriptive/map_cities_raw.png}
+    \caption{Distribution of cities in the \textit{Territorial Histories} dataset \citep{pt2}.
 Each dot represents a city.
 Some major cities are identified with labels.
-Polygons on the map do \textbf{not} represent historical borders; they merely represent regions by which cities are clustered in the \textit{Städtebuch}.
-The polygons are computed as convex hulls of the city locations in each region because geo-referenced borders of regions were not available --- this results in some overlap.}
-    \label{fig:map_cities}
+Polygons on the map do \textbf{not} represent historical borders; they merely represent regions by which cities are clustered in the \textit{Städtebuch}.}
+    \label{fig:map_cities_raw}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[scale=0.15]{paper/output/descriptive/map_cities_sample.png}
+    \caption{Distribution of cities in the baseline sample.
+Excludes cities that underwent more than two lifetime switches, and cities without any recorded construction activity.
+Each dot represents a city.
+Some major cities are identified with labels.
+Polygons on the map do \textbf{not} represent historical borders; they merely represent regions by which cities are clustered in the \textit{Städtebuch}.}
+    \label{fig:map_cities_sample}
 \end{figure}
@@
     \includegraphics[scale = 0.15]{paper/output/descriptive/map_conflict_NA_50y.png}
-    \caption{Share of observations remaining from each \textit{Städtebuch} region when controlling for conflict incidents in Equation \eqref{eq:baseline}.
-Note that this map shows fewer cities than Figure \ref{fig:map_cities} because Figure \ref{fig:map_cities} was produced from the full dataset, while this map only plots cities that satisfy the baseline sample restrictions.}
+    \caption{Share of observations from each \textit{Städtebuch} region with non-missing data on conflict incidents.
+The $Conflict$ variable is coded as missing for a period if it is earlier than the first recorded incident or later than the last recorded incident in a city.
+Observations with missing conflict data are excluded from the regressions in Columns (3) and (4) of Table \ref{tab:controls_50y}.}
     \label{fig:conflict_map}

### 2023-06-24 source/analysis/summarystats/MapCitiesSample.R
#27 fix legend on sample map

@@ -42,7 +42,7 @@ Main <- function(){
   bbox_new <- BiggerBoundingBox(stats_sf, scale_top = 0.1)
   
   # Define breaks myself (so that I can use the same ones in MapAllCities.R)
-  my.breaks <- c(0, 31, 61, 91, 121, 151)
+  my.breaks <- c(0, 31, 61, 91, 121, 151, 200)
   
   # Create basic map
   map1 <- tm_shape(stats_sf, 

### 2023-06-25 paper/paper.tex
#27 account for NAs

@@
 
+\paragraph{Territorial histories.}
+
 The most important There are detailed records of the territorial histories of cities (among many other things): we know to which lineage's territory a city belonged to at every time point in which it shows up in the data.
@@
 - time period recorded
-- missing values / full sequence of event years
-- dropNA construction -> sample size
 
-Construction: describe format / split and rationale. This can be two paragraphs.
+\paragraph{Event variables.}
 
-Conflict: describe format (just a dummy).
+I code such an "event" variable as missing for periods that are earlier than the first recorded event or later than the last
+recorded event in a city.
 
-I also build coarsened versions of this dataset in which I aggregate years into longer periods (10, 50, and 100 years).
-Not only does this aggregation make event studies nicer to look at, it also allows me to use more of the conflict and construction data as periods get longer: the 50-year and 100-year versions contain events that aren't recorded with a precise year, but only "beginning of the 16th century" and the like.
-While the 1-year and 10-year data builds contain BLANK construction events with sufficiently precise timing, the 50-year version contains BLANK and the 100-year version contains BLANK.
+Since I use some subset of construction as the outcome variable in all regressions, it is necessary to drop missing values.
+
+\paragraph{Aggregation.}
+
+I aggregate this yearly data into periods of different length.
+I build yearly, 10-yearly, 50-yearly and 100-yearly versions.
+I use the 50-yearly version throughout the regressions in the paper.
+Events are rare enough that this avoids losing too much information to having to drop missing values.
+Also, some events (conflict as well as construction) are recorded with uncertain timing --- descriptors like "beginning of the 16th century" rather than an exact year.
+In the aggregated version I include such events if their timing can be assigned with certainty to a 50-year period.
+
+\paragraph{Accounting for dropped observations.}
+
+The raw \textit{Territorial Histories} dataset starts out with 1,471,934 yearly observations.
+Two observations per city --- the first and the last year --- are lost while coding switches, for a total of 4,780.
+The remaining 1,467,154 yearly observations are aggregated into 30,980 observations with 50-year spacing.
+Of these I drop 16,110 that are not covered by the construction data.
+Finally, the baseline versions of all specifications in this paper restrict the sample to cities with no more than two lifetime switches.
+This will remove another 9,975 observations, leaving 6,635 in the baseline sample.
+
+Figure \ref{fig:map_cities_sample} plots the spatial distribution of the cities that make it into the baseline sample.
+The total number of cities has shrunken visibly compared to Figure \ref{fig:map_cities_raw} --- we can also see, however, that the shrinking appears to have been roughly uniform, preserving the proportions between regions.
+Only Brandenburg (containing Berlin) and Silesia (containing Breslau) seem to lose disproportionately many observations.
 
@@
 
+Conflict is not actually correlated with construction, but controlling for it has the side effect of dropping two thirds of all observations from the sample. This is because
+
 It doesn't make sense. It has to be something about the sample. Johannes must have been spot on.

### 2023-06-25 paper/paper.tex
#27 expand data section

@@
 
+A city-by-year panel of territorial history, taken from \cite{pt2}, forms the backbone of my dataset.
+For each city in each year it identifies the noble lineage to whose territory the city belonged in that year.
+This is what I
+
 The most important There are detailed records of the territorial histories of cities (among many other things): we know to which lineage's territory a city belonged to at every time point in which it shows up in the data.
@@
 
-I code such an "event" variable as missing for periods that are earlier than the first recorded event or later than the last
-recorded event in a city.
+Different types of construction
+
+Additionally, conflict
 
 Since I use some subset of construction as the outcome variable in all regressions, it is necessary to drop missing values.
+I code "event" variables as missing for periods that are earlier than the first recorded event or later than the last recorded event in a city.
 
@@
 
+The aggregated indicators for construction and conflict take the value 1 if at least one event was recorded during the period. The switches and rule types
+
 \paragraph{Accounting for dropped observations.}
@@
 The raw \textit{Territorial Histories} dataset starts out with 1,471,934 yearly observations.
-Two observations per city --- the first and the last year --- are lost while coding switches, for a total of 4,780.
-The remaining 1,467,154 yearly observations are aggregated into 30,980 observations with 50-year spacing.
+I lose two observations per city --- the first and the last year --- while coding switches, for a total of 4,780.
+I aggregate the remaining 1,467,154 yearly observations into 30,980 observations with 50-year spacing.
 Of these I drop 16,110 that are not covered by the construction data.

### 2023-06-25 paper/paper.tex
#27 big progress on data section

@@
 
-I use data from the \textit{Princes and Townspeople} database which is based on the \textit{Deutsches Städtebuch} \citep{keyser1939} and \cite{kobler2007}'s \textit{Historisches Lexikon der Deutschen Länder}.
-\cite{keyser1939} claim to  list "every settlement that at one point in the history of the Holy Roman Empire was awarded the status of a city" \citep[p.
+I use data from the \textit{Princes and Townspeople} collection which is based on the \textit{Deutsches Städtebuch} \citep{keyser1939} and \cite{kobler2007}'s \textit{Historisches Lexikon der Deutschen Länder}.
+Out of the six datasets currently published in the collection, I use those on territorial histories \citep{pt2}, construction activity \citep{pt5} and conflict incidents \citep{pt6} in my analysis.\footnote{I thank Davide Cantoni and Matthias Weigand for giving me access to unpublished yearly versions of the construction and territorial histories data.} I also use the geo-referenced city locations \citep{pt1} to create visualisations.
+
+From these sources I assemble a panel of cities.
+The unique unit of observation is the city-period.
+The ultimate dataset contains, in abstract terms, two types of variables: ones that code the \textit{territorial history} of cities, and ones that code the occurrence of certain \textit{events}.
+Records for each city begin in the 14th century at the earliest and end at the latest with the abolition of the German monarchy in the early 20th century.
+
+Figure \ref{fig:map_cities_raw} shows the spatial distribution of the 2,390 cities in the original \textit{Princes and Townspeople} data.
+The reporting boundaries are those of the German Empire in 1937 plus the Free City of Danzig; within those boundaries \cite{keyser1939} claim to list "every settlement that at one point in the history of the Holy Roman Empire was awarded the status of a city" \citep[p.
 2]{pt1}.
-The raw \textit{Princes and Townspeople} database contains those 2,390 cities.
-The reporting boundaries are those of the German Empire in 1937 plus the Free City of Danzig.
-The specific datasets I use are those on city locations \citep{pt1}, territorial histories \citep{pt2}, construction activity \citep{pt5} and conflict incidents \citep{pt6}.\footnote{I thank Davide Cantoni and Matthias Weigand for giving me access to unpublished yearly versions of the construction and territorial histories data.}
 
@@
 A city-by-year panel of territorial history, taken from \cite{pt2}, forms the backbone of my dataset.
-For each city in each year it identifies the noble lineage to whose territory the city belonged in that year.
-This is what I
+For each city and each year it identifies the noble lineage to whose territory the city belonged in that year.
+I use this information to identify years in which cities switch from one territory to another.
+The original data also records how each city came into the possession of the lineage ruling it.
+I combine the original 13 types of switches into 3 types: \textit{Conquests}, \textit{successions}, and \textit{other.} I code a switch as a conquest if the original data records it as either "conquest" or "acquisition by conflict", and as a succession if the record says "extinction of lineage" or "inheritance".
+The remaining categories are classified as \textit{other}.\footnote
+{
+The other categories are "no data or regular status", "personal union or marriage", "purchase", "foundation", "secularisation", "decision of the emperor", "gift", "exchange" and "mediatisation".
+}
 
-The most important There are detailed records of the territorial histories of cities (among many other things): we know to which lineage's territory a city belonged to at every time point in which it shows up in the data.
-Records begin in 1300 at the earliest and end at the latest with the abolition of the German monarchy and the foundation of the Weimar Republic in 1918.
+\paragraph{Events.}
 
-- unit of observation: city-year / city-period
-- precision of time points
-- aggregation of years
-- unbalanced panel
-- definition of territory holding
-- definition of switch in period
-- justification for using construction
-- different groupings of construction
-- scope of the data / regions included: 1937 germany
-- time period recorded
-
-\paragraph{Event variables.}
-
-Different types of construction
-
-Additionally, conflict
+Problems with construction -> To get more info out of it: Different types of construction
 
@@
 I code "event" variables as missing for periods that are earlier than the first recorded event or later than the last recorded event in a city.
+
+Additionally, conflict.
 

### 2023-06-25 source/analysis/summarystats/MapCitiesRaw.R
#27 tiny fixes, new map title

@@ -46,7 +46,7 @@ Main <- function(){
                 palette = "Greens",
                 breaks = my.breaks,
                 legend.reverse = TRUE) +
-    tm_layout(title = "Spatial distribution of cities - All cities",
+    tm_layout(title = "Spatial distribution of cities - Full set",
               title.size = 1.3,
               title.position = c("center", "top"),
               title.fontface = 2,

### 2023-06-25 source/build/Switches.R

@@ -50,4 +50,4 @@ Main <- function(){
   return(0)
 }
 
-test <- Main()
+Main()

### 2023-06-25 paper/paper.tex
Updates from Overleaf
#27 sketch events paragraph

@@
 From these sources I assemble a panel of cities.
-The unique unit of observation is the city-period.
+City-periods are the unique units of observation.
 The ultimate dataset contains, in abstract terms, two types of variables: ones that code the \textit{territorial history} of cities, and ones that code the occurrence of certain \textit{events}.
@@
 For each city and each year it identifies the noble lineage to whose territory the city belonged in that year.
-I use this information to identify years in which cities switch from one territory to another.
+I use this information to find years in which cities switch from one territory to another.
 The original data also records how each city came into the possession of the lineage ruling it.
@@
 {
-The other categories are "no data or regular status", "personal union or marriage", "purchase", "foundation", "secularisation", "decision of the emperor", "gift", "exchange" and "mediatisation".
+The other categories are: "no data or regular status", "personal union or marriage", "purchase", "foundation", "secularisation", "decision of the emperor", "gift", "exchange" and "mediatisation".
 }
@@
 \paragraph{Events.}
+
+The data on construction activity \citep{pt5} and conflict incidents \citep{pt6} resemble lists of events. The unit of observation is Each entry contains a
+
+I mainly care about construction events.
 
@@
     \caption{Share of observations from each \textit{Städtebuch} region with non-missing data on conflict incidents.
-The $Conflict$ variable is coded as missing for a period if it is earlier than the first recorded incident or later than the last recorded incident in a city.
+I code the $Conflict$ variable as missing for a period if it is earlier than the first recorded incident or later than the last recorded incident in a city.
 Observations with missing conflict data are excluded from the regressions in Columns (3) and (4) of Table \ref{tab:controls_50y}.}

### 2023-06-25 paper/paper.tex
#25 clarify equation 2 explanation

@@
 I simply replace the lead and lag dummies with the post-treatment indicator $\mathbf{1}(t \ge e^{NewState}_i)$.
-Also, this equation features a vector of city-year level control variables $\mathbf{X}_{it}$.
-In the baseline version of this specification I control only for the dummy $\mathbf{1}(S_{it} > 0)$ that indicates switches in the same period, exactly like in Equation \eqref{eq:sw22}.
+Another minor change is that this equation features a vector of city-year level control variables $\mathbf{X}_{it}$ --- however, in the baseline version of this specification I control only for the dummy $\mathbf{1}(S_{it} > 0)$ that indicates switches in the same period, exactly like in Equation \eqref{eq:sw22}.
 In a robustness check I will additionally control for conflict incidents.

### 2023-06-25 paper/paper.tex
#28 expand sw22 results section

@@
 Figure \ref{fig:SW_replication} plots the coefficients in Column (1) to visualise these dynamics.
+It looks similar to Figure BLANK, taken from \cite{schoenholzer2022}, except for the missing short-run drop.
+Note, however, that the outcome variable in their regression is population, which could be more sensitive to short-run effects --- after all, people may die or leave the city at an increased rate in times of political turmoil and conflict.
 
-By contrast, almost all the coefficients in Columns (2), (3) and (4) are insignificant, suggesting that neither of these specific subsets of construction is driving the results.
+The regression in Column (1) used all construction activity as the outcome, while Columns (2), (3), and (4) use subsets as defined in Section \ref{sec:data}.
+Almost all the coefficients in the last three columns are insignificant, suggesting that neither of these specific subsets of construction is driving the results.
 
@@
 new construction vs repairs (!!!)
+
+super low within-R2
 

### 2023-06-25 paper/paper.tex
#27 mostly finish data section

@@
 I use this information to find years in which cities switch from one territory to another.
+The median city in the original dataset switches territories 5.64 times in its history.
+
 The original data also records how each city came into the possession of the lineage ruling it.
@@
 } 
+Cities are ruled by conquering states 16\% of the time, by succeeding states 34\% of the time, and by states that gained possession of them in another way 50\% of the time.
 
-\paragraph{Events.}
+\paragraph{Construction activity.}
 
-The data on construction activity \citep{pt5} and conflict incidents \citep{pt6} resemble lists of events. The unit of observation is Each entry contains a
+The data on construction activity \citep{pt5} and conflict incidents \citep{pt6} both resemble lists of events.
+The unit of observation is the event, occurring in a specified city at a specified time point.
+Each entry contains additional information about the type of event and the range of uncertainty for the timing.
 
-I mainly care about construction events.
+I will regress construction activity on breaks in the territorial history of cities throughout the main part of this paper --- this is not self-explanatory.
+After all, the research question deals with the effect of switching states on urban economic development.
+I use construction as a proxy for development because it is available for most cities: out of the 2,390 cities in the \textit{Städtebuch}, 2,177 have at least one construction event linked to them.
+By contrast, only 207 cities from the \textit{Städtebuch} appear in \cite{bairoch1988}'s dataset of city populations.
 
-Problems with construction -> To get more info out of it: Different types of construction
+The record of construction events is unfortunately not exhaustive and mostly contains notable buildings --- churches, palaces, fortifications, malls, schools, libraries, opera houses, gardens.
+It contains residential buildings of some noble families and notable citizens but not of the general population.
+Even so, there are theoretical justifications for using construction activity as a proxy for growth: If buildings are a normal good, then demand for them should increase with income.
+While the types of buildings listed do seem to represent state and elite consumption more than general economic activity, the income of these groups is also dependent on the broader economy --- an upturn will increase tax revenues and by extension the disposable income of the state.
+On a basic level we would expect more new buildings, of any type, in cities that are booming (although the error term is likely large).
 
-Since I use some subset of construction as the outcome variable in all regressions, it is necessary to drop missing values.
-I code "event" variables as missing for periods that are earlier than the first recorded event or later than the last recorded event in a city.
+In an attempt to understand more about the effects I estimate, I use certain subsets of construction activity as the outcome in alternative specifications.
+There are 16 types of construction activity in the original data, which I combine into 4 facets: \textit{State construction} consists of palaces, castles, administrative and military buildings.
+\textit{Private construction} includes malls, economic (banks, factories, restaurants, and more) and private (residential) buildings.
+Lastly, \textit{construction of public goods} includes infrastructure, social, cultural, and educational buildings, memorials, gardens and parcs.\footnote
+{
+Churches, monasteries, clerical buildings and town halls are not included in any of the custom subsets of construction activity. They are, however, included in the "All construction" specifications.
+}
 
-Additionally, conflict.
+Because I drop city-period observations with missing construction data from the sample, it is worth briefly noting how I transform a list of events into a city-period panel of construction activity.
+First I group and count events that occur in the same city and period.
+Then I fill the observations in between the events with the value zero.
+Finally, I code values as missing for periods that are earlier than the first recorded event or later than the last recorded event in a city.
+I do not set these out-of-range values to zero because I have no way of knowing whether nothing actually happened, or whether the record simply does not cover these periods anymore.
+
+\paragraph{Conflict incidents.}
+
+I include conflict as a control variable in a robustness check.
+Conflict in my dataset is simply coded as an indicator whether a conflict incident of any type is recorded in this city and period.
+I code missing values exactly as I do for construction events.
 

### 2023-06-25 paper/paper.tex
#28 minimum baseline results

@@
 After 200 years this difference grows to 9 percentage points.
-These estimates are significant at the 5 percent and 1 percent level, respectively.
+These estimates are statistically significant at the 5\% and 1\% level, respectively.
 Figure \ref{fig:SW_replication} plots the coefficients in Column (1) to visualise these dynamics.
-It looks similar to Figure BLANK, taken from \cite{schoenholzer2022}, except for the missing short-run drop.
-Note, however, that the outcome variable in their regression is population, which could be more sensitive to short-run effects --- after all, people may die or leave the city at an increased rate in times of political turmoil and conflict.
 
@@
 Table \ref{tab:baseline_50y} shows estimates from Equation \eqref{eq:baseline}.
-This is a difference-in-differences specification with interactions to capture heterogeneous treatment effects.
+This is a difference-in-differences specification with added interactions to capture heterogeneous treatment effects.
 As before, different columns use different subsets of construction as the outcome variable.
@@
 I report estimates of the coefficient of interest $\theta_{conquest}$ in the second row.
+In Column (1) we can see that the probability of construction activity in a given period is 4 percentage points higher following a conquest vis-a-vis after a succession, but the difference is not statistically significant.
+
+The only subset of construction that shows a significant effect is state construction --- palaces, castles, military and administrative buildings.
+We can see in Column (2) that these are 5 percentage points \textit{less} likely to be built in a city after it was conquered; the effect is significant at the 5\% level.
 

### 2023-06-25 paper/paper.tex
#30 minimal discussion

@@
 
+The hypothesis was that conflict-driven selection for state power boosts urban growth.
+Given that my assumption about the difference in selectivity of conquests and successions holds true, the theory would have predicted a positive coefficient $\theta_{conquest}$.
+The results I obtain instead range from insignificant to negative.
+They offer no support to the hypothesis.
+
 % ######################################################
@@
 
-How robust this null result is to other specifications: I limit post-window to 200 years; I try Sun and Abraham, I use population data; I use different granularity.
+The results in Table \ref{tab:baseline_50y} corroborate part of my hypothesis: switching states within the Holy Roman Empire had a positive effect on the development of cities in the long run.
+However, the second part of my analysis yields a null result.
+I do not find evidence in support of the hypothesis that selection for state power is responsible for the positive effect of switching.
 
-differences within HRE may not have been large enough
+Table \ref{tab:controls_50y} shows estimates of Equation \eqref{eq:baseline} obtained with different sets of control variables.
+The outcome is $All \, construction$ in all regressions.
+Column (1) is identical to Column (1) in Table \ref{tab:baseline_50y} and controls only for switching in the same period.
+In Column (2) I remove this control.
+Columns (3) and (4) add conflict as a control variable, once in addition to switching and once without.
 
-successions may also select for state power
+Comparing Column (2) to (1), and (4) to (3), we can see that it makes no difference whether we control for switches in the same period or not; the estimates are barely distinguishable.
+When we control for conflict, however, the coefficient $\theta_{conquest}$ suddenly turns negative, highly significant and large in magnitude.
+If we trust these regressions, the difference between conquests and successions lowers probability of observing construction in a given year is 15 percentage points lower.
 
-different switch sizes for conquests vs extinctions; maybe larger switch sizes caused neglect (-) or integration (+)
+At first glance this may suggest that conflict is an omitted variable in Column (1), biasing the baseline estimate.
+A closer look at the coefficients on $Conflict$ in Columns (3) and (4) makes this seem unlikely, though: they are very close to zero and not significant by any stretch of the imagination.
+For conflict to be an omitted variable it would need to be correlated both with the post-$Conquest$ dummy and construction.
+The first correlation seems plausible --- once-conquered cities may continue to be highly contested, which would imply a positive relationship, or they may enjoy protection from further attacks by their powerful new owner, which would imply a negative relationship --- but the second correlation is zero in my sample.\footnote
+{
+The coefficients on $Conflict$ in Table \ref{tab:controls_50y} are conditional correlations.
+Simply computing the correlation coefficient between $Conflict$ and $All\,construction$ yields a value of -0.01.
+} The difference cannot be the result of omitted variable bias.
 
-new construction vs repairs (!!!)
+I investigate one alternative explanation: because there are many missing values in the $Conflict$ variable, more than two thirds of observations are dropped between Columns (2) and (3).
+If the distribution of missing values across observations is not random, the different estimates may be due to starkly different samples.
+Figure \ref{fig:conflict_map} visualises how including conflict in the regression affects the distribution of the sample across space.
+The colour of each region codes the share of observations from the baseline sample that remain after dropping observations with missing conflict data.
+If the spatial composition of the sample stayed exactly the same, the entire map would be a uniform shade.
+Instead we immediately see big discrepancies: Thuringia and most northern regions all but disappear from the sample, while the more darkly shaded west and south retain between 30\% and 60\% of their observations.
 
-super low within-R2
+Other things I did:
+\begin{itemize}
+    \item I limit post-window of the diff-in-diff to 200 years like in the event study; null results
+    \item try robust TWFE estimator by Sun and Abraham (2021); null results
+    \item use \cite{bairoch1988}'s population data as outcome; null results
+    \item regress on data aggregated to 1, 10, and 100 years instead; null results
+\end{itemize}
 
-table 3: why neg? conflict NA may be systematocally correlated
+Explanations for no effect:
+\begin{itemize}
+    \item states within the HRE may be too similar / competition between them not large enough
+    \item successions may also select for state power - maybe you get to enter more advantageous marriages if you are the leader of a powerful state
+    \item cities are normally conquered one at a time, while successions usually see many cities switch at the same time - maybe there is a crowding-out of the positive effect, e.g.
+because the ruler likes one of his newly acquired cities most and neglects the others
+    \item the data is just not that well suited to the task; within-R2 is super low in Table \ref{tab:baseline_50y}
 
-are some regions more common in the conflict NAs?
-
-The first explanation is that conflict is an important omitted variable, in the sense that it is correlated both with the treatment and with construction activity.
-
-If omitting  bias in $\theta_{conflict}$ from omitting conflict has the same sign as 
-$$\beta_{conflict} \cdot \frac{\mathbf{Cov}({Conflict},\: {Switching\: to\: another\: state \times Conquest})}{\mathbf{Var}(Conflict)}$$
-
-The whole term is negative.
-We can see in Columns (3) and (4) of Table \ref{tab:controls_50y} that $\beta_{conflict}$ has a positive sign.
-It follows that the correlation between must be negative (which makes sense - an owner with higher state power is likely better able to protect a city from attacks.
-
-But if $\beta_{conflict}$ is around zero, then why does it introduce bias at all?
-
-Conflict is not actually correlated with construction, but controlling for it has the side effect of dropping two thirds of all observations from the sample. This is because
-
-It doesn't make sense. It has to be something about the sample. Johannes must have been spot on.
+\end{itemize}
 

### 2023-06-25 source/analysis/summarystats/BasicStats.R
#27 update interactive summary scripts

@@ -1,18 +1,24 @@
 #!/usr/bin/env Rscript --vanilla
 
 library(tidyverse) |> suppressPackageStartupMessages()
+library(kableExtra) |> suppressPackageStartupMessages()
 
 setwd("~/GitHub/BA")
 
 source("source/utils/HandleCommandArgs.R")
+source("source/utils/DataPrepSuite.R")
+source("source/utils/PrepareBaselineData.R")
 source("source/analysis/summarystats/SummaryLib.R") # functions for analysis
+source("source/utils/PopBuild.R")
+
 
 Main <- function(){
-  # Read data
+  # Read 50-year build
   t <- HandleCommandArgs(default_length = 50)
   read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
   build <- read_csv(read_path, show_col_types = F)
   
+  # Read raw build (without aggregation and dropping NA c_all)
   raw_build <- read_csv("drive/derived/cities_switches.csv",
                         show_col_types = F) |> 
     rename(
@@ -20,18 +26,31 @@ Main <- function(){
       switches = switch) |> 
     mutate(conflict = 0) # just for convenience
   
+  # Produce baseline sample
+  sample <- PrepareBaselineData(build, max_switches=2,
+                                binarise_construction = F, binarise_switches = F) |> 
+    ungroup() # removing grouping is important
+  
+  
+  # Read pop build
+  pop_build <- AssemblePopBuild()
+  
   # Read city locations data for region-level stats
   locs <- read_csv("drive/raw/attributes/city_locations.csv", show_col_types = F) |> 
     select(city_id, region_id, name, nat)
   
-  dsets <- list("build" = build, "raw" = raw_build)
-  dsets |> map(GetBuildLevelInfo)
-  dsets |> map(GetCityLevelInfo)
-  dsets |> map(GetRegionLevelInfo, locs)
+  # Compute summary statistisc
+  dsets <- list("raw" = raw_build, "build" = build, "sample" = sample)
+  build_stats <- dsets |> map(GetBuildLevelInfo) |> UnifyStats(names(dsets))
+  city_stats <- dsets |> map(GetCityLevelInfo) |> UnifyStats(names(dsets))
+  region_stats <- dsets |> map(GetRegionLevelInfo, locs) |> UnifyStats(names(dsets))
 }
 
+# 
+# kbl(build_stats, format="latex",
[…]

### 2023-06-25 source/analysis/summarystats/SummaryLib.R

@@ -3,10 +3,10 @@ GetBuildLevelInfo <- function(build){
     summarise(
       min_year = min(period),
       max_year = max(period),
-      avg_year = mean(period),
       med_year = median(period),
       n_cities = n_distinct(city_id),
-      n_terrs = n_distinct(terr_id)
+      n_terrs = n_distinct(terr_id),
+      n_obs = n()
     )
 }
 
@@ -19,7 +19,10 @@ GetCityLevelInfo <- function(build){
       conflict = sum(conflict),
       across(starts_with("c_"), sum)) |> 
     summarise(
-      across(-c(city_id), \(col)(mean(col, na.rm=T)))
+      across(-c(city_id), list("mean" = \(col)(mean(col, na.rm=T)), 
+                               "median" = \(col)(median(col, na.rm=T))
+                               )
+             )
     )
 }
 
@@ -38,4 +41,11 @@ GetRegionLevelInfo <- function(build, locs){
       across(-c(region_id), mean)
     )
   return(stats)
+}
+
+UnifyStats <- function(stats, names_col){
+  stats |> 
+    bind_rows() |> 
+    add_column(list("build" = names_col), .after = 0) |> 
+    add_column(c(1, 50, 50), .after = 1)
 }

### 2023-06-25 paper/paper.tex
Updates from Overleaf

@@
 \usepackage[utf8]{inputenc}
-\usepackage[left=2cm,right=6cm,top=2.5cm,bottom=2.5cm]{geometry}  
+\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}  
 % set 11pt and right=6cm for submission
@@
 I study this question in the setting of the Holy Roman Empire, using historical data on construction activity and the territorial history of cities.
-I specifically test the hypothesis that competition between states, in which higher-quality states dominated, caused cities to benefit from permanent ownership changes.
-I find evidence in support of this.
+I specifically test the hypothesis that cities benefitted from permanent ownership changes because competition between states selected for higher-quality government.
+While I show that cities grow faster in the long run after switching states, I do not find evidence that conflict-driven selection played a role in this.
 \end{abstract}
@@
 
-The results in Table \ref{tab:baseline_50y} corroborate part of my hypothesis: switching states within the Holy Roman Empire had a positive effect on the development of cities in the long run.
-However, the second part of my analysis yields a null result.
+The results in Table \ref{tab:SW22_replication_50y} corroborate part of my hypothesis: switching states within the Holy Roman Empire had a positive effect on the development of cities in the long run.
+However, the second part of my analysis yields a null result: Table \ref{tab:baseline_50y} shows that the the positive overall effect we saw in Table \ref{tab:SW22_replication_50y} was not driven by conquests.
 I do not find evidence in support of the hypothesis that selection for state power is responsible for the positive effect of switching.
 
+I test the robustness of this result by estimating some alternative specifications.
 Table \ref{tab:controls_50y} shows estimates of Equation \eqref{eq:baseline} obtained with different sets of control variables.

### 2023-06-25 paper/paper.tex

@@
 \usepackage[utf8]{inputenc}
-\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}  
+\usepackage[left=3cm,right=3cm,top=2.5cm,bottom=2.5cm]{geometry}  
 % set 11pt and right=6cm for submission
@@
 \author{Elias Hadj Ammar}
-\date{May 2023}
+\date{June 27, 2023}
 
@@
 I study this question in the setting of the Holy Roman Empire, using historical data on construction activity and the territorial history of cities.
-I specifically test the hypothesis that cities benefitted from permanent ownership changes because competition between states selected for higher-quality government.
+I specifically test the hypothesis that cities benefitted from ownership changes because competition between states selected for higher-quality government.
 While I show that cities grow faster in the long run after switching states, I do not find evidence that conflict-driven selection played a role in this.
@@
 174).
-Wars brought death and devastation to the societies affected by them, but their effects also manifested on the political map --- they play out in a diverse ecosystem of polities that come and go and change size over time.
-Borders get redrawn and territory changes hands.
+Wars bring death and devastation to the societies affected by them, but their effects also manifest on the political map --- they play out in a diverse ecosystem of polities that come and go and change size over time.
+Borders are redrawn and territory changes hands.
 One can indeed imagine this as an evolutionary process: states compete for resources and powerful states eventually absorb weaker ones.
@@
 
-In this paper I study the role of this process in the economic development of Europe.
+In this paper I study the role of this process in the economic development of the Holy Roman Empire.
 In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development.
@@
 
-While I am able to replicate \cite{schoenholzer2022}'s result that the long-run effect of switching to another state is positive, I fail to identify a statistically significant effect of state power on urban development.
+While I am able to replicate \cite{schoenholzer2022}'s result that switching to another state has a positive effect one century later, I fail to identify a statistically significant effect of state power on urban development.
 This does not necessarily mean that conflict-driven evolution played no role in the rise of Europe --- several issues limit the statistical power of my test, which might prevent me from finding an effect even if there is one.
@@
 Each entry contains additional information about the type of event and the range of uncertainty for the timing.
+Because I drop city-period observations with missing construction data from the sample, it is worth briefly outlining how I transform a list of events into a city-period panel of construction activity.
+First I group and count events that occur in the same city and period.
+Then I fill the observations in between the events with the value zero.
+Finally, I code values as missing for periods that are earlier than the first recorded event or later than the last recorded event in a city.
+I do not set these out-of-range values to zero because I have no way of knowing whether nothing actually happened, or whether the record simply does not cover these periods anymore.
 
@@
 
-The record of construction events is unfortunately not exhaustive and mostly contains notable buildings --- churches, palaces, fortifications, malls, schools, libraries, opera houses, gardens.
+The record of construction events is unfortunately not exhaustive and mostly contains notable buildings --- churches, palaces, fortifications, malls, schools, libraries, opera houses, gardens, and the like.
 It contains residential buildings of some noble families and notable citizens but not of the general population.
@@
 While the types of buildings listed do seem to represent state and elite consumption more than general economic activity, the income of these groups is also dependent on the broader economy --- an upturn will increase tax revenues and by extension the disposable income of the state.
-On a basic level we would expect more new buildings, of any type, in cities that are booming (although the error term is likely large).
+On a basic level we would expect more new buildings, of any type, in cities that are booming (although the error term is likely large - we might also expect more new buildings in cities that were devastated by war or environmental catastrophes).
 
@@
 
-Because I drop city-period observations with missing construction data from the sample, it is worth briefly noting how I transform a list of events into a city-period panel of construction activity.
-First I group and count events that occur in the same city and period.
-Then I fill the observations in between the events with the value zero.
-Finally, I code values as missing for periods that are earlier than the first recorded event or later than the last recorded event in a city.
-I do not set these out-of-range values to zero because I have no way of knowing whether nothing actually happened, or whether the record simply does not cover these periods anymore.
-
 \paragraph{Conflict incidents.}
 
-I include conflict as a control variable in a robustness check.
+I include conflict incidents (sieges, plunders, civil unrest) as a control variable in a robustness check.
 Conflict in my dataset is simply coded as an indicator whether a conflict incident of any type is recorded in this city and period.
 I code missing values exactly as I do for construction events.
+
+Note that conflict \textit{incidents} are distinct from conflict as a driver of improvements in institutions.
+Conflict incidents are events recorded at the city level that often involve the destruction of life or property in that city.
+The statement "conflict causes weak states to disappear "
 
@@
 In the aggregated version I include such events if their timing can be assigned with certainty to a 50-year period.
+The
 
-The aggregated indicators for construction and conflict take the value 1 if at least one event was recorded during the period. The switches and rule types
+The switch indicators are aggregated by summing, to obtain the total number of switches that happen in a period, and the aggregated indicators for construction and conflict take the value 1 if at least one event was recorded during the period.
 
@@
 When we control for conflict, however, the coefficient $\theta_{conquest}$ suddenly turns negative, highly significant and large in magnitude.
-If we trust these regressions, the difference between conquests and successions lowers probability of observing construction in a given year is 15 percentage points lower.
+If we trust these regressions, the probability of observing construction in a different year increases by 15 percentage points \textit{less} after conquests relative to successions.
 
@@
 
+I conduct a number of other robustness checks.
+
 Other things I did:
@@
 
-Explanations for no effect:
+Other explanations for no effect:
 \begin{itemize}
@@
     \item the data is just not that well suited to the task; within-R2 is super low in Table \ref{tab:baseline_50y}
+    \item had I found an effect, it would have been difficult to show that
 
@@
 
+\section*{Appendix}
+
 \end{document}

### 2023-06-25 paper/paper.tex
#30 almost finish feedback from Lara + Mama

@@
 resources above subsistence.
-If the most successful states are the ones with the biggest surplus, selection should have made Europe increasingly prosperous.\footnote{The basic idea of Darwinian selection has previously been borrowed by economists - for example \cite{galor2002} do the literal version, where humans are subject to selection and accumulate favourable traits over the course of history.
-More infamously, \cite{clark2007} employs it in an attempt to explain why the Industrial Revolution took place in Europe, arguing that the rich out-bred the poor and passed favourable traits on to their offspring (in Europe but not elsewhere for some reason.)}
+If the most successful states are the ones with the biggest surplus, selection should have made Europe increasingly prosperous.\footnote
+{
+The basic idea of Darwinian selection has previously been borrowed by economists - for example by \cite{galor2002} who argue that humans are subject to selection and accumulate (economically) favourable traits over the course of history.
+More infamously, \cite{clark2007} employs it in an attempt to explain why the Industrial Revolution took place in Europe, arguing that at some point the European rich began to out-breed the poor and passed their favourable traits on to their offspring.
+}
 
@@
 These aristocratic families defined the territorial history of the HRE: territories were split by inheritances, joined by marriages and extinctions, and driven to war by struggles for dominance among the noble lineages ruling them.
-
-When I say "state" in the context of this paper I mean its early modern German equivalent --- the territory of a lineage of rulers, or the lineage itself.
+For this reason, when I say "state" in the context of this paper I mean its early modern German equivalent --- the territory of a lineage of rulers, or the lineage itself.
 
@@
 
-The record of construction events is unfortunately not exhaustive and mostly contains notable buildings --- churches, palaces, fortifications, malls, schools, libraries, opera houses, gardens, and the like.
+The historical record of construction events is unfortunately not exhaustive and mostly contains notable buildings --- churches, palaces, fortifications, malls, schools, libraries, opera houses, gardens, and the like.
 It contains residential buildings of some noble families and notable citizens but not of the general population.
@@
 While the types of buildings listed do seem to represent state and elite consumption more than general economic activity, the income of these groups is also dependent on the broader economy --- an upturn will increase tax revenues and by extension the disposable income of the state.
-On a basic level we would expect more new buildings, of any type, in cities that are booming (although the error term is likely large - we might also expect more new buildings in cities that were devastated by war or environmental catastrophes).
+On a basic level we would expect more new buildings, of any type, in cities that are booming (although the error term is likely large - in the short run we might also expect more new buildings in cities that were destroyed by war or environmental catastrophes).
 
@@
 Conflict incidents are events recorded at the city level that often involve the destruction of life or property in that city.
-The statement "conflict causes weak states to disappear "
+By contrast, in the hypothesis "conflict selects for states with higher-quality institutions" I mean conflict at the macro level, as a process that determines the fates of entire territories.
 
@@
 Of these I drop 16,110 that are not covered by the construction data.
-Finally, the baseline versions of all specifications in this paper restrict the sample to cities with no more than two lifetime switches.
+Finally, following \cite{schoenholzer2022}, for the baseline versions of all specifications I restrict the sample to cities with no more than two lifetime switches.\footnote
+{
+I relax this restriction in a robustness check.
+} 
 This will remove another 9,975 observations, leaving 6,635 in the baseline sample.
@@
     If a city undergoes multiple lasting switches, I create duplicate entries for each different $e^{NewState}_i$.
-I exclude cities that switch states (permanently or not) more than twice over the sample period.
-I report results obtained with different exclusion thresholds in Appendix BLANK.
 }
@@
 The coefficient $\pi_\tau$ for given $\tau$ is the difference in the probability of construction between two cities $i$ and $j$ with identical time-invariant characteristics, in the same period, who undergo the same kind of switching turmoil in this period.
-The only difference is that one city underwent a lasting switch to another state $\tau$ periods ago, while the other did not.
+The only difference is that one city switched permanently to another state $\tau$ periods ago, while the other did not.
 
@@
 In the best case, if the identifying assumption is satisfied, the estimates from the previous regression tell us what the effect of switching states is --- however, they cannot tell us why.
-In the following I attempt to identify the effect of state power, or more accurately, the effect of the difference in selection for state power between conquests and successions.
+In the following I attempt to estimate the effect of state power, or more accurately, the effect of the difference in selection for state power between conquests and successions, on construction activity.
 
@@
 
-For clarity I move from the event study to a basic difference-in-differences specification; this way there is only a single coefficient of interest.
+For clarity I have moved from the event study to a basic difference-in-differences specification; this way there is only a single coefficient of interest.
 I simply replace the lead and lag dummies with the post-treatment indicator $\mathbf{1}(t \ge e^{NewState}_i)$.
@@
 If this assumption holds true, the effect of state power should manifest itself in the difference between the trend of cities acquired by conquest and the trend of cities acquired by succession.
-Note that successions are the omitted category --- there is no interaction term between the post-treatment indicator and $\mathbf{1}(Succession_{it})$ in Equation \eqref{eq:baseline} --- so that the coefficient $\theta_{conquest}$ captures exactly what we want: the difference in trends after conquests vis-a-vis successions.
+Note that successions are the omitted category --- there is no interaction term between the post-treatment indicator and $\mathbf{1}(Succession_{it})$ in Equation \eqref{eq:baseline} --- so that the coefficient $\theta_{conquest}$ captures exactly what I want to estimate: the difference in trends after conquests vis-a-vis successions.
 
-The econometric assumption necessary to allow causal interpretation of $\theta_{conquest}$ is a different version of the parallel trends assumption.
+The econometric assumption that must hold true for $\theta_{conquest}$ to be a causal effect is a different version of the parallel trends assumption.
 In particular, it must hold that, had a conquered city i counterfactually been acquired by succession rather than conquest, it would have trended the same as the other cities acquired by succession.
@@
 
-The regression in Column (1) used all construction activity as the outcome, while Columns (2), (3), and (4) use subsets as defined in Section \ref{sec:data}.
-Almost all the coefficients in the last three columns are insignificant, suggesting that neither of these specific subsets of construction is driving the results.
+The regression in Column (1) uses all construction as the outcome, while Columns (2), (3), and (4) use the subsets of construction activity defined in Section \ref{sec:data}.
+Almost all the coefficients in the last three columns are insignificant, suggesting that neither of these specific subsets of construction is driving the overall results.
+Curiously, though, 50 years after switching there is a small, transient, but significant increase in construction of public goods.
+Since many of these would have been provided by the state, an explanation could be that rulers are initially reluctant to invest into their new territory --- the coefficient for the 50 years immediately following the switch ($\tau = 0$) is insignificant but negative --- but catch up once their rule is established.
 
@@
 When we control for conflict, however, the coefficient $\theta_{conquest}$ suddenly turns negative, highly significant and large in magnitude.
-If we trust these regressions, the probability of observing construction in a different year increases by 15 percentage points \textit{less} after conquests relative to successions.
+If we trust these regressions, the probability of observing construction in a given year increases by 15 percentage points \textit{less} after conquests relative to successions.
 
@@
     \item the data is just not that well suited to the task; within-R2 is super low in Table \ref{tab:baseline_50y}
-    \item had I found an effect, it would have been difficult to show that
+\end{itemize}
 
+Improvements / future research:
+\begin{itemize}
+    \item differences in timing - do early switchers in my sample experience different effects than late switchers?
+    \item lag controls because construction takes a while and the years probably record completion
 \end{itemize}

### 2023-06-25 paper/paper.tex
#30 expand on the time horizon

@@
 City-periods are the unique units of observation.
-The ultimate dataset contains, in abstract terms, two types of variables: ones that code the \textit{territorial history} of cities, and ones that code the occurrence of certain \textit{events}.
-Records for each city begin in the 14th century at the earliest and end at the latest with the abolition of the German monarchy in the early 20th century.
+The ultimate dataset contains, in abstract terms, two types of variables: ones that code the territorial history of cities, and ones that code the incidence of construction and conflict events.
+Records for each city begin in 1300 at the earliest and end at the latest in 1800 (after which no construction events are recorded).
+The median city is observed over a time span of 290 years.
 
@@
 
-The historical record of construction events is unfortunately not exhaustive and mostly contains notable buildings --- churches, palaces, fortifications, malls, schools, libraries, opera houses, gardens, and the like.
+The historical record of construction events is unfortunately not exhaustive and mostly contains notable buildings --- churches, palaces, castles, malls, schools, libraries, opera houses, gardens, and the like.
 It contains residential buildings of some noble families and notable citizens but not of the general population.

### 2023-06-26 paper/paper.tex
#30 finish review of v0.9(.1)
#29 commit before big changes
#29 clarify exposition of results

@@
 As before, different columns use different subsets of construction as the outcome variable.
+Since successions are omitted from the interaction terms, the coefficient $\pi$ on the post-treatment dummy $Switch\,to\,another\,state$ captures the treatment effect of switching states via succession.
+The interaction terms are relative to that: the coefficient $\theta_{conquest}$ on $Switch\,to\,another\,state \times Conquest$ is the difference between the treatment effect of switching via succession, and the treatment effect of switching via conquest.
 
-I report estimates of the coefficient of interest $\theta_{conquest}$ in the second row.
-In Column (1) we can see that the probability of construction activity in a given period is 4 percentage points higher following a conquest vis-a-vis after a succession, but the difference is not statistically significant.
+In Column (1) we can see that the probability of construction activity in a given period is 4 percentage points higher following a conquest vis-a-vis following a succession.
+The treatment effect of a conquest (compared to no treatment) is $0.02 + 0.04 = 0.06$, while the treatment effect of a succession is $0.02$.
+Crucially, though, the difference between them is not statistically significant.
 
-The only subset of construction that shows a significant effect is state construction --- palaces, castles, military and administrative buildings.
+The only subset of construction for which the difference in trends is significant is state construction --- palaces, castles, military and administrative buildings.
 We can see in Column (2) that these are 5 percentage points \textit{less} likely to be built in a city after it was conquered; the effect is significant at the 5\% level.
+BLANK.
 
@@
 Given that my assumption about the difference in selectivity of conquests and successions holds true, the theory would have predicted a positive coefficient $\theta_{conquest}$.
-The results I obtain instead range from insignificant to negative.
+The estimates I obtain instead range from insignificant to negative.
 They offer no support to the hypothesis.
@@
 
-I test the robustness of this result by estimating some alternative specifications.
-Table \ref{tab:controls_50y} shows estimates of Equation \eqref{eq:baseline} obtained with different sets of control variables.
+I test the robustness of this second result by estimating Equation \eqref{eq:baseline} with different control variables $\textbf{X}_{it}$.
+We can see in Table \ref{tab:controls_50y} shows estimates of  obtained with different sets of control variables.
 The outcome is $All \, construction$ in all regressions.

### 2023-06-26 paper/paper.tex
#29 remove repetition

@@
 Crucially, though, the difference between them is not statistically significant.
+Given that my assumption about the difference in selectivity of conquests and successions holds true, we would have expected the treatment effect of conquests to be significantly higher.
 
-The only subset of construction for which the difference in trends is significant is state construction --- palaces, castles, military and administrative buildings.
+The only subset of construction for which the difference in trends is indeed significant is state construction --- palaces, castles, military and administrative buildings.
 We can see in Column (2) that these are 5 percentage points \textit{less} likely to be built in a city after it was conquered; the effect is significant at the 5\% level.
 BLANK.
-
-The hypothesis was that conflict-driven selection for state power boosts urban growth.
-Given that my assumption about the difference in selectivity of conquests and successions holds true, the theory would have predicted a positive coefficient $\theta_{conquest}$.
-The estimates I obtain instead range from insignificant to negative.
-They offer no support to the hypothesis.
 
@@
 The results in Table \ref{tab:SW22_replication_50y} corroborate part of my hypothesis: switching states within the Holy Roman Empire had a positive effect on the development of cities in the long run.
-However, the second part of my analysis yields a null result: Table \ref{tab:baseline_50y} shows that the the positive overall effect we saw in Table \ref{tab:SW22_replication_50y} was not driven by conquests.
+However, the second part of my analysis yields a null result: Table \ref{tab:baseline_50y} shows that the the positive overall effect we saw in Table \ref{tab:SW22_replication_50y} was not in fact driven by conquests.
 I do not find evidence in support of the hypothesis that selection for state power is responsible for the positive effect of switching.

### 2023-06-26 paper/paper.tex
#29 tighten existing discussion

@@
 \title{Conflict Driven Evolution and Urban Development: \\ Evidence from the Holy Roman Empire\footnote{
-I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.
-I thank Matthias Weigand in particular for providing me with his own data and for pointing me towards the paper by \cite{schoenholzer2022}.
+I thank Lara Andermann, Leonie Bielefeld, Daniel Bursian, Davide Cantoni, Emilio Esguerra, Filip Milojevic, Johannes Vorbach and Matthias Weigand for helpful comments and feedback.
 }
@@
 The results in Table \ref{tab:SW22_replication_50y} corroborate part of my hypothesis: switching states within the Holy Roman Empire had a positive effect on the development of cities in the long run.
-However, the second part of my analysis yields a null result: Table \ref{tab:baseline_50y} shows that the the positive overall effect we saw in Table \ref{tab:SW22_replication_50y} was not in fact driven by conquests.
-I do not find evidence in support of the hypothesis that selection for state power is responsible for the positive effect of switching.
+However, the null result in the rest of the analysis fails to convince us that conquests drove the positive overall effect.
+I find weak evidence at best for the hypothesis that selection for state power played a role in the economic development of European cities.
+In this section I check whether the null result in Table \ref{tab:baseline_50y} is robust to alternative specifications and discuss its meaning.
 
-I test the robustness of this second result by estimating Equation \eqref{eq:baseline} with different control variables $\textbf{X}_{it}$.
-We can see in Table \ref{tab:controls_50y} shows estimates of  obtained with different sets of control variables.
-The outcome is $All \, construction$ in all regressions.
+As a first check I include different controls $\textbf{X}_{it}$ in Equation \eqref{eq:baseline}.
+Table \ref{tab:controls_50y} shows the results of this exercise.
 Column (1) is identical to Column (1) in Table \ref{tab:baseline_50y} and controls only for switching in the same period.
 In Column (2) I remove this control.
-Columns (3) and (4) add conflict as a control variable, once in addition to switching and once without.
+Columns (3) and (4) add conflict incidents as a control variable, once in addition to switching and once without.
+The outcome is $All \, construction$ in all regressions.
 
 Comparing Column (2) to (1), and (4) to (3), we can see that it makes no difference whether we control for switches in the same period or not; the estimates are barely distinguishable.
-When we control for conflict, however, the coefficient $\theta_{conquest}$ suddenly turns negative, highly significant and large in magnitude.
-If we trust these regressions, the probability of observing construction in a given year increases by 15 percentage points \textit{less} after conquests relative to successions.
+When we control for conflict incidents, however, the coefficient $\theta_{conquest}$ suddenly turns negative, highly significant and large in magnitude.
+After a conquest the probability of observing construction in a given period now increases by 15 percentage points \textit{less} than after a succession.
 
-At first glance this may suggest that conflict is an omitted variable in Column (1), biasing the baseline estimate.
-A closer look at the coefficients on $Conflict$ in Columns (3) and (4) makes this seem unlikely, though: they are very close to zero and not significant by any stretch of the imagination.
-For conflict to be an omitted variable it would need to be correlated both with the post-$Conquest$ dummy and construction.
-The first correlation seems plausible --- once-conquered cities may continue to be highly contested, which would imply a positive relationship, or they may enjoy protection from further attacks by their powerful new owner, which would imply a negative relationship --- but the second correlation is zero in my sample.\footnote
+While at first this result may seem like evidence for omitted variable bias in the baseline estimate --- the "correct" estimate being the negative one in Column (3) --- a closer look at the coefficients on $Conflict$ makes this seem unlikely: they are close to zero and not significant by any stretch of the imagination.
+To cause bias conflict incidents would need to be correlated both with the post-$Conquest$ dummy and with construction, but the correlation with construction is empirically zero.\footnote
 {
 The coefficients on $Conflict$ in Table \ref{tab:controls_50y} are conditional correlations.
-Simply computing the correlation coefficient between $Conflict$ and $All\,construction$ yields a value of -0.01.
-} The difference cannot be the result of omitted variable bias.
-
-I investigate one alternative explanation: because there are many missing values in the $Conflict$ variable, more than two thirds of observations are dropped between Columns (2) and (3).
-If the distribution of missing values across observations is not random, the different estimates may be due to starkly different samples.
-Figure \ref{fig:conflict_map} visualises how including conflict in the regression affects the distribution of the sample across space.
-The colour of each region codes the share of observations from the baseline sample that remain after dropping observations with missing conflict data.
+Simply computing the correlation coefficient between $Conflict$ and $All\,construction$ yields a value of $-0.01$.
+} 
+Rather, I suggest that the coefficients are different because they are estimated on different samples: more than two thirds of observations are lost between Columns (2) and (3) because of missing values in the $Conflict$ variable.
+Figure \ref{fig:conflict_map} shows that the remaining observations follow quite a different distribution across space: The colour of each region codes the share of observations from the baseline sample that are included in the conflict sample.
 If the spatial composition of the sample stayed exactly the same, the entire map would be a uniform shade.
 Instead we immediately see big discrepancies: Thuringia and most northern regions all but disappear from the sample, while the more darkly shaded west and south retain between 30\% and 60\% of their observations.
+
+If the difference between the coefficients in (1) and (3) is due to the different spatial distribution of the sample, that has the interesting implication of spatially heterogeneous effects. BLANK.
 

### 2023-06-26 paper/paper.tex
#29 finish spatial heterogeneity part

@@
 \cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled the spread of innovation in medieval China would not have survived long in a more fragmented state system.
-Fragmentation also plays a central role in \cite{cervellati2022}'s theory.
+Geography and fragmentation also play central roles in \cite{cervellati2022}'s theory of European state history.
 
@@
 However, they do not show explicitly how these improvements in state quality are driven by selection through conflict.
-They use a dataset with a bigger geographical scope (all of Europe) but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is only available in intervals of 50 or 100 years.
+They use a dataset with a bigger geographical scope (all of Europe) but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is available for fewer cities and fewer years.
 
@@
 
-While at first this result may seem like evidence for omitted variable bias in the baseline estimate --- the "correct" estimate being the negative one in Column (3) --- a closer look at the coefficients on $Conflict$ makes this seem unlikely: they are close to zero and not significant by any stretch of the imagination.
-To cause bias conflict incidents would need to be correlated both with the post-$Conquest$ dummy and with construction, but the correlation with construction is empirically zero.\footnote
+While this result may seem like evidence for omitted variable bias in the baseline estimate, that explanation is ultimately not plausible.
+To cause bias in Column (1) conflict incidents would need to be correlated both with the post-$Conquest$ dummy and with construction --- however, we can see that the latter is not the case: the coefficients on $Conflict$ in Columns (3) and (4) are tiny and not significant by any stretch of the imagination.\footnote
 {
 The coefficients on $Conflict$ in Table \ref{tab:controls_50y} are conditional correlations.
-Simply computing the correlation coefficient between $Conflict$ and $All\,construction$ yields a value of $-0.01$.
+Simply computing the correlation coefficient between the dummies for conflict incidents and for construction yields a value of $-0.01$.
 } 
-Rather, I suggest that the coefficients are different because they are estimated on different samples: more than two thirds of observations are lost between Columns (2) and (3) because of missing values in the $Conflict$ variable.
-Figure \ref{fig:conflict_map} shows that the remaining observations follow quite a different distribution across space: The colour of each region codes the share of observations from the baseline sample that are included in the conflict sample.
-If the spatial composition of the sample stayed exactly the same, the entire map would be a uniform shade.
-Instead we immediately see big discrepancies: Thuringia and most northern regions all but disappear from the sample, while the more darkly shaded west and south retain between 30\% and 60\% of their observations.
+Rather, I suggest that the estimates for $\theta_{conquest}$ are different because they are estimated on different samples.
+More than two thirds of observations are lost between Columns (2) and (3) because of missing values in the $Conflict$ variable.
+Figure \ref{fig:conflict_map} shows that the remaining observations follow quite a different distribution across space --- if the spatial composition of the smaller sample was the same, the entire map would be a uniform shade.
+Instead we see big discrepancies: the west, the south, and Brandenburg retain 30\% to 60\% of their observations, while many northern and eastern regions all but vanish from the sample.
 
-If the difference between the coefficients in (1) and (3) is due to the different spatial distribution of the sample, that has the interesting implication of spatially heterogeneous effects. BLANK.
+If we accept that the difference between the $\theta_{conquest}$ coefficients in (1) and (3) is due to the different spatial distribution of the sample, that has the interesting implication of spatially heterogeneous treatment effects.
+The null effect in (1) could be an average of the negative effect in the southwest and a hypothetical positive effect in the northeast.
+Outsiders could play a role, in a similar vein to \cite{levine2021, levine2022}'s theory --- the south and the west are bordered by the Habsburg Empire and France, two stable major powers, while the northeast is somewhat sheltered by the North Sea, the Baltic Sea, and the territory that comprises 1937 Poland, which was fragmented throughout much of its history.\footnote
+{
+Of course, England and Sweden would attack via the North and Baltic seas, but it seems a reasonable assumption that competition for territory would be more intense between states that share a land border.
+} 
+\cite{cervellati2022} BLANK. I leave this for future research to explore.
 

### 2023-06-26 paper/paper.tex
#29 fix typo + tighten in data part

@@
 In the aggregated version I include such events if their timing can be assigned with certainty to a 50-year period.
-The
 
@@
 
-The raw \textit{Territorial Histories} dataset starts out with 1,471,934 yearly observations.
-I lose two observations per city --- the first and the last year --- while coding switches, for a total of 4,780.
-I aggregate the remaining 1,467,154 yearly observations into 30,980 observations with 50-year spacing.
+The raw \textit{Territorial Histories} dataset starts out with 1,471,934 yearly observations, of which I lose a small number while coding switches.
+I aggregate the rest into 30,980 observations with 50-year spacing.
 Of these I drop 16,110 that are not covered by the construction data.
@@
 } 
-This will remove another 9,975 observations, leaving 6,635 in the baseline sample.
+This removes another 9,975 observations, leaving 6,635 in the baseline sample.
 

### 2023-06-26 paper/output/regressions/robustness_50y.tex
#29 add more robustness table

\begin{table}[htbp]
   \caption{\label{tab:robustness_50y} Heterogeneous effects of switching: More robustness}
   \centering
   \begin{tabular}{lcccc}
      \tabularnewline \midrule \midrule
      Dependent Variable: & \multicolumn{4}{c}{All construction}\\
                                                 & Baseline & 4 Switches   & All switches & Window \\   
      Model:                                     & (1)      & (2)          & (3)          & (4)\\  
      \midrule
      \emph{Variables}\\
      Switch to another state                    & 0.02     & 0.04$^{***}$ & 0.02$^{***}$ & 0.02\\   
                                                 & (0.03)   & (0.01)       & (0.01)       & (0.02)\\   
      Switch to another state $\times$ Conquest  & 0.04     & 0.03         & -0.01        & 0.04\\   
                                                 & (0.04)   & (0.02)       & (0.01)       & (0.04)\\   
      Switch to another state $\times$ Other     & 0.01     & -0.02        & 0.00         & 0.01\\   
                                                 & (0.03)   & (0.01)       & (0.01)       & (0.03)\\   
      Switching indicator                        & -0.01    & -0.02        & -0.01        & -0.01\\   
                                                 & (0.02)   & (0.02)       & (0.01)       & (0.02)\\   
      \midrule
      \emph{Fixed-effects}\\
      City                                       & Yes      & Yes          & Yes          & Yes\\  
      Period                                     & Yes      & Yes          & Yes          & Yes\\  
      \midrule
      \emph{Fit statistics}\\
      Observations                               & 6,635    & 12,979       & 31,754       & 6,635\\  
      R$^2$                                      & 0.3553   & 0.3664       & 0.3518       & 0.3553\\  
      Within R$^2$                               & 0.0009   & 0.0019       & 0.0003       & 0.0009\\  
      \midrule \midrule

   \end{tabular}

   \par \raggedright 
   Note: Table presents results of estimation equation \eqref{eq:baseline}. Column (1) uses the baseline sample, Column (2) includes cities with up to 4 lifetime switches, Column (3) includes all cities regardless of the number of lifetime switches, and Column (4) uses a post-treatment indicator that is only active in the first 200 years after the treatment. Dependent variable is "All construction" in  all regressions.The switch type "Succession" is omitted as the reference category. Yearly data was aggregated into periods of 50 years. Observations are at the city-period level. Standard errors are  clustered at the city level. *, **, and *** denote significance on the 10 percent, 5 percent, and 1 percent  level, respectively.
\end{table}

### 2023-06-26 source/analysis/robustness/Robustness.R

@@ -1,0 +1,104 @@
+#!/usr/bin/env Rscript --vanilla
+
+library(tidyverse) |> suppressPackageStartupMessages()
+library(fixest) |> suppressPackageStartupMessages()
+
+setwd("~/GitHub/BA")
+
+source("source/utils/HandleCommandArgs.R")
+source("source/utils/DataPrepSuite.R")
+source("source/utils/PrepareBaselineData.R")
+source("source/utils/PrepareWindowData.R")
+source("source/utils/etableDefaults.R")
+
+Main <- function(){
+  
+  # Read data
+  t <- HandleCommandArgs(default_length = 50)
+  read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
+  build <- read_csv(read_path, show_col_types = F)
+  
+  # Select different samples and process data
+  dat_baseline <- PrepareBaselineData(build,
+                     max_switches = 2,
+                     binarise_construction = T,
+                     binarise_switches = T)
+  
+  dat_4switches <- PrepareBaselineData(build,
+                     max_switches = 4,
+                     binarise_construction = T,
+                     binarise_switches = T)
+  
+  dat_allswitches <- PrepareBaselineData(build,
+                     max_switches = 100,
+                     binarise_construction = T,
+                     binarise_switches = T)
+
+  dat_window <- PrepareWindowData(build,
+                     max_switches = 2,
+                     years_post = 200,
+                     binarise_construction = T,
+                     binarise_switches = T)
+  
+  # Estimate different specifications
+  mod_baseline <- fixest::feols(
+    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat_baseline)
+  
+  mod_4switches <- fixest::feols(
+    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat_4switches)
+  
+  mod_allswitches <- fixest::feols(
+    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat_allswitches)
+  
+  mod_window <- fixest::feols(
+    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat_window)
+  
+  # Produce regression table and export to LaTeX
+  etableDefaults()
+  
[…]

### 2023-06-26 source/utils/PrepareWindowData.R

@@ -1,0 +1,41 @@
+PrepareWindowData <- function(build, max_switches, years_post,
+                        binarise_construction, binarise_switches){
+  
+  max_switches <- 2
+  years_post <- 200
+  
+  # Filter out cities with too many lifetime switches
+  selected <- FilterBySwitches(build, max_switches)
+  
+  # Identify across-period switches
+  with_e_dummies <- AddEAnother(selected)
+  
+  # Add TREAT x POST dummies D
+  with_D <- AddTreatXPost(with_e_dummies)
+  
+  # Binarise construction outcomes, if specified
+  if (binarise_construction) { with_D <- with_D |> BinariseOutcomes() }
+  
+  # Binarise no. of switches (control), if specified
+  if (binarise_switches) { with_D <- with_D |> BinariseSwitches() }
+  
+  # Process type change into treatment type
+  with_types <- ProcessTypeChange(with_D)
+  
+  # Set D to 0 if treatment was more than 200 years ago
+  with_window <- with_types |>
+    mutate(D = case_when(
+      period <= treat_time + years_post ~ D,
+      period > treat_time + years_post ~ 0
+    )) |> 
+    replace_na(list(D = 0))
+  
+  # Rearrange columns (for readability)
+  clean <- with_window |> select(
+    city_id, treat_time, period, terr_id, switches, e_another, D, 
+    rule_conquest, rule_succession, rule_other,
+    conflict, c_all, c_state, c_private, c_public
+  )
+  
+  return(clean)
+}

### 2023-06-26 source/analysis/robustness/Robustness.R
#29 cut caption of table 4

@@ -64,8 +64,7 @@ Main <- function(){
   \\eqref{eq:baseline}. Column (1) uses the baseline sample, Column (2) includes cities
   with up to 4 lifetime switches, Column (3) includes all cities regardless of the number
   of lifetime switches, and Column (4) uses a post-treatment indicator that is only active
-  in the first 200 years after the treatment. Dependent variable is \"All construction\" in 
-  all regressions.The switch type \"Succession\" is omitted as the reference category.", 
+  in the first 200 years after the treatment.", 
   PeriodInsert(t), "Observations are at the city-period level. Standard errors are 
   clustered at the city level.", SignifInsert(), sep = " ") |> 
     str_replace_all("\n ", "")

### 2023-06-26 paper/output/regressions/robustness_50y.tex
#29 cut caption of table 4
#29 fix typo in header

@@
       Dependent Variable: & \multicolumn{4}{c}{All construction}\\
-                                                 & Baseline & 4 Switches   & All switches & Window \\   
+                                                 & Baseline & 4 switches   & All switches & Window \\   
       Model:                                     & (1)      & (2)          & (3)          & (4)\\  
@@
 Column (1) uses the baseline sample, Column (2) includes cities with up to 4 lifetime switches, Column (3) includes all cities regardless of the number of lifetime switches, and Column (4) uses a post-treatment indicator that is only active in the first 200 years after the treatment.
-Dependent variable is "All construction" in  all regressions.The switch type "Succession" is omitted as the reference category.
 Yearly data was aggregated into periods of 50 years.

### 2023-06-26 source/analysis/robustness/Robustness.R

@@ -69,7 +69,7 @@ Main <- function(){
   clustered at the city level.", SignifInsert(), sep = " ") |> 
     str_replace_all("\n ", "")
   
-  header <- c("Baseline", "4 Switches", "All switches", "Window")
+  header <- c("Baseline", "4 switches", "All switches", "Window")
   
   tex_output <- etable(mod_baseline, mod_4switches, 
                        mod_allswitches, mod_window,

### 2023-06-26 paper/paper.tex
#29 include extra table
#29 appendix numbering
#29 finish More robustness

@@
 {
+This is to avoid creating an excessive amount of duplicate observations.
+Allowing more switches skews the sample towards cities that switch often.
+Keeping all cities and using only e.g.
+the first switch for each city skews it in favour of early switches.
 I relax this restriction in a robustness check.
@@
 While this result may seem like evidence for omitted variable bias in the baseline estimate, that explanation is ultimately not plausible.
-To cause bias in Column (1) conflict incidents would need to be correlated both with the post-$Conquest$ dummy and with construction --- however, we can see that the latter is not the case: the coefficients on $Conflict$ in Columns (3) and (4) are tiny and not significant by any stretch of the imagination.\footnote
+To cause bias in Column (1) conflict incidents would need to be correlated both with the post-conquest dummy and with construction --- however, we can see that the latter is not the case: the coefficients on $Conflict$ in Columns (3) and (4) are tiny and not significant by any stretch of the imagination.\footnote
 {
@@
 The null effect in (1) could be an average of the negative effect in the southwest and a hypothetical positive effect in the northeast.
-Outsiders could play a role, in a similar vein to \cite{levine2021, levine2022}'s theory --- the south and the west are bordered by the Habsburg Empire and France, two stable major powers, while the northeast is somewhat sheltered by the North Sea, the Baltic Sea, and the territory that comprises 1937 Poland, which was fragmented throughout much of its history.\footnote
+Outsiders could play a role, in a similar vein to \cite{levine2021, levine2022}'s theory: the south and the west are bordered by the Habsburg Empire and France, two stable major powers, while the northeast is somewhat sheltered by the North Sea, the Baltic Sea, and the territory that comprises 1937 Poland, which was fragmented throughout much of its history.\footnote
 {
@@
 } 
-\cite{cervellati2022} BLANK. I leave this for future research to explore.
+BLANK \cite{cervellati2022}. I leave this for future research to explore.
 
-I conduct a number of other robustness checks.
+I conduct a number of other robustness checks, which are summarised in the Appendix.
+Table \ref{tab:robustness_50y} contains regressions using different definitions of the sample and treatment.
+Columns (2) and (3) use larger samples that include cities with a higher number of lifetime switches.
+Column (4) shows the results of a regression in which the post-treatment dummy reverts back to 0 after 200 years.
+With this I intend to replicate the same event window as in Equation \eqref{eq:sw22} and to zero in on the effect over the first two centuries following a switch.
+While increasing the cutoff for lifetime switches increases the significance of the overall treatment effect (likely because the sample contains many duplicates of cities that switch often), none of these alternative specifications affect the estimates of the interactions.
 
@@
 
+\clearpage
 \section*{Appendix}
 
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+
+\input{paper/output/regressions/robustness_50y}
+
 \end{document}

### 2023-06-26 paper/output/regressions/robustness_50y.tex
#29 change window header

@@
       Dependent Variable: & \multicolumn{4}{c}{All construction}\\
-                                                 & Baseline & 4 switches   & All switches & Window \\   
+                                                 & Baseline & 4 switches   & All switches & 200 years \\   
       Model:                                     & (1)      & (2)          & (3)          & (4)\\  

### 2023-06-26 source/analysis/robustness/Robustness.R

@@ -69,7 +69,7 @@ Main <- function(){
   clustered at the city level.", SignifInsert(), sep = " ") |> 
     str_replace_all("\n ", "")
   
-  header <- c("Baseline", "4 switches", "All switches", "Window")
+  header <- c("Baseline", "4 switches", "All switches", "200 years")
   
   tex_output <- etable(mod_baseline, mod_4switches, 
                        mod_allswitches, mod_window,

### 2023-06-26 paper/paper.tex
#29 remove additional robustness bullets for now

@@
 
-Other things I did:
-\begin{itemize}
-    \item I limit post-window of the diff-in-diff to 200 years like in the event study; null results
-    \item try robust TWFE estimator by Sun and Abraham (2021); null results
-    \item use \cite{bairoch1988}'s population data as outcome; null results
-    \item regress on data aggregated to 1, 10, and 100 years instead; null results
-\end{itemize}
-
 Other explanations for no effect:

### 2023-06-26 paper/paper.tex
#29 add pretty scene change macro

@@
 \usepackage{booktabs}
+
+\newcommand{\decoline}{% pretty "scene change" macro
+  \par
+  \vspace{\baselineskip}
+  \hfill
+  {\Large*}
+  \hfill
+  \vspace{\baselineskip}
+\par}
 
@@
 
-If we accept that the difference between the $\theta_{conquest}$ coefficients in (1) and (3) is due to the different spatial distribution of the sample, that has the interesting implication of spatially heterogeneous treatment effects.
-The null effect in (1) could be an average of the negative effect in the southwest and a hypothetical positive effect in the northeast.
+If we accept that the disagreement between the interaction coefficients in (1) and (3) is due to the different spatial distribution of the sample, that has the interesting implication of spatially heterogeneous differences in treatment effects.
+The null difference in (1) could be an average of the negative difference in the southwest and a hypothetical positive difference in the northeast.
 Outsiders could play a role, in a similar vein to \cite{levine2021, levine2022}'s theory: the south and the west are bordered by the Habsburg Empire and France, two stable major powers, while the northeast is somewhat sheltered by the North Sea, the Baltic Sea, and the territory that comprises 1937 Poland, which was fragmented throughout much of its history.\footnote
@@
 With this I intend to replicate the same event window as in Equation \eqref{eq:sw22} and to zero in on the effect over the first two centuries following a switch.
-While increasing the cutoff for lifetime switches increases the significance of the overall treatment effect (likely because the sample contains many duplicates of cities that switch often), none of these alternative specifications affect the estimates of the interactions.
+While increasing the cutoff for lifetime switches increases the significance of the overall treatment effect (likely because the sample contains many duplicates of cities that switch often), none of these alternative specifications give the coefficient we care about statistical significance.
+
+\decoline
+
+To sum these results up, I have found fairly robust non-evidence for the hypothesis that selection for state power drove city growth in the Holy Roman Empire between 1300 and 1800.
 

### 2023-06-26 paper/paper.tex
#29 discuss attenuation bias

@@
 \usepackage[utf8]{inputenc}
-\usepackage[left=3cm,right=3cm,top=2.5cm,bottom=2.5cm]{geometry}  
+\usepackage[left=2cm,right=6cm,top=2.5cm,bottom=2.5cm]{geometry}  
 % set 11pt and right=6cm for submission
@@
   \par
-  \vspace{\baselineskip}
+  \vspace{0.33\baselineskip}
   \hfill
-  {\Large*}
+  {\large* * *}
   \hfill
-  \vspace{\baselineskip}
+  \vspace{0.33\baselineskip}
 \par}
@@
 439).
-It is worth highlighting how the authors model war between states: you win with state power, and state power is simply free resources, i.e.
-resources above subsistence.
+It is worth highlighting how the authors model war between states: you win them with state power, and state power is simply aggregate production in excess of the subsistence level.
 If the most successful states are the ones with the biggest surplus, selection should have made Europe increasingly prosperous.\footnote
@@
 
-The closest relative to this paper is \cite{schoenholzer2022} who investigate how switching between states affects the development of cities.
+The closest relative of this paper is \cite{schoenholzer2022} who investigate how switching between states affects the development of cities.
 They find that switching states affects population negatively in the short run but positively in the long run.
@@
 
-To sum these results up, I have found fairly robust non-evidence for the hypothesis that selection for state power drove city growth in the Holy Roman Empire between 1300 and 1800.
+To summarize these results: I find robust non-evidence for the hypothesis that selection for state power drove city growth in the Holy Roman Empire between 1300 and 1800.
+The question is, can we trust this non-finding?
+
+The key economic assumption underlying the identification strategy for the effect of state power was that states that acquire a city by conquest are more powerful on average than states that acquire a city by succession.
+If this assumption does not hold, then the difference between the treatment effect of a conquest and the treatment effect of a succession is just that: a heterogeneity with no obvious meaning.
+
+I justify this assumption with the fact that a conqueror is, by definition, the winner of an armed conflict.
+State power is not the only determinant of winning armed conflicts, though: luck plays a central role, as do many other time-varying factors for which I cannot control.
+If the correlation between state power differentials (the independent variable whose effects we want to measure) and winning conflicts (how we measure it) is small then we essentially have measurement error in the independent variable.
+This leads to attenuation bias.
+The estimate for $\theta_conquest$ has the expected positive sign in my baseline regression --- it just lacks statistical significance.
+If we could regress differences in construction trends directly on observed differences in state power, we would sidestep this measurement error and likely find larger and significant effects.
+
+On the flip side, winning armed conflicts is also not the only consequence of state power. 
+That the 
+It could even be the opposite: while successions are about familial ties on a superficial level, how this network of familial ties links different lineages is not necessarily uncorrelated with state power.
+It is a possibility that
+
+On top of this basic assumption about the interpretation of the independent variable come econometric assumptions.
+
+External validity
 

### 2023-06-26 paper/paper.tex
#29 remove J(i,t)

@@
 In all cases this refers to how the state currently owning the city came into power --- for example $Conquest_{it} = 1$ means that the state that holds city $i$ at time $t$ acquired city $i$ by conquest.
-It will be useful to adopt one final bit of notation from \cite{schoenholzer2022} and let $\mathbf{J}(i, t)$ denote the state ruling city $i$ at time $t$.
 
 A key economic assumption is needed, even before the econometric ones, for these interaction terms to tell us anything about state power.
-It is this: the state power of a city's owner $\mathbf{J}(i, t)$ is correlated with the manner in which the city was acquired by that owner.
+It is this: the power of the state owning city $i$ is correlated with the manner in which the state acquired city $i$.
 In particular, I assume that conquests are more strongly positively correlated with state power than successions.

### 2023-06-26 paper/paper.tex
#29 version 0.9.3

@@
 \usepackage[utf8]{inputenc}
-\usepackage[left=2cm,right=6cm,top=2.5cm,bottom=2.5cm]{geometry}  
+\usepackage[left=3cm,right=3cm,top=2.5cm,bottom=2.5cm]{geometry}  
 % set 11pt and right=6cm for submission
@@
 State power is not the only determinant of winning armed conflicts, though: luck plays a central role, as do many other time-varying factors for which I cannot control.
-If the correlation between state power differentials (the independent variable whose effects we want to measure) and winning conflicts (how we measure it) is small then we essentially have measurement error in the independent variable.
-This leads to attenuation bias.
-The estimate for $\theta_conquest$ has the expected positive sign in my baseline regression --- it just lacks statistical significance.
+On the flip side, winning armed conflicts is also not the only consequence of state power.
+Importantly, the rulers of more powerful states may be able to make more advantageous matches for their children --- dynastic marriages that eventually result in the expansion of their territory.
+Exhibit A in this argument is the House of Habsburg: they became a dominant power in Europe at the turn of the 16th century when strategic marriages made them rulers of Burgundy and Castille, but Habsburgs had been the Holy Roman Emperors for decades prior.
+
+These arguments could imply that the difference between conquering and inheriting a city does not really capture state power differentials --- that my measure does not actually contain much information about the independent variable.
+If conquests are at best a noisy measure of state power, we essentially have measurement error in the independent variable.
+Accordingly, we should expect attenuation bias to push the result towards zero.
+The estimate for $\theta_{conquest}$ has the expected positive sign in my baseline regression --- it just lacks statistical significance.
 If we could regress differences in construction trends directly on observed differences in state power, we would sidestep this measurement error and likely find larger and significant effects.
 
-On the flip side, winning armed conflicts is also not the only consequence of state power. 
-That the 
-It could even be the opposite: while successions are about familial ties on a superficial level, how this network of familial ties links different lineages is not necessarily uncorrelated with state power.
-It is a possibility that
-
-On top of this basic assumption about the interpretation of the independent variable come econometric assumptions.
-
-External validity
+On top of this basic assumption about the interpretation of the independent variable come econometric assumptions. BLANK. External validity
 
@@
     \item states within the HRE may be too similar / competition between them not large enough
-    \item successions may also select for state power - maybe you get to enter more advantageous marriages if you are the leader of a powerful state
     \item cities are normally conquered one at a time, while successions usually see many cities switch at the same time - maybe there is a crowding-out of the positive effect, e.g.

### 2023-06-26 paper/paper.tex
#29 dump metrics facts, update intro

@@
 While I am able to replicate \cite{schoenholzer2022}'s result that switching to another state has a positive effect one century later, I fail to identify a statistically significant effect of state power on urban development.
-This does not necessarily mean that conflict-driven evolution played no role in the rise of Europe --- several issues limit the statistical power of my test, which might prevent me from finding an effect even if there is one.
-One of them is that the data I use may not be fully adequate for answering the research question.
+This does not necessarily mean that conflict-driven evolution played no role in the rise of Europe --- there are fundamental issues with the data and the approach which limit the statistical power of my analysis.
+I argue that the effect would likely be significant in the absence of these issues.
 
@@
 The econometric assumption that must hold true for $\theta_{conquest}$ to be a causal effect is a different version of the parallel trends assumption.
-In particular, it must hold that, had a conquered city i counterfactually been acquired by succession rather than conquest, it would have trended the same as the other cities acquired by succession.
+In particular, it must hold that, had a conquered city $i$ counterfactually been acquired by succession rather than conquest, it would have trended the same as the other cities acquired by succession.
 BLANK.
@@
 The estimate for $\theta_{conquest}$ has the expected positive sign in my baseline regression --- it just lacks statistical significance.
-If we could regress differences in construction trends directly on observed differences in state power, we would sidestep this measurement error and likely find larger and significant effects.
+If we could regress differences in construction trends on directly observed differences in state power, we would sidestep this measurement error and likely find larger and significant effects.
 
-On top of this basic assumption about the interpretation of the independent variable come econometric assumptions. BLANK. External validity
+Econometrics come on top of this basic assumption about the interpretation of the independent variable.
+The parallel trends assumption is: Had a conquered city $i$ counterfactually been acquired by succession rather than conquest, it would have trended the same as the other cities acquired by succession.
+
+The first implication of that is that cities did not select into treatment based on uncontrolled-for characteristics that . 
+There are no uncontrolled-for characteristics that made a city more or less likely to be conquered and also had an effect on construction activity.
+Note that I control for all time-invariant characteristics of cities in the city fixed effect.
+
+The second implication is there is nothing else happening at the same time as the conquest that affects conquered vs. inherited cities differently.
+
+If conquests had an effect on construction activity not just through state power but also through the number of cities that also switched - maybe switching to a new state alone is better than switching with all cities of an entire territory.
+
+External validity
 

### 2023-06-26 paper/paper.tex
#29 version 0.9.4

@@
 Econometrics come on top of this basic assumption about the interpretation of the independent variable.
-The parallel trends assumption is: Had a conquered city $i$ counterfactually been acquired by succession rather than conquest, it would have trended the same as the other cities acquired by succession.
+We can interpret the interaction coefficient $\theta_{conquest}$ as a causal effect if the following holds: had a conquered city $i$ counterfactually been acquired by succession rather than conquest, it would have trended the same as the other cities acquired by succession.
 
-The first implication of that is that cities did not select into treatment based on uncontrolled-for characteristics that . 
-There are no uncontrolled-for characteristics that made a city more or less likely to be conquered and also had an effect on construction activity.
+This assumption requires that there are no uncontrolled-for variables that are correlated with being conquered and also have an effect on construction activity (while not being outcomes of state power themselves).
 Note that I control for all time-invariant characteristics of cities in the city fixed effect.
+However, there is one thing that comes to mind: maybe switching to another state had a more positive effect on a city's development if no other cities switched to the same state at the same time.
+A single city may be more easy for the new state to integrate into its economy than many, or alternatively, rulers who win multiple cities at the same time may invest only in their favourite cities and neglect the rest.
 
-The second implication is there is nothing else happening at the same time as the conquest that affects conquered vs. inherited cities differently.
+The entire territory goes to another state (or multiple) when a ruling lineage goes extinct, but conquests usually happen one at a time.
+The average number of simultaneous switches is likely much larger for successions than for conquests.
+If conquests had an effect on construction activity not just through the change in state power but also through the number of cities that switched simultaneously, then the coefficient $\theta_{conquest}$ no longer captures only the causal effect of switching.
+I believe this is one of the remaining threats to the internal validity of my regression analysis.
+It should be possible to compute the number of cities that switch to the same state as city $i$ in period $t$ and then to control for it, which would solve the problem.
+I leave this for a future extension of the paper.
 
-If conquests had an effect on construction activity not just through state power but also through the number of cities that also switched - maybe switching to a new state alone is better than switching with all cities of an entire territory.
-
-External validity
-
-Other explanations for no effect:
-\begin{itemize}
-    \item states within the HRE may be too similar / competition between them not large enough
-    \item cities are normally conquered one at a time, while successions usually see many cities switch at the same time - maybe there is a crowding-out of the positive effect, e.g.
-because the ruler likes one of his newly acquired cities most and neglects the others
-    \item the data is just not that well suited to the task; within-R2 is super low in Table \ref{tab:baseline_50y}
-\end{itemize}
+External validity BLANK
+Switches barely explain variation in construction activity (super low within-R2) - possible solution: add lags for controls BLANK
 
@@
     \item differences in timing - do early switchers in my sample experience different effects than late switchers?
-    \item lag controls because construction takes a while and the years probably record completion
+    \item further explore differences across regions
+    \item better measures of state power
 \end{itemize}

### 2023-06-26 paper/paper.tex
#29 begin external validity, cite CY HHE

@@
 
-The entire territory goes to another state (or multiple) when a ruling lineage goes extinct, but conquests usually happen one at a time.
+An entire territory goes to another state (or multiple) when a ruling lineage goes extinct, but conquests usually happen one at a time.
 The average number of simultaneous switches is likely much larger for successions than for conquests.
@@
 I believe this is one of the remaining threats to the internal validity of my regression analysis.
-It should be possible to compute the number of cities that switch to the same state as city $i$ in period $t$ and then to control for it, which would solve the problem.
-I leave this for a future extension of the paper.
+It should be possible to compute the number of cities that switch to the same state as city $i$ in period $t$ and then to control for it in Equation \ref{eq:baseline}, which would solve the problem.
+I leave this for a future extension of this paper.
 
-External validity BLANK
-Switches barely explain variation in construction activity (super low within-R2) - possible solution: add lags for controls BLANK
+Having discussed issues with internal validity now, I will briefly talk about external validity. For this purpose it is helpful to first put the goal of my paper into context. 
+\cite{cantoni2021}
+
+I doubt that we can learn anything about present conflicts or present growth rates from my results.
+
+In terms of \cite{levine2021, levine2022}'s model of state systems, most international conflict nowadays is defined by the presence of multiple extremely strong outsiders who act to keep the balance of power in the status quo.
+
+% ######################################################
+
+\section{Conclusion} \label{sec:conclusion}
 
@@
 \end{itemize}
-
-% ######################################################
-
-\section{Conclusion} \label{sec:conclusion}
 

### 2023-06-26 paper/paper.tex
#29 implement Daniel feedback 1/

@@
 Each entry contains additional information about the type of event and the range of uncertainty for the timing.
-Because I drop city-period observations with missing construction data from the sample, it is worth briefly outlining how I transform a list of events into a city-period panel of construction activity.
+Because I drop city-period observations with missing construction data from the sample, it is worth outlining how I transform a list of events into a city-period panel of construction activity.
 First I group and count events that occur in the same city and period.
@@
 In an attempt to understand more about the effects I estimate, I use certain subsets of construction activity as the outcome in alternative specifications.
-There are 16 types of construction activity in the original data, which I combine into 4 facets: \textit{State construction} consists of palaces, castles, administrative and military buildings.
+There are 16 types of construction activity in the original data, which I combine into 3 facets: \textit{State construction} consists of palaces, castles, administrative and military buildings.
 \textit{Private construction} includes malls, economic (banks, factories, restaurants, and more) and private (residential) buildings.
@@
 
-A key economic assumption is needed, even before the econometric ones, for these interaction terms to tell us anything about state power.
-It is this: the power of the state owning city $i$ is correlated with the manner in which the state acquired city $i$.
+A key economic assumption is needed, even before the econometric ones, for these interaction terms to tell us anything about state power: the power of the state owning city $i$ is correlated with the manner in which the state acquired city $i$.
 In particular, I assume that conquests are more strongly positively correlated with state power than successions.

### 2023-06-26 paper/output/regressions/controls_50y.tex
#29 include check in controls table

@@
    \centering
-   \begin{tabular}{lcccc}
+   \begin{tabular}{lccccc}
       \tabularnewline \midrule \midrule
-      Dependent Variable: & \multicolumn{4}{c}{All construction}\\
-      Model:                                     & (1)    & (2)    & (3)           & (4)\\  
+      Dependent Variable: & \multicolumn{5}{c}{All construction}\\
+      Model:                                     & (1)    & (2)    & (3)           & (4)          & (5)\\  
       \midrule
       \emph{Variables}\\
-      Switch to another state                    & 0.02   & 0.02   & -0.01         & 0.03\\   
-                                                 & (0.03) & (0.02) & (0.05)        & (0.05)\\   
-      Switch to another state $\times$ Conquest  & 0.04   & 0.04   & -0.15$^{***}$ & -0.14$^{**}$\\   
-                                                 & (0.04) & (0.04) & (0.05)        & (0.06)\\   
-      Switch to another state $\times$ Other     & 0.01   & 0.01   & 0.05          & 0.05\\   
-                                                 & (0.03) & (0.03) & (0.05)        & (0.05)\\   
-      Switching indicator                        & -0.01  &        & -0.09         &   \\   
-                                                 & (0.02) &        & (0.06)        &   \\   
-      Conflict                                   &        &        & 0.01          & 0.01\\   
-                                                 &        &        & (0.03)        & (0.03)\\   
+      Switch to another state                    & 0.02   & 0.02   & -0.01         & 0.03         & -0.01\\   
+                                                 & (0.03) & (0.02) & (0.05)        & (0.05)       & (0.05)\\   
+      Switch to another state $\times$ Conquest  & 0.04   & 0.04   & -0.15$^{***}$ & -0.14$^{**}$ & -0.15$^{***}$\\   
+                                                 & (0.04) & (0.04) & (0.05)        & (0.06)       & (0.05)\\   
+      Switch to another state $\times$ Other     & 0.01   & 0.01   & 0.05          & 0.05         & 0.05\\   
+                                                 & (0.03) & (0.03) & (0.05)        & (0.05)       & (0.05)\\   
+      Switching indicator                        & -0.01  &        & -0.09         &              & -0.09\\   
+                                                 & (0.02) &        & (0.06)        &              & (0.06)\\   
+      Conflict                                   &        &        & 0.01          & 0.01         &   \\   
+                                                 &        &        & (0.03)        & (0.03)       &   \\   
       \midrule
       \emph{Fixed-effects}\\
-      City                                       & Yes    & Yes    & Yes           & Yes\\  
-      Period                                     & Yes    & Yes    & Yes           & Yes\\  
+      City                                       & Yes    & Yes    & Yes           & Yes          & Yes\\  
+      Period                                     & Yes    & Yes    & Yes           & Yes          & Yes\\  
       \midrule
       \emph{Fit statistics}\\
-      Observations                               & 6,635  & 6,635  & 1,955         & 1,955\\  
-      R$^2$                                      & 0.3553 & 0.3553 & 0.4245        & 0.4230\\  
-      Within R$^2$                               & 0.0009 & 0.0008 & 0.0071        & 0.0045\\  
+      Observations                               & 6,635  & 6,635  & 1,955         & 1,955        & 1,955\\  
+      R$^2$                                      & 0.3553 & 0.3553 & 0.4245        & 0.4230       & 0.4244\\  
+      Within R$^2$                               & 0.0009 & 0.0008 & 0.0071        & 0.0045       & 0.0069\\  
       \midrule \midrule
@@
    Note: Table presents results of estimation equation \eqref{eq:baseline} but using different control variables.
-The switch type  "Succession" is omitted as the reference category.
+Columns (3)-(5) drop observations with missing data on conflict incidents from the sample.
 Yearly data was aggregated into periods of 50 years.

### 2023-06-26 source/analysis/robustness/Controls.R

@@ -7,6 +7,7 @@ setwd("~/GitHub/BA")
 
 source("source/utils/HandleCommandArgs.R")
 source("source/utils/DataPrepSuite.R")
+source("source/utils/PrepareBaselineData.R")
 source("source/utils/etableDefaults.R")
 
 Main <- function(){
@@ -17,10 +18,13 @@ Main <- function(){
   build <- read_csv(read_path, show_col_types = F)
   
   # Select sample and process data (exact steps documented below)
-  dat <- PrepareData(build,
+  dat <- PrepareBaselineData(build,
                      max_switches = 2,
                      binarise_construction = T,
                      binarise_switches = T)
+  
+  dat_noconflict <- dat |> 
+    drop_na(conflict)
   
   # Estimate Equation (2) with different sets of controls
   mod_baseline <- fixest::feols(
@@ -39,13 +43,16 @@ Main <- function(){
     c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + conflict | city_id + period, 
     data = dat)
   
+  mod_conf_check <- fixest::feols(
+    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
+    data = dat_noconflict)
   
   # Produce regression table and export to LaTeX
   etableDefaults()
   
   note <- paste("Note: Table presents results of estimation equation
-  \\eqref{eq:baseline} but using different control variables. The switch type 
-  \"Succession\" is omitted as the reference category.", PeriodInsert(t), 
+  \\eqref{eq:baseline} but using different control variables. Columns (3)-(5) drop
+  observations with missing data on conflict incidents from the sample.", PeriodInsert(t), 
   "Observations are at the city-period level. The dependent variable is an 
   indicator that takes the value 1 if any construction activity was recorded. 
   Standard errors are clustered at the city level.", SignifInsert(), sep = " ") |> 
@@ -53,6 +60,7 @@ Main <- function(){
   
   tex_output <- etable(mod_baseline, mod_noswitches,
                        mod_conflict, mod_conf_nosw,
+                       mod_conf_check,
                        tex=TRUE,
                        title="Heterogeneous effects of switching: Robustness",
                        label = sprintf("tab:controls_%iy", t),
[…]

### 2023-06-26 paper/paper.tex
#29 implement Daniel's feedback 2/

@@
 Specifically, if we consider the universe of all switching events, the average increase in state power from the old to the new owner would be higher for the conquests than for the successions.
-The reasoning is that conflict --- as in a conquest --- selects the new state for power, but conflict-free successions, depending mainly on dynastic marriages, are more "random".
+The reasoning is that more state power is necessary to conquer a city, which generally involves winning an armed conflict, than to inherit it from an extinct lineage, which mainly requires favourable marriages.
 This assumption allows me to use conquests as a stand-in for an increase in state power.
-Several other assumptions are needed to interpret my results causally.
+Additional assumptions are needed to interpret my results causally.
 
@@
 } 
-Cities are ruled by conquering states 16\% of the time, by succeeding states 34\% of the time, and by states that gained possession of them in another way 50\% of the time.
+On average, cities spend 16\% of their recorded lifespan under the rule of states that conquered them, 34\% under states that inherited them from an extinct lineage, and 50\% under states that gained possession of them in another way.
 
@@
 
-I aggregate this yearly data into periods of different length.
-I build yearly, 10-yearly, 50-yearly and 100-yearly versions.
-I use the 50-yearly version throughout the regressions in the paper.
-Events are rare enough that this avoids losing too much information to having to drop missing values.
-Also, some events (conflict as well as construction) are recorded with uncertain timing --- descriptors like "beginning of the 16th century" rather than an exact year.
-In the aggregated version I include such events if their timing can be assigned with certainty to a 50-year period.
+I aggregate this yearly data into periods with a length of 50 years.
+This is primarily because many events (conflict as well as construction) are recorded with uncertain timing, i.e.
+descriptors like "beginning of the 16th century" rather than an exact year --- I can use such events in the aggregated version if their timing can be assigned with certainty to a 50-year period.
+A pleasant side effect of using aggregated data is that coefficient plots and tables are easier to read.
 

### 2023-06-26 paper/paper.tex
#29 finish adding Col (5) and feedback 3/3

@@
 Columns (3) and (4) add conflict incidents as a control variable, once in addition to switching and once without.
+Finally, Column (5) runs the baseline specification from (1) on the smaller sample used in (3) and (4).
 The outcome is $All \, construction$ in all regressions.
@@
 While this result may seem like evidence for omitted variable bias in the baseline estimate, that explanation is ultimately not plausible.
-To cause bias in Column (1) conflict incidents would need to be correlated both with the post-conquest dummy and with construction --- however, we can see that the latter is not the case: the coefficients on $Conflict$ in Columns (3) and (4) are tiny and not significant by any stretch of the imagination.\footnote
-{
-The coefficients on $Conflict$ in Table \ref{tab:controls_50y} are conditional correlations.
-Simply computing the correlation coefficient between the dummies for conflict incidents and for construction yields a value of $-0.01$.
-} 
-Rather, I suggest that the estimates for $\theta_{conquest}$ are different because they are estimated on different samples.
-More than two thirds of observations are lost between Columns (2) and (3) because of missing values in the $Conflict$ variable.
-Figure \ref{fig:conflict_map} shows that the remaining observations follow quite a different distribution across space --- if the spatial composition of the smaller sample was the same, the entire map would be a uniform shade.
+Column (5) shows that the disagreement between the estimates does not come from controlling for conflict, but from the sample composition: more than two thirds of observations are lost between Columns (2) and (3) because of missing values in the $Conflict$ variable.
+When I run the baseline specification on this partial sample, I obtain the exact same estimates as in Column (3) --- even without controlling for conflict.
+
+Figure \ref{fig:conflict_map} shows that the observations that remain in the partial sample follow quite a different distribution across space --- if the spatial composition of the partial sample was the same as that of the original sample, the entire map would be a uniform shade.
 Instead we see big discrepancies: the west, the south, and Brandenburg retain 30\% to 60\% of their observations, while many northern and eastern regions all but vanish from the sample.
 
-If we accept that the disagreement between the interaction coefficients in (1) and (3) is due to the different spatial distribution of the sample, that has the interesting implication of spatially heterogeneous differences in treatment effects.
-The null difference in (1) could be an average of the negative difference in the southwest and a hypothetical positive difference in the northeast.
+Together, Table \ref{tab:controls_50y} and Figure \ref{fig:conflict_map} have the interesting implication of spatially heterogeneous differences in treatment effects.
+The null difference in Column (1) could be an average of the negative difference in the southwest and a hypothetical positive difference in the northeast.
 Outsiders could play a role, in a similar vein to \cite{levine2021, levine2022}'s theory: the south and the west are bordered by the Habsburg Empire and France, two stable major powers, while the northeast is somewhat sheltered by the North Sea, the Baltic Sea, and the territory that comprises 1937 Poland, which was fragmented throughout much of its history.\footnote

### 2023-06-26 paper/paper.tex
#29 v0.9.5

@@
 
-Having discussed issues with internal validity now, I will briefly talk about external validity. For this purpose it is helpful to first put the goal of my paper into context. 
-\cite{cantoni2021}
-
-I doubt that we can learn anything about present conflicts or present growth rates from my results.
-
-In terms of \cite{levine2021, levine2022}'s model of state systems, most international conflict nowadays is defined by the presence of multiple extremely strong outsiders who act to keep the balance of power in the status quo.
+Having discussed issues with internal validity now, I will briefly talk about external validity.
+The hypothesis I test is a prediction of the theory of conflict-driven evolution.
+While the theory itself can be applied widely --- \cite{levine2021} use it to explain stylised facts about pre-modern Europe, pre-modern India, the present-day United Kingdom and Spain, and 20th-century China --- my empirical results only test whether the theory can explain what happens in my dataset.
 
@@
 \section{Conclusion} \label{sec:conclusion}
+
+Accounting for the possibility of attenuation bias from measurement error makes the null result look less unexpected and more like suggestive evidence.
 

### 2023-06-27 paper/paper.tex
#29 finish robustness (ext validity)

@@
 Having discussed issues with internal validity now, I will briefly talk about external validity.
-The hypothesis I test is a prediction of the theory of conflict-driven evolution.
-While the theory itself can be applied widely --- \cite{levine2021} use it to explain stylised facts about pre-modern Europe, pre-modern India, the present-day United Kingdom and Spain, and 20th-century China --- my empirical results only test whether the theory can explain what happens in my dataset.
+We cannot infer from my results that the theory of conflict-driven evolution is true or false.
+I test a general prediction of the theory in the Holy Roman Empire between 1300 and 1800, which as a setting is fairly unique.
+It features a vast array of fragmented states that compete and go to war with each other, but are also part of the same empire and likely to share many institutions.
+Religious conflict peaks in the sample period, with the Thirty Years War taking place from 1618-1648, and alliances are often based on the shared confession of rulers.
+Spain, by contrast, was a large monolithic state by the end of the Reconquista in 1492, in which the Protestant reformation never really took off.
+
+Based on these unique features of the setting we should be wary even of generalising my results to the rest of Europe.\footnote
+{
+\cite{schoenholzer2022} fill this gap partially, given that their sample comprises all of Europe and they at least find a similar overall effect of switching states on long-run development.
+}
+Nonetheless, the results have value outside of understanding the history of the Holy Roman Empire: as a puzzle piece.
+A main objective of this paper was to test a theory.
+Theories become more widely accepted when researchers repeatedly test and confirm their predictions; they are discredited when their predictions stubbornly disagree with reality.
+Even the result that there is no evidence for the prediction is valuable because it points towards limitations of the theory.
 
@@
 \section{Conclusion} \label{sec:conclusion}
+
+What do we learn from this paper?
 

### 2023-06-27 paper/paper.tex
#31 first third of conclusion

@@
 
+Early modern Europe was a uniquely warlike place.
+A recent theory models wars and conflict as a Darwinian selection mechanism: strong states absorb the territory of weak ones, which gradually filters out weaker states.
+If this is true, territory that changes hands during conflict will generally tend to move into the possession of a stronger state and benefit from its strength.
+
+In this paper I test this prediction of the theory in the setting of the Holy Roman Empire.
+I want to understand the role that the process of conflict-driven evolution played in the economic development of Europe.
+I use data compiled from an encyclopedia of 2,390 historically German cities, covering the years 1300 to 1800.
+I use records of construction activity as a proxy for economic activity on the city level.
+
+I run panel regressions with two-way fixed effects and staggered adoption on this dataset.
+First I estimate the dynamic effects of switching to another state.
+A significant effect manifests one century after a switch and becomes larger over time; after two centuries, we are 9 percentage points more likely to observe construction in cities that switched states.
+
+I then estimate heterogeneous treatment effects of different types of switches.
+In particular, I care about the difference between the treatment effect of conquests and the treatment effect of successions: if we assume that conquests are more strongly associated with state power than successions, then we can interpret a switch by conquests as a proxy for an increase in state power.
+
 What do we learn from this paper?
@@
 
+One could do more to explore the idiosyncrasies of the setting. 
 Improvements / future research:

### 2023-06-27 paper/paper.tex
#31 almost finish conclusion

@@
 I specifically test the hypothesis that cities benefitted from ownership changes because competition between states selected for higher-quality government.
-While I show that cities grow faster in the long run after switching states, I do not find evidence that conflict-driven selection played a role in this.
+While I show that cities grow faster in the long run after switching states, the evidence that conflict-driven selection played a role in this is less compelling.
 \end{abstract}
@@
 {
-The basic idea of Darwinian selection has previously been borrowed by economists - for example by \cite{galor2002} who argue that humans are subject to selection and accumulate (economically) favourable traits over the course of history.
-More infamously, \cite{clark2007} employs it in an attempt to explain why the Industrial Revolution took place in Europe, arguing that at some point the European rich began to out-breed the poor and passed their favourable traits on to their offspring.
+The basic idea of Darwinian selection has previously been borrowed by economists - for example by \cite{galor2002} who argue that humans are subject to selection and accumulate (economically) favourable traits over the course of all history.
+More infamously, \cite{clark2007} employs it in an attempt to explain why the Industrial Revolution took place in Europe, arguing that at some point the European rich began to out-breed the poor and passed their middle-class values on to their offspring.
 }
@@
 However, they do not show explicitly how these improvements in state quality are driven by selection through conflict.
-They use a dataset with a bigger geographical scope (all of Europe) but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is available for fewer cities and fewer years.
+Their data has a bigger geographical scope (all of Europe) but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is available for fewer cities and fewer years.
 
@@
 This does not necessarily mean that conflict-driven evolution played no role in the rise of Europe --- there are fundamental issues with the data and the approach which limit the statistical power of my analysis.
-I argue that the effect would likely be significant in the absence of these issues.
+The most important one is attenuation bias from measurement error in the independent variable, and I argue that the effect may well be significant in its absence.
 
@@
 
-Early modern Europe was a uniquely warlike place.
+Early modern Europe was a uniquely war-torn place.
 A recent theory models wars and conflict as a Darwinian selection mechanism: strong states absorb the territory of weak ones, which gradually filters out weaker states.
-If this is true, territory that changes hands during conflict will generally tend to move into the possession of a stronger state and benefit from its strength.
+If this is true, territory that changes hands during conflict will generally tend to move into the possession of a stronger state and subsequently benefit from its strength.
 
@@
 I want to understand the role that the process of conflict-driven evolution played in the economic development of Europe.
-I use data compiled from an encyclopedia of 2,390 historically German cities, covering the years 1300 to 1800.
-I use records of construction activity as a proxy for economic activity on the city level.
+I use data compiled from an encyclopedia of 2,390 historically German cities, covering the years 1300 to 1800 and featuring records of construction activity as a city-level proxy for economic activity.
+I run panel regressions with two-way fixed effects and staggered adoption on this dataset.
 
-I run panel regressions with two-way fixed effects and staggered adoption on this dataset.
 First I estimate the dynamic effects of switching to another state.
 A significant effect manifests one century after a switch and becomes larger over time; after two centuries, we are 9 percentage points more likely to observe construction in cities that switched states.
+I do not find a pre-trend.
 
 I then estimate heterogeneous treatment effects of different types of switches.
-In particular, I care about the difference between the treatment effect of conquests and the treatment effect of successions: if we assume that conquests are more strongly associated with state power than successions, then we can interpret a switch by conquests as a proxy for an increase in state power.
+In particular, I care about the difference between the treatment effect of conquests and the treatment effect of successions.
+If we assume that conquered cities tend to move into the possession of stronger states than inherited cities, we can interpret the difference in treatment effects as the effect of state power.
+The literature suggests that state power is positively linked to growth-fostering institutions like fiscal capacity; we would therefore expect it to have a significant positive effect on economic (construction) activity.
 
-What do we learn from this paper?
+I indeed obtain a positive estimate for the heterogeneity --- conquests increase the probability of observing construction activity by 4 percentage points more than successions --- but it is not statistically significant.
+This finding remains unchanged in most robustness checks.
+However, because the contrast between conquests and successions with respect to state power is not very sharp, all these estimates likely have attenuation bias working against them.
+Accounting for this possibility makes the null result seem less unexpected, and more like suggestive evidence: it means that the coefficient would be bigger and possibly significant if I had a more accurate measure of state power to regress on.
 
-Accounting for the possibility of attenuation bias from measurement error makes the null result look less unexpected and more like suggestive evidence.
+The main takeaway is that the effect of switching to another state on urban development is positive in the long run, and that I find tentative empirical support for the theory of conflict-driven selection.
 
-One could do more to explore the idiosyncrasies of the setting. 
-Improvements / future research:
-\begin{itemize}
-    \item differences in timing - do early switchers in my sample experience different effects than late switchers?
-    \item further explore differences across regions
-    \item better measures of state power
-\end{itemize}
+It would be a big improvement on my research strategy to find a clearer measure of state power, to check whether measurement error really does bias my estimates downward.
+\cite{cantoni2023} construct a novel dataset of territory-level institutions which one could potentially use for that purpose.
+Another improvement would be to control for the number of cities switching to another state at the same time in Equation \eqref{eq:baseline}, as described in the Discussion.
+Third, an event study version of Equation \eqref{eq:baseline} may be able to better tease out significant effects if the effect of state power takes some time to phase in.
+Finally, recent literature (\cite{baker2022} provide an overview) has questioned the validity of standard two-way fixed effects regressions with staggered adoption and time-varying treatment --- one could improve my results by using one of the robust estimators proposed by \cite{sunab2021} or \cite{cengiz2019}, for example.
+
+Future research in the setting of my thesis could explore the spatial heterogeneity for which I find evidence in Table \ref{tab:controls_50y} and Figure \ref{fig:conflict_map}.
+It would be interesting to learn why the coefficient on the conquest interaction is so dramatically different when the sample is skewed towards the southern and western regions.
+One could also take a much deeper dive into \cite{levine2021}'s theory and use the detailed geospatial and territorial histories data to test other implications of their model.
+
+Better understanding how the European state system evolved can deepen our understanding of social and economic history and of the origins of efficient institutions.
 

### 2023-06-27 paper/paper.tex
#31 finish conclusion on PERIOD

@@
 
-Better understanding how the European state system evolved can deepen our understanding of social and economic history and of the origins of efficient institutions.
+While the research question may not seem hugely relevant for the present-day Western world, I want to close, briefly, with the Democratic Republic of the Congo (DRC).
+This African state's central government is weak and large parts of its territory are controlled by independent armed groups \citep[p.
+34]{sierra2020}.
+Recent research has studied the DRC to investigate questions about state formation \citep{sierra2020}, taxation and state capacity \citep{balan2022}, and conflict \citep{henn2021}.
+One could argue that we are witnessing economic history "in real time" there.\footnote
+{
+This thought is not my own; Davide Cantoni said it to me in conversation.
+} 
+My point is that a better understanding of how the European state system evolved may not only deepen our knowledge of the origins of institutions --- it could also inform our perspective on regions of the world that do not have a stable state system at present.
 

### 2023-06-27 paper/paper.tex
#31 v1.0
#32 add title page
#32 add declaration

@@
 \begin{document}
+
+\begin{center}\uppercase{Ludwig-Maximilians-Universität München}\end{center}
+\begin{center}\uppercase{Seminar für Wirtschaftsgeschichte}\end{center}
+
+\vspace{3cm}
+
+\begin{center}
+    {\centering\huge{{Conflict-Driven Evolution and Urban Development:\\
+Evidence from the Holy Roman Empire
+}}}
+\end{center}
+\date{\vspace{-5ex}}
+\thispagestyle{empty}
+
+%%%Block Masterarbeit:
+\begin{center}
+\begin{large}
+\begin{Large}
+Bachelorarbeit\\
+\end{Large}
+im Studiengang 'Volkswirtschaftslehre' \\
+\end{large}
+\end{center}
+\begin{center}
+vorgelegt von\\
+\begin{large}
+Elias Hadj Ammar\\
+\end{large}
+\end{center}
+\vspace{1cm}
+\begin{center}
+Betreuerin:\\
+\begin{large}
+Leonie Bielefeld\\
+\end{large}
+\end{center}
+%%%Ende Block Masterarbeit
+
+\begin{center}
+\begin{large}
+München, den 27. Juni 2023 \\
+\end{large}
+\end{center}
+
+\vspace{1,5cm}
+
+\begin{center}
+\url{Elias.Ammar@campus.lmu.de}\\
+Matrikelnr.: 12113698 \\
+\end{center}
+
+\clearpage
+
 \onehalfspacing
@@
 
+\clearpage
+\section*{Declaration}
+
+Hiermit versichere ich, dass ich die vorliegende Arbeit selbständig und ohne Benutzung anderer als der angegebenen Hilfsmittel angefertigt, noch nicht einer anderen Prüfungsbehörde vorgelegt und noch nicht veröffentlicht habe.
+
+\vspace{3cm}
+München, den 27.06.2023
+
 \end{document}

### 2023-06-27 source/analysis/robustness/SunabBasic.R
#32 housekeep repo

@@ -1,70 +1,0 @@
-#!/usr/bin/env Rscript --vanilla
-
-library(tidyverse) |> suppressPackageStartupMessages()
-library(fixest) |> suppressPackageStartupMessages()
-
-setwd("~/GitHub/BA")
-
-source("source/utils/HandleCommandArgs.R")
-source("source/utils/DataPrepSuite.R")
-source("source/utils/etableDefaults.R")
-
-Main <- function(){
-  
-  # Read data
-  t <- HandleCommandArgs(default_length = 50)
-  read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
-  build <- read_csv(read_path, show_col_types = F)
-  
-  # Select sample and process data (exact steps documented below)
-  dat <- PrepareData(build,
-                     max_switches = 2,
-                     binarise_construction = T,
-                     binarise_switches = T)
-  
-  mod <- fixest::feols(
-    c_all ~ sunab(treat_time, period) + switches | city_id + period, 
-    data = dat
-  )
-  
-  summary(mod, agg="att")
-
-  iplot(mod)  
-}
-
-
-PrepareData <- function(build, max_switches,
-                        binarise_construction, binarise_switches){
-  # Copied from baseline/Baseline.R
-  
-  
-  # Filter out cities with too many lifetime switches
-  selected <- FilterBySwitches(build, max_switches)
-  
-  # Identify across-period switches
-  with_e_dummies <- AddEAnother(selected)
-  
-  # Add TREAT x POST dummies D
-  with_D <- AddTreatXPost(with_e_dummies)
-  
-  # Binarise construction outcomes, if specified
-  if (binarise_construction) { with_D <- with_D |> BinariseOutcomes() }
-  
-  # Binarise no. of switches (control), if specified
-  if (binarise_switches) { with_D <- with_D |> BinariseSwitches() }
-  
-  # Process type change into treatment type
-  with_types <- ProcessTypeChange(with_D)
-  
-  # Rearrange columns (for readability)
-  clean <- with_types |> select(
-    city_id, treat_time, period, terr_id, switches, e_another, D, 
-    rule_conquest, rule_succession, rule_other,
-    conflict, c_all, c_state, c_private, c_public
-  )
-  
-  return(clean)
-}
-
-
-Main()

### 2023-06-27 source/analysis/robustness/Window.R

@@ -1,130 +1,0 @@
-#!/usr/bin/env Rscript --vanilla
-
-library(tidyverse) |> suppressPackageStartupMessages()
-library(fixest) |> suppressPackageStartupMessages()
-
-setwd("~/GitHub/BA")
-
-source("source/utils/HandleCommandArgs.R")
-source("source/utils/DataPrepSuite.R")
-source("source/utils/etableDefaults.R")
-
-Main <- function(){
-  
-  # Read data
-  t <- HandleCommandArgs(default_length = 50)
-  read_path <- sprintf("drive/derived/cities_data_%iy.csv", t)
-  build <- read_csv(read_path, show_col_types = F)
-  
-  # Select sample and process data (exact steps documented below)
-  dat <- PrepareData(build,
-                     max_switches = 2,
-                     years_post = 200,
-                     binarise_construction = T,
-                     binarise_switches = T)
-  
-  # Estimate my baseline interaction equation
-  mod_all_basic <- fixest::feols(
-    c_all ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
-    data = dat)
-  
-  mod_state_basic <- fixest::feols(
-    c_state ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
-    data = dat)
-  
-  mod_priv_basic <- fixest::feols(
-    c_private ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
-    data = dat)
-  
-  mod_pub_basic <- fixest::feols(
-    c_public ~ D + i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + switches | city_id + period, 
-    data = dat)
-  
-  # mod_all_conflict <- fixest::feols(
-  #   c_all ~ i(rule_conquest, D, ref=0) + i(rule_other, D, ref=0) + D + switches + conflict | city_id + period, 
-  #   data = dat)
-  
-  
-  # Produce regression table and export to LaTeX
-  etableDefaults()
-  
-  note <- paste("Note: Table presents results of estimation equation
-  \\eqref{eq:baseline}. The switch type \"Succession\" is omitted as the 
-  reference category.", PeriodInsert(t), "Observations are at the city-period 
-  level. Dependent variables are indicators that take the value 1 if 
-  construction activity of the respective type was recorded. Standard errors are 
[…]

### 2023-06-27 paper/paper.tex
#32 remove blanks (compiles in 7607)

@@
 In particular, I assume that conquests are more strongly positively correlated with state power than successions.
-Support for this assumption comes from \cite{levine2021}'s model of conflict-driven evolution.
-BLANK.
-I will discuss its validity at length later on.
+The idea is that conquering a city means winning it in a conflict --- which is easier if you are a powerful state --- while inheriting it only requires that you, or even your distant ancestors, entered a fortunate marriage.
+I will discuss the validity of this assumption at length later on.
 
-If this assumption holds true, the effect of state power should manifest itself in the difference between the trend of cities acquired by conquest and the trend of cities acquired by succession.
+If the assumption holds true, the effect of state power should manifest itself in the difference between the trend of cities acquired by conquest and the trend of cities acquired by succession.
 Note that successions are the omitted category --- there is no interaction term between the post-treatment indicator and $\mathbf{1}(Succession_{it})$ in Equation \eqref{eq:baseline} --- so that the coefficient $\theta_{conquest}$ captures exactly what I want to estimate: the difference in trends after conquests vis-a-vis successions.
@@
 In particular, it must hold that, had a conquered city $i$ counterfactually been acquired by succession rather than conquest, it would have trended the same as the other cities acquired by succession.
-BLANK.
 
@@
 
-The only subset of construction for which the difference in trends is indeed significant is state construction --- palaces, castles, military and administrative buildings.
+The only subset of construction for which the difference in treatment effects is indeed significant is state construction --- palaces, castles, military and administrative buildings.
 We can see in Column (2) that these are 5 percentage points \textit{less} likely to be built in a city after it was conquered; the effect is significant at the 5\% level.
-BLANK.
+It is hard to explain this fact with state power; however, keeping in mind that this is the effect of conquests relative to successions, it could be that rulers who inherit a city feel a bigger desire to impress and make their presence known to the townspeople than someone who just marched an army into it.
 
@@
 } 
-BLANK \cite{cervellati2022}. I leave this for future research to explore.
+However, I leave this topic for future research to explore.
 

### 2023-06-27 paper/paper.tex
v1.1

@@
 One can indeed imagine this as an evolutionary process: states compete for resources and powerful states eventually absorb weaker ones.
-As a result, traits that increase state power --- the ability to win conflicts against rivals --- get selected for, while unfavourable traits are doomed to disappear \citep{levine2021}.
+As a result, traits that increase state power --- the ability to win conflicts against rivals --- are selected for, while unfavourable traits are doomed to disappear \citep{levine2021}.
 
@@
 \cite{north1970}, \cite{delong1993}, \cite{ajr2001}).
-From this vast literature one theme is especially relevant here: the institution of fiscal capacity, or the state's ability to collect taxes.
+From this vast literature one theme may be especially relevant here: the institution of fiscal capacity, or the state's ability to collect taxes.
 \cite{tilly1985} argues that the main reason for states to collect taxes is to be able to wage war.
 In other words, fiscal capacity and state power are tightly linked --- so if conflict selects for state power, then it also selects for fiscal capacity.
-\cite{gennaioli2015} and \cite{cantoni2023} show that conflict drove state consolidation.
+\cite{gennaioli2015} and \cite{cantoni2023} show that conflict drove state consolidation, with the latter set of authors using the same data as this paper.
 \cite{dincecco2012} show that fiscal capacity had a positive effect on growth.
@@
 While I am able to replicate \cite{schoenholzer2022}'s result that switching to another state has a positive effect one century later, I fail to identify a statistically significant effect of state power on urban development.
-This does not necessarily mean that conflict-driven evolution played no role in the rise of Europe --- there are fundamental issues with the data and the approach which limit the statistical power of my analysis.
+This does not necessarily mean that conflict-driven evolution played no role in the rise of Europe, though --- I recognise some issues with the data and the approach which limit the statistical power of my analysis.
 The most important one is attenuation bias from measurement error in the independent variable, and I argue that the effect may well be significant in its absence.
@@
 The rest of the paper proceeds as follows.
-Section \ref{sec:background} gives essential historical background information.
+Section \ref{sec:background} gives some very brief historical background information.
 Section \ref{sec:data} describes the data used in the analysis.
@@
 2]{pt1}.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[scale=0.15]{paper/output/descriptive/map_cities_raw.png}
+    \caption{Distribution of cities in the \textit{Territorial Histories} dataset \citep{pt2}.
+Each dot represents a city.
+Some major cities are identified with labels.
+Polygons on the map do \textbf{not} represent historical borders; they merely represent regions by which cities are clustered in the \textit{Städtebuch}.}
+    \label{fig:map_cities_raw}
+\end{figure}
 
@@
 Only Brandenburg (containing Berlin) and Silesia (containing Breslau) seem to lose disproportionately many observations.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[scale=0.15]{paper/output/descriptive/map_cities_sample.png}
+    \caption{Distribution of cities in the baseline sample.
+Excludes cities that underwent more than two lifetime switches, and cities without any recorded construction activity.
+Each dot represents a city.
+Some major cities are identified with labels.
+Polygons on the map do \textbf{not} represent historical borders; they merely represent regions by which cities are clustered in the \textit{Städtebuch}.}
+    \label{fig:map_cities_sample}
+\end{figure}
 
@@
 
+\input{paper/output/regressions/SW22_replication_50y}
+
 We can see in Column (1) that switching states has no immediate effect on the probability of recorded construction activity.
@@
 Figure \ref{fig:SW_replication} plots the coefficients in Column (1) to visualise these dynamics.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[scale = 0.3]{paper/output/regressions/SW22_replication_50y.png}
+    \caption{Dynamic effects of switching to another state.
+Coefficient estimates from Table \ref{eq:sw22}, Column (1).
+The outcome variable is the probability that construction activity of any type is observed in the period.
+The period beginning 50 years before the switch is omitted.}
+    \label{fig:SW_replication}
+\end{figure}
 
@@
 The interaction terms are relative to that: the coefficient $\theta_{conquest}$ on $Switch\,to\,another\,state \times Conquest$ is the difference between the treatment effect of switching via succession, and the treatment effect of switching via conquest.
+
+\input{paper/output/regressions/baseline_50y}
 
@@
 
+\input{paper/output/regressions/controls_50y}
+
 Comparing Column (2) to (1), and (4) to (3), we can see that it makes no difference whether we control for switches in the same period or not; the estimates are barely distinguishable.
@@
 When I run the baseline specification on this partial sample, I obtain the exact same estimates as in Column (3) --- even without controlling for conflict.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[scale = 0.15]{paper/output/descriptive/map_conflict_NA_50y.png}
+    \caption{Share of observations from each \textit{Städtebuch} region with non-missing data on conflict incidents.
+I code the $Conflict$ variable as missing for a period if it is earlier than the first recorded incident or later than the last recorded incident in a city.
+Observations with missing conflict data are excluded from the regressions in Columns (3) and (4) of Table \ref{tab:controls_50y}.}
+    \label{fig:conflict_map}
+\end{figure}
 
@@
 
-I conduct a number of other robustness checks, which are summarised in the Appendix.
+I conduct a number of other robustness checks, which are included as an Appendix.
 Table \ref{tab:robustness_50y} contains regressions using different definitions of the sample and treatment.
@@
 
-While the research question may not seem hugely relevant for the present-day Western world, I want to close, briefly, with the Democratic Republic of the Congo (DRC).
-This African state's central government is weak and large parts of its territory are controlled by independent armed groups \citep[p.
+While the research question may not seem hugely relevant for the present-day Western world, I want to conclude by talking, briefly, about the Democratic Republic of the Congo (DRC).
+This African state has a weak central government and large parts of its territory are controlled by independent armed groups that are in many ways like small proto-states.
+\citep[p.
 34]{sierra2020}.
-Recent research has studied the DRC to investigate questions about state formation \citep{sierra2020}, taxation and state capacity \citep{balan2022}, and conflict \citep{henn2021}.
+Recent research has studied state formation \citep{sierra2020}, taxation and state capacity \citep{balan2022}, and conflict \citep{henn2021} in the setting of the DRC.
 One could argue that we are witnessing economic history "in real time" there.\footnote
 {
-This thought is not my own; Davide Cantoni said it to me in conversation.
+This thought is not my own; I got it from Davide Cantoni.
 } 
-My point is that a better understanding of how the European state system evolved may not only deepen our knowledge of the origins of institutions --- it could also inform our perspective on regions of the world that do not have a stable state system at present.
+In this sense a better understanding of how the European state system evolved may not only deepen our knowledge of the origins of institutions --- it could also inform our perspective, and those of policy makers, on regions of the world that do not have a stable state system at present.
 
@@
 \newpage
-\section*{Tables and figures}
 
-\subsection*{Figures}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[scale=0.15]{paper/output/descriptive/map_cities_raw.png}
-    \caption{Distribution of cities in the \textit{Territorial Histories} dataset \citep{pt2}.
-Each dot represents a city.
-Some major cities are identified with labels.
-Polygons on the map do \textbf{not} represent historical borders; they merely represent regions by which cities are clustered in the \textit{Städtebuch}.}
-    \label{fig:map_cities_raw}
-\end{figure}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[scale=0.15]{paper/output/descriptive/map_cities_sample.png}
-    \caption{Distribution of cities in the baseline sample.
-Excludes cities that underwent more than two lifetime switches, and cities without any recorded construction activity.
-Each dot represents a city.
-Some major cities are identified with labels.
-Polygons on the map do \textbf{not} represent historical borders; they merely represent regions by which cities are clustered in the \textit{Städtebuch}.}
-    \label{fig:map_cities_sample}
-\end{figure}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[scale = 0.4]{paper/output/regressions/SW22_replication_50y.png}
-    \caption{Dynamic effects of switching to another state.
-Coefficient estimates from Table \ref{eq:sw22}, Column (1).
-The outcome variable is the probability that construction activity of any type is observed in the period.
-The period beginning 50 years before the switch is omitted.}
-    \label{fig:SW_replication}
-\end{figure}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[scale = 0.15]{paper/output/descriptive/map_conflict_NA_50y.png}
-    \caption{Share of observations from each \textit{Städtebuch} region with non-missing data on conflict incidents.
-I code the $Conflict$ variable as missing for a period if it is earlier than the first recorded incident or later than the last recorded incident in a city.
-Observations with missing conflict data are excluded from the regressions in Columns (3) and (4) of Table \ref{tab:controls_50y}.}
-    \label{fig:conflict_map}
-\end{figure}
-
-\clearpage
-\subsection*{Tables}
-
-\input{paper/output/regressions/SW22_replication_50y}
-
-\input{paper/output/regressions/baseline_50y}
-
-\input{paper/output/regressions/controls_50y}
-
-\clearpage
 \section*{Appendix}

### 2023-06-27 paper/paper.tex
#32 three-author citations

@@
 A longstanding tradition in economics emphasises the role of institutional factors in long-run growth (e.g.
-\cite{north1970}, \cite{delong1993}, \cite{ajr2001}).
+\cite{north1970}, \cite{delong1993}, \cite*{ajr2001}).
 From this vast literature one theme may be especially relevant here: the institution of fiscal capacity, or the state's ability to collect taxes.
@@
 In other words, fiscal capacity and state power are tightly linked --- so if conflict selects for state power, then it also selects for fiscal capacity.
-\cite{gennaioli2015} and \cite{cantoni2023} show that conflict drove state consolidation, with the latter set of authors using the same data as this paper.
+\cite{gennaioli2015} and \cite*{cantoni2023} show that conflict drove state consolidation, with the latter set of authors using the same data as this paper.
 \cite{dincecco2012} show that fiscal capacity had a positive effect on growth.
@@
 However, they do not show explicitly how these improvements in state quality are driven by selection through conflict.
-Their data has a bigger geographical scope (all of Europe) but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is available for fewer cities and fewer years.
+Their data has a bigger geographical scope (all of Europe) but lower resolution: they use urban population from \cite*{bairoch1988} as a proxy for development, which is available for fewer cities and fewer years.
 
@@
 Though ruled by the Holy Roman Emperor at the Imperial level, it consisted of many small and large territories with their own regional rulers.
-Before Germany first became a republic in 1918, it is not helpful to think of "states" in the modern sense: while territories may have existed as political entities on paper, they were ruled by aristocrats who handled power in their familial structures \citep[pp.
+Before Germany first became a republic in 1918, it is not helpful to think of "states" in the modern sense: while territories may have existed as political entities on paper, they were ruled by aristocrats who handled power in their familial structures \citep*[pp.
 2-3]{pt2}.
@@
 I use data from the \textit{Princes and Townspeople} collection which is based on the \textit{Deutsches Städtebuch} \citep{keyser1939} and \cite{kobler2007}'s \textit{Historisches Lexikon der Deutschen Länder}.
-Out of the six datasets currently published in the collection, I use those on territorial histories \citep{pt2}, construction activity \citep{pt5} and conflict incidents \citep{pt6} in my analysis.\footnote{I thank Davide Cantoni and Matthias Weigand for giving me access to unpublished yearly versions of the construction and territorial histories data.} I also use the geo-referenced city locations \citep{pt1} to create visualisations.
+Out of the six datasets currently published in the collection, I use those on territorial histories \citep*{pt2}, construction activity \citep{pt5} and conflict incidents \citep{pt6} in my analysis.\footnote{I thank Davide Cantoni and Matthias Weigand for giving me access to unpublished yearly versions of the construction and territorial histories data.} I also use the geo-referenced city locations \citep*{pt1} to create visualisations.
 
@@
 Figure \ref{fig:map_cities_raw} shows the spatial distribution of the 2,390 cities in the original \textit{Princes and Townspeople} data.
-The reporting boundaries are those of the German Empire in 1937 plus the Free City of Danzig; within those boundaries \cite{keyser1939} claim to list "every settlement that at one point in the history of the Holy Roman Empire was awarded the status of a city" \citep[p.
+The reporting boundaries are those of the German Empire in 1937 plus the Free City of Danzig; within those boundaries \cite{keyser1939} claim to list "every settlement that at one point in the history of the Holy Roman Empire was awarded the status of a city" \citep*[p.
 2]{pt1}.
@@
     \includegraphics[scale=0.15]{paper/output/descriptive/map_cities_raw.png}
-    \caption{Distribution of cities in the \textit{Territorial Histories} dataset \citep{pt2}.
+    \caption{Distribution of cities in the \textit{Territorial Histories} dataset \citep*{pt2}.
 Each dot represents a city.
@@
 
-A city-by-year panel of territorial history, taken from \cite{pt2}, forms the backbone of my dataset.
+A city-by-year panel of territorial history, taken from \cite*{pt2}, forms the backbone of my dataset.
 For each city and each year it identifies the noble lineage to whose territory the city belonged in that year.
@@
 I use construction as a proxy for development because it is available for most cities: out of the 2,390 cities in the \textit{Städtebuch}, 2,177 have at least one construction event linked to them.
-By contrast, only 207 cities from the \textit{Städtebuch} appear in \cite{bairoch1988}'s dataset of city populations.
+By contrast, only 207 cities from the \textit{Städtebuch} appear in \cite*{bairoch1988}'s dataset of city populations.
 
@@
 
-Because the \textit{Städtebuch} data compiled by \cite{pt2} records exactly how each city was acquired by the territory holding it, I am able to estimate heterogeneous effects for different types of ownership changes.\footnote
+Because the \textit{Städtebuch} data compiled by \cite*{pt2} records exactly how each city was acquired by the territory holding it, I am able to estimate heterogeneous effects for different types of ownership changes.\footnote
 {
@@
 It would be a big improvement on my research strategy to find a clearer measure of state power, to check whether measurement error really does bias my estimates downward.
-\cite{cantoni2023} construct a novel dataset of territory-level institutions which one could potentially use for that purpose.
+\cite*{cantoni2023} construct a novel dataset of territory-level institutions which one could potentially use for that purpose.
 Another improvement would be to control for the number of cities switching to another state at the same time in Equation \eqref{eq:baseline}, as described in the Discussion.
 Third, an event study version of Equation \eqref{eq:baseline} may be able to better tease out significant effects if the effect of state power takes some time to phase in.
-Finally, recent literature (\cite{baker2022} provide an overview) has questioned the validity of standard two-way fixed effects regressions with staggered adoption and time-varying treatment --- one could improve my results by using one of the robust estimators proposed by \cite{sunab2021} or \cite{cengiz2019}, for example.
+Finally, recent literature (\cite*{baker2022} provide an overview) has questioned the validity of standard two-way fixed effects regressions with staggered adoption and time-varying treatment --- one could improve my results by using one of the robust estimators proposed by \cite{sunab2021} or \cite{cengiz2019}, for example.
 

### 2023-06-27 paper/paper.tex
#32 final commit after submission

@@
 
-\title{Conflict Driven Evolution and Urban Development: \\ Evidence from the Holy Roman Empire\footnote{
+\title{Conflict-Driven Evolution and Urban Development: \\ Evidence from the Holy Roman Empire\footnote{
 I thank Lara Andermann, Leonie Bielefeld, Daniel Bursian, Davide Cantoni, Emilio Esguerra, Filip Milojevic, Johannes Vorbach and Matthias Weigand for helpful comments and feedback.
@@
 \begin{center}
-    {\centering\huge{{Conflict-Driven Evolution and Urban Development:\\
+    {\centering\Large{{Conflict-Driven Evolution and Urban Development:\\
+
 Evidence from the Holy Roman Empire
@@
 In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development.
-If the effect is positive, then the conflict driven evolution of states may partially explain the eventual ascendancy of Europe: by selecting for military success, Europe's violent history may have accidentally produced fiscal capacity, efficient legal systems, and prospering middle classes.
+If the effect is positive, then the conflict-driven evolution of states may partially explain the eventual ascendancy of Europe: by selecting for military success, Europe's violent history may have accidentally produced fiscal capacity, efficient legal systems, and prospering middle classes.
 
-I build my paper on the theory of conflict driven evolution, outlined above, developed by \cite{levine2013, levine2021, levine2022}.
+I build my paper on the theory of conflict-driven evolution, outlined above, developed by \cite{levine2013, levine2021, levine2022}.
 Its main contribution is to explain why hegemonies arose in East Asia while balances of power could prevail in Europe: unlike China, the major powers of continental Europe were frequently preyed upon by strong outsiders --- the Vikings, the Swedes, and the English --- which prevented any one state from growing too strong (\citealp{levine2021}, p.
@@
 
-\cite{diamond1997} and \cite{landes1969, landes2006} have argued for the economic benefits of political fragmentation and state competition. \cite{diamond1997} asserts that geography 
-remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history.
+\cite{diamond1997} and \cite{landes1969, landes2006} have argued for the economic benefits of political fragmentation and state competition.
+\cite{diamond1997} asserts that the geographic remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history.
 \cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled the spread of innovation in medieval China would not have survived long in a more fragmented state system.
@@
 I justify this assumption with the fact that a conqueror is, by definition, the winner of an armed conflict.
-State power is not the only determinant of winning armed conflicts, though: luck plays a central role, as do many other time-varying factors for which I cannot control.
+State power is not the only determinant of winning armed conflicts, though: luck plays a central role, as do other time-varying factors for which I cannot control.
 On the flip side, winning armed conflicts is also not the only consequence of state power.
 Importantly, the rulers of more powerful states may be able to make more advantageous matches for their children --- dynastic marriages that eventually result in the expansion of their territory.
-Exhibit A in this argument is the House of Habsburg: they became a dominant power in Europe at the turn of the 16th century when strategic marriages made them rulers of Burgundy and Castille, but Habsburgs had been the Holy Roman Emperors for decades prior.
+Exhibit A in this argument is the House of Habsburg: they became a dominant power in Europe at the turn of the 16th century when strategic marriages made them rulers of Burgundy and Castille, but Habsburgs had already been the Holy Roman Emperors for decades prior.
 
@@
 While the research question may not seem hugely relevant for the present-day Western world, I want to conclude by talking, briefly, about the Democratic Republic of the Congo (DRC).
-This African state has a weak central government and large parts of its territory are controlled by independent armed groups that are in many ways like small proto-states.
-\citep[p.
+This African state has a weak central government and large parts of its territory are controlled by independent armed groups that are in many ways like small proto-states \citep[p.
 34]{sierra2020}.

### 2023-06-28 paper/slides.tex
#33 add slides.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass{beamer}

\mode<presentation> 
{

\usetheme{Madrid}

}
\usepackage{graphicx} % we need this so we can add figures
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
\usepackage{natbib} % we need this so we can use citation and bib properly
\usepackage{hyperref} % allows you to add hyperlink
\usepackage{amsmath} % allows you to use most mathematical features
\usepackage{setspace} % allows you to change the line spacing
\usepackage{longtable}
\usepackage{xr}
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
\usepackage{longtable}
\usepackage{threeparttable}
\usepackage{ragged2e}

\title[Kolloquium]{Conflict-Driven Evolution and Urban Development} % The short title appears at the bottom of every slide, the full title is only on the title page
\subtitle{Evidence from the Holy Roman Empire}

\author{Elias Hadj Ammar} % Your name

\institute[LMU] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
LMU Munich\\ % Your institution for the title page
\medskip
\textit{Elias.Ammar@campus.lmu.de} % Your email address
}
\date{August 3, 2023} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

% no table of contents

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
\section{Motivation} % 
%------------------------------------------------

\begin{frame}
\frametitle{Motivation}

\end{frame}

\begin{frame}
\frametitle{Motivation}

\end{frame}

%------------------------------------------------
\section{Outline} % 
%------------------------------------------------

\begin{frame}
\frametitle{Summary:}
\justifying
\begin{itemize}
\item \textbf{Research Question}: It should be a question, not a statement here. If you think the author has asked multiple questions, pick up to three important ones. 
\item \textbf{Main Methods}: Briefly mention their main identification strategy. What data did they use. 
\item \textbf{Main findings}: pick up to three findings. You can use key numbers here if you think they are important. Like this policy has reduced 10\% pollution.(Just an example)
\end{itemize}

\end{frame}

%------------------------------------------------

\begin{frame}
\frametitle{Positive View: }

\begin{block}{Point 3}
Here briefly mention one main positive view
\end{block}

if you want to add graphs, it is the same procedure as when you write your essay. Make sure you adjust the size so your classmates can see it properly.
\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.3\textwidth]{paper/output/descriptive/map_cities_sample.png}
    \caption{The composition of migrants and the provision of common schools; Source: me}.
    \label{table:my_label}
\end{figure}

\end{frame}

%------------------------------------------------

\begin{frame}
\frametitle{Positive View: }
You can also draw a table yourself

\begin{table}
\begin{tabular}{l l l}
\toprule
\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
\midrule
Treatment 1 & 0.0003262 & 0.562 \\
Treatment 2 & 0.0015681 & 0.910 \\
Treatment 3 & 0.0009271 & 0.296 \\
\bottomrule
\end{tabular}
\caption{Table caption}
\end{table}

\end{frame}

%------------------------------------------------
\section{End}% 
%------------------------------------------------

\begin{frame}
\frametitle{References}
\justifying

\bibliography{paper/references}
\bibliographystyle{chicago}

\end{frame}

%------------------------------------------------

\begin{frame}
\Huge{\centerline{The End}}
\end{frame}

%----------------------------------------------------------------------------------------

%------------------------------------------------
\section{Appendix}% 
%------------------------------------------------

\end{document}

### 2023-06-28 source/slides/MapTerrs.R
#33 add terr maps

@@ -16,92 +16,30 @@ Main <- function(){
     select(city_id, region_id, name, nat)
   locs_spatial <- st_read("drive/raw/attributes/city_borders")
   
+  # Create map
+  CleanTerrsMap(1700, terrs, locs_spatial)
   
-  BasicTerrsMapOneYear(1700, terrs, locs_spatial)
+  # Create and save maps at different years
+  time_points <- c(1300, 1500, 1800, 1917)
   
-  # Get snapshot of city ownership in 1500
-  dat1500_sf <- GetGeodataOneYear(1500, terrs, locs_spatial)
-  
-  
-  # Make a spatial dataset of region polygons
-  terrs1500 <- unique(dat1500_sf$terr_id) |> sort() |> `names<-`("terr_id")
-  
-  polygons <- terrs1500 |> 
-    purrr::map(\(id)(TerrIDtoUnion(id, dat1500_sf))) |> 
-    bind_rows()
-  
-  # Bind the cities to region polygons (how about a spatial join?)
-  terrs1500_sf <- bind_cols(terrs1500, polygons) |> 
-    rename(terr_id = ...1) |> 
-    st_as_sf()
-  
-  ### Create heatmap of city counts
-  # Increase the bounding box of the spatial data to leave space for title
-  bbox_new <- BiggerBoundingBox(terrs1500_sf, scale_top = 0.1)
-  
-  # Create basic map
-  map1 <- tm_shape(terrs1500_sf, 
-                   bbox = bbox_new)+
-    tm_polygons(col="terr_id", 
-                title="No. of cities") +
-    tm_layout(title = "Spatial distribution of cities - Full set",
-              title.size = 1.3,
-              title.position = c("center", "top"),
-              title.fontface = 2,
-              legend.title.size = 1.3,
-              legend.text.size = 1)
-  
-  map1
-  
-  # Add an unlabelled dot for each city in the build
-  allpoints_sf <- locs_spatial |> 
-    filter(city_id %in% unique(build$city_id))
-  
-  map2 <- map1 + 
-    tm_shape(allpoints_sf)+
-    tm_dots(alpha = 0.3)
-  
-  
-  # Add some city names as reference points
-  citynames <- c("Muenchen", "Berlin", "Koenigsberg Pr.", "Koeln", 
-                 "Hannover", "Breslau", "Wuerzburg", "Dresden",
[…]

### 2023-06-28 source/slides/TerrMapUtils.R

@@ -6,11 +6,34 @@ TerrIDtoUnion <- function(id, locs_spatial){
 }
 
 
+GetPolygonData<- function(terrs_oneyear){
+  
+  # Make a spatial dataset of region polygons
+  terrs_t <- unique(terrs_oneyear$terr_id) |> sort() |> `names<-`("terr_id")
+  
+  polygons <- terrs_t |> 
+    purrr::map(\(id)(TerrIDtoUnion(id, terrs_oneyear))) |> 
+    bind_rows()
+  
+  # Bind the cities to region polygons
+  terrs_t_sf <- bind_cols(terrs_t, polygons) |> 
+    rename(terr_id = ...1)
+  
+  # Add info on city counts (to later recode singleton terrs)
+  with_city_counts <- terrs_oneyear |> 
+    group_by(terr_id) |> 
+    summarise(n_cities = n_distinct(city_id)) |> 
+    left_join(terrs_t_sf, by="terr_id") |> 
+    st_as_sf()
+  
+  return(with_city_counts)
+}
+
+
 GetGeodataOneYear <- function(t, terrs, locs_spatial){
   terrs_oneyear <- terrs |> 
     filter(year == t) |> 
-    mutate(terr_id = if_else(city_status != 0, "00_FREE", terr_id)) |> 
-    select(city_id, terr_id, city_status) |> 
+    select(city_id, terr_id) |> 
     left_join(locs_spatial, by="city_id") |> 
     st_as_sf()
   

### 2023-06-29 paper/slides.tex
Updates from Overleaf
#33 finish motivation slides

@@
 \usepackage{ragged2e}
+\usepackage{caption}
+\usepackage{subcaption} % for subfigures
 
@@
 
+\begin{itemize}
+    \item Great Divergence: Europe vs. China and the rest
+    \item One difference: balance of power, competition between states
+\end{itemize}
+
+\end{frame}
+
+% the motivation needs to be clear even when the audience doesn't know yet what the question is!
+
+\begin{frame}
+\frametitle{Motivation}
+
+\begin{figure}
+     \centering
+     \begin{subfigure}[b]{0.3\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{paper/output/slides/map_terrs_1300.png}
+         \caption{1300}
+         \label{fig:map1300}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.3\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{paper/output/slides/map_terrs_1800.png}
+         \caption{1800}
+         \label{fig:map1800}
+     \end{subfigure}
+    \hfill
+     \begin{subfigure}[b]{0.3\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{paper/output/slides/map_terrs_1917.png}
+         \caption{1917}
+         \label{fig:map1917}
+     \end{subfigure}
+        \caption{Territorial history of the German Empire}
+        \label{fig:terrmaps}
+\end{figure}
+
 \end{frame}
@@
 \frametitle{Motivation}
+
+\begin{itemize}
+    \item states struggle for resources (=territory)
+    \item some states survive and grow, others get eaten
+   \item survival of the fittest: selection for strength / power\footnote[frame,1]{\cite{levine2021}}
+\end{itemize}
 
@@
 \begin{frame}
-\frametitle{Summary:}
+\frametitle{This paper}
 \justifying
+$\rightarrow$ \textbf{Question}: 
+Did this process play a role in the rise of Europe?
+\medskip
 \begin{itemize}
-\item \textbf{Research Question}: It should be a question, not a statement here. If you think the author has asked multiple questions, pick up to three important ones. 
-\item \textbf{Main Methods}: Briefly mention their main identification strategy. What data did they use. 
-\item \textbf{Main findings}: pick up to three findings. You can use key numbers here if you think they are important. Like this policy has reduced 10\% pollution.(Just an example)
+\item Wars = weak states getting eaten by better ones
+\item Maybe "stronger" states are richer (institutions?) 
+\item Darwinian selection caused state quality in Europe to constantly improve
 \end{itemize}
